### 2008-01-29 trunk/SAiP/Module1/H1/doc/paper.tex
Import from the assignment.

\documentclass[a4paper,10pt]{article}

\usepackage[latin1]{inputenc}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{url}
\usepackage{times}
\usepackage{rotating}

\begin{document}

\title{Exercise 1: Software Architecture Description of the HS07 System}

\author{
  $<<$Names$>>$\\
  Department of Computer Science, University of Aarhus\\
  Aabogade 34, 8200 {\AA}rhus N, Denmark\\
  \makeatletter
  \texttt{<<Group name>>}\\
  \texttt{<<Student IDs>>}\\
  \texttt{\{..., ...\}@daimi.au.dk}
}

\date{$<<$Date: 2007/11/05 22:51:26 $>>$}

\maketitle

% =====================================================================
\begin{abstract}
  The HS07 system implements a closed-loop control of the heating in a
  private home. It monitors thermometers in the home, and based on
  measurements HS07 adjusts radiators in the home. This report gives a
  software architecture description of an architectural prototype of
  the HS07 system. The techniques used for architectural description
  are taken from \cite{christensen2004archdesc}.
\end{abstract}

% =====================================================================
\section{Introduction}

Figure~\ref{fig:hs07} shows a schematic overview of HS07 in a
home. The home may be accessed by the home owner from the outside
through the HS07 gateway. The HS07 gateway also monitors and controls
the home.
\begin{figure}[!htb]
\centerline{\epsfig{figure=figures/hs07,scale=0.4 }}
\caption{HS07 in a home}
\label{fig:hs07}
\end{figure}

HS07 includes sensor and actuator hardware which runs on an embedded Java virtual
machine with standard software.

% =====================================================================
\section{Architectural Requirements}

For our purposes there is one main use case for the HS07 system:
\begin{quote}
  \emph{Control Temperature}: The gateway collects measurements from
  thermometers and reports this to radiators that then control the
  temperature.
\end{quote}

The major driving qualities attributes of the HS07 system
are\footnote{These qualities will be operationalized in Exercise 2}:

\begin{itemize}
\item \emph{Performance.} HS07 should be performant so that a large
  number of thermometers and radiators may be part of the system.
\item \emph{Modifiability.} It must be possible to modify HS07 to
  include new types of sensors and actuators.
  \item $<<$Extra quality requirement that you consider important $>>$
\end{itemize}

% =====================================================================
\section{Architectural Description}

% ---------------------------------------------------------------------
\subsection{Module Viewpoint}

$<<$Describe HS07 from this viewpoint using UML. Where appropriate
provide textual supplements$>>$

% ---------------------------------------------------------------------
\subsection{Component \& Connector Viewpoint}

$<<$Describe HS07 from this viewpoint using UML. Where appropriate
provide textual supplements$>>$

% ---------------------------------------------------------------------
\subsection{Allocation Viewpoint}

$<<$Describe HS07 from this viewpoint using UML. Where appropriate
provide textual supplements$>>$

% =====================================================================
\section{Discussion}
$<<$What are the strengths and limitations of this approach to
architectural description judging from this case?$>>$

$<<$Are there aspects of the software architecture that have not been
properly described?$>>$

$<<$...$>>$

% =====================================================================
\section{...}

% =====================================================================
\bibliography{paper}
\bibliographystyle{apalike}

\end{document}

### 2008-02-02 Module2/chapter/abstract-factory/iteration-0-with-show/src/paystation/domain/ReceiptImpl.java
Adding the exercise source code

@@ -1,0 +1,33 @@
+package paystation.domain;
+
+import java.io.PrintStream;
+import java.util.*;
+
+/** Implementation of Receipt.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class ReceiptImpl implements Receipt {
+  private int value;
+  public ReceiptImpl(int value) { this.value = value; }
+  public int value() { return value;}
+  public void show(PrintStream stream) {
+    String valuestring = ""+value;
+    if ( valuestring.length() == 1 ) { valuestring = "00"+valuestring; }
+    if ( valuestring.length() == 2 ) { valuestring = "0"+valuestring; }
+    Calendar now = GregorianCalendar.getInstance();
+    String hour = ""+now.get(Calendar.HOUR_OF_DAY);
+    if ( hour.length() == 1 ) { hour = "0"+hour; }
+    String min = ""+now.get(Calendar.MINUTE);
+    if ( min.length() == 1 ) { min = "0"+min; }
+    String nowstring = hour+":"+min;
+
+    stream.println("-------------------------------------------------");
+    stream.println("-------  P A R K I N G   R E C E I P T    -------");
+    stream.println("                Value "+valuestring
+                   +" minutes.               ");
+    stream.println("              Car parked at "+nowstring);
+    stream.println("-------------------------------------------------");
+  }
+}

### 2008-02-02 Module2/chapter/abstract-factory/iteration-0/src/paystation/domain/WeekendDecisionStrategy.java

@@ -1,0 +1,10 @@
+package paystation.domain;
+
+/** A strategy for deciding whether it is weekend or not.
+    Author: (c) Henrik Bærbak Christensen 2007
+*/
+
+public interface WeekendDecisionStrategy {
+  public boolean isWeekend();
+}
+

### 2008-02-02 Module2/chapter/abstract-factory/iteration-0/test/paystation/domain/TestAll.java

@@ -1,0 +1,23 @@
+package paystation.domain;
+
+import org.junit.runner.RunWith;
+import org.junit.runners.Suite;
+
+@RunWith(Suite.class)
+  @Suite.SuiteClasses({ TestPayStation.class,
+                          TestLinearRate.class,
+                          TestProgressiveRate.class,
+                          TestAlternatingRateStrategy.class})
+  
+/** Suite to run all test cases for PayStation.
+    Author: (c) Henrik Bærbak Christensen 2007 */
+
+
+public class TestAll {
+  /** This wrapper is only required for running the old JUnit 3.8
+   * graphical user interface on new JUnit 4 test cases */
+  public static junit.framework.Test suite() {
+    return new junit.framework.JUnit4TestAdapter(TestAll.class);
+  } 
+
+}

### 2008-02-02 Module2/chapter/abstract-factory/iteration-1/src/paystation/domain/PayStationFactory.java

@@ -1,0 +1,17 @@
+package paystation.domain;
+
+/** The factory for creating the objects that configure
+    a pay station for the particular town to operate in.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public interface PayStationFactory {
+  /** Create an instance of the rate strategy to use. */
+  public RateStrategy createRateStrategy();
+
+  /** Create an instance of the receipt.
+      @param the number of minutes parking time the receipt is valid for.
+  */
+  public Receipt createReceipt( int parkingTime );
+}

### 2008-02-02 Module2/chapter/abstract-factory/iteration-1/src/paystation/domain/StandardReceipt.java

@@ -1,0 +1,33 @@
+package paystation.domain;
+
+import java.io.PrintStream;
+import java.util.*;
+
+/** Implementation of Receipt.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class StandardReceipt implements Receipt {
+  private int value;
+  public StandardReceipt(int value) { this.value = value; }
+  public int value() { return value;}
+  public void show(PrintStream stream) {
+    String valuestring = ""+value;
+    if ( valuestring.length() == 1 ) { valuestring = "00"+valuestring; }
+    if ( valuestring.length() == 2 ) { valuestring = "0"+valuestring; }
+    Calendar now = GregorianCalendar.getInstance();
+    String hour = ""+now.get(Calendar.HOUR_OF_DAY);
+    if ( hour.length() == 1 ) { hour = "0"+hour; }
+    String min = ""+now.get(Calendar.MINUTE);
+    if ( min.length() == 1 ) { min = "0"+min; }
+    String nowstring = hour+":"+min;
+
+    stream.println("-------------------------------------------------");
+    stream.println("-------  P A R K I N G   R E C E I P T    -------");
+    stream.println("                Value "+valuestring
+                   +" minutes.               ");
+    stream.println("              Car parked at "+nowstring);
+    stream.println("-------------------------------------------------");
+  }
+}

### 2008-02-02 Module2/chapter/abstract-factory/iteration-1/test/paystation/domain/TestPayStation.java

@@ -1,0 +1,124 @@
+package paystation.domain;
+
+import java.io.*;
+
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Pay Station that 
+    are not related to rate calculations.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestPayStation {
+  PayStation ps;
+  /** Fixture for pay station testing. */
+  @Before
+  public void setUp() {
+    PayStationFactory factory = new TestTownFactory();
+    ps = new PayStationImpl( factory );
+  }
+
+  /** Test that all three types of coins are accepted */
+  @Test
+  public void testAllCoins() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    assertEquals( 5+10+25 , ps.readDisplay() ); 
+  }
+  /** Testing for illegal coin entry. */
+  @Test(expected=IllegalCoinException.class)
+  public void testEnterIllegalCoin() throws IllegalCoinException {
+    ps.addPayment(17);
+  }
+  /** Test buying. */
+  @Test public void testBuy() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    Receipt receipt;
+    receipt = ps.buy();
+    assertNotNull( receipt );
+    assertEquals( 5+10+25 , receipt.value() );
+  }
+  /** Test buying for 100 cents */
+  @Test public void testBuy100cent() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    ps.addPayment(25);
+    ps.addPayment(25);
+
+    Receipt receipt;
+    receipt = ps.buy();
+    assertEquals((3*5+10+3*25), receipt.value() );
+  }
+  /** Test receipt */
+  @Test public void testReceiptValue() {
+    Receipt receipt = new StandardReceipt(30);
+    assertEquals( 30, receipt.value() );
+  }
+  /** Test that buy clears for a new transaction */
+  @Test public void testClearingAfterBuy() throws IllegalCoinException {
+    ps.addPayment(25);
+    ps.buy(); // don't care about result
+    assertEquals( 0 , ps.readDisplay() ); 
+  }
+  /** Test that cancel resets the display */
[…]

### 2008-02-02 Module2/chapter/abstract-factory/iteration-2/src/paystation/domain/PayStationImpl.java

@@ -1,0 +1,53 @@
+package paystation.domain;
+
+/** Implementation of the pay station.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+public class PayStationImpl implements PayStation {
+  private int insertedSoFar;
+  private int timeBought;
+
+  /** the strategy for rate calculations */
+  private RateStrategy rateStrategy;
+  /** the factory that defines strategies */
+  private PayStationFactory factory;
+
+  /** Construct a pay station.
+      @param factory the factory to produce strategies and receipts
+  */
+  public PayStationImpl( PayStationFactory factory ) {
+    this.factory = factory;
+    this.rateStrategy = factory.createRateStrategy();
+    resetTransaction();
+  }
+  public void addPayment( int coinValue ) throws IllegalCoinException {
+    switch ( coinValue ) {
+    case 5: 
+    case 10: 
+    case 25: break;  
+    default: 
+      throw new IllegalCoinException("Invalid coin: "+coinValue+" cent.");
+    }
+    insertedSoFar += coinValue;
+    timeBought = rateStrategy.calculateTime(insertedSoFar);
+  }
+  public int readDisplay() {
+    return timeBought;
+  }
+  public Receipt buy() {
+    Receipt r = factory.createReceipt(timeBought);
+    resetTransaction();
+    return r;
+  }
+  public void cancel() {
+    resetTransaction();
+  }
+  public int timeBought() {
+    return timeBought;
+  }
+
+  /** resets the present transaction */
+  private void resetTransaction() {
+    timeBought = insertedSoFar = 0;
+  }    
+}
+

### 2008-02-02 Module2/chapter/abstract-factory/iteration-2/src/paystation/domain/Receipt.java

@@ -1,0 +1,21 @@
+package paystation.domain;
+
+import java.io.PrintStream;
+
+/** This interface models the receipt that the customer receives from
+    the pay station.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public interface Receipt {
+  /** return the number of minutes this receipt is valid for.
+   *  @return number of minutes parking time
+   */
+  int value();
+  
+  /** prints the receipt on the given PrintStream.
+   */
+  void show(PrintStream stream);
+}
+

### 2008-02-02 Module2/chapter/abstract-factory/iteration-2/src/paystation/domain/StandardReceipt.java

@@ -1,0 +1,46 @@
+package paystation.domain;
+
+import java.io.PrintStream;
+import java.util.*;
+
+/** Implementation of Receipt. 
+    Paramerized via a special constructor to enable
+    (fake) bar code printing.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class StandardReceipt implements Receipt {
+  private int value;
+  private boolean doPrintBarCode;
+  /** Create a receipt.
+   * @param value the number of minutes this receipt is valid for.
+   * @param doPrintBarCode if true then an additional bar code
+   * is printed on the receipt 
+   */
+  public StandardReceipt(int value, boolean doPrintBarCode) 
+	{ this.value = value; this.doPrintBarCode = doPrintBarCode; }
+  public StandardReceipt(int value) { this(value, false); }
+  public int value() { return value;}
+  public void show(PrintStream stream) {
+    String valuestring = ""+value;
+    if ( valuestring.length() == 1 ) { valuestring = "00"+valuestring; }
+    if ( valuestring.length() == 2 ) { valuestring = "0"+valuestring; }
+    Calendar now = GregorianCalendar.getInstance();
+    String hour = ""+now.get(Calendar.HOUR_OF_DAY);
+    if ( hour.length() == 1 ) { hour = "0"+hour; }
+    String min = ""+now.get(Calendar.MINUTE);
+    if ( min.length() == 1 ) { min = "0"+min; }
+    String nowstring = hour+":"+min;
+
+    stream.println("-------------------------------------------------");
+    stream.println("-------  P A R K I N G   R E C E I P T    -------");
+    stream.println("                Value "+valuestring
+                   +" minutes.               ");
+    stream.println("              Car parked at "+nowstring);
+    if ( doPrintBarCode ) {
+      stream.println( "||  ||| | || || |||| || ||| | || | | |||| | || |");
+    }
+    stream.println("-------------------------------------------------");
+  }
+}

### 2008-02-02 Module2/chapter/abstract-factory/iteration-3/src/paystation/domain/AlphaTownFactory.java

@@ -1,0 +1,15 @@
+package paystation.domain;
+
+/** The factory for AlphaTown.
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class AlphaTownFactory implements PayStationFactory {
+  public RateStrategy createRateStrategy() {
+    return new LinearRateStrategy();
+  }
+
+  public Receipt createReceipt( int parkingTime ) {
+    return new StandardReceipt(parkingTime);
+  }
+}

### 2008-02-02 Module2/chapter/abstract-factory/iteration-3/test/paystation/domain/TestAll.java

@@ -1,0 +1,24 @@
+package paystation.domain;
+
+import org.junit.runner.RunWith;
+import org.junit.runners.Suite;
+
+@RunWith(Suite.class)
+  @Suite.SuiteClasses({ TestPayStation.class,
+                          TestConfiguration.class,
+                          TestLinearRate.class,
+                          TestProgressiveRate.class,
+                          TestAlternatingRateStrategy.class})
+  
+/** Suite to run all test cases for PayStation.
+    Author: (c) Henrik Bærbak Christensen 2007 */
+
+
+public class TestAll {
+  /** This wrapper is only required for running the old JUnit 3.8
+   * graphical user interface on new JUnit 4 test cases */
+  public static junit.framework.Test suite() {
+    return new junit.framework.JUnit4TestAdapter(TestAll.class);
+  } 
+
+}

### 2008-02-02 Module2/chapter/strategy/compositional/iteration-0/Receipt.java

@@ -1,0 +1,18 @@
+/** This interface models the receipt that the customer receives from
+    the pay station.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public interface Receipt {
+  /** return the number of minutes this receipt is valid for.
+   *  @return number of minutes parking time
+   */
+  int value();
+  
+  /** "shows" the receipt by writing text to the 
+   *  shell/prompt.
+   */
+  void show();
+}
+

### 2008-02-02 Module2/chapter/strategy/compositional/iteration-2/TestProgressiveRate.java

@@ -1,0 +1,25 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the BetaTown pay station with progressive rate policy.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestProgressiveRate {
+  PayStation ps;
+  /** Fixture for pay station testing. */
+  @Before
+  public void setUp() {
+    ps = new PayStationImpl(new LinearRateStrategy());
+  }
+
+  /** Test a single hour parking */
+  @Test public void testOneHour() throws IllegalCoinException { 
+    // First hour: $1.5
+    ps.addPayment( 25 ); ps.addPayment( 25 );
+    ps.addPayment( 25 ); ps.addPayment( 25 ); 
+
+    ps.addPayment( 25 ); ps.addPayment( 25 ); 
+
+    assertEquals( 60 /*minutes*/, ps.readDisplay() ); 
+  }
+}

### 2008-02-02 Module2/chapter/strategy/compositional/iteration-3/TestProgressiveRate.java

@@ -1,0 +1,41 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the BetaTown pay station with progressive rate policy.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestProgressiveRate {
+  PayStation ps;
+  /** Fixture for pay station testing. */
+  @Before
+  public void setUp() {
+    ps = new PayStationImpl(new ProgressiveRateStrategy());
+  }
+
+  /** Test a single hour parking */
+  @Test public void testOneHour() throws IllegalCoinException { 
+    // First hour: $1.5
+    addOneDollar();
+    addHalfDollar();
+
+    assertEquals( 60 /*minutes*/, ps.readDisplay() ); 
+  }
+ 
+  /** Test two hours parking */
+  @Test public void testTwoHour() throws IllegalCoinException { 
+    // Two hours: $1.5+2.0
+    addOneDollar();
+    addOneDollar();
+    addOneDollar();
+    addHalfDollar();
+
+    assertEquals( 2 * 60 /*minutes*/ , ps.readDisplay() ); 
+  }
+ 
+  private void addHalfDollar() throws IllegalCoinException  {
+    ps.addPayment( 25 ); ps.addPayment( 25 ); 
+  }
+  private void addOneDollar() throws IllegalCoinException {
+    addHalfDollar(); addHalfDollar();
+  }
+}

### 2008-02-02 Module2/chapter/strategy/compositional/iteration-last/TestProgressiveRate.java

@@ -1,0 +1,86 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the BetaTown pay station with progressive rate policy.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestProgressiveRate {
+  PayStation ps;
+  /** Fixture for pay station testing. */
+  @Before
+  public void setUp() {
+    ps = new PayStationImpl(new ProgressiveRateStrategy());
+  }
+
+  /** Test a single hour parking */
+  @Test public void testOneHour() throws IllegalCoinException { 
+    // First hour: $1.5
+    addHalfDollar();
+    addOneDollar();
+
+    assertEquals( 60 /*minutes*/, ps.readDisplay() ); 
+  }
+   /** Test a 1.5 hour parking */
+  @Test public void testOneHourHalf() throws IllegalCoinException { 
+    // First hour: $1.5, next half hour = $1
+    addTwoDollar();
+    addHalfDollar();
+      
+    assertEquals( 60 + 30 /*minutes*/, ps.readDisplay() ); 
+  }
+  /** Test two hours parking */
+  @Test public void testTwoHour() throws IllegalCoinException { 
+    // Two hours: $1.5+2.0
+    addTwoDollar();
+    addOneDollar();
+    addHalfDollar();
+
+    assertEquals( 2 * 60 /*minutes*/ , ps.readDisplay() ); 
+  }
+  /** Test two and a half hour parking */
+  @Test public void testTwoHourHalf() throws IllegalCoinException { 
+    // Two and a half hours: $1.5+2.0+ (3.0/2)
+    
+    // $1.5 = 1st hour
+    addOneDollar(); addHalfDollar();
+    // $2.0 = 2nd hour
+    addTwoDollar();
+    // $1.5 = half hour in 2-3 hour interval
+    addOneDollar(); addHalfDollar();
+    
+    assertEquals( 2 * 60 + 30 /*minutes*/ , ps.readDisplay() ); 
+  }
+  /** Test three hours */
+  @Test public void testThreeHour() throws IllegalCoinException { 
+    // Three hours: $1.5+2.0+3.0 = $6.5
+    addFiveDollar();
+    addOneDollar();
+    addHalfDollar();
+    
+    assertEquals( 3 * 60 /*minutes*/ , ps.readDisplay() ); 
+  }
+  /** Test four hours */
[…]

### 2008-02-02 Module2/chapter/strategy/compositional/separate-testing/One2OneRateStrategy.java

@@ -1,0 +1,13 @@
+/** A one-cent-one-minute rate calculation strategy to
+    simplify testing the responsibilities of the pay
+    station that are not related to rate calculation.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class One2OneRateStrategy implements RateStrategy {
+  public int calculateTime( int amount ) {
+    return amount;
+  }
+}
+

### 2008-02-02 Module2/chapter/strategy/compositional/separate-testing/TestPayStation.java

@@ -1,0 +1,77 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Pay Station that 
+    are not related to rate calculations.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestPayStation {
+  PayStation ps;
+  /** Fixture for pay station testing. */
+  @Before
+  public void setUp() {
+    ps = new PayStationImpl(new One2OneRateStrategy());
+  }
+
+  /** Test that all three types of coins are accepted */
+  @Test
+  public void testAllCoins() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    assertEquals( 5+10+25 , ps.readDisplay() ); 
+  }
+  /** Testing for illegal coin entry. */
+  @Test(expected=IllegalCoinException.class)
+  public void testEnterIllegalCoin() throws IllegalCoinException {
+    ps.addPayment(17);
+  }
+  /** Test buying. */
+  @Test public void testBuy() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    Receipt receipt;
+    receipt = ps.buy();
+    assertNotNull( receipt );
+    assertEquals( 5+10+25 , receipt.value() );
+  }
+  /** Test buying for 100 cents */
+  @Test public void testBuy100cent() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    ps.addPayment(25);
+    ps.addPayment(25);
+
+    Receipt receipt;
+    receipt = ps.buy();
+    assertEquals((3*5+10+3*25), receipt.value() );
+  }
+  /** Test receipt */
+  @Test public void testReceiptValue() {
+    Receipt receipt = new ReceiptImpl(30);
+    assertEquals( 30, receipt.value() );
+  }
+  /** Test that buy clears for a new transaction */
+  @Test public void testClearingAfterBuy() throws IllegalCoinException {
+    ps.addPayment(25);
+    ps.buy(); // don't care about result
+    assertEquals( 0 , ps.readDisplay() ); 
+  }
+  /** Test that cancel resets the display */
+  @Test public void testCancel() throws IllegalCoinException {
+    ps.addPayment(25);
[…]

### 2008-02-02 Module2/chapter/strategy/parametric/IllegalCoinException.java

@@ -1,0 +1,8 @@
+/** Instances of this class represents an exception raised
+    when an illegal coin is encountered in pay station.
+    
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+public class IllegalCoinException extends Exception {
+  public IllegalCoinException( String e ) { super(e); }
+}

### 2008-02-02 Module2/chapter/strategy/parametric/TestAlphaTown.java

@@ -1,0 +1,81 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Pay Station system.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestAlphaTown {
+  PayStation ps;
+  /** Fixture for pay station testing. */
+  @Before
+  public void setUp() {
+    ps = new PayStationImpl(PayStationImpl.Town.ALPHATOWN);
+  }
+
+  /** Testing that a nickel gives two minutes parking time */
+  @Test
+  public void testEnterNickel() throws IllegalCoinException {
+    ps.addPayment( 5 );
+    assertEquals( 2, ps.readDisplay() ); 
+  }
+  /** Testing that a quarter gives ten minutes parking time */
+  @Test
+  public void testEnterQuarter() throws IllegalCoinException {
+    ps.addPayment( 25 );
+    assertEquals( 10, ps.readDisplay() ); 
+  }
+  /** Testing for illegal coin entry. */
+  @Test(expected=IllegalCoinException.class)
+  public void testEnterIllegalCoin() throws IllegalCoinException {
+    ps.addPayment(17);
+  }
+  /** Test buying. */
+  @Test public void testBuy() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    Receipt receipt;
+    receipt = ps.buy();
+    assertNotNull( receipt );
+    assertEquals((5+10+25) / 5 * 2 , receipt.value() );
+  }
+  /** Test buying for 100 cents */
+  @Test public void testBuy100cent() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    ps.addPayment(25);
+    ps.addPayment(25);
+
+    Receipt receipt;
+    receipt = ps.buy();
+    assertEquals((3*5+10+3*25) / 5 * 2 , receipt.value() );
+  }
+  /** Test receipt */
+  @Test public void testReceiptValue() {
+    Receipt receipt = new ReceiptImpl(30);
+    assertEquals( 30, receipt.value() );
+  }
+  /** Test that buy clears for a new transaction */
+  @Test public void testClearingAfterBuy() throws IllegalCoinException {
+    ps.addPayment(25);
+    ps.buy(); // don't care about result
[…]

### 2008-02-02 Module2/chapter/strategy/polymorphic/TestBetaTown.java

@@ -1,0 +1,86 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Pay Station system.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestBetaTown {
+  PayStation ps;
+  /** Fixture for pay station testing. */
+  @Before
+  public void setUp() {
+    ps = new PayStationProgressiveRate();
+  }
+
+  /** Test a single hour parking */
+  @Test public void testOneHour() throws IllegalCoinException { 
+    // First hour: $1.5
+    addHalfDollar();
+    addOneDollar();
+
+    assertEquals( 60 /*minutes*/, ps.readDisplay() ); 
+  }
+   /** Test a 1.5 hour parking */
+  @Test public void testOneHourHalf() throws IllegalCoinException { 
+    // First hour: $1.5, next half hour = $1
+    addTwoDollar();
+    addHalfDollar();
+      
+    assertEquals( 60 + 30 /*minutes*/, ps.readDisplay() ); 
+  }
+  /** Test two hours parking */
+  @Test public void testTwoHour() throws IllegalCoinException { 
+    // Two hours: $1.5+2.0
+    addTwoDollar();
+    addOneDollar();
+    addHalfDollar();
+
+    assertEquals( 2 * 60 /*minutes*/ , ps.readDisplay() ); 
+  }
+  /** Test two and a half hour parking */
+  @Test public void testTwoHourHalf() throws IllegalCoinException { 
+    // Two and a half hours: $1.5+2.0+ (3.0/2)
+    
+    // $1.5 = 1st hour
+    addOneDollar(); addHalfDollar();
+    // $2.0 = 2nd hour
+    addTwoDollar();
+    // $1.5 = half hour in 2-3 hour interval
+    addOneDollar(); addHalfDollar();
+    
+    assertEquals( 2 * 60 + 30 /*minutes*/ , ps.readDisplay() ); 
+  }
+  /** Test three hours */
+  @Test public void testThreeHour() throws IllegalCoinException { 
+    // Three hours: $1.5+2.0+3.0 = $6.5
+    addFiveDollar();
+    addOneDollar();
+    addHalfDollar();
+    
+    assertEquals( 3 * 60 /*minutes*/ , ps.readDisplay() ); 
+  }
+  /** Test four hours */
+  @Test public void testFourHour() throws IllegalCoinException { 
[…]

### 2008-02-02 Module2/chapter/tdd/iteration-0/PayStationImpl.java

@@ -1,0 +1,26 @@
+/** Implementation of the pay station.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class PayStationImpl implements PayStation {
+
+  public void addPayment( int coinValue ) throws IllegalCoinException {
+  }
+  
+  public int readDisplay() {
+    return 0;
+  }
+  
+  public Receipt buy() {
+    return null;
+  }
+
+  public void cancel() {
+  }
+  
+  public int timeBought() {
+    return 0;
+  }
+}
+

### 2008-02-02 Module2/chapter/tdd/iteration-0/PlainTest.java

@@ -1,0 +1,21 @@
+/** A testing tool written from scratch.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class PlainTest {
+  public static void main(String[] args) {
+    PayStation ps;
+    ps = new PayStationImpl();
+    // Test entering 5 cents
+    try {
+      ps.addPayment( 5 );
+    } catch ( IllegalCoinException e ) {
+      System.out.println("Pay station did not accept valid coin");
+    }
+    if ( ps.readDisplay() != 2 ) {
+      System.out.println("5 cents does not give 2 min parking time.");
+    }
+    // ... fill in more tests
+  }
+}

### 2008-02-02 Module2/chapter/tdd/iteration-2/TestPayStation.java

@@ -1,0 +1,33 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Pay Station system.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestPayStation {
+  PayStation ps;
+  /** Fixture for pay station testing. */
+  @Before
+  public void setUp() {
+    ps = new PayStationImpl();
+  }
+
+  /** Testing that a nickel gives two minutes parking time */
+  @Test
+  public void testEnterNickel() throws IllegalCoinException {
+    ps.addPayment( 5 );
+    assertEquals( 2, ps.readDisplay() ); 
+  }
+  /** Testing that a quarter gives ten minutes parking time */
+  @Test
+  public void testEnterQuarter() throws IllegalCoinException {
+    ps.addPayment( 25 );
+    assertEquals( 10, ps.readDisplay() ); 
+  }
+
+  /** This wrapper is only required for running the old JUnit 3.8
+   * graphical user interface on new JUnit 4 test cases */
+  public static junit.framework.Test suite() {
+    return new junit.framework.JUnit4TestAdapter(TestPayStation.class);
+  } 
+}

### 2008-02-02 Module2/chapter/tdd/iteration-3/PayStationImpl.java

@@ -1,0 +1,27 @@
+/** Implementation of the pay station.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+public class PayStationImpl implements PayStation {
+  private int insertedSoFar;
+  public void addPayment( int coinValue ) throws IllegalCoinException {
+    switch ( coinValue ) {
+    case 5: 
+    case 10: 
+    case 25: break;  
+    default: 
+      throw new IllegalCoinException("Invalid coin: "+coinValue+" cent.");
+    }
+    insertedSoFar += coinValue;
+  }
+  public int readDisplay() {
+    return insertedSoFar / 5 * 2;
+  }
+  public Receipt buy() {
+    return null;
+  }
+  public void cancel() {
+  }
+  public int timeBought() {
+    return 0;
+  }
+}
+

### 2008-02-02 Module2/chapter/tdd/iteration-7/TestPayStation.java

@@ -1,0 +1,73 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Pay Station system.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestPayStation {
+  PayStation ps;
+  /** Fixture for pay station testing. */
+  @Before
+  public void setUp() {
+    ps = new PayStationImpl();
+  }
+
+  /** Testing that a nickel gives two minutes parking time */
+  @Test
+  public void testEnterNickel() throws IllegalCoinException {
+    ps.addPayment( 5 );
+    assertEquals( 2, ps.readDisplay() ); 
+  }
+  /** Testing that a quarter gives ten minutes parking time */
+  @Test
+  public void testEnterQuarter() throws IllegalCoinException {
+    ps.addPayment( 25 );
+    assertEquals( 10, ps.readDisplay() ); 
+  }
+  /** Testing for illegal coin entry. */
+  @Test(expected=IllegalCoinException.class)
+  public void testEnterIllegalCoin() throws IllegalCoinException {
+    ps.addPayment(17);
+  }
+  /** Test buying. */
+  @Test public void testBuy() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    Receipt receipt;
+    receipt = ps.buy();
+    assertNotNull( receipt );
+    assertEquals((5+10+25) / 5 * 2 , receipt.value() );
+  }
+  /** Test buying for 100 cents */
+  @Test public void testBuy100cent() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    ps.addPayment(25);
+    ps.addPayment(25);
+
+    Receipt receipt;
+    receipt = ps.buy();
+    assertEquals((3*5+10+3*25) / 5 * 2 , receipt.value() );
+  }
+  /** Test receipt */
+  @Test public void testReceiptValue() {
+    Receipt receipt = new ReceiptImpl(30);
+    assertEquals( 30, receipt.value() );
+  }
+  /** Test that buy clears for a new transaction */
+  @Test public void testClearing() throws IllegalCoinException {
+    ps.addPayment(25);
+    ps.buy(); // don't care about result
+    assertEquals( 0 , ps.readDisplay() ); 
+  }
+  
[…]

### 2008-02-02 Module2/chapter/tdd/iteration-last/PayStation.java

@@ -1,0 +1,47 @@
+/** The business logic of a Parking Pay Station.
+
+    Responsibilities:
+
+    1) Accept payment;
+    2) Calculate parking time based on payment;
+    3) Know earning, parking time bought;
+    4) Issue receipts;
+    5) Handle buy and cancel events.
+
+    The interface also acts as a Facade of the pay station
+    domain handling system with regards to the graphical
+    user interface / hardware side of the station.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public interface PayStation {
+  /** Insert coin into the pay station and adjusts state accordingly.
+   *  @param coinValue is an integer value representing the coin in
+   *  cent. That is a quarter is coinValue=25, etc.
+   */
+  public void addPayment( int coinValue ) throws IllegalCoinException;
+  
+  /** Read the machine's display showing the amount of parking time
+   *  bought so far based on inserted payment.
+   * @return the number to display on the pay station display
+   */
+  public int readDisplay();
+  
+  /** Buy parking time. Terminate the ongoing transaction and
+   * return a parking receipt. A non-null object is always returned.
+   * @return a valid parking receipt object. 
+   */
+  public Receipt buy();
+
+  /** Cancel the present transaction. Resets the machine for a
+   *  new transaction. 
+   */
+  public void cancel();
+  
+  /** Return the time bought so far in the present transaction.
+   * @return time bought so far in minutes.
+   */
+  public int timeBought();
+}
+

### 2008-02-02 Module2/chapter/tdd/iteration-last/PayStationImpl.java

@@ -1,0 +1,41 @@
+/** Implementation of the pay station.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+public class PayStationImpl implements PayStation {
+  private int insertedSoFar;
+  private int timeBought;
+
+  public PayStationImpl() {
+    resetTransaction();
+  }
+  public void addPayment( int coinValue ) throws IllegalCoinException {
+    switch ( coinValue ) {
+    case 5: 
+    case 10: 
+    case 25: break;  
+    default: 
+      throw new IllegalCoinException("Invalid coin: "+coinValue+" cent.");
+    }
+    insertedSoFar += coinValue;
+    timeBought = insertedSoFar / 5 * 2;
+  }
+  public int readDisplay() {
+    return timeBought;
+  }
+  public Receipt buy() {
+    Receipt r = new ReceiptImpl(timeBought);
+    resetTransaction();
+    return r;
+  }
+  public void cancel() {
+    resetTransaction();
+  }
+  public int timeBought() {
+    return timeBought;
+  }
+
+  /** resets the present transaction */
+  private void resetTransaction() {
+    timeBought = insertedSoFar = 0;
+  }    
+}
+

### 2008-02-02 Module2/chapter/tdd/iteration-last/TestPayStation.java

@@ -1,0 +1,87 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Pay Station system.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestPayStation {
+  PayStation ps;
+  /** Fixture for pay station testing. */
+  @Before
+  public void setUp() {
+    ps = new PayStationImpl();
+  }
+
+  /** Testing that a nickel gives two minutes parking time */
+  @Test
+  public void testEnterNickel() throws IllegalCoinException {
+    ps.addPayment( 5 );
+    assertEquals( 2, ps.readDisplay() ); 
+  }
+  /** Testing that a quarter gives ten minutes parking time */
+  @Test
+  public void testEnterQuarter() throws IllegalCoinException {
+    ps.addPayment( 25 );
+    assertEquals( 10, ps.readDisplay() ); 
+  }
+  /** Testing for illegal coin entry. */
+  @Test(expected=IllegalCoinException.class)
+  public void testEnterIllegalCoin() throws IllegalCoinException {
+    ps.addPayment(17);
+  }
+  /** Test buying. */
+  @Test public void testBuy() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    Receipt receipt;
+    receipt = ps.buy();
+    assertNotNull( receipt );
+    assertEquals((5+10+25) / 5 * 2 , receipt.value() );
+  }
+  /** Test buying for 100 cents */
+  @Test public void testBuy100cent() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    ps.addPayment(25);
+    ps.addPayment(25);
+
+    Receipt receipt;
+    receipt = ps.buy();
+    assertEquals((3*5+10+3*25) / 5 * 2 , receipt.value() );
+  }
+  /** Test receipt */
+  @Test public void testReceiptValue() {
+    Receipt receipt = new ReceiptImpl(30);
+    assertEquals( 30, receipt.value() );
+  }
+  /** Test that buy clears for a new transaction */
+  @Test public void testClearingAfterBuy() throws IllegalCoinException {
+    ps.addPayment(25);
+    ps.buy(); // don't care about result
+    assertEquals( 0 , ps.readDisplay() ); 
+  }
[…]

### 2008-02-02 Module2/chapter/test-stub/iteration-0/src/paystation/domain/Receipt.java

@@ -1,0 +1,20 @@
+package paystation.domain;
+
+/** This interface models the receipt that the customer receives from
+    the pay station.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public interface Receipt {
+  /** return the number of minutes this receipt is valid for.
+   *  @return number of minutes parking time
+   */
+  int value();
+  
+  /** "shows" the receipt by writing text to the 
+   *  shell/prompt.
+   */
+  void show();
+}
+

### 2008-02-02 Module2/chapter/test-stub/iteration-1/test/paystation/domain/TestGammaWeekday.java

@@ -1,0 +1,22 @@
+package paystation.domain;
+
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Gammatown rate strategy on a weekday.
+    Author: (c) Henrik Bærbak Christensen 2007 */
+
+public class TestGammaWeekday {
+  RateStrategy rs;
+  @Before
+  public void setUp() {
+    rs = new AlternatingRateStrategy( new LinearRateStrategy(),
+                                      new ProgressiveRateStrategy(),
+                                      new ClockBasedDecisionStrategy() );
+  }
+
+  /** Test for 500 cent */
+  @Test public void test500cent() { 
+    assertEquals( 500 / 5 * 2 /*minutes*/ , rs.calculateTime(500) ); 
+  }
+}

### 2008-02-02 Module2/chapter/test-stub/iteration-1/test/paystation/domain/TestGammaWeekend.java

@@ -1,0 +1,22 @@
+package paystation.domain;
+
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Gammatown rate strategy on a weekend.
+    Author: (c) Henrik Bærbak Christensen 2007 */
+
+public class TestGammaWeekend {
+  RateStrategy rs;
+  @Before
+  public void setUp() {
+    rs = new AlternatingRateStrategy( new LinearRateStrategy(),
+                                      new ProgressiveRateStrategy(),
+                                      new ClockBasedDecisionStrategy() );
+  }
+
+  /** Test for 500 cent */
+  @Test public void test500cent() { 
+    assertEquals( 2 * 60 + 30 /*minutes*/ , rs.calculateTime(500) ); 
+  }
+}

### 2008-02-02 Module2/chapter/test-stub/iteration-2/src/paystation/domain/AlternatingRateStrategy.java

@@ -1,0 +1,29 @@
+package paystation.domain;
+
+/** An alternating rate calculation strategy that
+    uses two rate strategies, one for weekdays and
+    another for weekends.
+
+    @author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class AlternatingRateStrategy implements RateStrategy {
+  RateStrategy weekendStrategy, weekdayStrategy, currentState;
+  WeekendDecisionStrategy decisionStrategy;
+  public AlternatingRateStrategy( RateStrategy weekdayStrategy,
+                                  RateStrategy weekendStrategy,
+                                  WeekendDecisionStrategy decisionStrategy ) {
+    this.weekdayStrategy = weekdayStrategy;
+    this.weekendStrategy = weekendStrategy;
+    this.decisionStrategy = decisionStrategy;
+    this.currentState = null;
+  }
+  public int calculateTime( int amount ) {
+    if ( decisionStrategy.isWeekend() ) {
+      currentState = weekendStrategy; 
+    } else {
+      currentState = weekdayStrategy;
+    }
+    return currentState.calculateTime( amount );
+  }
+}

### 2008-02-02 Module2/exercise/abstract-factory/DoorAlarm/src/dooralarm/domain/MessageDatabase.java

@@ -1,0 +1,13 @@
+package dooralarm.domain;
+
+/** MessageDatabase is a database of messages in a particular language.
+
+    Author: Henrik Bærbak Christensen
+*/
+
+public interface MessageDatabase {
+  public String getFrameTitle();
+  public String getWelcomeMessage();
+  public String getDeniedMessage();
+  public String getGrantedMessage();
+}

### 2008-02-02 Module2/exercise/compositional/PayStation/src/paystation/domain/AlphaTownFactory.java

@@ -1,0 +1,16 @@
+package paystation.domain;
+
+/** Factory for Alphatown.
+
+    Author: (c) Henrik Bærbak Christensen 2006.
+*/
+
+public class AlphaTownFactory implements PayStationFactory {
+  public Receipt createReceipt( int parkingTime ) { 
+    return new ReceiptImpl(parkingTime); 
+  }
+  public RateStrategy createRateStrategy() { 
+    return new LinearRateStrategy(); 
+  }
+}
+

### 2008-02-02 Module2/exercise/compositional/PayStation/src/paystation/domain/BarCodeReceipt.java

@@ -1,0 +1,31 @@
+package paystation.domain;
+
+import java.util.*;
+
+/** Implementation of the receipt with a bar code.
+
+    The bar code is faked in this implementation.
+
+    author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class BarCodeReceipt implements Receipt {
+  private int value;
+  public BarCodeReceipt(int value) { this.value = value; }
+  public int value() { return value;}
+  public void show() {
+    System.out.println( "-------------------------------------------------");
+    System.out.println( "-------  P A R K I N G   R E C E I P T  ---------");
+    System.out.println( "    Value "+ value+" minutes." );
+    Calendar now = GregorianCalendar.getInstance();
+    now.add(Calendar.MINUTE, value);
+    String end = 
+      "" + now.get(Calendar.HOUR_OF_DAY) + 
+      ":" + now.get(Calendar.MINUTE);
+    System.out.println( "    Valid until "+ end );
+    // A "fake-it" bar code on the receipt
+    System.out.println( "||  ||| | || | | || | ||| ||||| | ||| ||  || | ||");
+    System.out.println( "-------------------------------------------------");
+  }
+}
+

### 2008-02-02 Module2/exercise/compositional/PayStation/src/paystation/domain/GammaTownFactory.java

@@ -1,0 +1,19 @@
+package paystation.domain;
+
+/** Factory for Gammatown.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class GammaTownFactory implements PayStationFactory {
+  public Receipt createReceipt( int parkingTime ) { 
+    return new ReceiptImpl(parkingTime); 
+  }
+  public RateStrategy createRateStrategy() { 
+    RateStrategy rs = 
+      new CompositeRateStrategy( new LinearRateStrategy(),
+                                 new ProgressiveRateStrategy() );
+    return rs; 
+  }
+}
+

### 2008-02-02 Module2/exercise/compositional/PayStation/src/paystation/domain/PayStation.java

@@ -1,0 +1,48 @@
+package paystation.domain;
+
+/** The business logic of a Parking Pay Station.
+
+    Responsibilities:
+
+    1) Accept payment;
+    2) Calculate parking time based on payment;
+    3) Know earning, parking time bought;
+    4) Issue receipts;
+    5) Handle buy and cancel transactions.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public interface PayStation {
+  /** Insert coin into the pay station and adjusts state accordingly.
+     @param coinValue is an integer value representing the coin in
+     cent. That is a quarter is coinValue=25, etc.
+  */
+  public void addPayment( int coinValue ) throws IllegalCoinException;
+    
+  /** Read the machine's display showing the amount of parking time
+     bought so far based on inserted payment.
+     @return the number to display on the pay station display
+  */
+  public int readDisplay();
+  
+  /** Return a parking receipt.
+     @return a valid parking receipt object. 
+  */
+  public Receipt pushBuyButton();
+
+  /** Cancel the present transaction. Resets the machine for a
+      new transaction. 
+  */
+  public void pushCancelButton();
+    
+  /** Return the amount of payment entered since last call of
+      empty().
+  */
+  public int amountEarned();
+
+  /** Empty the machine. The amount earned is reset */
+  public void empty();
+
+}
+

### 2008-02-02 Module2/exercise/compositional/PayStation/src/paystation/domain/PayStationImpl.java

@@ -1,0 +1,53 @@
+package paystation.domain;
+
+/** Implementation of a Parking Pay Station.
+
+    author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class PayStationImpl implements PayStation {
+  private int insertedSoFar;
+  private int timeBought;
+  
+  /** the strategy for rate calculations */
+  private RateStrategy rateStrategy;
+  /** the factory that defines strategies */
+  private PayStationFactory factory;
+
+  /** Construct a pay station.
+      @param factory the factory to produce 
+      strategies and receipts
+  */
+  public PayStationImpl( PayStationFactory factory ) {
+    this.factory = factory;
+    this.rateStrategy = factory.createRateStrategy();
+  }
+  public void addPayment( int coinValue ) throws IllegalCoinException {
+    switch ( coinValue ) {
+    case 5: 
+    case 10: 
+    case 25: break;
+    default: 
+      throw new IllegalCoinException("Invalid coin: "+coinValue+" cent.");
+    }
+    insertedSoFar += coinValue;
+    timeBought = rateStrategy.calculateTime(insertedSoFar);
+  } 
+  public int readDisplay() {
+    return timeBought;
+  }
+  public Receipt pushBuyButton() {
+    Receipt r = factory.createReceipt(timeBought);
+    timeBought = insertedSoFar = 0;
+    return r;
+  }
+  public void pushCancelButton() {
+    timeBought = insertedSoFar = 0;
+  }
+  public int amountEarned() {
+    return 0;
+  }
+  public void empty() {
+  }
+}
+

### 2008-02-02 Module2/exercise/compositional/PayStation/src/paystation/domain/ProgressiveRateStrategy.java

@@ -1,0 +1,23 @@
+package paystation.domain;
+
+/** A progressive rate calculation strategy.
+
+    author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class ProgressiveRateStrategy implements RateStrategy {
+  public int calculateTime( int amount ) {
+    int time = 0;
+    if ( amount >= 150+200 ) { // from 2nd hour onwards
+      amount -= 350;
+      time = 120 /*min*/ + amount / 5;
+    } else if ( amount >= 150 ) { // from 1st to 2nd hour
+      amount -= 150;
+      time = 60 /*min*/ + amount * 3 / 10;
+    } else { // up to 1st hour
+      time = amount * 2 / 5;
+    }
+    return time;
+  }
+}
+

### 2008-02-02 Module2/exercise/compositional/PayStation/src/paystation/domain/Receipt.java

@@ -1,0 +1,20 @@
+package paystation.domain;
+
+/** This interface models the receipt that the customer receives from
+    the pay station.
+
+    author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public interface Receipt {
+  /** return the number of minutes this receipt is valid for.
+      @return number of minutes parking time
+  */
+  int value();
+  
+  /** "shows" the receipt by writing text to the 
+      shell/prompt.
+  */
+  void show();
+}
+

### 2008-02-02 Module2/exercise/compositional/PayStation/src/paystation/domain/ReceiptImpl.java

@@ -1,0 +1,28 @@
+package paystation.domain;
+
+import java.util.*;
+
+/** Implementation of the receipt interface.
+
+    author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class ReceiptImpl implements Receipt {
+  private int value;
+  public ReceiptImpl(int value) { this.value = value; }
+  public int value() { return value;}
+  public void show() {
+    System.out.println( "-------------------------------------------------");
+    System.out.println( "-------  P A R K I N G   R E C E I P T  ---------");
+    System.out.println( "    Value "+ value+" minutes." );
+    Calendar now = GregorianCalendar.getInstance();
+    now.add(Calendar.MINUTE, value);
+    String end = 
+      "" + now.get(Calendar.HOUR_OF_DAY) + 
+      ":" + now.get(Calendar.MINUTE);
+    System.out.println( "    Valid until "+ end );
+    System.out.println( "-------------------------------------------------");
+    
+  }
+}
+

### 2008-02-02 Module2/exercise/compositional/PayStation/test/paystation/domain/TestAll.java

@@ -1,0 +1,13 @@
+package paystation.domain;
+
+import junit.framework.*;
+
+public class TestAll extends TestCase {
+
+  public static Test suite() {
+    TestSuite ts = new TestSuite();
+    ts.addTest(TestPayStationAlpha.suite());
+    ts.addTest(TestPayStationBeta.suite());
+    return ts;    
+  }
+}

### 2008-02-02 Module2/exercise/compositional/PayStation/test/paystation/domain/TestPayStationAlpha.java

@@ -1,0 +1,79 @@
+package paystation.domain;
+
+import junit.framework.*;
+
+public class TestPayStationAlpha extends TestCase {
+
+  private PayStation ps;
+  public void setUp() {
+    PayStationFactory factory = new AlphaTownFactory();
+    ps = new PayStationImpl( factory );
+  }
+  public void testEnterNickel() { 
+    try {
+      ps.addPayment( 5 );
+    } catch ( IllegalCoinException e ) {
+      fail("Pay station did not accept valid coin");
+    }
+    assertEquals( 2, ps.readDisplay() ); 
+  }
+  public void testEnterQuarter() { 
+    try {
+      ps.addPayment( 25 );
+    } catch ( IllegalCoinException e ) {
+      fail("Pay station did not accept valid coin");
+    }
+    assertEquals( 25 / 5 * 2 , ps.readDisplay() ); 
+    // 2 minutes per 5 cents
+  }
+  public void testEnterIllegalCoin() {
+    try {
+      ps.addPayment(17);
+      fail("Pay station accepted a 17cent coin");
+    } catch ( IllegalCoinException e ) {}
+  }
+  public void testBuy() {
+    try {
+      ps.addPayment(5);
+      ps.addPayment(10);
+      ps.addPayment(25);
+    } catch ( IllegalCoinException e ) {
+      fail("Pay station did not accept valid coin");
+    }
+    Receipt receipt;
+    receipt = ps.pushBuyButton();
+    assertTrue( receipt != null );
+    assertEquals((5+10+25) / 5 * 2 , receipt.value() );
+  }
+  public void testBuy100cent() {
+    try {
+      ps.addPayment(5);
+      ps.addPayment(5);
+      ps.addPayment(5);
+      ps.addPayment(10);
+      ps.addPayment(25);
+      ps.addPayment(25);
+      ps.addPayment(25);
+    } catch ( IllegalCoinException e ) {
+      fail("Pay station did not accept valid coin");
+    }
+    Receipt receipt;
+    receipt = ps.pushBuyButton();
+    assertTrue( receipt != null );
+    assertEquals( 100 / 5 * 2 , receipt.value() );
+  }
+  public void testReceiptValue() {
+    Receipt receipt = new ReceiptImpl(30);
+    assertEquals( 30, receipt.value() );
+  }
+  public void testClearing() {
+    try {
+      ps.addPayment(25);
+    } catch ( IllegalCoinException e ) {}
[…]

### 2008-02-02 Module2/exercise/compositional/PayStation/test/paystation/domain/TestPayStationBeta.java

@@ -1,0 +1,104 @@
+package paystation.domain;
+
+import junit.framework.*;
+
+public class TestPayStationBeta extends TestCase {
+
+  private PayStation ps;
+  public void setUp() {
+    ps = new PayStationImpl( new BetaTownFactory() );
+  }
+  public void testFirstHour() { 
+    // First hour: $1.5
+    try {
+      ps.addPayment( 25 );
+      ps.addPayment( 25 );
+
+      ps.addPayment( 25 );
+      ps.addPayment( 25 ); 
+
+      ps.addPayment( 25 ); 
+      ps.addPayment( 25 ); 
+    } catch ( IllegalCoinException e ) {
+      fail("Pay station did not accept valid coin");
+    }
+    assertEquals( 60 /*minutes*/, ps.readDisplay() ); 
+  }
+  public void testSecondHour1() { 
+    // Two hours: $1.5+2.0
+    try {
+      ps.addPayment( 25 ); ps.addPayment( 25 );
+      ps.addPayment( 25 ); ps.addPayment( 25 );
+      ps.addPayment( 25 ); ps.addPayment( 25 );
+
+      ps.addPayment( 25 ); ps.addPayment( 25 );
+      ps.addPayment( 25 ); ps.addPayment( 25 );
+      ps.addPayment( 25 ); ps.addPayment( 25 );
+      ps.addPayment( 25 ); ps.addPayment( 25 );
+
+    } catch ( IllegalCoinException e ) {
+      fail("Pay station did not accept valid coin");
+    }
+    assertEquals( 120 /*minutes*/ , ps.readDisplay() ); 
+  }
+  public void testSecondHour2() { 
+    // Two and a half hours: $1.5+2.0+ (2.0/2)
+    try {
+      // $1.5 = 1st hour
+      ps.addPayment( 25 ); ps.addPayment( 25 );
+      ps.addPayment( 25 ); ps.addPayment( 25 );
+      ps.addPayment( 25 ); ps.addPayment( 25 );
+
+      // $2.0 = 2nd hour
+      ps.addPayment( 25 ); ps.addPayment( 25 );
+      ps.addPayment( 25 ); ps.addPayment( 25 );
+      ps.addPayment( 25 ); ps.addPayment( 25 );
+      ps.addPayment( 25 ); ps.addPayment( 25 );
+
+      // $1.5 = half hour in 2-3 hour interval
+      ps.addPayment( 25 ); ps.addPayment( 25 );
[…]

### 2008-02-02 Module2/exercise/tdd/backgammon-state/BackgammonState.java

@@ -1,0 +1,13 @@
+/** The set of states that a backgammon game can be in 
+
+  Author: Henrik Bærbak Christensen 2006
+*/
+
+public enum BackgammonState {
+  StartState, // Game not started
+  DiceShakeState, // Dice must be thrown
+  RedMoveState, // Red must move checker
+  BlackMoveState, // Black must move checker
+  EndState, // Game has been won
+}
+  

### 2008-02-02 Module2/exercise/tdd/backgammon-state/BackgammonStateMachineImpl.java

@@ -1,0 +1,15 @@
+/** A statemachine for handling backgammon.
+
+   Author: Henrik Bærbak Christensen 2006
+*/
+
+public class BackgammonStateMachineImpl implements BackgammonStateMachine {
+  public boolean newGame() { return true; }
+  public boolean rollDice() { return true; }
+  public boolean moveChecker( boolean isRedPlayer ) {
+    return true;
+  }
+  public BackgammonState getState() {
+    return BackgammonState.StartState;
+  }
+}

### 2008-02-02 Module2/exercise/tdd/met-report/MetReportFormatter.java

@@ -1,0 +1,8 @@
+/** Formatter for MET REPORT string representations.
+
+    Author: Henrik Bærbak Christensen 2006 
+*/
+
+public interface MetReportFormatter {
+  public String format_dddff( TwoMinuteMeanWind w );
+}

### 2008-02-02 Module2/exercise/tdd/met-report/TwoMinuteMeanWind.java

@@ -1,0 +1,24 @@
+/** Two minute mean wind.
+    A simple data object containing mean wind data.
+
+    Author: Henrik Bærbak Christensen 2006 
+*/
+
+public class TwoMinuteMeanWind {
+  /** true iff sensor readings are valid */
+  public boolean valid;
+  /** two minute mean wind direction; the value
+      is between 0 and 359 degrees. */
+  public int direction;
+  /** two minute mean wind speed in knots; the
+      value is between 0.0 and 99.9 */
+  public double speed;
+  /** low extreme wind direction in the
+      two minute time span; range [0;359] degrees.
+      low < high always. */
+  public int low;
+  /** high extreme wind direction in the
+      two minute time span; range [0;359] degrees.
+      low < high always. */
+  public int high;
+} 

### 2008-02-02 Module2/exercise/tdd/seven-segment/SevenSegment.java

@@ -1,0 +1,18 @@
+/** Java interface for seven segment LED display
+
+    Responsibility: To enable turning on and off
+    individual segments.
+
+    Author: Henrik Bærbak Christensen 2006
+*/
+
+public interface SevenSegment {
+  /** turn a LED on or off.
+      @param led the number of the LED. Range is 0 to 6. The LEDs are
+      numbered top to bottom, left to right. That is, the top,
+      horizontal, LED is 0, the top left LED is 1, etc.
+      @param on if true the LED is turned on otherwise it is turned
+      off.
+  */
+  void setLED(int led, boolean on);
+}

### 2008-02-02 Module2/exercise/tdd/seven-segment/TestNumberDisplay.java

@@ -1,0 +1,18 @@
+import junit.framework.*;
+
+/** Test cases for NumberDisplay.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+
+public class TestNumberDisplay extends TestCase {
+  public void setUp() {
+  }
+  public void testOne() {
+    assertEquals( false, true );
+  }
+  public static Test suite() {
+    return new TestSuite(TestNumberDisplay.class);
+  }
+}

### 2008-02-02 trunk/Module2/chapter/abstract-factory/iteration-0/test/paystation/domain/TestPayStation.java
Adding the exercise source code

@@ -1,0 +1,79 @@
+package paystation.domain;
+
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Pay Station that 
+    are not related to rate calculations.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestPayStation {
+  PayStation ps;
+  /** Fixture for pay station testing. */
+  @Before
+  public void setUp() {
+    ps = new PayStationImpl(new One2OneRateStrategy());
+  }
+
+  /** Test that all three types of coins are accepted */
+  @Test
+  public void testAllCoins() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    assertEquals( 5+10+25 , ps.readDisplay() ); 
+  }
+  /** Testing for illegal coin entry. */
+  @Test(expected=IllegalCoinException.class)
+  public void testEnterIllegalCoin() throws IllegalCoinException {
+    ps.addPayment(17);
+  }
+  /** Test buying. */
+  @Test public void testBuy() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    Receipt receipt;
+    receipt = ps.buy();
+    assertNotNull( receipt );
+    assertEquals( 5+10+25 , receipt.value() );
+  }
+  /** Test buying for 100 cents */
+  @Test public void testBuy100cent() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    ps.addPayment(25);
+    ps.addPayment(25);
+
+    Receipt receipt;
+    receipt = ps.buy();
+    assertEquals((3*5+10+3*25), receipt.value() );
+  }
+  /** Test receipt */
+  @Test public void testReceiptValue() {
+    Receipt receipt = new ReceiptImpl(30);
+    assertEquals( 30, receipt.value() );
+  }
+  /** Test that buy clears for a new transaction */
+  @Test public void testClearingAfterBuy() throws IllegalCoinException {
+    ps.addPayment(25);
+    ps.buy(); // don't care about result
+    assertEquals( 0 , ps.readDisplay() ); 
+  }
+  /** Test that cancel resets the display */
+  @Test public void testCancel() throws IllegalCoinException {
[…]

### 2008-02-02 trunk/Module2/chapter/abstract-factory/iteration-1/test/paystation/domain/TestAlternatingRateStrategy.java

@@ -1,0 +1,28 @@
+package paystation.domain;
+
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Gammatown alternating rate strategy.
+    Author: (c) Henrik Bærbak Christensen 2007 */
+
+public class TestAlternatingRateStrategy {
+
+  /** Test for 500 cent during weekends */
+  @Test public void test500centWeekend() { 
+    RateStrategy rs;
+    rs = new AlternatingRateStrategy( new LinearRateStrategy(),
+                                      new ProgressiveRateStrategy(),
+                                      new FixedDecisionStrategy(true) );
+    assertEquals( 2 * 60 + 30 /*minutes*/ , rs.calculateTime(500) ); 
+  }
+
+  /** Test for 500 cent during weekdays */
+  @Test public void test500centWeekday() { 
+    RateStrategy rs;
+    rs = new AlternatingRateStrategy( new LinearRateStrategy(),
+                                      new ProgressiveRateStrategy(),
+                                      new FixedDecisionStrategy(false) );
+    assertEquals( 500 / 5 * 2 /*minutes*/ , rs.calculateTime(500) ); 
+  }
+}

### 2008-02-02 trunk/Module2/chapter/abstract-factory/iteration-2/src/paystation/domain/BetaTownFactory.java

@@ -1,0 +1,15 @@
+package paystation.domain;
+
+/** The factory for BetaTown.
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class BetaTownFactory implements PayStationFactory {
+  public RateStrategy createRateStrategy() {
+    return new ProgressiveRateStrategy();
+  }
+
+  public Receipt createReceipt( int parkingTime ) {
+    return new StandardReceipt(parkingTime, true);
+  }
+}

### 2008-02-02 trunk/Module2/chapter/abstract-factory/iteration-2/test/paystation/domain/TestConfiguration.java

@@ -1,0 +1,33 @@
+package paystation.domain;
+
+import java.io.*;
+
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Pay Station configurations.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestConfiguration {
+  PayStation ps;
+
+  /** Test BetaTown receipts with bar codes */
+  @Test public void testBetaTownReceipt() throws IllegalCoinException {
+    ps = new PayStationImpl( new BetaTownFactory() );
+    addOneDollar(ps);
+    Receipt receipt = ps.buy();
+
+    ByteArrayOutputStream baos = new ByteArrayOutputStream();
+    PrintStream ps = new PrintStream(baos);
+    receipt.show(ps);
+    String output = baos.toString();
+    String[] lines = output.split("\n");
+
+    assertEquals( 6, lines.length );
+  }
+
+  private void addOneDollar(PayStation ps) throws IllegalCoinException {
+    ps.addPayment(25); ps.addPayment(25);
+    ps.addPayment(25); ps.addPayment(25);
+  }
+}

### 2008-02-02 trunk/Module2/chapter/build-management/iteration-4/src/paystation/domain/ProgressiveRateStrategy.java

@@ -1,0 +1,20 @@
+package paystation.domain;
+
+/** A progressive rate calculation strategy.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+public class ProgressiveRateStrategy implements RateStrategy {
+  public int calculateTime( int amount ) {
+    int time = 0;
+    if ( amount >= 150+200 ) { // from 2nd hour onwards
+      amount -= 350;
+      time = 120 /*min*/ + amount / 5;
+    } else if ( amount >= 150 ) { // from 1st to 2nd hour
+      amount -= 150;
+      time = 60 /*min*/ + amount * 3 / 10;
+    } else { // up to 1st hour
+      time = amount * 2 / 5;
+    }
+    return time;
+  }
+}
+

### 2008-02-02 trunk/Module2/chapter/state/compositional/src/paystation/domain/AlternatingRateStrategy.java

@@ -1,0 +1,37 @@
+package paystation.domain;
+
+import java.util.*;
+
+/** An alternating rate calculation strategy that
+    uses two rate strategies, one for weekdays and
+    another for weekends.
+
+    @author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class AlternatingRateStrategy implements RateStrategy {
+  RateStrategy weekendStrategy, weekdayStrategy, currentState;
+  public AlternatingRateStrategy( RateStrategy weekdayStrategy,
+                                  RateStrategy weekendStrategy ) {
+    this.weekdayStrategy = weekdayStrategy;
+    this.weekendStrategy = weekendStrategy;
+    this.currentState = null;
+  }
+  public int calculateTime( int amount ) {
+    if ( isWeekend() ) {
+      currentState = weekendStrategy; 
+    } else {
+      currentState = weekdayStrategy;
+    }
+    return currentState.calculateTime( amount );
+  }
+  private boolean isWeekend() {
+    Date d = new Date();
+    Calendar c = new GregorianCalendar();
+    c.setTime(d);
+    int dayOfWeek = c.get(Calendar.DAY_OF_WEEK);
+    return ( dayOfWeek == Calendar.SATURDAY 
+             || 
+             dayOfWeek == Calendar.SUNDAY);
+  }
+}

### 2008-02-02 trunk/Module2/chapter/state/iteration-0/src/paystation/domain/LinearRateStrategy.java

@@ -1,0 +1,13 @@
+package paystation.domain;
+
+/** A linear rate calculation strategy.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class LinearRateStrategy implements RateStrategy {
+  public int calculateTime( int amount ) {
+    return amount * 2 / 5;
+  }
+}
+

### 2008-02-02 trunk/Module2/chapter/state/iteration-0/src/paystation/domain/ReceiptImpl.java

@@ -1,0 +1,13 @@
+package paystation.domain;
+
+/** Implementation of Receipt.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class ReceiptImpl implements Receipt {
+  private int value;
+  public ReceiptImpl(int value) { this.value = value; }
+  public int value() { return value;}
+  public void show() {}
+}

### 2008-02-02 trunk/Module2/chapter/state/iteration-0/test/paystation/domain/TestProgressiveRate.java

@@ -1,0 +1,46 @@
+package paystation.domain;
+
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the BetaTown progressive rate algorithm.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestProgressiveRate {
+  RateStrategy rs;
+  @Before
+  public void setUp() {
+    rs = new ProgressiveRateStrategy();
+  }
+
+  /** Test a single hour parking */
+  @Test public void testOneHour() { 
+    // First hour: $1.5
+    assertEquals( 60 /*minutes*/, rs.calculateTime(150) ); 
+  }
+  /** Test a 1.5 hour parking */
+  @Test public void testOneHourHalf() { 
+    // First hour: $1.5, next half hour = $1   
+    assertEquals( 60 + 30 /*minutes*/, rs.calculateTime(250) ); 
+  }
+  /** Test two hours parking */
+  @Test public void testTwoHour()  { 
+    // Two hours: $1.5+2.0
+    assertEquals( 2 * 60 /*minutes*/ , rs.calculateTime(350) ); 
+  }
+  /** Test two and a half hour parking */
+  @Test public void testTwoHourHalf() { 
+    // Two and a half hours: $1.5+2.0+ (3.0/2)
+    assertEquals( 2 * 60 + 30 /*minutes*/ , rs.calculateTime(500) ); 
+  }
+  /** Test three hours */
+  @Test public void testThreeHour() { 
+    // Three hours: $1.5+2.0+3.0 = $6.5
+    assertEquals( 3 * 60 /*minutes*/ , rs.calculateTime(650) ); 
+  }
+  /** Test four hours */
+  @Test public void testFourHour() { 
+    // Three hours: $1.5+2.0+ 2 * 3.0 = $9.5
+    assertEquals( 4 * 60 /*minutes*/ , rs.calculateTime(950) ); 
+  }
+}

### 2008-02-02 trunk/Module2/chapter/strategy/compositional/iteration-1/TestPayStation.java

@@ -1,0 +1,87 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Pay Station system.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestPayStation {
+  PayStation ps;
+  /** Fixture for pay station testing. */
+  @Before
+  public void setUp() {
+    ps = new PayStationImpl( new LinearRateStrategy() );
+  }
+
+  /** Testing that a nickel gives two minutes parking time */
+  @Test
+  public void testEnterNickel() throws IllegalCoinException {
+    ps.addPayment( 5 );
+    assertEquals( 2, ps.readDisplay() ); 
+  }
+  /** Testing that a quarter gives ten minutes parking time */
+  @Test
+  public void testEnterQuarter() throws IllegalCoinException {
+    ps.addPayment( 25 );
+    assertEquals( 10, ps.readDisplay() ); 
+  }
+  /** Testing for illegal coin entry. */
+  @Test(expected=IllegalCoinException.class)
+  public void testEnterIllegalCoin() throws IllegalCoinException {
+    ps.addPayment(17);
+  }
+  /** Test buying. */
+  @Test public void testBuy() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    Receipt receipt;
+    receipt = ps.buy();
+    assertNotNull( receipt );
+    assertEquals((5+10+25) / 5 * 2 , receipt.value() );
+  }
+  /** Test buying for 100 cents */
+  @Test public void testBuy100cent() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    ps.addPayment(25);
+    ps.addPayment(25);
+
+    Receipt receipt;
+    receipt = ps.buy();
+    assertEquals((3*5+10+3*25) / 5 * 2 , receipt.value() );
+  }
+  /** Test receipt */
+  @Test public void testReceiptValue() {
+    Receipt receipt = new ReceiptImpl(30);
+    assertEquals( 30, receipt.value() );
+  }
+  /** Test that buy clears for a new transaction */
+  @Test public void testClearingAfterBuy() throws IllegalCoinException {
+    ps.addPayment(25);
+    ps.buy(); // don't care about result
+    assertEquals( 0 , ps.readDisplay() ); 
[…]

### 2008-02-02 trunk/Module2/chapter/strategy/compositional/iteration-2/PayStationImpl.java

@@ -1,0 +1,45 @@
+/** Implementation of the pay station.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+public class PayStationImpl implements PayStation {
+  private int insertedSoFar;
+  private int timeBought;
+
+  /** the strategy for rate calculations */
+  private RateStrategy rateStrategy;
+
+  public PayStationImpl(RateStrategy rateStrategy) {
+    this.rateStrategy = rateStrategy;
+    resetTransaction();
+  }
+  public void addPayment( int coinValue ) throws IllegalCoinException {
+    switch ( coinValue ) {
+    case 5: 
+    case 10: 
+    case 25: break;  
+    default: 
+      throw new IllegalCoinException("Invalid coin: "+coinValue+" cent.");
+    }
+    insertedSoFar += coinValue;
+    timeBought = rateStrategy.calculateTime(insertedSoFar);
+  }
+  public int readDisplay() {
+    return timeBought;
+  }
+  public Receipt buy() {
+    Receipt r = new ReceiptImpl(timeBought);
+    resetTransaction();
+    return r;
+  }
+  public void cancel() {
+    resetTransaction();
+  }
+  public int timeBought() {
+    return timeBought;
+  }
+
+  /** resets the present transaction */
+  private void resetTransaction() {
+    timeBought = insertedSoFar = 0;
+  }    
+}
+

### 2008-02-02 trunk/Module2/chapter/strategy/compositional/iteration-2/TestAll.java

@@ -1,0 +1,19 @@
+import org.junit.runner.RunWith;
+import org.junit.runners.Suite;
+
+@RunWith(Suite.class)
+  @Suite.SuiteClasses({ TestPayStation.class,
+                        TestProgressiveRate.class})
+  
+/** Suite to run all test cases for PayStation.
+    Author: (c) Henrik Bærbak Christensen 2007 */
+
+
+public class TestAll {
+  /** This wrapper is only required for running the old JUnit 3.8
+   * graphical user interface on new JUnit 4 test cases */
+  public static junit.framework.Test suite() {
+    return new junit.framework.JUnit4TestAdapter(TestAll.class);
+  } 
+
+}

### 2008-02-02 trunk/Module2/chapter/strategy/compositional/iteration-3/ProgressiveRateStrategy.java

@@ -1,0 +1,15 @@
+/** A progressive rate calculation strategy.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+public class ProgressiveRateStrategy implements RateStrategy {
+  public int calculateTime( int amount ) {
+    int time = 0;
+    if ( amount >= 150 ) { // from 1st to 2nd hour
+      amount -= 150;
+      time = 60 /*min*/ + amount * 3 / 10;
+    } else { // up to 1st hour
+      time = amount * 2 / 5;
+    }
+    return time;
+  }
+}
+

### 2008-02-02 trunk/Module2/chapter/strategy/compositional/iteration-last/LinearRateStrategy.java

@@ -1,0 +1,11 @@
+/** A linear rate calculation strategy.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class LinearRateStrategy implements RateStrategy {
+  public int calculateTime( int amount ) {
+    return amount * 2 / 5;
+  }
+}
+

### 2008-02-02 trunk/Module2/chapter/strategy/compositional/iteration-last/ProgressiveRateStrategy.java

@@ -1,0 +1,18 @@
+/** A progressive rate calculation strategy.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+public class ProgressiveRateStrategy implements RateStrategy {
+  public int calculateTime( int amount ) {
+    int time = 0;
+    if ( amount >= 150+200 ) { // from 2nd hour onwards
+      amount -= 350;
+      time = 120 /*min*/ + amount / 5;
+    } else if ( amount >= 150 ) { // from 1st to 2nd hour
+      amount -= 150;
+      time = 60 /*min*/ + amount * 3 / 10;
+    } else { // up to 1st hour
+      time = amount * 2 / 5;
+    }
+    return time;
+  }
+}
+

### 2008-02-02 trunk/Module2/chapter/strategy/compositional/separate-testing/RateStrategy.java

@@ -1,0 +1,20 @@
+/** This interface defines the responsibility of
+    calculating parking time based upon the
+    payment accepted by a pay station. It
+    plays the Strategy role of the Strategy pattern.
+
+    Instances implementing this interface defines
+    a specific rate structure.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public interface RateStrategy {
+  /** return the number of minutes parking time the
+      provided amount of payment is valid for.
+      @param amount payment in some currency.
+      @return number of minutes parking time
+  */
+  public int calculateTime( int amount );
+}
+

### 2008-02-02 trunk/Module2/chapter/strategy/compositional/separate-testing/TestAll.java

@@ -1,0 +1,20 @@
+import org.junit.runner.RunWith;
+import org.junit.runners.Suite;
+
+@RunWith(Suite.class)
+  @Suite.SuiteClasses({ TestPayStation.class,
+                          TestLinearRate.class,
+                          TestProgressiveRate.class})
+  
+/** Suite to run all test cases for PayStation.
+    Author: (c) Henrik Bærbak Christensen 2007 */
+
+
+public class TestAll {
+  /** This wrapper is only required for running the old JUnit 3.8
+   * graphical user interface on new JUnit 4 test cases */
+  public static junit.framework.Test suite() {
+    return new junit.framework.JUnit4TestAdapter(TestAll.class);
+  } 
+
+}

### 2008-02-02 trunk/Module2/chapter/strategy/compositional/separate-testing/TestLinearRate.java

@@ -1,0 +1,23 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Alphatown rate policy
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestLinearRate {
+  RateStrategy rs;
+
+  @Before
+  public void setUp() {
+    rs = new LinearRateStrategy();
+  }
+
+  /** Test 35 cents */
+  @Test public void test35() {
+    assertEquals( 35 / 5 * 2, rs.calculateTime(35) ); 
+  }
+   /** Test a 100 cents */
+  @Test public void test100() throws IllegalCoinException { 
+    assertEquals( 100 / 5 * 2, rs.calculateTime(100) ); 
+  }
+}

### 2008-02-02 trunk/Module2/chapter/strategy/compositional/separate-testing/TestProgressiveRate.java

@@ -1,0 +1,44 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the BetaTown progressive rate algorithm.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestProgressiveRate {
+  RateStrategy rs;
+  @Before
+  public void setUp() {
+    rs = new ProgressiveRateStrategy();
+  }
+
+  /** Test a single hour parking */
+  @Test public void testOneHour() { 
+    // First hour: $1.5
+    assertEquals( 60 /*minutes*/, rs.calculateTime(150) ); 
+  }
+  /** Test a 1.5 hour parking */
+  @Test public void testOneHourHalf() { 
+    // First hour: $1.5, next half hour = $1   
+    assertEquals( 60 + 30 /*minutes*/, rs.calculateTime(250) ); 
+  }
+  /** Test two hours parking */
+  @Test public void testTwoHour()  { 
+    // Two hours: $1.5+2.0
+    assertEquals( 2 * 60 /*minutes*/ , rs.calculateTime(350) ); 
+  }
+  /** Test two and a half hour parking */
+  @Test public void testTwoHourHalf() { 
+    // Two and a half hours: $1.5+2.0+ (3.0/2)
+    assertEquals( 2 * 60 + 30 /*minutes*/ , rs.calculateTime(500) ); 
+  }
+  /** Test three hours */
+  @Test public void testThreeHour() { 
+    // Three hours: $1.5+2.0+3.0 = $6.5
+    assertEquals( 3 * 60 /*minutes*/ , rs.calculateTime(650) ); 
+  }
+  /** Test four hours */
+  @Test public void testFourHour() { 
+    // Three hours: $1.5+2.0+ 2 * 3.0 = $9.5
+    assertEquals( 4 * 60 /*minutes*/ , rs.calculateTime(950) ); 
+  }
+}

### 2008-02-02 trunk/Module2/chapter/strategy/parametric/PayStationImpl.java

@@ -1,0 +1,61 @@
+/** Implementation of the pay station.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class PayStationImpl implements PayStation {
+  private int insertedSoFar;
+  private int timeBought;
+
+  public enum Town { ALPHATOWN, BETATOWN }
+  private Town town;
+
+  public PayStationImpl( Town town ) {
+    this.town = town;
+    resetTransaction();
+  }
+
+  public void addPayment( int coinValue ) throws IllegalCoinException {
+    switch ( coinValue ) {
+    case 5: 
+    case 10: 
+    case 25: break;  
+    default: 
+      throw new IllegalCoinException("Invalid coin: "+coinValue+" cent.");
+    }
+    insertedSoFar += coinValue;
+    if ( town == Town.ALPHATOWN ) {
+      timeBought = insertedSoFar / 5 * 2;
+    } else if ( town == Town.BETATOWN ) {
+      int amount = insertedSoFar;
+      timeBought = 0;
+      if ( amount >= 150+200 ) { // from 2nd hour onwards
+        amount -= 350;
+        timeBought = 120 /*min*/ + amount / 5;
+      } else if ( amount >= 150 ) { // from 1st to 2nd hour
+        amount -= 150;
+        timeBought = 60 /*min*/ + amount * 3 / 10;
+      } else { // up to 1st hour
+        timeBought = amount * 2 / 5;
+      }
+    }
+  }
+  public int readDisplay() {
+    return timeBought;
+  }
+  public Receipt buy() {
+    Receipt r = new ReceiptImpl(timeBought);
+    resetTransaction();
+    return r;
+  }
+  public void cancel() {
+    resetTransaction();
+  }
+  public int timeBought() {
+    return timeBought;
+  }
+
+  /** resets the present transaction */
+  private void resetTransaction() {
+    timeBought = insertedSoFar = 0;
+  }    
+}
+

### 2008-02-02 trunk/Module2/chapter/strategy/parametric/TestAll.java

@@ -1,0 +1,19 @@
+import org.junit.runner.RunWith;
+import org.junit.runners.Suite;
+
+@RunWith(Suite.class)
+  @Suite.SuiteClasses({ TestAlphaTown.class,
+                        TestBetaTown.class} )
+  
+/** Suite to run all test cases for PayStation.
+    Author: (c) Henrik Bærbak Christensen 2007 */
+
+
+public class TestAll {
+  /** This wrapper is only required for running the old JUnit 3.8
+   * graphical user interface on new JUnit 4 test cases */
+  public static junit.framework.Test suite() {
+    return new junit.framework.JUnit4TestAdapter(TestAll.class);
+  } 
+
+}

### 2008-02-02 trunk/Module2/chapter/strategy/polymorphic/TestAlphaTown.java

@@ -1,0 +1,81 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Pay Station system.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestAlphaTown {
+  PayStation ps;
+  /** Fixture for pay station testing. */
+  @Before
+  public void setUp() {
+    ps = new PayStationImpl();
+  }
+
+  /** Testing that a nickel gives two minutes parking time */
+  @Test
+  public void testEnterNickel() throws IllegalCoinException {
+    ps.addPayment( 5 );
+    assertEquals( 2, ps.readDisplay() ); 
+  }
+  /** Testing that a quarter gives ten minutes parking time */
+  @Test
+  public void testEnterQuarter() throws IllegalCoinException {
+    ps.addPayment( 25 );
+    assertEquals( 10, ps.readDisplay() ); 
+  }
+  /** Testing for illegal coin entry. */
+  @Test(expected=IllegalCoinException.class)
+  public void testEnterIllegalCoin() throws IllegalCoinException {
+    ps.addPayment(17);
+  }
+  /** Test buying. */
+  @Test public void testBuy() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    Receipt receipt;
+    receipt = ps.buy();
+    assertNotNull( receipt );
+    assertEquals((5+10+25) / 5 * 2 , receipt.value() );
+  }
+  /** Test buying for 100 cents */
+  @Test public void testBuy100cent() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    ps.addPayment(25);
+    ps.addPayment(25);
+
+    Receipt receipt;
+    receipt = ps.buy();
+    assertEquals((3*5+10+3*25) / 5 * 2 , receipt.value() );
+  }
+  /** Test receipt */
+  @Test public void testReceiptValue() {
+    Receipt receipt = new ReceiptImpl(30);
+    assertEquals( 30, receipt.value() );
+  }
+  /** Test that buy clears for a new transaction */
+  @Test public void testClearingAfterBuy() throws IllegalCoinException {
+    ps.addPayment(25);
+    ps.buy(); // don't care about result
[…]

### 2008-02-02 trunk/Module2/chapter/tdd/iteration-1/TestPayStation.java

@@ -1,0 +1,24 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Pay Station system.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class TestPayStation {
+
+  /** Testing that a nickel gives two minutes parking time */
+  @Test
+  public void testEnterNickel() throws IllegalCoinException {
+    PayStation ps = new PayStationImpl();
+    ps.addPayment( 5 );
+    assertEquals( 2, ps.readDisplay() ); 
+  }
+
+  /** This wrapper is only required for running the old JUnit 3.8
+   * graphical user interface on new JUnit 4 test cases */
+  public static junit.framework.Test suite() {
+    return new junit.framework.JUnit4TestAdapter(TestPayStation.class);
+  } 
+}

### 2008-02-02 trunk/Module2/chapter/tdd/iteration-2/PayStationImpl.java

@@ -1,0 +1,20 @@
+/** Implementation of the pay station.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+public class PayStationImpl implements PayStation {
+  private int insertedSoFar;
+  public void addPayment( int coinValue ) throws IllegalCoinException {
+    insertedSoFar += coinValue;
+  }
+  public int readDisplay() {
+    return insertedSoFar / 5 * 2;
+  }
+  public Receipt buy() {
+    return null;
+  }
+  public void cancel() {
+  }
+  public int timeBought() {
+    return 0;
+  }
+}
+

### 2008-02-02 trunk/Module2/chapter/tdd/iteration-4/TestPayStation.java

@@ -1,0 +1,48 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Pay Station system.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestPayStation {
+  PayStation ps;
+  /** Fixture for pay station testing. */
+  @Before
+  public void setUp() {
+    ps = new PayStationImpl();
+  }
+
+  /** Testing that a nickel gives two minutes parking time */
+  @Test
+  public void testEnterNickel() throws IllegalCoinException {
+    ps.addPayment( 5 );
+    assertEquals( 2, ps.readDisplay() ); 
+  }
+  /** Testing that a quarter gives ten minutes parking time */
+  @Test
+  public void testEnterQuarter() throws IllegalCoinException {
+    ps.addPayment( 25 );
+    assertEquals( 10, ps.readDisplay() ); 
+  }
+  /** Testing for illegal coin entry. */
+  @Test(expected=IllegalCoinException.class)
+  public void testEnterIllegalCoin() throws IllegalCoinException {
+    ps.addPayment(17);
+  }
+  /** Test buying. */
+  @Test public void testBuy() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    Receipt receipt;
+    receipt = ps.buy();
+    assertNotNull( receipt );
+    assertEquals((5+10+25) / 5 * 2 , receipt.value() );
+  }
+
+  /** This wrapper is only required for running the old JUnit 3.8
+   * graphical user interface on new JUnit 4 test cases */
+  public static junit.framework.Test suite() {
+    return new junit.framework.JUnit4TestAdapter(TestPayStation.class);
+  } 
+}

### 2008-02-02 trunk/Module2/chapter/tdd/iteration-6/TestPayStation.java

@@ -1,0 +1,67 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Pay Station system.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+
+public class TestPayStation {
+  PayStation ps;
+  /** Fixture for pay station testing. */
+  @Before
+  public void setUp() {
+    ps = new PayStationImpl();
+  }
+
+  /** Testing that a nickel gives two minutes parking time */
+  @Test
+  public void testEnterNickel() throws IllegalCoinException {
+    ps.addPayment( 5 );
+    assertEquals( 2, ps.readDisplay() ); 
+  }
+  /** Testing that a quarter gives ten minutes parking time */
+  @Test
+  public void testEnterQuarter() throws IllegalCoinException {
+    ps.addPayment( 25 );
+    assertEquals( 10, ps.readDisplay() ); 
+  }
+  /** Testing for illegal coin entry. */
+  @Test(expected=IllegalCoinException.class)
+  public void testEnterIllegalCoin() throws IllegalCoinException {
+    ps.addPayment(17);
+  }
+  /** Test buying. */
+  @Test public void testBuy() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    Receipt receipt;
+    receipt = ps.buy();
+    assertNotNull( receipt );
+    assertEquals((5+10+25) / 5 * 2 , receipt.value() );
+  }
+  /** Test buying for 100 cents */
+  @Test public void testBuy100cent() throws IllegalCoinException {
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(5);
+    ps.addPayment(10);
+    ps.addPayment(25);
+    ps.addPayment(25);
+    ps.addPayment(25);
+
+    Receipt receipt;
+    receipt = ps.buy();
+    assertEquals((3*5+10+3*25) / 5 * 2 , receipt.value() );
+  }
+  /** Test receipt */
+  @Test public void testReceiptValue() {
+    Receipt receipt = new ReceiptImpl(30);
+    assertEquals( 30, receipt.value() );
+  }
+
+  /** This wrapper is only required for running the old JUnit 3.8
+   * graphical user interface on new JUnit 4 test cases */
+  public static junit.framework.Test suite() {
+    return new junit.framework.JUnit4TestAdapter(TestPayStation.class);
+  } 
+}

### 2008-02-02 trunk/Module2/chapter/tdd/iteration-7/PayStationImpl.java

@@ -1,0 +1,32 @@
+/** Implementation of the pay station.
+    Author: (c) Henrik Bærbak Christensen 2006 */
+public class PayStationImpl implements PayStation {
+  private int insertedSoFar;
+  private int timeBought;
+
+  public void addPayment( int coinValue ) throws IllegalCoinException {
+    switch ( coinValue ) {
+    case 5: 
+    case 10: 
+    case 25: break;  
+    default: 
+      throw new IllegalCoinException("Invalid coin: "+coinValue+" cent.");
+    }
+    insertedSoFar += coinValue;
+    timeBought = insertedSoFar / 5 * 2;
+  }
+  public int readDisplay() {
+    return timeBought;
+  }
+  public Receipt buy() {
+    Receipt r = new ReceiptImpl(timeBought);
+    timeBought = insertedSoFar = 0;
+    return r;
+  }
+  public void cancel() {
+  }
+  public int timeBought() {
+    return 0;
+  }
+}
+

### 2008-02-02 trunk/Module2/chapter/test-stub/iteration-0/src/paystation/domain/One2OneRateStrategy.java

@@ -1,0 +1,15 @@
+package paystation.domain;
+
+/** A one-cent-one-minute rate calculation strategy to
+    simplify testing the responsibilities of the pay
+    station that are not related to rate calculation.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class One2OneRateStrategy implements RateStrategy {
+  public int calculateTime( int amount ) {
+    return amount;
+  }
+}
+

### 2008-02-02 trunk/Module2/chapter/test-stub/iteration-0/test/paystation/domain/TestGammaWeekend.java

@@ -1,0 +1,21 @@
+package paystation.domain;
+
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Gammatown rate strategy on a weekend.
+    Author: (c) Henrik Bærbak Christensen 2007 */
+
+public class TestGammaWeekend {
+  RateStrategy rs;
+  @Before
+  public void setUp() {
+    rs = new AlternatingRateStrategy( new LinearRateStrategy(),
+                                      new ProgressiveRateStrategy() );
+  }
+
+  /** Test for 500 cent */
+  @Test public void test500cent() { 
+    assertEquals( 2 * 60 + 30 /*minutes*/ , rs.calculateTime(500) ); 
+  }
+}

### 2008-02-02 trunk/Module2/chapter/test-stub/iteration-2/src/paystation/domain/ClockBasedDecisionStrategy.java

@@ -1,0 +1,20 @@
+package paystation.domain;
+
+import java.util.*;
+
+/** A weekend decicion strategy based on the operating system clock.
+    Author: (c) Henrik Bærbak Christensen 2007
+*/
+
+public class ClockBasedDecisionStrategy implements WeekendDecisionStrategy {
+  public boolean isWeekend() {
+    Date d = new Date();
+    Calendar c = new GregorianCalendar();
+    c.setTime(d);
+    int dayOfWeek = c.get(Calendar.DAY_OF_WEEK);
+    return ( dayOfWeek == Calendar.SATURDAY 
+             || 
+             dayOfWeek == Calendar.SUNDAY);
+  }
+}
+

### 2008-02-02 trunk/Module2/chapter/test-stub/iteration-2/test/paystation/domain/TestGammaWeekend.java

@@ -1,0 +1,18 @@
+package paystation.domain;
+
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Testcases for the Gammatown rate strategy on a weekend.
+    Author: (c) Henrik Bærbak Christensen 2007 */
+
+public class TestGammaWeekend {
+  /** Test for 500 cent */
+  @Test public void test500centWeekend() { 
+    RateStrategy rs;
+    rs = new AlternatingRateStrategy( new LinearRateStrategy(),
+                                      new ProgressiveRateStrategy(),
+                                      new FixedDecisionStrategy(true) );
+    assertEquals( 2 * 60 + 30 /*minutes*/ , rs.calculateTime(500) ); 
+  }
+}

### 2008-02-02 trunk/Module2/exercise/abstract-factory/DoorAlarm/src/dooralarm/domain/DanishMessageDatabase.java

@@ -1,0 +1,13 @@
+package dooralarm.domain;
+
+/** DanishMessageDatabase: Messages in Danish.
+
+    Author: Henrik Bærbak Christensen
+*/
+
+public class DanishMessageDatabase implements MessageDatabase {
+  public String getFrameTitle() { return "Døralarm"; }
+  public String getWelcomeMessage() { return "Indtast kode:"; }
+  public String getDeniedMessage() { return "Afvist"; }
+  public String getGrantedMessage() { return "Godkendt"; }
+}

### 2008-02-02 trunk/Module2/exercise/compositional/PayStation/src/paystation/domain/CompositeRateStrategy.java

@@ -1,0 +1,37 @@
+package paystation.domain;
+
+import java.util.*;
+
+/** A composite rate calculation strategy that
+    uses two rate strategies, one for weekdays and
+    another for weekends.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class CompositeRateStrategy implements RateStrategy {
+  RateStrategy weekendStrategy, weekdayStrategy;
+  public CompositeRateStrategy( RateStrategy weekdayStrategy,
+                                RateStrategy weekendStrategy ) {
+    this.weekdayStrategy = weekdayStrategy;
+    this.weekendStrategy = weekendStrategy;
+  }
+  public int calculateTime( int amount ) {
+    int time;
+    if ( isWeekend() ) {
+      time = weekendStrategy.calculateTime( amount );
+    } else {
+      time = weekdayStrategy.calculateTime( amount );
+    }
+    return time;
+  }
+  private boolean isWeekend() {
+    Date d = new Date();
+    Calendar c = new GregorianCalendar();
+    c.setTime(d);
+    int dayOfWeek = c.get(Calendar.DAY_OF_WEEK);
+    return ( dayOfWeek == Calendar.SATURDAY 
+             || 
+             dayOfWeek == Calendar.SUNDAY);
+  }
+}

### 2008-02-02 trunk/Module2/exercise/compositional/PayStation/src/paystation/domain/LoggingPayStation.java

@@ -1,0 +1,23 @@
+package paystation.domain;
+
+/** A decorator of a PayStation that logs coin entries.
+
+    author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public class LoggingPayStation implements PayStation {
+  private PayStation paystation;
+  public LoggingPayStation( PayStation ps ) {
+    paystation = ps;
+  }
+  public void addPayment( int coinValue ) throws IllegalCoinException {
+    System.out.println( ""+coinValue+" cents. Time (now)" );
+    paystation.addPayment( coinValue );
+  }
+  public int readDisplay() { return paystation.readDisplay(); }
+  public Receipt pushBuyButton() { return paystation.pushBuyButton(); }
+  public void pushCancelButton() { paystation.pushCancelButton(); }
+  public int amountEarned() { return paystation.amountEarned(); }
+  public void empty() { paystation.empty(); }
+}
+

### 2008-02-02 trunk/Module2/exercise/compositional/PayStation/src/paystation/domain/PayStationFactory.java

@@ -1,0 +1,18 @@
+package paystation.domain;
+
+/** Abstract Factory for the Pay Station.
+
+    Author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public interface PayStationFactory {
+  /** Create instances of Receipts.
+      @param parkingTime of minutes parking time the receipt is valid for.
+  */
+  public Receipt createReceipt( int parkingTime );
+    
+  /** Create instance of rate strategy
+  */
+  public RateStrategy createRateStrategy();
+}
+

### 2008-02-02 trunk/Module2/exercise/compositional/PayStation/src/paystation/domain/WeekdayDeterminationStrategy.java

@@ -1,0 +1,13 @@
+package paystation.domain;
+
+import java.util.*;
+
+/** The strategy that determines whether it is weekend or not.
+
+    author: (c) Henrik Bærbak Christensen 2006
+*/
+
+public interface WeekdayDeterminationStrategy {
+  /** return true if it is presently a weekend day */
+  boolean isWeekend();
+}

### 2008-02-02 trunk/Module2/exercise/compositional/PayStation/src/paystation/gui/PayStationGUI.java

@@ -1,0 +1,175 @@
+package paystation.gui;
+
+import javax.swing.*;
+import java.awt.*;
+import java.awt.event.*;
+
+import paystation.domain.*;
+import paystation.util.*;
+
+/** A crude graphical user interface for testing the pay station.
+
+    Author: Henrik Bærbak Christense 2006
+*/
+public class PayStationGUI extends JFrame {
+
+  public static void main(String[] args) {
+    new PayStationGUI();
+  }
+
+  /** The "digital display" where readings are shown */
+  FourDigitDisplay display;
+  /** The domain pay station that the gui interacts with */
+  PayStation payStation;
+
+  /** Create the GUI */
+  public PayStationGUI() {
+    super("PayStation GUI" );
+
+    payStation = new PayStationImpl( new AlphaTownFactory() ); 
+
+    // all the gui setup
+    JFrame.setDefaultLookAndFeelDecorated(true);
+    setLocation( 100, 20 );
+    setDefaultCloseOperation(JFrame.EXIT_ON_CLOSE);
+
+     
+    Container cpane = getContentPane(); 
+
+    cpane.setLayout( new BorderLayout() );
+
+    cpane.add( createCoinInputPanel(), BorderLayout.EAST );
+    cpane.add( createButtonPanel(), BorderLayout.SOUTH );
+
+    cpane.add( createDisplayPanel(), BorderLayout.CENTER );
+
+    setJMenuBar( createAllMenus() );
+
+    pack();
+    setVisible(true);
+  }
+
+  /** Update the digital display with whatever the
+      pay station domain shows */
+  private void updateDisplay() {
+    display.setDisplay( payStation.readDisplay() );
+  }
+  
+  /** Create the coin input panel */
+  private JComponent createCoinInputPanel() {
+    Box p = new Box( BoxLayout.Y_AXIS );
+    p.add( defineButton( " 5 c", "5" ));
+    p.add( defineButton( "10 c", "10" ));
+    p.add( defineButton( "25 c", "25" ));
+    return p;
+  }
+  
+  /** The button action listener that reacts on clicking the
+      coin buttons */
+  private ActionListener buttonActionListener = new ActionListener() {
+      public void actionPerformed(ActionEvent e) {
+        String s = e.getActionCommand();
+        int coin = Integer.parseInt(s);
+        try {
+          payStation.addPayment( coin );
+        } catch (IllegalCoinException exc) {
+          // illegal coins just do nothing.
+        }
+        updateDisplay();
+      }
+    };
+  
[…]

### 2008-02-02 trunk/Module2/exercise/tdd/backgammon-state/TestBackgammonState.java

@@ -1,0 +1,28 @@
+import org.junit.*;
+import static org.junit.Assert.*;
+
+/** Test cases for Backgammon state exercise
+
+    Author: (c) Henrik Bærbak Christensen 2007
+*/
+
+
+public class TestBackgammonState {
+  private BackgammonStateMachine bs;
+
+  /** fixture for testing the Backgammon statemachine */
+  @Before
+  public void setUp() {
+    bs = new BackgammonStateMachineImpl();
+  }
+  
+  @Test
+  public void testInitialisation() {
+    assertEquals( BackgammonState.StartState, bs.getState() );
+  }
+  /** This wrapper is only required for running the old JUnit 3.8
+   * graphical user interface on new JUnit 4 test cases */
+  public static junit.framework.Test suite() {
+    return new junit.framework.JUnit4TestAdapter(TestBackgammonState.class);
+  } 
+}

### 2008-02-04 SAiP/Module1/H1/doc/paper.tex
Added figures to the document.

--- 2008-01-29 trunk/SAiP/Module1/H1/doc/paper.tex
@@
 \author{
-  $<<$Names$>>$\\
+  Anders H Poder, Jesper Dalberg, Lars Kringelbach\\\\
   Department of Computer Science, University of Aarhus\\
-  Aabogade 34, 8200 {\AA}rhus N, Denmark\\
+  Aabogade 34, 8200 {\AA}rhus N, Denmark\\\\
   \makeatletter
-  \texttt{<<Group name>>}\\
-  \texttt{<<Student IDs>>}\\
-  \texttt{\{..., ...\}@daimi.au.dk}
+  \texttt{Group 11. Kilo}\\
+  \texttt{19951439, 20074976, 20074842}\\
+  \texttt{\{ahp, jdalberg, u074842\}@daimi.au.dk}
 }
@@
 
-$<<$Describe HS07 from this viewpoint using UML. Where appropriate
-provide textual supplements$>>$
+The module viewpoint of the HS07 system.
+
+\begin{figure}[!htb]
+\center {\includegraphics[viewport=0 10 600 550,scale=0.4]{figures/mv.pdf}}
+\caption{HS07 Package Diagram}
+\label{fig:mv}
+\end{figure}
+
+The figure shows the packages of the HS07 system, and how they interact 
+with java packages used.
 
@@
 
-$<<$Describe HS07 from this viewpoint using UML. Where appropriate
-provide textual supplements$>>$
+The Component \& Connecter view consists of an Active Objects diagram and a
+sequence diagram. The Active Objects diagram shows the active objects of 
+the HS07 system, and how they interact.
+
+\begin{figure}[!htb]
+\center {\includegraphics[viewport=0 10 600 420,scale=0.5]{figures/cc_ao.pdf}}
+\caption{HS07 Active Objects}
+\label{fig:cc_ao}
+\end{figure}
+The sequence diagram illustrates the timing of the interactions.
+\begin{figure}[!htb]
+\includegraphics[viewport=0 10 500 650,scale=0.3]{figures/sequence.pdf}
+\caption{HS07 Sequence Diagram}
+\label{fig:sequence}
+\end{figure}
 
@@
 
-$<<$Describe HS07 from this viewpoint using UML. Where appropriate
-provide textual supplements$>>$
+The allocation viewpoint illustrates how components are deployed im actaual processes
+within the HS07 system.
+
+\begin{figure}[!htb]
+\center {\includegraphics[viewport=0 10 700 350,scale=0.4]{figures/deployment.pdf}}
+\caption{HS07 Allocation Diagram}
+\label{fig:allocation}
+\end{figure}
 

### 2008-02-04 trunk/SAiP/Module1/H1/doc/paper.tex
Title page updated with names etc.
dates

@@
 \author{
-  $<<$Names$>>$\\
+  Anders H Poder, Jesper Dalberg, Lars Kringelbach\\\\
   Department of Computer Science, University of Aarhus\\
-  Aabogade 34, 8200 {\AA}rhus N, Denmark\\
+  Aabogade 34, 8200 {\AA}rhus N, Denmark\\\\
   \makeatletter
-  \texttt{<<Group name>>}\\
-  \texttt{<<Student IDs>>}\\
-  \texttt{\{..., ...\}@daimi.au.dk}
+  \texttt{Group 11 - Kilo}\\
+  \texttt{19951439, 20074976, 20074842}\\
+  \texttt{\{ahp, jdalberg, u074842\}@daimi.au.dk}
 }
 
-\date{$<<$Date: 2007/11/05 22:51:26 $>>$}
+\date{$Date$}
 

### 2008-02-04 SAiP/Module1/H1/doc/paper.tex
svn keyword testing
a bit of discussion stuff

@@
   \makeatletter
-  \texttt{Group 11. Kilo}\\
+  \texttt{Group 11 - Kilo}\\
   \texttt{19951439, 20074976, 20074842}\\
@@
 
-\date{$<<$Date: 2007/11/05 22:51:26 $>>$}
+\date{2008-02-04 13:32}
 
@@
 
-The allocation viewpoint illustrates how components are deployed im actaual processes
+The allocation viewpoint illustrates how components are deployed in actual processes
 within the HS07 system.
@@
 
+When performing reverse engineering to achieve an architectural description, the dangers
+posed seem to be finding the correct balance in the choice of detail level. It
+is easy to get to much information, cluttering the purpose of the 
+AD, which in a basic sense is to clarify and present at wider understanding of
+the system.
+
+If you are a systems developer, a reverse engineered AD might help you understand 
+the code better.
+
+Experienced architects might recognize patterns emerging from the process, and 
+immediately see possible design flaws in the implementation of these, thus
+being able to help improve the next version of the system.
+
 $<<$Are there aspects of the software architecture that have not been
@@
 
-$<<$...$>>$
+$<<For the architectural description above, discuss what (if anything) should be changed or added for it to comply with the IEEE recommended practice for architectural description>>$
+
+$<<Consider the definition of software architecture by [Perry and Wolf, 1992]. Discuss what the 'elements', 'form', and 'rationale' according to this definition would be for the HS07 system>>$
+
+\cite{perrywolf1992} defines "elements" as three different classes, data elements, processing elements and connecting elements.
+In HS07 the data elements are temperatures, the processing element is the gateway and the connecting element is HTTP.
+The "form", as described by Perry and Wolf, is the choice of the observer pattern in HS07, and the "rationale" for doing so are defined by the Achitectural Requirements described elsewhere in this document.
 

### 2008-02-04 trunk/SAiP/Module1/H1/doc/paper.tex
Date hardcoded :(
More discussion detail

@@
 
-\date{$Date$}
+\date{2008-02-04 13:32}
 
@@
 \section{Discussion}
+\subsection{Strength and Limitations of the approach}
 $<<$What are the strengths and limitations of this approach to
@@
 
+When using the approach described in \cite{christensen2004archdesc} with reverse 
+engineering to achieve an architectural description, the dangers posed seem to 
+be finding the correct balance in the choice of detail level. It
+is easy to get to much information, cluttering the purpose of the 
+AD, which in a basic sense is to clarify and present at wider understanding of
+the system.
+
+Experienced architects might recognize patterns emerging from the process, and 
+immediately see possible design flaws in the implementation of these, thus
+being able to help improve the quality of any evolving system.
+
+The purpose of doing a reverse engineered AD must either be to evolve
+the system of to extract clarification. Care should be taken to achieve either
+goal. 
+\subsection{Whats missing?}
 $<<$Are there aspects of the software architecture that have not been
 properly described?$>>$
+Stakeholders and their concerns. This encompasses both users/aquirers and developers. We also need to discuss the feasibility and maintainablity of the system.
 
-$<<$...$>>$
+\subsection{IEEE Conformance}
+
+$<<$For the architectural description above, discuss what (if anything) should be changed or added for it to comply with the IEEE recommended practice for architectural description$>>$
+
+The following are the problems identified in order to comply with \cite{ieeerecommendedpractice}.
+
+Clause 5.2: The stakeholders and their precise concerns need to be identified for each view.
+
+Clause 5.5: Consistency checks across all existing views.
+
+\subsection{Perry \& Wolf Considerations}
+$<<$Consider the definition of software architecture by [Perry and Wolf, 1992]. Discuss what the 'elements', 'form', and 'rationale' according to this definition would be for the HS07 system$>>$
+
+\cite{perrywolf1992} defines "elements" as three different classes, data elements, processing elements and connecting elements.
+In HS07 the data elements are temperatures, the processing element is the gateway and the connecting element is HTTP.
+The "form", as described by Perry and Wolf, is the choice of the observer pattern in HS07, and the "rationale" for doing so are defined by the Achitectural Requirements described elsewhere in this document.
 

### 2008-02-04 SAiP/Module1/H1/doc/paper.tex
Added a new MV figure

@@
 \begin{figure}[!htb]
-\center {\includegraphics[viewport=0 10 600 550,scale=0.4]{figures/mv.pdf}}
+\center {\includegraphics[viewport=0 10 600 500,scale=0.5]{figures/mv.pdf}}
 \caption{HS07 Package Diagram}
@@
 
-When performing reverse engineering to achieve an architectural description, the dangers
-posed seem to be finding the correct balance in the choice of detail level. It
+When using the approach described in \cite{christensen2004archdesc} with reverse 
+engineering to achieve an architectural description, the dangers posed seem to 
+be finding the correct balance in the choice of detail level. It
 is easy to get to much information, cluttering the purpose of the 
@@
 
-If you are a systems developer, a reverse engineered AD might help you understand 
-the code better.
-
 Experienced architects might recognize patterns emerging from the process, and 
 immediately see possible design flaws in the implementation of these, thus
-being able to help improve the next version of the system.
+being able to help improve the quality of any evolving system.
 
+The purpose of doing a reverse engineered AD must either be to evolve
+the system of to extract clarification. Care should be taken to achieve either
+goal. 
+\subsection{Whats missing?}
 $<<$Are there aspects of the software architecture that have not been
 properly described?$>>$
+Stakeholders and their concerns. This encompasses both users/aquirers and developers. We also need to discuss the feasibility and maintainablity of the system.
 
-$<<For the architectural description above, discuss what (if anything) should be changed or added for it to comply with the IEEE recommended practice for architectural description>>$
+\subsection{IEEE Conformance}
 
-$<<Consider the definition of software architecture by [Perry and Wolf, 1992]. Discuss what the 'elements', 'form', and 'rationale' according to this definition would be for the HS07 system>>$
+$<<$For the architectural description above, discuss what (if anything) should be changed or added for it to comply with the IEEE recommended practice for architectural description$>>$
+
+The following are the problems identified in order to comply with \cite{ieeerecommendedpractice}.
+
+Clause 5.2: The stakeholders and their precise concerns need to be identified for each view.
+
+Clause 5.5: Consistency checks across all existing views.
+
+\subsection{Perry \& Wolf Considerations}
+$<<$Consider the definition of software architecture by [Perry and Wolf, 1992]. Discuss what the 'elements', 'form', and 'rationale' according to this definition would be for the HS07 system$>>$
 

### 2008-02-04 SAiP/Module1/H1/doc/paper.tex
Adding comments to sequence diagram

@@
 \end{figure}
-The sequence diagram illustrates the timing of the interactions.
+The sequence diagram illustrates the protocol of the interactions involved in the setup and the execution of the Gateway.
+It consists of a registration sequence and an infinite loop of retrieving the temperature of all registered themometers and nitifying all observers (in this case the Radiators).
+The example is slightly simplified to improve readability, as the registerThermometer, registerObserver, getTemperature and notify is actually network calls performed using the Invoker and the respective servers.
 \begin{figure}[!htb]

### 2008-02-04 trunk/SAiP/Module1/H1/doc/paper.tex
Angle set to 90 on sequence diagram

@@
 \end{figure}
-The sequence diagram illustrates the timing of the interactions.
+The sequence diagram illustrates the protocol of the interactions involved in the setup and the execution of the Gateway.
+It consists of a registration sequence and an infinite loop of retrieving the temperature of all registered themometers and nitifying all observers (in this case the Radiators).
+The example is slightly simplified to improve readability, as the registerThermometer, registerObserver, getTemperature and notify is actually network calls performed using the Invoker and the respective servers.
 \begin{figure}[!htb]
-\includegraphics[viewport=0 10 500 650,scale=0.3]{figures/sequence.pdf}
+\includegraphics[angle=90,scale=1.2]{figures/sequence.pdf}
 \caption{HS07 Sequence Diagram}

### 2008-02-04 SAiP/Module1/H1/doc/paper.tex
Module View and Component and Connector updated plus small changes.

@@
 
-The figure shows the packages of the HS07 system, and how they interact 
-with java packages used.
+Figure~\ref{fig:mv} shows the packages of the HS07 system, and how they interact 
+with java packages used. The Sensor, Gateway and Actuator packages do not have
+any knowledge of each other. They are all services that are able to respond to
+HTTP queries. They uses the Invoker to interact with the other modules' HTTP
+server. The interactions between the components are described in section
+\ref{sec:component}. The Java packages Net, Jetty and Http are used for the
+HTTP servers and HTTP requests.
 
@@
 \subsection{Component \& Connector Viewpoint}
+\label{sec:component}
 
 The Component \& Connecter view consists of an Active Objects diagram and a
-sequence diagram. The Active Objects diagram shows the active objects of 
-the HS07 system, and how they interact.
+sequence diagram.
+
+The Active Objects diagram in figure \ref{fig:cc_ao} shows the active objects
+of the HS07 system, and how they interact. The Gateway accesses the thermometers
+by requesting the temperature repeatedly from all registered TemperatureServices.
+The measured temperature is then published to all registered observers. The
+RadiatorService and ThermometerService registers to the Gateway as Observers and
+Servers respectively through the GatewayService. It is also possible to register
+to the Gateway from an internet connection.
 
@@
 \begin{figure}[!htb]
-\includegraphics[viewport=0 10 500 650,scale=0.3]{figures/sequence.pdf}
+\includegraphics[angle=90,scale=1.2]{figures/sequence.pdf}
 \caption{HS07 Sequence Diagram}
@@
 \end{figure}
+\clearpage
 
@@
 \subsection{Strength and Limitations of the approach}
-$<<$What are the strengths and limitations of this approach to
-architectural description judging from this case?$>>$
+%$<<$What are the strengths and limitations of this approach to architectural description judging from this case?$>>$
 
@@
 be finding the correct balance in the choice of detail level. It
-is easy to get to much information, cluttering the purpose of the 
+is easy to get too much information, cluttering the purpose of the 
 AD, which in a basic sense is to clarify and present at wider understanding of
@@
 The purpose of doing a reverse engineered AD must either be to evolve
-the system of to extract clarification. Care should be taken to achieve either
+the system or to extract clarification. Care should be taken to achieve either
 goal. 
 \subsection{Whats missing?}
-$<<$Are there aspects of the software architecture that have not been
-properly described?$>>$
+%$<<$Are there aspects of the software architecture that have not been properly described?$>>$
 Stakeholders and their concerns. This encompasses both users/aquirers and developers. We also need to discuss the feasibility and maintainablity of the system.
@@
 
-$<<$For the architectural description above, discuss what (if anything) should be changed or added for it to comply with the IEEE recommended practice for architectural description$>>$
+%$<<$For the architectural description above, discuss what (if anything) should be changed or added for it to comply with the IEEE recommended practice for architectural description$>>$
 
@@
 \subsection{Perry \& Wolf Considerations}
-$<<$Consider the definition of software architecture by [Perry and Wolf, 1992]. Discuss what the 'elements', 'form', and 'rationale' according to this definition would be for the HS07 system$>>$
+%$<<$Consider the definition of software architecture by [Perry and Wolf, 1992]. Discuss what the 'elements', 'form', and 'rationale' according to this definition would be for the HS07 system$>>$
 
@@
 % =====================================================================
-\section{...}
-
-% =====================================================================
 \bibliography{paper}

### 2008-02-05 trunk/SAiP/Module1/H1/doc/paper.tex
\clearpage added after the large sequence diagram. This maybe needs to be removed if it does not look good in the end.
Merged Discussion.txt into the doc

@@
   include new types of sensors and actuators.
-  \item $<<$Extra quality requirement that you consider important $>>$
+%  \item $<<$Extra quality requirement that you consider important $>>$
 \end{itemize}
@@
 \subsection{Strength and Limitations of the approach}
-$<<$What are the strengths and limitations of this approach to
-architectural description judging from this case?$>>$
+%$<<$What are the strengths and limitations of this approach to architectural description judging from this case?$>>$
+\subsection{Strengths}
+\begin{itemize}
+\item UML is a language that is unambigious and known by many.
+\item The three viewpoints are tried and proven, and due to this fact many developers are used to thinking in this way and can generate an understanding of the system based on these views.
+ \item Having a well-understandable architecture to go by, makes it easier for multiple developers to work in parallel with a common understanding of the priorities inherint in the architecture.
+\end{itemize}
 
-When using the approach described in \cite{christensen2004archdesc} with reverse 
-engineering to achieve an architectural description, the dangers posed seem to 
-be finding the correct balance in the choice of detail level. It
-is easy to get to much information, cluttering the purpose of the 
-AD, which in a basic sense is to clarify and present at wider understanding of
-the system.
+\subsection{Limitations}
+\begin{itemize}
+\item Just like with any other language it only makes sense to those who know the language. 
+\item Choosing three views to focus on, due to the fact that they fit most of the time, means that the focus will sometimes be wrong.
+\end{itemize}
 
-Experienced architects might recognize patterns emerging from the process, and 
-immediately see possible design flaws in the implementation of these, thus
-being able to help improve the quality of any evolving system.
-
-The purpose of doing a reverse engineered AD must either be to evolve
-the system of to extract clarification. Care should be taken to achieve either
-goal. 
-\subsection{Whats missing?}
-$<<$Are there aspects of the software architecture that have not been
-properly described?$>>$
-Stakeholders and their concerns. This encompasses both users/aquirers and developers. We also need to discuss the feasibility and maintainablity of the system.
+\subsection{Not covered aspects}
+%$<<$Are there aspects of the software architecture that have not been properly described?$>>$
+\begin{itemize}
+\item \emph{Stakeholders and their concerns}. This encompasses both users/aquirers and developers. \item \emph{Ffeasibility}
+\item \emph{Maintainablity}
+\item \emph{Security} How is the Gateway protected? What protection is there against attack and inproper use? 
+\item \emph {Safety} How is synchronization done?
+What insurances is there that the system cannot mailfunction and overheat a radiator or turn on a radiator which has been turned off deliberately because flammable material is placed near it?
+\item \emph{Availability} What is the up-time of the HTTP server that the system run on, both embedded and PC?
+\item \emph{Testability} What is the tesability of the components? How easily is it to replace the actual HTTP servers with test doubles?
+\end{itemize}
 

### 2008-02-05 SAiP/Module1/H1/doc/paper.tex
Updating the sequence diagram with some variable names and fixing type-o in LaTex.

@@
 \end{figure}
-The sequence diagram illustrates the protocol of the interactions involved in the setup and the execution of the Gateway.
-It consists of a registration sequence and an infinite loop of retrieving the temperature of all registered themometers and nitifying all observers (in this case the Radiators).
-The example is slightly simplified to improve readability, as the registerThermometer, registerObserver, getTemperature and notify is actually network calls performed using the Invoker and the respective servers.
+
+The sequence diagram illustrates the protocol of the interactions involved in the 
+setup and the execution of the Gateway. It consists of a registration sequence and 
+an infinite loop of retrieving the temperature of all registered themometers and 
+notifying all observers (in this case the Radiators). The example is slightly 
+simplified to improve readability, as the registerThermometer, registerObserver, 
+getTemperature and notify is actually network calls performed using the Invoker and 
+the respective servers.
+
+\clearpage
+
 \begin{figure}[!htb]
-\includegraphics[angle=90,scale=1.2]{figures/sequence.pdf}
+\center {\includegraphics[angle=270,scale=0.5]{figures/sequence.pdf}}
 \caption{HS07 Sequence Diagram}
@@
 \end{figure}
-\clearpage
 
@@
 \begin{figure}[!htb]
-\center {\includegraphics[viewport=0 10 700 350,scale=0.4]{figures/deployment.pdf}}
+\center {\includegraphics[viewport=0 10 700 350,scale=0.5]{figures/deployment.pdf}}
 \caption{HS07 Allocation Diagram}
@@
 %$<<$What are the strengths and limitations of this approach to architectural description judging from this case?$>>$
+\subsection{Strengths}
+\begin{itemize}
+\item UML is a language that is unambigious and known by many.
+\item The three viewpoints are tried and proven, and due to this fact many developers are used to thinking in this way and can generate an understanding of the system based on these views.
+ \item Having a well-understandable architecture to go by, makes it easier for multiple developers to work in parallel with a common understanding of the priorities inherint in the architecture.
+\end{itemize}
 
-When using the approach described in \cite{christensen2004archdesc} with reverse 
-engineering to achieve an architectural description, the dangers posed seem to 
-be finding the correct balance in the choice of detail level. It
-is easy to get too much information, cluttering the purpose of the 
-AD, which in a basic sense is to clarify and present at wider understanding of
-the system.
+\subsection{Limitations}
+\begin{itemize}
+\item Just like with any other language it only makes sense to those who know the language. 
+\item Choosing three views to focus on, due to the fact that they fit most of the time, means that the focus will sometimes be wrong.
+\end{itemize}
 
-Experienced architects might recognize patterns emerging from the process, and 
-immediately see possible design flaws in the implementation of these, thus
-being able to help improve the quality of any evolving system.
-
-The purpose of doing a reverse engineered AD must either be to evolve
-the system or to extract clarification. Care should be taken to achieve either
-goal. 
-\subsection{Whats missing?}
+\subsection{Not covered aspects}
 %$<<$Are there aspects of the software architecture that have not been properly described?$>>$
-Stakeholders and their concerns. This encompasses both users/aquirers and developers. We also need to discuss the feasibility and maintainablity of the system.
+\begin{itemize}
+\item \emph{Stakeholders and their concerns}. This encompasses both users/aquirers and developers. \item \emph{Ffeasibility}
+\item \emph{Maintainablity}
+\item \emph{Security} How is the Gateway protected? What protection is there against attack and inproper use? 
+\item \emph {Safety} How is synchronization done?
+What insurances is there that the system cannot mailfunction and overheat a radiator or turn on a radiator which has been turned off deliberately because flammable material is placed near it?
+\item \emph{Availability} What is the up-time of the HTTP server that the system run on, both embedded and PC?
+\item \emph{Testability} What is the tesability of the components? How easily is it to replace the actual HTTP servers with test doubles?
+\end{itemize}
 

### 2008-02-05 trunk/SAiP/Module1/H1/doc/paper.tex
Document update

@@
 \section{Architectural Requirements}
+\label{sec:requirements}
 
@@
 
-The allocation viewpoint illustrates how components are deployed in actual processes
-within the HS07 system.
+The allocation viewpoint illustrates how components are deployed in actual
+processes within the HS07 system.
+
+Figure~\ref{fig:allocation_pc} shows the setup as specified in the build-file
+where all HS07 processes are located on the same node. As can be seen in the
+figure the gateway can be accessed from one or more PC through an internet
+connection.
 
 \begin{figure}[!htb]
-\center {\includegraphics[viewport=0 10 700 350,scale=0.4]{figures/deployment.pdf}}
-\caption{HS07 Allocation Diagram}
-\label{fig:allocation}
+\center {\includegraphics[scale=0.4]{figures/deployment.pdf}}
+\caption{HS07 Allocation Diagram showing deployment on a PC}
+\label{fig:allocation_pc}
+\end{figure}
+
+Figure~\ref{fig:allocation_actual} shows a more realistic example of a deployment
+where thermometers and radiators are seperate nodes that are connected to
+the gateway through a LAN connection.
+
+\begin{figure}[!htb]
+\center {\includegraphics[viewport=110 200 600 420, scale=0.4]{figures/deployment_actual.pdf}}
+\caption{HS07 Allocation Diagram showing deployment on seperate nodes}
+\label{fig:allocation_actual}
 \end{figure}
@@
 \begin{itemize}
-\item \emph{Stakeholders and their concerns}. This encompasses both users/aquirers and developers. \item \emph{Ffeasibility}
+\item \emph{Stakeholders and their concerns}. This encompasses both users/aquirers and developers.
+\item \emph{Feasibility}
 \item \emph{Maintainablity}
@@
 
-\cite{perrywolf1992} defines "elements" as three different classes, data elements, processing elements and connecting elements.
-In HS07 the data elements are temperatures, the processing element is the gateway and the connecting element is HTTP.
-The "form", as described by Perry and Wolf, is the choice of the observer pattern in HS07, and the "rationale" for doing so are defined by the Achitectural Requirements described elsewhere in this document.
+\cite{perrywolf1992} defines "elements" as three different classes, data elements,
+processing elements and connecting elements. In HS07 the data elements are
+temperatures, the processing element is the gateway and the connecting element
+is HTTP. The "form", as described by Perry and Wolf, is the choice of the observer
+pattern in HS07.
+
+The "rationale" described in \cite{perrywolf1992} captures the reasons for choices
+made when defining the architecture. In this architectural description a rationale
+would e.g. document the choice of using HTTP as the protocol between the components. A
+Rationale is not part of this document.
 

### 2008-02-06 trunk/SAiP/Module1/H1/doc/paper.tex
Update of the IEEE Conformance section and other smaller changes.

@@
 \usepackage{rotating}
+
+\renewcommand{\descriptionlabel}[1]
+    {\hspace{\labelsep}\emph{#1}}
 
@@
 
-\date{2008-02-04 13:32}
+\date{2008-02-06}
 
@@
 % ---------------------------------------------------------------------
-\subsection{Module Viewpoint}
+\subsection{Module View}
 
-The module viewpoint of the HS07 system.
+This section contains the module view of the HS07 system. This view
+is based on the Module Viewpoint as defined in \cite{christensen2004archdesc}.
 
@@
 % ---------------------------------------------------------------------
-\subsection{Component \& Connector Viewpoint}
+\subsection{Component \& Connector View}
 \label{sec:component}
+
+This section contains the Component \& Connecter view of the HS07 system.
+This view is based on the Component \& Connecter viewpoint as defined in
+\cite{christensen2004archdesc}.
 
@@
 % ---------------------------------------------------------------------
-\subsection{Allocation Viewpoint}
+\subsection{Allocation View}
 
-The allocation viewpoint illustrates how components are deployed in actual
+This section contains the allocation view of the HS07 system. This view
+is based on the Allocation Viewpoint as defined in \cite{christensen2004archdesc}.
+
+The allocation view illustrates how components are deployed in actual
 processes within the HS07 system.
@@
 %$<<$Are there aspects of the software architecture that have not been properly described?$>>$
-\begin{itemize}
-\item \emph{Stakeholders and their concerns}. This encompasses both users/aquirers and developers.
-\item \emph{Feasibility}
-\item \emph{Maintainablity}
-\item \emph{Security} How is the Gateway protected? What protection is there against attack and inproper use? 
-\item \emph {Safety} How is synchronization done?
+\begin{description}
+\item[Stakeholders and their concerns] This encompasses both users/aquirers and developers.
+\item[Feasibility]
+\item[Maintainablity]
+\item[Security] How is the Gateway protected? What protection is there against attack and inproper use? 
+\item[Safety] How is synchronization done?
 What insurances is there that the system cannot mailfunction and overheat a radiator or turn on a radiator which has been turned off deliberately because flammable material is placed near it?
-\item \emph{Availability} What is the up-time of the HTTP server that the system run on, both embedded and PC?
-\item \emph{Testability} What is the tesability of the components? How easily is it to replace the actual HTTP servers with test doubles?
-\end{itemize}
+\item[Availability] What is the up-time of the HTTP server that the system run on, both embedded and PC?
+\item[Testability] What is the testability of the components? How easily is it to replace the actual HTTP servers with test doubles?
+\end{description}
 
@@
 
-The following are the problems identified in order to comply with \cite{ieeerecommendedpractice}.
+This section describes how this architectural description conforms to recommended practice
+that is defined in \cite{ieeerecommendedpractice}.
 
-Clause 5.2: The stakeholders and their precise concerns need to be identified for each view.
-
-Clause 5.5: Consistency checks across all existing views.
+\begin{description}
+	\item[Clause 5.1 Architectural documentation] This report does not include a change history
+		and a glossary as defined in the recommended practice. (XXX: scope, context?)
+	\item[Clause 5.2 Identification of stakeholders and concers] The stakeholders of this system and
+		their precise concerns have not been identified in this report. This must be done in order to
+		conform to the recommended practice.
+	\item[Clause 5.3 Selection of architectural viewpoints] The viewpoints used in this architectural
+		description are defined in \cite{christensen2004archdesc}.
+	\item[Clause 5.4 Architectural views] This report contains the following architectural views: Module
+		View, Component \& Connector View and Allocation View.
+	\item[Clause 5.5 Consistency among architectural views] There is no consistency check across the views
+		in this report as specified in the recommended practice.
+	\item[Clause 5.6 Architectural rationale] There is no rationale for the architectural decisions
+		that have made in this report.
+\end{description}
 

### 2008-02-06 trunk/SAiP/Module1/H1/doc/paper.tex
Final updates from LKR

@@
 
-The major driving qualities attributes of the HS07 system
+The major driving quality attributes of the HS07 system
 are\footnote{These qualities will be operationalized in Exercise 2}:
@@
 
-Figure~\ref{fig:mv} shows the packages of the HS07 system, and how they interact 
-with java packages used. The Sensor, Gateway and Actuator packages do not have
+Figure~\ref{fig:mv} shows the packages of the HS07 system, and how they depend
+on java packages. The Sensor, Gateway and Actuator packages do not have
 any knowledge of each other. They are all services that are able to respond to
-HTTP queries. They uses the Invoker to interact with the other modules' HTTP
-server. The interactions between the components are described in section
+HTTP queries. They use the Invoker to interact with the HTTP server in the
+other modules. The interactions between the components are described in section
 \ref{sec:component}. The Java packages Net, Jetty and Http are used for the
@@
 getTemperature and notify is actually network calls performed using the Invoker and 
-the respective servers.
+the respective servers. The figure shows an example with only one thermometer and
+one radiator.
 
@@
 
-Figure~\ref{fig:allocation_pc} shows the setup as specified in the build-file
+Figure~\ref{fig:allocation_pc} shows an example setup
 where all HS07 processes are located on the same node. As can be seen in the
-figure the gateway can be accessed from one or more PC through an internet
+figure the gateway can be accessed from one or more PCs through an internet
 connection.
+
+\clearpage
 
@@
 \begin{figure}[!htb]
-\center {\includegraphics[viewport=110 200 600 420, scale=0.4]{figures/deployment_actual.pdf}}
+\center {\includegraphics[viewport=250 200 600 420, scale=0.4]{figures/deployment_actual.pdf}}
 \caption{HS07 Allocation Diagram showing deployment on seperate nodes}
@@
 %$<<$Are there aspects of the software architecture that have not been properly described?$>>$
-\begin{description}
-\item[Stakeholders and their concerns] This encompasses both users/aquirers and developers.
-\item[Feasibility]
-\item[Maintainablity]
-\item[Security] How is the Gateway protected? What protection is there against attack and inproper use? 
-\item[Safety] How is synchronization done?
+\begin{itemize}
+\item \emph{Stakeholders and their concerns} This encompasses both users/aquirers and developers.
+\item \emph{Feasibility}
+\item \emph{Maintainablity}
+\item \emph{Security} How is the Gateway protected? What protection is there against attack and inproper use? 
+\item \emph{Safety} How is synchronization done?
 What insurances is there that the system cannot mailfunction and overheat a radiator or turn on a radiator which has been turned off deliberately because flammable material is placed near it?
-\item[Availability] What is the up-time of the HTTP server that the system run on, both embedded and PC?
-\item[Testability] What is the testability of the components? How easily is it to replace the actual HTTP servers with test doubles?
-\end{description}
+\item \emph{Availability} What is the up-time of the HTTP server that the system run on, both embedded and PC?
+\item \emph{Testability} What is the testability of the components? How easily is it to replace the actual HTTP servers with test doubles?
+\end{itemize}
 
@@
 
-This section describes how this architectural description conforms to recommended practice
+This section describes how this architectural description conforms to the recommended practice
 that is defined in \cite{ieeerecommendedpractice}.

### 2008-02-06 trunk/SAiP/Module1/H1/doc/paper.tex
Comment about scope and context added.

@@
 \begin{description}
-	\item[Clause 5.1 Architectural documentation] This report does not include a change history
-		and a glossary as defined in the recommended practice. (XXX: scope, context?)
+	\item[Clause 5.1 Architectural documentation] This report does not include a description of the
+		scope of this document, a change history and a glossary as defined in the recommended practice.
+		The context is briefly covered in the introduction.
 	\item[Clause 5.2 Identification of stakeholders and concers] The stakeholders of this system and

### 2008-02-19 SAiP/Module1/H2/table.tex
Adding LaTex document with the first scenario begun

\documentclass[10pt,a4paper]{article}

\usepackage{times}
\usepackage[latin1]{inputenc}
\usepackage{rotating}
\usepackage{multirow}

\begin{document}

\begin{table}[!htp]
\begin{center}
\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & After an average user has used the system for less than one hour e or she has mastered the basic functionality \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} usability & \\
  \hline
  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
  & {\bfseries Source:} &  \\
  \cline{2-3}
  & {\bfseries Stimulus:} &  \\
  \cline{2-3}
  & {\bfseries Artifact} &  \\
  \cline{2-3}
  & {\bfseries Environment:} &  \\
  \cline{2-3}
  & {\bfseries Response:} &  \\
  \cline{2-3}
  & {\bfseries Response Measure:} & \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} &  \\
  \hline
  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
  \hline
\end{tabular}
\caption{Scenario refinement for HS07-...}
\end{center}
\end{table}

\end{document}

### 2008-02-20 SAiP/Module1/H2/table.tex
Updating with more tables

@@
   \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} usability & \\
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Usability \\
   \hline
   \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
-  & {\bfseries Source:} &  \\
+  & {\bfseries Source:} & The average user \\
   \cline{2-3}
-  & {\bfseries Stimulus:} &  \\
+  & {\bfseries Stimulus:} & Begin using \\
   \cline{2-3}
-  & {\bfseries Artifact} &  \\
+  & {\bfseries Artifact} & The system \\
   \cline{2-3}
-  & {\bfseries Environment:} &  \\
+  & {\bfseries Environment:} & Normal operation \\
   \cline{2-3}
-  & {\bfseries Response:} &  \\
+  & {\bfseries Response:} & Understand basic functionality \\
   \cline{2-3}
-  & {\bfseries Response Measure:} & \\
+  & {\bfseries Response Measure:} & Less than 1 hour \\
   \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} &  \\
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & What constitutes basic functionality? What is an average user? \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} & Update rather than simply translate manual to suite different cultures \\
+  \hline
+\end{tabular}
+\caption{Scenario refinement for HS07}
+\end{center}
+\end{table}
+
+\begin{table}[!htp]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The user adds a new actuator and the system configures to use the new hardware automatically within 1 minute \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability, Usability, Performance \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The user \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Attaches a new actuator \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The system \\
+  \cline{2-3}
+  & {\bfseries Environment:} & Maintanance operation \\
+  \cline{2-3}
+  & {\bfseries Response:} & New hardware is automatically added to the system configuration and accessible remotely \\
+  \cline{2-3}
+  & {\bfseries Response Measure:} & Within one minute \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & Is the user allwed to install now components himself? \\
   \hline
@@
 \end{tabular}
-\caption{Scenario refinement for HS07-...}
+\caption{Scenario refinement for HS07}
+\end{center}
+\end{table}
+
+\begin{table}[!htp]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The user changes the preferred temperature from a web page and the system adapts to it \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Usability,  Availability \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The user \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Remotely change the temperature \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The system \\
+  \cline{2-3}
+  & {\bfseries Environment:} & Normal operation \\
+  \cline{2-3}
+  & {\bfseries Response:} & Begin adjusting to temperature \\
+  \cline{2-3}
+  & {\bfseries Response Measure:} & Within 5 seconds \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & When the the temperature adjustment begin? When the command is received by the radiator? \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
+  \hline
+\end{tabular}
+\caption{Scenario refinement for HS07}
+\end{center}
+\end{table}
+
+\begin{table}[!htp]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & A sensor stops responding \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifyability \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & A sensor \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Stops communicating with gatekeper \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The system \\
+  \cline{2-3}
+  & {\bfseries Environment:} & Normal operation \\
+  \cline{2-3}
+  & {\bfseries Response:} & Removes sensor from configuration and display error message for user \\
+  \cline{2-3}
+  & {\bfseries Response Measure:} & Within 30 seconds \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & How is it know that it stops responding? Heartbeat? \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
+  \hline
+\end{tabular}
+\caption{Scenario refinement for HS07}
 \end{center}

### 2008-02-20 SAiP/Module1/H2/paper.tex
import from wiki

--- 2008-02-06 trunk/SAiP/Module1/H1/doc/paper.tex
@@
 \usepackage{rotating}
+\usepackage{multirow}
 
@@
 
+% ---------------------------------------------------------------------
+\section{Quality Attributes}
+The following chapters documents the selection of the Quality Attributes that we 
+select for the HS07 system. The selection of QA is split into a 3-part process.
+\begin{itemize}
+\item Brainstorming - we think up scenarios
+\item Selection - we chose which scenarios are interesting.
+\item Refinement - we refine the selected scenarios
+\end{itemize}
+\subsection{Brainstorming}
+In this part of the process we just attempt to think of as many possible scenarios for the HS07 system as we possibly can.
+
+\subsection{Scenarios}
+
+\begin{enumerate}
+\item A sensor stops responding.
+\item Someone leaves a window open in an ice storm, and the sensors makes the actuators overheat.
+\item Someone accidentally puts their stiletto through the Ethernet switch, disabling networking 
+parts of the system.
+\item After an average user has used the system for less than one hour e or she has mastered the basic functionality
+\item The user adds a new actuator and the system configures to use the new hardware automatically within 1 minute.
+\item The user logs on externally through the gateway and has access to the home control within 1 second of verifying 95\% of the time (5 seconds 4\% of the time, more than 5 seconds (unscheduled down-time) less than 1\%) is connected and has access
+\item The user attempts to turn radiator above normal safety-level from outside the house and receive an error back explaining the infraction.
+The system stores a log of the episode and refuse to execute the command.
+\item An attacker attempts to log on using an invalid password and is thrown off. The next attempt will take longer (progressive log-on time). The system stores a log of the failed attempt.
+\item Developer updates and tests the code base with a new device type in 1 day.
+\item A technical installer installs a system at a customer in no more than 4 hours.
+\item Developer updates the code base to a new requirement in less than 30 man-hours
+\item Configuration manager installs an upgrade remotely at a customer in no more than 30min
+\item Tester/debugger extract debugging information directly from running system at customer.
+\item The system handles 50 sensors and 50 actuators.
+\item The system handles up to 50 rooms/groups with one or more sensors and actuators in each group.
+\item The user changes the preferred temperature from a web page and the system adapts to it.
+\item The customer requests that the system runs on other hardware (e.g. less power consuming hardware).
+\item The customer requests that the system interfaces with other sensors/actuators?
+\item The developer modifies the sensor poll interval at build time.
+\item The system turns off all actuators when sensors cannot be probed.
+\item The end user adds a sensor/actuator at runtime which registers automatically.
+\end{enumerate}
+
+\subsection{Selection}
+Here we put the scenarios to the vote. This is done by emulation of a real-life situation where the actual stakeholder would
+be part of the process. We take on the roles of the stakeholders in their absence, and chose which scenarios would be more
+interesting for us.
+\subsection{Stakeholders and interests}
+\begin{itemize}
+\item User (homeowner)
+\begin{itemize}
+\item Easy to use and configure
+\item Always available
+\item Safe (will not make an error and burn down the house)
+\item Secure (other people cannot control my home)
+\end{itemize}
+\item Producer
+\begin{itemize}
+\item Cheap
+\item Production line (modifiability for other uses)
+\end{itemize}
+\item Technical installer
+\begin{itemize}
+\item Easily deployable (installable)
+\end{itemize}
+\item Developer (?)
+\begin{itemize}
+\item Easily adjustable to new requirements
+\item Scalable
+\item Remote update (upgrading online)
+\item Remote maintenance (debugging online)
+\item Automatically testable (regression test)
+\item Developing in parallel
+\end{itemize}
+\item Insurance company
+\begin{itemize}
+\item Safe with respect to fires (radiators) and break ins (door), ...
+\end{itemize}
+\end{itemize}
+
+\subsection{The Vote}
+\subsection{First vote}
+\begin{tabular}{|l|c|c|c|c|c|c|}
+\hline
+Person & Vote 1 & Vote 2 & Vote 3 & Vote 4 & Vote 5 & Vote 6\\
+\hline
+Jesper& & & & & & \\
+Lars& & & & & & \\
+Anders&2&4& & & & \\
+\hline
+\end{tabular}
+\subsection{Second vote}
+\begin{tabular}{|l|c|c|c|c|c|c|}
+\hline
+Person & Vote 1 & Vote 2 & Vote 3 & Vote 4 & Vote 5 & Vote 6\\
+\hline
+Jesper& & & & & & \\
+Lars& & & & & & \\
+Anders& & & & & & \\
+\hline
+\end{tabular}
+
+\subsection{Refinement}
+The process of refining the chosen scenarios centers around specification of various elements of the scenario. Stimulus, Response, Response Measurement. 
+\begin{table}[!htp]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & After an average user has used the system for less than one hour e or she has mastered the basic functionality \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} usability & \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} &  \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} &  \\
+  \cline{2-3}
+  & {\bfseries Artifact} &  \\
+  \cline{2-3}
+  & {\bfseries Environment:} &  \\
+  \cline{2-3}
+  & {\bfseries Response:} &  \\
+  \cline{2-3}
+  & {\bfseries Response Measure:} & \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} &  \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
+  \hline
+\end{tabular}
+\caption{Scenario refinement for HS07-...}
+\end{center}
+\end{table}
+
 % =====================================================================

### 2008-02-21 trunk/SAiP/Module1/H2/paper.tex
more in line with the problem description

--- 2008-02-20 SAiP/Module1/H2/paper.tex
@@
 
-\title{Exercise 1: Software Architecture Description of the HS07 System}
+\title{Exercise 2: Quality Attributes and Architectural Design of the HS07 System}
 
@@
 
-\date{2008-02-06}
+\date{2008-02-21}
 
@@
   measurements HS07 adjusts radiators in the home. This report gives a
-  software architecture description of an architectural prototype of
-  the HS07 system. The techniques used for architectural description
-  are taken from \cite{christensen2004archdesc}.
+  quality attribute analysis and discussion and a architectural design of 
+  the HS07 system. The techniques used for both quality attributes and 
+  architectural design are taken from \cite{bass2003sa}.
 \end{abstract}
@@
 The major driving quality attributes of the HS07 system
-are\footnote{These qualities will be operationalized in Exercise 2}:
+are:
 
@@
   include new types of sensors and actuators.
-%  \item $<<$Extra quality requirement that you consider important $>>$
+\item \emph{Usability.} It must be easy to use the gateway user interface.
+  include new types of sensors and actuators.
+\item \emph{Availability.} The system must be responsive and available for 
+both users and hardware elements.
+\item \emph{Safety.} The system must take care not to overheat the actuators.
 \end{itemize}
@@
 % =====================================================================
-\section{Architectural Description}
-
-% ---------------------------------------------------------------------
-\subsection{Module View}
-
-This section contains the module view of the HS07 system. This view
-is based on the Module Viewpoint as defined in \cite{christensen2004archdesc}.
-
-\begin{figure}[!htb]
-\center {\includegraphics[viewport=0 10 600 500,scale=0.5]{figures/mv.pdf}}
-\caption{HS07 Package Diagram}
-\label{fig:mv}
-\end{figure}
-
-Figure~\ref{fig:mv} shows the packages of the HS07 system, and how they depend
-on java packages. The Sensor, Gateway and Actuator packages do not have
-any knowledge of each other. They are all services that are able to respond to
-HTTP queries. They use the Invoker to interact with the HTTP server in the
-other modules. The interactions between the components are described in section
-\ref{sec:component}. The Java packages Net, Jetty and Http are used for the
-HTTP servers and HTTP requests.
-
-% ---------------------------------------------------------------------
-\subsection{Component \& Connector View}
-\label{sec:component}
-
-This section contains the Component \& Connecter view of the HS07 system.
-This view is based on the Component \& Connecter viewpoint as defined in
-\cite{christensen2004archdesc}.
-
-The Component \& Connecter view consists of an Active Objects diagram and a
-sequence diagram.
-
-The Active Objects diagram in figure \ref{fig:cc_ao} shows the active objects
-of the HS07 system, and how they interact. The Gateway accesses the thermometers
-by requesting the temperature repeatedly from all registered TemperatureServices.
-The measured temperature is then published to all registered observers. The
-RadiatorService and ThermometerService registers to the Gateway as Observers and
-Servers respectively through the GatewayService. It is also possible to register
-to the Gateway from an internet connection.
-
-\begin{figure}[!htb]
-\center {\includegraphics[viewport=0 10 600 420,scale=0.5]{figures/cc_ao.pdf}}
-\caption{HS07 Active Objects}
-\label{fig:cc_ao}
-\end{figure}
-
-The sequence diagram illustrates the protocol of the interactions involved in the 
-setup and the execution of the Gateway. It consists of a registration sequence and 
-an infinite loop of retrieving the temperature of all registered themometers and 
-notifying all observers (in this case the Radiators). The example is slightly 
-simplified to improve readability, as the registerThermometer, registerObserver, 
-getTemperature and notify is actually network calls performed using the Invoker and 
-the respective servers. The figure shows an example with only one thermometer and
-one radiator.
-
-\clearpage
-
-\begin{figure}[!htb]
-\center {\includegraphics[angle=270,scale=0.5]{figures/sequence.pdf}}
-\caption{HS07 Sequence Diagram}
-\label{fig:sequence}
-\end{figure}
-
-% ---------------------------------------------------------------------
-\subsection{Allocation View}
-
-This section contains the allocation view of the HS07 system. This view
-is based on the Allocation Viewpoint as defined in \cite{christensen2004archdesc}.
-
-The allocation view illustrates how components are deployed in actual
-processes within the HS07 system.
-
-Figure~\ref{fig:allocation_pc} shows an example setup
-where all HS07 processes are located on the same node. As can be seen in the
-figure the gateway can be accessed from one or more PCs through an internet
-connection.
-
-\clearpage
-
-\begin{figure}[!htb]
-\center {\includegraphics[scale=0.4]{figures/deployment.pdf}}
-\caption{HS07 Allocation Diagram showing deployment on a PC}
-\label{fig:allocation_pc}
-\end{figure}
-
-Figure~\ref{fig:allocation_actual} shows a more realistic example of a deployment
-where thermometers and radiators are seperate nodes that are connected to
-the gateway through a LAN connection.
-
-\begin{figure}[!htb]
-\center {\includegraphics[viewport=250 200 600 420, scale=0.4]{figures/deployment_actual.pdf}}
-\caption{HS07 Allocation Diagram showing deployment on seperate nodes}
-\label{fig:allocation_actual}
-\end{figure}
-
-% ---------------------------------------------------------------------
 \section{Quality Attributes}
@@
 Lars& & & & & & \\
-Anders&2&4& & & & \\
+Anders&2/7&4&6/16&8&14&5/21\\
 \hline
@@
 % =====================================================================
-\section{Discussion}
-\subsection{Strength and Limitations of the approach}
-%$<<$What are the strengths and limitations of this approach to architectural description judging from this case?$>>$
-\subsection{Strengths}
-\begin{itemize}
-\item UML is a language that is unambigious and known by many.
-\item The three viewpoints are tried and proven, and due to this fact many developers are used to thinking in this way and can generate an understanding of the system based on these views.
- \item Having a well-understandable architecture to go by, makes it easier for multiple developers to work in parallel with a common understanding of the priorities inherint in the architecture.
-\end{itemize}
-
-\subsection{Limitations}
-\begin{itemize}
-\item Just like with any other language it only makes sense to those who know the language. 
-\item Choosing three views to focus on, due to the fact that they fit most of the time, means that the focus will sometimes be wrong.
-\end{itemize}
-
-\subsection{Not covered aspects}
-%$<<$Are there aspects of the software architecture that have not been properly described?$>>$
-\begin{itemize}
-\item \emph{Stakeholders and their concerns} This encompasses both users/aquirers and developers.
-\item \emph{Feasibility}
-\item \emph{Maintainablity}
-\item \emph{Security} How is the Gateway protected? What protection is there against attack and inproper use? 
-\item \emph{Safety} How is synchronization done?
-What insurances is there that the system cannot mailfunction and overheat a radiator or turn on a radiator which has been turned off deliberately because flammable material is placed near it?
-\item \emph{Availability} What is the up-time of the HTTP server that the system run on, both embedded and PC?
-\item \emph{Testability} What is the testability of the components? How easily is it to replace the actual HTTP servers with test doubles?
-\end{itemize}
-
-\subsection{IEEE Conformance}
-
-%$<<$For the architectural description above, discuss what (if anything) should be changed or added for it to comply with the IEEE recommended practice for architectural description$>>$
-
-This section describes how this architectural description conforms to the recommended practice
-that is defined in \cite{ieeerecommendedpractice}.
-
-\begin{description}
-	\item[Clause 5.1 Architectural documentation] This report does not include a description of the
-		scope of this document, a change history and a glossary as defined in the recommended practice.
-		The context is briefly covered in the introduction.
-	\item[Clause 5.2 Identification of stakeholders and concers] The stakeholders of this system and
-		their precise concerns have not been identified in this report. This must be done in order to
-		conform to the recommended practice.
-	\item[Clause 5.3 Selection of architectural viewpoints] The viewpoints used in this architectural
-		description are defined in \cite{christensen2004archdesc}.
-	\item[Clause 5.4 Architectural views] This report contains the following architectural views: Module
-		View, Component \& Connector View and Allocation View.
-	\item[Clause 5.5 Consistency among architectural views] There is no consistency check across the views
-		in this report as specified in the recommended practice.
-	\item[Clause 5.6 Architectural rationale] There is no rationale for the architectural decisions
-		that have made in this report.
-\end{description}
-
-\subsection{Perry \& Wolf Considerations}
-%$<<$Consider the definition of software architecture by [Perry and Wolf, 1992]. Discuss what the 'elements', 'form', and 'rationale' according to this definition would be for the HS07 system$>>$
-
-\cite{perrywolf1992} defines "elements" as three different classes, data elements,
-processing elements and connecting elements. In HS07 the data elements are
-temperatures, the processing element is the gateway and the connecting element
-is HTTP. The "form", as described by Perry and Wolf, is the choice of the observer
-pattern in HS07.
-
-The "rationale" described in \cite{perrywolf1992} captures the reasons for choices
-made when defining the architecture. In this architectural description a rationale
-would e.g. document the choice of using HTTP as the protocol between the components. A
-Rationale is not part of this document.
-
+\section{Evaluation}
+% =====================================================================
+\section{Architectural Design}
 % =====================================================================

### 2008-02-21 SAiP/Module1/H2/paper.tex
Updating with new votes
Vote1 for Jesper (User)

@@
 \subsection{First vote}
-\begin{tabular}{|l|c|c|c|c|c|c|}
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
 \hline
-Person & Vote 1 & Vote 2 & Vote 3 & Vote 4 & Vote 5 & Vote 6\\
+Person & Stakeholder Role & Vote 1 & Vote 2 & Vote 3 & Vote 4 & Vote 5 & Vote 6\\
 \hline
-Jesper& & & & & & \\
-Lars& & & & & & \\
-Anders&2&4& & & & \\
+Jesper&User&4&5&6/16&20&21&2/7\\
+Lars& & & & & & & \\
+Anders& &2/7&4&6/16&8&14&5/21\\
 \hline
@@
 \subsection{Second vote}
-\begin{tabular}{|l|c|c|c|c|c|c|}
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
 \hline
-Person & Vote 1 & Vote 2 & Vote 3 & Vote 4 & Vote 5 & Vote 6\\
+Person & Stakeholder Role & Vote 1 & Vote 2 & Vote 3 & Vote 4 & Vote 5 & Vote 6\\
 \hline
-Jesper& & & & & & \\
-Lars& & & & & & \\
-Anders& & & & & & \\
+Jesper& & & & & & & \\
+Lars& & & & & & & \\
+Anders& & & & & & & \\
 \hline

### 2008-02-24 SAiP/Module1/H2/paper.tex
Wiki sync

@@
 \item \emph{Usability.} It must be easy to use the gateway user interface.
-  include new types of sensors and actuators.
 \item \emph{Availability.} The system must be responsive and available for 
@@
 \section{Quality Attributes}
+\subsection{Quality Attributes in the HS07 system}
+\subsection{Availability}
+As the HS07 relay on distributed components receiving and transmitting information, as well as remote user access over unreliable communication media outside of the systems control, availability has a high priority.
+Timely availability
+Hard requirements as to timeing of responses from the individual services are not that important in the HS07 system.
+Whether an actuator responds in 1 second or 5 seconds is not that important, just like the internet link in the gateway will not sease to function if its response time is e.g.
+doubled.
+\subsection{Usability}
+The HS07 system must be used by "ordinary" users (e.g.
+non-technical), and a high degree of usability must therefore be inherent in the architecture to support rapid user interface modification as well as implementing functionality that may aid the user during use.
+\subsection{Performance}
+Performance and availability are related in the sense that increasing performance may increase timely availability, yet as the timely availability was deemed of low priority, this part of the performance quality attirbute is also of little importance.
+Another part of performance is related to footprint and memory usage rather than execution speed, and as the system is based on smaller sensors and actuators, this part of performance becomes very important.
+It is also related to another important quality; cost.
+The system must be able to handle enough sensors and actuators to control a full house.
+\subsection{Modifiability}
+As the system is based on dynamically replacable sensors and actuators of varying type and specification, it is imparative that the architecture supports these varying specification, and modifiability therefore become vital for the system.
+\subsection{Security}
+There are always at least two levels to security; percieved security and actual security.
+As long as the sensors and actuators are controlling non-vital functions, e.g.
+heating, lights, etc.
+it may be sufficient with perceived security as an attacker may not endanger safety and have no financial motive.
+On the other hand if the system is extended to control door-locks and house survailance, then actual security becomes a must.
+Perceived security is important to dissuade the fun-time intruders and reassure the customer.
+\subsection{Testability}
+As the system is highly modifiable, regression-testing is highly-important.
+Also the requirement of availability creates extra demands on testability, as it is required to test many fault scenarios to ensure propper recovery (if this is the fault-handling mechanism of choice).
+
+\subsection{Quality Attribute Selection}
 The following chapters documents the selection of the Quality Attributes that we 
@@
 \end{itemize}
-
+\subsection{Identical scenarios}
+Due to the distributed location of the participants in the brainstorming multiple "similar" scenarios was created.
+In order to avoid having the same scenario represented twice with a slightly different wording, the below list show the scenarios which will be considered identical, meaning that their votes will be accumulated.
+\begin{itemize}
+\item 2. and 7.
+\item 6. and 16.
+\item 10. and 17.
+\item 9. and 18.
+\item 1. and 20.
+\item 5. and 21.
+\end{itemize}
 \subsection{The Vote}
@@
 Jesper&User&4&5&6/16&20&21&2/7\\
-Lars& & & & & & & \\
-Anders& &2/7&4&6/16&8&14&5/21\\
+Lars&Developer&19&19&5/21&9/18&11&13\\
+Anders&Producer&2/7&4&6/16&8&14&5/21\\
 \hline
@@
 \hline
-Jesper& & & & & & & \\
-Lars& & & & & & & \\
-Anders& & & & & & & \\
+Jesper&Developer&11&19&9&11&13&6/16\\
+Lars&User&1/20&4&6/16&15&4&5/21\\
+Anders&Producer&9&11&11&12&17&9/18\\
 \hline
 \end{tabular}
-
+\subsection{Voting Results}
+This is the result of the vote in a prioritized order.
+\begin{enumerate}
+\item 9/18 (6 votes)
+\item 11 (6 votes)
+\item 5/21 (4 votes)
+\item 19 (3 votes)
+\item 4 (3 votes)
+\item 13 (3 votes)
+\item 1/20 (2 votes)
+\item 2/7 (2 votes)
+\item 6/16 (2 votes)
+\item 10/17 (2 votes)
+\item 12 (2 votes)
+\item 15 (1 votes)
+\item 3 (0 votes)
+\item 8 (0 votes)
+\item 14 (0 votes)
+\end{enumerate}
 \subsection{Refinement}
 The process of refining the chosen scenarios centers around specification of various elements of the scenario. Stimulus, Response, Response Measurement.
-\begin{table}[!htp]
-\begin{center}
-\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
-  \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & After an average user has used the system for less than one hour e or she has mastered the basic functionality \\
-  \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} usability & \\
-  \hline
-  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
-  & {\bfseries Source:} &  \\
-  \cline{2-3}
-  & {\bfseries Stimulus:} &  \\
-  \cline{2-3}
-  & {\bfseries Artifact} &  \\
-  \cline{2-3}
-  & {\bfseries Environment:} &  \\
-  \cline{2-3}
-  & {\bfseries Response:} &  \\
-  \cline{2-3}
-  & {\bfseries Response Measure:} & \\
-  \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} &  \\
-  \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
-  \hline
-\end{tabular}
-\caption{Scenario refinement for HS07-...}
-\end{center}
-\end{table}
 

### 2008-02-24 trunk/SAiP/Module1/H2/paper.tex
Fixed copy paste error.
Voting results updated

@@
 \end{itemize}
-
+\subsection{Identical scenarios}
+Due to the distributed location of the participants in the brainstorming multiple "similar" scenarios was created.
+In order to avoid having the same scenario represented twice with a slightly different wording, the below list show the scenarios which will be considered identical, meaning that their votes will be accumulated.
+\begin{itemize}
+\item 2. and 7.
+\item 6. and 16.
+\item 10. and 17.
+\item 9. and 18.
+\item 1. and 20.
+\item 5. and 21.
+\end{itemize}
 \subsection{The Vote}
 \subsection{First vote}
-\begin{tabular}{|l|c|c|c|c|c|c|}
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
 \hline
-Person & Vote 1 & Vote 2 & Vote 3 & Vote 4 & Vote 5 & Vote 6\\
+Person & Stakeholder Role & Vote 1 & Vote 2 & Vote 3 & Vote 4 & Vote 5 & Vote 6\\
 \hline
-Jesper& & & & & & \\
-Lars& & & & & & \\
-Anders&2/7&4&6/16&8&14&5/21\\
+Jesper&User&4&5&6/16&20&21&2/7\\
+Lars&Developer&19&19&5/21&9/18&11&13\\
+Anders&Producer&2/7&4&6/16&8&14&5/21\\
 \hline
@@
 \subsection{Second vote}
-\begin{tabular}{|l|c|c|c|c|c|c|}
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
 \hline
-Person & Vote 1 & Vote 2 & Vote 3 & Vote 4 & Vote 5 & Vote 6\\
+Person & Stakeholder Role & Vote 1 & Vote 2 & Vote 3 & Vote 4 & Vote 5 & Vote 6\\
 \hline
-Jesper& & & & & & \\
-Lars& & & & & & \\
-Anders& & & & & & \\
+Jesper&Developer&11&19&9&11&13&6/16\\
+Lars&User&1/20&4&6/16&15&4&5/21\\
+Anders&Producer&9&6/16&11&12&17&9/18\\
 \hline
 \end{tabular}
-
+\subsection{Voting Results}
+This is the result of the vote in a prioritized order.
+\begin{enumerate}
+\item 9/18 (6 votes)
+\item 11 (5 votes)
+\item 5/21 (4 votes)
+\item 6/16 (4 votes)
+\item 19 (3 votes)
+\item 4 (3 votes)
+\item 13 (2 votes)
+\item 1/20 (2 votes)
+\item 2/7 (2 votes)
+\item 10/17 (2 votes)
+\item 12 (2 votes)
+\item 15 (1 votes)
+\item 3 (0 votes)
+\item 8 (0 votes)
+\item 14 (0 votes)
+\end{enumerate}
 \subsection{Refinement}
 The process of refining the chosen scenarios centers around specification of various elements of the scenario. Stimulus, Response, Response Measurement.
-\begin{table}[!htp]
-\begin{center}
-\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
-  \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & After an average user has used the system for less than one hour e or she has mastered the basic functionality \\
-  \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} usability & \\
-  \hline
-  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
-  & {\bfseries Source:} &  \\
-  \cline{2-3}
-  & {\bfseries Stimulus:} &  \\
-  \cline{2-3}
-  & {\bfseries Artifact} &  \\
-  \cline{2-3}
-  & {\bfseries Environment:} &  \\
-  \cline{2-3}
-  & {\bfseries Response:} &  \\
-  \cline{2-3}
-  & {\bfseries Response Measure:} & \\
-  \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} &  \\
-  \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
-  \hline
-\end{tabular}
-\caption{Scenario refinement for HS07-...}
-\end{center}
-\end{table}
 

### 2008-02-24 SAiP/Module1/H2/table.tex
Updating with needed scenarios

@@
 \begin{document}
-
-\begin{table}[!htp]
-\begin{center}
-\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
-  \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & After an average user has used the system for less than one hour e or she has mastered the basic functionality \\
-  \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Usability \\
-  \hline
-  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
-  & {\bfseries Source:} & The average user \\
-  \cline{2-3}
-  & {\bfseries Stimulus:} & Begin using \\
-  \cline{2-3}
-  & {\bfseries Artifact} & The system \\
-  \cline{2-3}
-  & {\bfseries Environment:} & Normal operation \\
-  \cline{2-3}
-  & {\bfseries Response:} & Understand basic functionality \\
-  \cline{2-3}
-  & {\bfseries Response Measure:} & Less than 1 hour \\
-  \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & What constitutes basic functionality? What is an average user? \\
-  \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} & Update rather than simply translate manual to suite different cultures \\
-  \hline
-\end{tabular}
-\caption{Scenario refinement for HS07}
-\end{center}
-\end{table}
 
@@
 \end{tabular}
-\caption{Scenario refinement for HS07}
+\caption{Scenario refinement for HS07 - Secnerio 5/21}
 \end{center}
@@
   \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & When the the temperature adjustment begin? When the command is received by the radiator? \\
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & When do the temperature adjustment begin? When the command is received by the radiator? \\
   \hline
@@
 \end{tabular}
-\caption{Scenario refinement for HS07}
+\caption{Scenario refinement for HS07 - Scenario 6/16}
 \end{center}
@@
   \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & A sensor stops responding \\
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & Developer updates the code base to a new requirement in less than 30 man-hours \\
   \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifyability \\
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability, Testability \\
   \hline
   \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
-  & {\bfseries Source:} & A sensor \\
+  & {\bfseries Source:} & The developer \\
   \cline{2-3}
-  & {\bfseries Stimulus:} & Stops communicating with gatekeper \\
+  & {\bfseries Stimulus:} & Updates code for new requirement \\
   \cline{2-3}
-  & {\bfseries Artifact} & The system \\
+  & {\bfseries Artifact} & The code base \\
   \cline{2-3}
-  & {\bfseries Environment:} & Normal operation \\
+  & {\bfseries Environment:} & Developement time \\
   \cline{2-3}
-  & {\bfseries Response:} & Removes sensor from configuration and display error message for user \\
+  & {\bfseries Response:} & Modification is made with no sideeffect (all unit-/integrationtest pass) \\
   \cline{2-3}
-  & {\bfseries Response Measure:} & Within 30 seconds \\
+  & {\bfseries Response Measure:} & Less than 30 man-hours \\
   \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & How is it know that it stops responding? Heartbeat? \\
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & What limits the allowed requirement changes? \\
   \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} & Developer is assumed to know the system architecture \\
   \hline
 \end{tabular}
-\caption{Scenario refinement for HS07}
+\caption{Scenario refinement for HS07 - Scenario 11}
 \end{center}

### 2008-02-24 trunk/SAiP/Module1/H2/table.tex
Scenario added: Developer updates and tests the code base with a new device type in 1 week

--- 2008-02-24 SAiP/Module1/H2/table.tex
@@
 \begin{document}
+
+\begin{table}[!htp]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & Developer updates and tests the code base with a new device type in 1 week. \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The developer \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Whishes to add a new device type \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The code \\
+  \cline{2-3}
+  & {\bfseries Environment:} & At design time \\
+  \cline{2-3}
+  & {\bfseries Response:} & Modification is made with no side effects \\
+  \cline{2-3}
+  & {\bfseries Response Measure:} & Modification is made and tested within 1 week \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & Does the device use implemented HTTP protocol or do we need an adaptor. Is it necessary to make structual modifications to the user interface. \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
+  \hline
+\end{tabular}
+\caption{Scenario refinement for HS07}
+\end{center}
+\end{table}
 

### 2008-02-24 SAiP/Module1/H2/paper.tex
Tables completed

@@
 \item \emph{Usability.} It must be easy to use the gateway user interface.
+  include new types of sensors and actuators.
 \item \emph{Availability.} The system must be responsive and available for 
 both users and hardware elements.
-\item \emph{Safety.} The system must take care not to overheat the actuators.
 \end{itemize}
@@
 \item An attacker attempts to log on using an invalid password and is thrown off. The next attempt will take longer (progressive log-on time). The system stores a log of the failed attempt.
-\item Developer updates and tests the code base with a new device type in 1 day.
+\item Developer updates and tests the code base with a new device type in 1 week.
 \item A technical installer installs a system at a customer in no more than 4 hours.
@@
 Lars&User&1/20&4&6/16&15&4&5/21\\
-Anders&Producer&9&11&11&12&17&9/18\\
+Anders&Producer&9&6/16&11&12&17&9/18\\
 \hline
@@
 \item 9/18 (6 votes)
-\item 11 (6 votes)
+\item 11 (5 votes)
 \item 5/21 (4 votes)
+\item 6/16 (4 votes)
 \item 19 (3 votes)
 \item 4 (3 votes)
-\item 13 (3 votes)
+\item 13 (2 votes)
 \item 1/20 (2 votes)
 \item 2/7 (2 votes)
-\item 6/16 (2 votes)
 \item 10/17 (2 votes)
@@
 The process of refining the chosen scenarios centers around specification of various elements of the scenario. Stimulus, Response, Response Measurement. 
+\begin{table}[!htp]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & Developer updates and tests the code base with a new device type in 1 week. \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The developer \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Whishes to add a new device type \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The code \\
+  \cline{2-3}
+  & {\bfseries Environment:} & At design time \\
+  \cline{2-3}
+  & {\bfseries Response:} & Modification is made with no side effects \\
+  \cline{2-3}
+  & {\bfseries Response \mbox{Measure:}} & Modification is made and tested within 1 week \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & Does the device use implemented HTTP protocol or do we need an adaptor. Is it necessary to make structual modifications to the user interface. \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
+  \hline
+\end{tabular}
+\caption{Scenario refinement for HS07 - Scenario 9}
+\end{center}
+\end{table}
+
+\begin{table}[!htp]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The user adds a new actuator and the system configures to use the new hardware automatically within 1 minute \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability, Usability, Performance \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The user \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Attaches a new actuator \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The system \\
+  \cline{2-3}
+  & {\bfseries Environment:} & Maintanance operation \\
+  \cline{2-3}
+  & {\bfseries Response:} & New hardware is automatically added to the system configuration and accessible remotely \\
+  \cline{2-3}
+  & {\bfseries Response \mbox{Measure:}} & Within one minute \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & Is the user allwed to install now components himself? \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
+  \hline
+\end{tabular}
+\caption{Scenario refinement for HS07 - Secnerio 5/21}
+\end{center}
+\end{table}
+
+\begin{table}[!htp]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The user changes the preferred temperature from a web page and the system adapts to it \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Usability,  Availability \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The user \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Remotely change the temperature \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The system \\
+  \cline{2-3}
+  & {\bfseries Environment:} & Normal operation \\
+  \cline{2-3}
+  & {\bfseries Response:} & Begin adjusting to temperature \\
+  \cline{2-3}
+  & {\bfseries Response \mbox{Measure:}} & Within 5 seconds \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & When do the temperature adjustment begin? When the command is received by the radiator? \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
+  \hline
+\end{tabular}
+\caption{Scenario refinement for HS07 - Scenario 6/16}
+\end{center}
+\end{table}
+
+\begin{table}[!htp]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & Developer updates the code base to a new requirement in less than 30 man-hours \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability, Testability \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The developer \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Updates code for new requirement \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The code base \\
+  \cline{2-3}
+  & {\bfseries Environment:} & Developement time \\
+  \cline{2-3}
+  & {\bfseries Response:} & Modification is made with no sideeffect (all unit-/integrationtest pass) \\
+  \cline{2-3}
+  & {\bfseries Response \mbox{Measure:}} & Less than 30 man-hours \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & What limits the allowed requirement changes? \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} & Developer is assumed to know the system architecture \\
+  \hline
+\end{tabular}
+\caption{Scenario refinement for HS07 - Scenario 11}
+\end{center}
+\end{table}
 
@@
 \section{Evaluation}
+
 % =====================================================================

### 2008-02-24 trunk/SAiP/Module1/H2/paper.tex

@@
 both users and hardware elements.
-\item \emph{Safety.} The system must take care not to overheat the actuators.
 \end{itemize}
@@
 
+\begin{table}[!htp]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & Developer updates and tests the code base with a new device type in 1 week. \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability, Testability  \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The developer \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Whishes to add a new device type \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The code \\
+  \cline{2-3}
+  & {\bfseries Environment:} & At design time \\
+  \cline{2-3}
+  & {\bfseries Response:} & Modification is made with no side effects (all unit-/integrationtest pass)\\
+  \cline{2-3}
+  & {\bfseries Response \mbox{Measure:}} & Modification is made and tested within 1 week \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & Does the device use implemented HTTP protocol or do we need an adaptor?
+Is it necessary to make structual modifications to the user interface?
+Does the change have architectural impact?
+\\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} & Developer is assumed to know the system architecture  \\
+  \hline
+\end{tabular}
+\caption{Refined Scenario \#1 - QA Scenario 9/11/18}
+\end{center}
+\end{table}
+
+\begin{table}[!htp]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The user adds a new actuator and the system configures to use the new hardware automatically within 1 minute \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability, Usability, Performance \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The user \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Attaches a new actuator \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The system \\
+  \cline{2-3}
+  & {\bfseries Environment:} & Maintanance operation \\
+  \cline{2-3}
+  & {\bfseries Response:} & New hardware is automatically added to the system configuration and accessible remotely \\
+  \cline{2-3}
+  & {\bfseries Response \mbox{Measure:}} & Within one minute \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & Is the user allwed to install now components himself? \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
+  \hline
+\end{tabular}
+\caption{Refined Scenario \#2 - QA Scenario 5/21}
+\end{center}
+\end{table}
+
+\begin{table}[!htp]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The user changes the preferred temperature from a web page and the system adapts to it \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Usability,  Availability \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The user \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Remotely change the temperature \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The system \\
+  \cline{2-3}
+  & {\bfseries Environment:} & Normal operation \\
+  \cline{2-3}
+  & {\bfseries Response:} & Begin adjusting to temperature \\
+  \cline{2-3}
+  & {\bfseries Response \mbox{Measure:}} & Within 5 seconds \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & When do the temperature adjustment begin? When the command is received by the radiator? \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
+  \hline
+\end{tabular}
+\caption{Refined Scenario \#3 - QA Scenario 6/16}
+\end{center}
+\end{table}
+
 % =====================================================================
 \section{Evaluation}
+The compliance of the Architectural Description of HS07 \cite{SAiP1H1} for the selected refined scenarios
+is discussed in the following chapters.
+\subsection{Refined Scenario \#1}
+The new device type can either be a new thermometer/actuator or a completely new type of device, i.e. a door.
+
+\subsection{Adding new sensor/actuator}
+Localized to changes within the given sensor or actuator type. Update by modification, supported by the architecture now. Not in gateway, which is Update by addition in this case.
+
+\subsection{Adding different types}
+Update by addition will need to be built into the architecture in the sensor and actuator subsystems to encompass new types apart from thermometers and radiators.
+Radiators and thermometers are hardcoded into the gateway, it will need to be more abstract.
+
+We have chosen to focus in the integration of new types of thermometers and actuators.
+
+\subsection{Refined Scenario \#2}
+Because of the Client/Server architecture of the gateway, adding new actuators is currently possible without architectural changes.
+With the current design, only actuators of the radiator type can be added.
+
+\subsection{Refined Scenario \#3}
+In order to have the system encompass a user interface that allows a user to change the preferred temperature from a web page, the gateway subsystem needs to become the Model part of a MVC pattern.
+
+This functionality is not part of the current architecture, even though it supports commands on the gateway. At least one new command needs to be implemented on the gateway.
+
 % =====================================================================
 \section{Architectural Design}
+
+\subsection{Refined Scenario \#1}
+\subsection{Tactics}
+\subsection{Modifiablity}
+\subsection{Testability}
+\subsection{Refined Scenario \#3}
+\subsection{Tactics}
+\subsection{Availability}
+We will use a passive redundancy system for the http service in order to handle the response measurement. 
+We consider the risk of hardware failure to small to use hardware redundancy, given the expected low amount of traffic on the system.
+\subsection{Usability}
+The web interface must...
+
 % =====================================================================

### 2008-02-24 trunk/SAiP/Module1/H2/paper.tex
Availability and Usability tactics described for scenario #3.

@@
   \cline{2-3}
-  & {\bfseries Response:} & Begin adjusting to temperature \\
+  & {\bfseries Response:} & Acknowledge of operation from gateway \\
   \cline{2-3}
-  & {\bfseries Response \mbox{Measure:}} & Within 5 seconds \\
+  & {\bfseries Response \mbox{Measure:}} & Within 5 seconds 98,9% of the time, within 30 seconds 1% of the time and more than 30 seconds 0,1% of the time (unshceduled down-time) \\
   \hline
-  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & When do the temperature adjustment begin? When the command is received by the radiator? \\
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & When do the temperature adjustment begin? When is the command is received by the radiator? \\
   \hline
@@
 \subsection{Refined Scenario \#3}
+Scenario \#3 requires the HS07 system to provide a web interface that can be used to change the temperature in the house.
+The response requirements sets some requirements to the availability of the http service.
 \subsection{Tactics}
 \subsection{Availability}
-We will use a passive redundancy system for the http service in order to handle the response measurement. 
-We consider the risk of hardware failure to small to use hardware redundancy, given the expected low amount of traffic on the system.
+As the user may experience non-responsiveness 0,1% of the time it is believed that this can be done with a single hardware unit (no HW-redundancy).
+At the same time the requirement of 30 seconds 1% of the time is sufficient time to restart the web-service, and it is believed sufficient to have \emph{passive redundancy}.
+Finally it is believed that an availability of 5 seconds 98,9% of the times, require a form of redundancy, \emph{passive redundancy} is chosen.
+As an example of a design may be mentioned the Apache web-server design, which is redundant processes.
+A number of processes is spawned to handle HTTP/GET requests, and it uses some kind of fault detection scheme to detect if one become non-responsive.
+If this happens it is simply killed and an identical process is spawned in its place.
+
 \subsection{Usability}
-The web interface must...
+A web interface must be provided to support the requirement that the user can change the temperature from a web page.
+In order to maintain the modifiability of the system, a Model-View-Controller pattern will be used to separate the web interface from the gateway.
+A web backend process will be added to limit the web access to the model thereby hiding irrelevant information.
+The model needs to maintain all information about the system that is relevant to the user.
 

### 2008-02-24 trunk/SAiP/Module1/H2/paper.tex
Adding 11 to the identical scenarios
Adding a little more about availability

@@
 \item 10. and 17.
-\item 9. and 18.
+\item 9. and 11. and 18.
 \item 1. and 20.
@@
 \subsection{Availability}
-As the user may experience non-responsiveness 0,1% of the time it is believed that this can be done with a single hardware unit (no HW-redundancy).
-At the same time the requirement of 30 seconds 1% of the time is sufficient time to restart the web-service, and it is believed sufficient to have \emph{passive redundancy}.
-Finally it is believed that an availability of 5 seconds 98,9% of the times, require a form of redundancy, \emph{passive redundancy} is chosen.
+As the user may experience non-responsiveness 0,1% of the time it is believed that this can be done with a single hardware unit (no HW-reduadance, e.g.
+\emph{voting}).
+At the same time the requirement of 30 seconds 1% of the time is sufficient time to restart the web-service, and it is therefore not required to have \emph{active redundancy}.
+Finally it is believed that an availability of 5 seconds 98,9% of the times, require a form of redundancy, and for this reason \emph{passive redundancy} is chosen.
 As an example of a design may be mentioned the Apache web-server design, which is redundant processes.

### 2008-02-24 trunk/SAiP/Module1/H2/paper.tex
Adding Modifiability

--- 2008-02-24 trunk/SAiP/Module1/H2/paper.tex
@@
   \cline{2-3}
-  & {\bfseries Response \mbox{Measure:}} & Within 5 seconds 98,9% of the time, within 30 seconds 1% of the time and more than 30 seconds 0,1% of the time (unshceduled down-time) \\
+  & {\bfseries Response \mbox{Measure:}} & Within 5 seconds 98,9\% of the time, within 30 seconds 1\% of the time and more than 30 seconds 0,1\% of the time (unshceduled down-time) \\
   \hline
@@
 \subsection{Modifiablity}
+\begin{itemize}
+\item New Thermometer/Radiator implementation
+\begin{itemize}
+\item In the present implementation the ThemometerService has a hard-coded reference to Themometer, and having a new operation in the implementation is bad design if modifiability is desired, as it promoted change by modification.
+Changing the ThemometerService to use e.g.
+AbstractFactory to create the given Themometer instance will allow a new Themometer to be implemented and deployed by merely adding a class and a factory, which is modification by addition and then writing a main for the device, which uses this factory and the generic ThermometerService.
+At present the getController returns an object and the AbstractFactory may be implemented to also return this.
+It would however be an advantage to define a Thermometer interface in order to ensure that the implementations meets the protocol (has getTemperature method).
+This imply several tactics, as it \emph{localize modifications} to the temperature implementation.
+If a Temperature interface is created, then this uses \emph{anticipate expected changes} as the interface must contain the methods, which is expected to be needed, not just now but in the future.
+\emph{Generalize the module} is used in the fact that Objects are exchanged between Gateway and sensor/actuator, giving the Gateway to work on any instance of a class.
+\item The use of HTTP between Gatwway, sensor and actuator removes the syntactical dependency replacing it with runtime checking (by the Invoker), with the advantages and disadvantages it has.
+There is also a strong \emph{Location dependency} from all sensore/actuators to the gateway, as they must know where it is located (no discovery).
+It may be updated with a DNS-lookup to make this weaker.
+\item As the Gateway must support all sensors/actuators in its run algorithm there is a \emph{QoS dependency} that this must not take longer than the period by which the run-method is expected to iterate.
+\item The Gateway already support \emph{defer binding time} as the actuator/sensor registers with the Gateway at runtime.
+\end{itemize}
+\item New sensor/actuator type
+\begin{itemize}
+\item The current implementation of the Gateway supports \emph{defer binding time}, yet as the Gateway run method has a hardcoded dependency on the semantics of the radiator and thermometer (through the getTemperature and notifyObserver command syntax) it is not possible to simply register other types.
+The Gateway would have to use \emph{Generalize the module} in order to allow more general actuators and sensors to register, as opposed to only those that have radiator and thermometer semantics.
+\end{itemize}
+\end{itemize}
 \subsection{Testability}
@@
 \subsection{Availability}
-As the user may experience non-responsiveness 0,1% of the time it is believed that this can be done with a single hardware unit (no HW-redundancy).
-At the same time the requirement of 30 seconds 1% of the time is sufficient time to restart the web-service, and it is believed sufficient to have \emph{passive redundancy}.
-Finally it is believed that an availability of 5 seconds 98,9% of the times, require a form of redundancy, \emph{passive redundancy} is chosen.
+As the user may experience non-responsiveness 0,1\% of the time it is believed that this can be done with a single hardware unit (no HW-reduadance, e.g.
+\emph{voting}).
+At the same time the requirement of 30 seconds 1\% of the time is sufficient time to restart the web-service, and it is therefore not required to have \emph{active redundancy}.
+Finally it is believed that an availability of 5 seconds 98,9\% of the times, require a form of redundancy, and for this reason \emph{passive redundancy} is chosen.
 As an example of a design may be mentioned the Apache web-server design, which is redundant processes.

### 2008-02-24 SAiP/Module1/H2/paper.tex
Votes updated.

--- 2008-02-24 SAiP/Module1/H2/paper.tex
@@
 \end{itemize}
-\item Developer (?)
+\item Developer
 \begin{itemize}
@@
 \begin{itemize}
-\item 2. and 7.
-\item 6. and 16.
-\item 10. and 17.
-\item 9. and 18.
-\item 1. and 20.
-\item 5. and 21.
+\item 2 and 7
+\item 6 and 16
+\item 10 and 17
+\item 9 and 11 and 18
+\item 1 and 20
+\item 5 and 21
 \end{itemize}
@@
 Jesper&User&4&5&6/16&20&21&2/7\\
-Lars&Developer&19&19&5/21&9/18&11&13\\
-Anders&Producer&2/7&4&6/16&8&14&5/21\\
+Lars&Developer&19&19&5/21&9/11/18&9/11/18&13\\
+Anders&Producer&2/7&9/11/18&9/11/18&12&10/17&10/17\\
 \hline
@@
 \hline
-Jesper&Developer&11&19&9&11&13&6/16\\
+Jesper&Developer&9/11/18&9/11/18&9/11/18&19&13&6/16\\
 Lars&User&1/20&4&6/16&15&4&5/21\\
-Anders&Producer&9&11&11&12&17&9/18\\
+Anders&Producer&9/11/18&6/16&5/21&5/21&12&10/17\\
 \hline
@@
 \begin{enumerate}
-\item 9/18 (6 votes)
-\item 11 (6 votes)
-\item 5/21 (4 votes)
+\item 9/11/18 (8 votes)
+\item 5/21 (6 votes)
+\item 6/16 (4 votes)
 \item 19 (3 votes)
 \item 4 (3 votes)
-\item 13 (3 votes)
+\item 13 (2 votes)
 \item 1/20 (2 votes)
 \item 2/7 (2 votes)
-\item 6/16 (2 votes)
-\item 10/17 (2 votes)
+\item 10/17 (3 votes)
 \item 12 (2 votes)
@@
 
+\begin{table}[!htp]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & Developer updates and tests the code base with a new device type in 1 week. \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability, Testability  \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The developer \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Whishes to add a new device type \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The code \\
+  \cline{2-3}
+  & {\bfseries Environment:} & At design time \\
+  \cline{2-3}
+  & {\bfseries Response:} & Modification is made with no side effects (all unit-/integrationtest pass)\\
+  \cline{2-3}
+  & {\bfseries Response \mbox{Measure:}} & Modification is made and tested within 1 week \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & Does the device use implemented HTTP protocol or do we need an adaptor?
+Is it necessary to make structual modifications to the user interface?
+Does the change have architectural impact?
+\\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} & Developer is assumed to know the system architecture  \\
+  \hline
+\end{tabular}
+\caption{Refined Scenario \#1 - QA Scenario 9/11/18}
+\end{center}
+\end{table}
+
+\begin{table}[!htp]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The user adds a new actuator and the system configures to use the new hardware automatically within 1 minute \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Modifiability, Usability, Performance \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The user \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Attaches a new actuator \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The system \\
+  \cline{2-3}
+  & {\bfseries Environment:} & Maintanance operation \\
+  \cline{2-3}
+  & {\bfseries Response:} & New hardware is automatically added to the system configuration and accessible remotely \\
+  \cline{2-3}
+  & {\bfseries Response \mbox{Measure:}} & Within one minute \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & Is the user allwed to install now components himself? \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
+  \hline
+\end{tabular}
+\caption{Refined Scenario \#2 - QA Scenario 5/21}
+\end{center}
+\end{table}
+
+\begin{table}[!htp]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The user changes the preferred temperature from a web page and the system adapts to it \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Usability,  Availability \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The user \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Remotely change the temperature \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The system \\
+  \cline{2-3}
+  & {\bfseries Environment:} & Normal operation \\
+  \cline{2-3}
+  & {\bfseries Response:} & Acknowledge of operation from gateway \\
+  \cline{2-3}
+  & {\bfseries Response \mbox{Measure:}} & Within 5 seconds 98,9\% of the time, within 30 seconds 1\% of the time and more than 30 seconds 0,1\% of the time (unshceduled down-time) \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & When do the temperature adjustment begin? When is the command is received by the radiator? \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} &  \\
+  \hline
+\end{tabular}
+\caption{Refined Scenario \#3 - QA Scenario 6/16}
+\end{center}
+\end{table}
+
 % =====================================================================
 \section{Evaluation}
+The compliance of the Architectural Description of HS07 \cite{SAiP1H1} for the selected refined scenarios
+is discussed in the following chapters.
+\subsection{Refined Scenario \#1}
+The new device type can either be a new thermometer/actuator or a completely new type of device, i.e. a door.
+
+\subsection{Adding new sensor/actuator}
+Localized to changes within the given sensor or actuator type. Update by modification, supported by the architecture now. Not in gateway, which is Update by addition in this case.
+
+\subsection{Adding different types}
+Update by addition will need to be built into the architecture in the sensor and actuator subsystems to encompass new types apart from thermometers and radiators.
+Radiators and thermometers are hardcoded into the gateway, it will need to be more abstract.
+
+We have chosen to focus in the integration of new types of thermometers and actuators.
+
+\subsection{Refined Scenario \#2}
+Because of the Client/Server architecture of the gateway, adding new actuators is currently possible without architectural changes.
+With the current design, only actuators of the radiator type can be added.
+
+\subsection{Refined Scenario \#3}
+In order to have the system encompass a user interface that allows a user to change the preferred temperature from a web page, the gateway subsystem needs to become the Model part of a MVC pattern.
+
+This functionality is not part of the current architecture, even though it supports commands on the gateway. At least one new command needs to be implemented on the gateway.
+
 % =====================================================================
 \section{Architectural Design}
+
+\subsection{Refined Scenario \#1}
+\subsection{Tactics}
+\subsection{Modifiablity}
+\begin{itemize}
+\item New Thermometer/Radiator implementation
+\begin{itemize}
+\item In the present implementation the ThemometerService has a hard-coded reference to Themometer, and having a new operation in the implementation is bad design if modifiability is desired, as it promoted change by modification.
+Changing the ThemometerService to use e.g.
+AbstractFactory to create the given Themometer instance will allow a new Themometer to be implemented and deployed by merely adding a class and a factory, which is modification by addition and then writing a main for the device, which uses this factory and the generic ThermometerService.
+At present the getController returns an object and the AbstractFactory may be implemented to also return this.
+It would however be an advantage to define a Thermometer interface in order to ensure that the implementations meets the protocol (has getTemperature method).
+This imply several tactics, as it \emph{localize modifications} to the temperature implementation.
+If a Temperature interface is created, then this uses \emph{anticipate expected changes} as the interface must contain the methods, which is expected to be needed, not just now but in the future.
+\emph{Generalize the module} is used in the fact that Objects are exchanged between Gateway and sensor/actuator, giving the Gateway to work on any instance of a class.
+\item The use of HTTP between Gatwway, sensor and actuator removes the syntactical dependency replacing it with runtime checking (by the Invoker), with the advantages and disadvantages it has.
+There is also a strong \emph{Location dependency} from all sensore/actuators to the gateway, as they must know where it is located (no discovery).
+It may be updated with a DNS-lookup to make this weaker.
+\item As the Gateway must support all sensors/actuators in its run algorithm there is a \emph{QoS dependency} that this must not take longer than the period by which the run-method is expected to iterate.
+\item The Gateway already support \emph{defer binding time} as the actuator/sensor registers with the Gateway at runtime.
+\end{itemize}
+\item New sensor/actuator type
+\begin{itemize}
+\item The current implementation of the Gateway supports \emph{defer binding time}, yet as the Gateway run method has a hardcoded dependency on the semantics of the radiator and thermometer (through the getTemperature and notifyObserver command syntax) it is not possible to simply register other types.
+The Gateway would have to use \emph{Generalize the module} in order to allow more general actuators and sensors to register, as opposed to only those that have radiator and thermometer semantics.
+\end{itemize}
+\end{itemize}
+\subsection{Testability}
+\subsection{Refined Scenario \#3}
+Scenario \#3 requires the HS07 system to provide a web interface that can be used to change the temperature in the house.
+The response requirements sets some requirements to the availability of the http service.
+\subsection{Tactics}
+\subsection{Availability}
+As the user may experience non-responsiveness 0,1\% of the time it is believed that this can be done with a single hardware unit (no HW-reduadance, e.g.
+\emph{voting}).
+At the same time the requirement of 30 seconds 1\% of the time is sufficient time to restart the web-service, and it is therefore not required to have \emph{active redundancy}.
+Finally it is believed that an availability of 5 seconds 98,9\% of the times, require a form of redundancy, and for this reason \emph{passive redundancy} is chosen.
+As an example of a design may be mentioned the Apache web-server design, which is redundant processes.
+A number of processes is spawned to handle HTTP/GET requests, and it uses some kind of fault detection scheme to detect if one become non-responsive.
+If this happens it is simply killed and an identical process is spawned in its place.
+
+\subsection{Usability}
+A web interface must be provided to support the requirement that the user can change the temperature from a web page.
+In order to maintain the modifiability of the system, a \emph{Model-View-Controller} pattern will be used to separate the web interface from the gateway.
+A web backend process will be added to limit the web access to the model thereby hiding irrelevant information.
+The model needs to maintain all information about the system that is relevant to the user - \emph{Maintain a model of the system}.
+
 % =====================================================================

### 2008-02-24 trunk/Module2/exercise/abstract-factory/DoorAlarm/src/dooralarm/gui/DoorAlarm.java

@@ -67,11 +67,11 @@ public class DoorAlarm extends JFrame {
     keyCodeListener = new KeyCodeListener();
 
 	// ********** Solution 1 **************** //
-//    JPanel keyboard = factory.createKeyboardPanel(keyCodeListener);
+    JPanel keyboard = factory.createKeyboardPanel(keyCodeListener);
 	// ********** End Solution 1 **************** //
 
 	// ********** Solution 2 **************** //
-    JPanel keyboard = createKeyboardPanel();
+//    JPanel keyboard = createKeyboardPanel();
 	// ********** End Solution 2 **************** //
 
     panel.add( keyboard );

### 2008-02-24 trunk/SAiP/Module1/H2/HS07/src/dk/atisa/hs07/sensor/SensorFactory.java
Updating Sensor to use AbstractFactory

@@ -1,0 +1,5 @@
+package dk.atisa.hs07.sensor;
+
+public interface SensorFactory {
+	Object createSensor();
+}

### 2008-02-24 trunk/SAiP/Module1/H2/HS07/src/dk/atisa/hs07/sensor/ThermometerService.java

@@ -14,7 +14,7 @@ import dk.atisa.hs07.Service;
 public class ThermometerService extends Service {
 	
 	public Object getController() {
-		return new Thermometer();
+		return sensorFactory.createSensor();
 	}
 	
 	/**
@@ -23,18 +23,21 @@ public class ThermometerService extends Service {
 	 * @param thisLocation
 	 * @throws Exception
 	 */
-	public ThermometerService(String gatewayLocation, String thisLocation) throws Exception {
+	public ThermometerService(String gatewayLocation, String thisLocation, SensorFactory sf) throws Exception {
 		super(thisLocation);
 		URL url = new URL(thisLocation);
+		sensorFactory = sf;
 		Invoker.invoke(gatewayLocation, "registerThermometer", "location", url.toString());
 		System.out.println("Started thermometer service at " + url);
 	}
 
 	public static void main(String[] args) throws Exception {
 		URL baseUrl = new URL(args[1]);
+		SensorFactory sf = new RandomThermometerFactory();
 		for (int i = 0; i < Integer.parseInt(args[2]); i++) {
 			String location = "http://" + baseUrl.getHost() + ":" + (baseUrl.getPort() + i) + "/";
-			new ThermometerService(args[0], location);
+			new ThermometerService(args[0], location, sf);
 		}
 	}
+	private SensorFactory sensorFactory;
 }

### 2008-02-24 SAiP/Module1/H2/HS07/src/dk/atisa/hs07/sensor/RandomThermometer.java
Updating Sensor to use AbstractFactory

@@ -6,7 +6,7 @@ package dk.atisa.hs07.sensor;
  * @author Klaus Marius Hansen, klaus.m.hansen@daimi.au.dk
  *
  */
-public class Thermometer {
+public class RandomThermometer {
 	private double  temperature = 20;
 
 	/**

### 2008-02-24 SAiP/Module1/H2/HS07/src/dk/atisa/hs07/sensor/RandomThermometerFactory.java

@@ -1,0 +1,7 @@
+package dk.atisa.hs07.sensor;
+
+public class RandomThermometerFactory implements SensorFactory {
+	public Object createSensor() {
+	    return new RandomThermometer();
+	}
+}

### 2008-02-24 trunk/SAiP/Module1/H2/HS07/src/dk/atisa/hs07/Service.java
Updating so the service may be started after creating. Performing callbacks from base class to sub-class in constructor is very bad form, as member variables are not initialized

@@ -22,6 +22,25 @@ public abstract class Service {
 	 * @throws Exception
 	 */
 	public Service(String location) throws Exception {
+		start(location);
+	}
+
+	/**
+	 * Creates an un-started HS07 service
+	 * 
+	 * @param location of the service
+	 * @throws Exception
+	 */
+	public Service() {
+	}
+
+	/**
+	 * Starts an HS07 service using the Jetty servlet container
+	 * 
+	 * @param location of the service
+	 * @throws Exception
+	 */
+	public void start(String location) throws Exception {
 		Server server = new Server(new URL(location).getPort());    
 		Context root = new Context(server, "/", Context.SESSIONS);
 		Servlet servlet = new ProtocolServlet(getController());

### 2008-02-24 trunk/SAiP/Module1/H2/HS07/src/dk/atisa/hs07/sensor/ThermometerService.java

@@ -24,9 +24,9 @@ public class ThermometerService extends Service {
 	 * @throws Exception
 	 */
 	public ThermometerService(String gatewayLocation, String thisLocation, SensorFactory sf) throws Exception {
-		super(thisLocation);
 		URL url = new URL(thisLocation);
 		sensorFactory = sf;
+		start(thisLocation);
 		Invoker.invoke(gatewayLocation, "registerThermometer", "location", url.toString());
 		System.out.println("Started thermometer service at " + url);
 	}

### 2008-02-24 trunk/SAiP/Module1/H2/HS07/src/dk/atisa/hs07/gateway/Gateway.java
Added the model package

@@ -19,6 +19,12 @@ public class Gateway implements Runnable {
 	private Collection<String> thermometers = new LinkedList<String>();
 	private Collection<String> observers = new LinkedList<String>();
 	
+	private String modelLocation = "";
+	
+	public Gateway( String _modelLocation ) {
+		modelLocation = _modelLocation;
+	}
+	
 	/**
 	 * Register a thermometer which is used for estimating temperature
 	 * 
@@ -45,9 +51,9 @@ public class Gateway implements Runnable {
 	 * @throws MalformedURLException
 	 * @throws IOException
 	 */
-	public void notifyObservers(double temperature) throws MalformedURLException, IOException {
+	public void notifyObservers(double temperature, double targetTemperature) throws MalformedURLException, IOException {
 		for (String location : observers) {
-			Invoker.invoke(location, "notify", "temperatur___e", "" + temperature);
+			Invoker.invoke(location, "notify", "temperatur___e", "" + temperature, "targettem___p", "" + targetTemperature);
 		}
 	}
 
@@ -65,6 +71,7 @@ public class Gateway implements Runnable {
 		}
 		return sum/thermometers.size();
 	}
+	
 
 	/**
 	 * Run the control algorithm which is decentralized in this case, i.e.,
@@ -75,9 +82,14 @@ public class Gateway implements Runnable {
 			try {
 				Thread.sleep(SLEEP_TIME);
 				if (thermometers.size() > 0) {
+					/* Get the target temperature from the model */
+					double targetTemperature = Double.parseDouble(Invoker.invoke(modelLocation, "getTargetTemperature"));
 					double temperature = getTemperature();
+
 					System.out.println("Average temperature: " + temperature + " for " + thermometers.size() + " thermometer(s)");
-					notifyObservers(temperature);
+					notifyObservers(temperature,targetTemperature);
+
+					Invoker.invoke(modelLocation, "setAverageTemperature", "averageTem___p", ""+temperature);
 				}
 			} catch (Exception e) {
 				e.printStackTrace();

### 2008-02-24 SAiP/Module1/H2/HS07/src/dk/atisa/hs07/actuator/Radiator.java
Added the model package

@@ -7,28 +7,23 @@ package dk.atisa.hs07.actuator;
  *
  */
 public class Radiator {
-	/**
-	 * Maximum temperature for control algorithm
-	 */
-	public static final double MAX_TEMPERATURE = 20.5;
-	/**
-	 * Minimum temperature for control algorithm
-	 */
-	public static final double MIN_TEMPERATURE = 19.5;
+	/* Temperature spans a centered with .5 degrees variance */
+	public static final double TEMPERATURE_VARIANCE = 0.5;
 
 	private boolean state = false;
-
+	
 	/**
 	 * Run the control algorithm upon notification of temperature change
 	 * 
 	 * @param _temperature
 	 */
-	public void notify(String _temperature) {
+	public void notify(String _temperature, String _targetTemperature) {
+		double currentTarget = Double.parseDouble(_targetTemperature);
 		double temperature = Double.parseDouble(_temperature);
-		if (temperature < MIN_TEMPERATURE) {
+		if (temperature < currentTarget - TEMPERATURE_VARIANCE) {
 			System.out.println("Turn on radiator");
 			setState(true);
-		} else if (temperature > MAX_TEMPERATURE) {
+		} else if (temperature > currentTarget + TEMPERATURE_VARIANCE) {
 			System.out.println("Turn off radiator");
 			setState(false);
 		} 

### 2008-02-24 SAiP/Module1/H2/HS07/src/dk/atisa/hs07/gateway/GatewayService.java

@@ -14,6 +14,8 @@ import dk.atisa.hs07.Service;
  *
  */
 public class GatewayService extends Service {
+	private static String modelUrl = "";
+	
 	public GatewayService(String gatewayLocation) throws Exception {
 		super(gatewayLocation);
 		System.out.println("Started gateway service at " + gatewayLocation);
@@ -24,12 +26,13 @@ public class GatewayService extends Service {
 	 * algorithm continuously
 	 */
 	public Object getController() {
-		Runnable controller = new Gateway();
+		Runnable controller = new Gateway(modelUrl);
 		new Thread(controller).start();
 		return controller;
 	}
 	
 	public static void main(String[] args) throws Exception {
+		modelUrl = args[1];
 		new GatewayService(args[0]);
 	}
 }

### 2008-02-24 SAiP/Module1/H2/HS07/src/dk/atisa/hs07/model/Model.java
The actual model package.

@@ -1,0 +1,42 @@
+package dk.atisa.hs07.model;
+
+import java.io.IOException;
+import java.net.MalformedURLException;
+
+/**
+ * The model of the system as seen from an external source
+ * 
+ * @author Jesper Dalberg, jdalberg@daimi.au.dk
+ *
+ */
+public class Model {
+	/* Target Temperature - as controlled by the user through the web interface */
+	private double targetTemperature = 20.0;
+	
+	/* Store the average temperature for the web interface */
+	private double avgTemperature = 0;
+		
+	/* externally accessable functions */
+
+	/* Gets the current target temperature from the gateway */
+	public double getTargetTemperature() throws MalformedURLException, IOException {
+		System.out.println("getTargetTemperature: " + targetTemperature);
+		return targetTemperature;
+	}
+	
+	public double getAverageTemperature() throws MalformedURLException, IOException {
+		System.out.println("getAverageTemperature: " + avgTemperature);
+		return avgTemperature;
+	}
+	
+	/* Only invoked from the gateway */
+	public void setAverageTemperature( String _avgTemperature ) throws MalformedURLException, IOException {
+		System.out.println("setAverageTemperature: "+_avgTemperature);
+		avgTemperature = Double.parseDouble( _avgTemperature );
+	}
+
+	public void setTargetTemperature( String _targetTemperature ) throws MalformedURLException, IOException {
+		System.out.println("setTargetTemperature: "+_targetTemperature);
+		targetTemperature = Double.parseDouble(_targetTemperature);
+	}
+}

### 2008-02-24 SAiP/Module1/H2/HS07/src/dk/atisa/hs07/model/ModelService.java

@@ -1,0 +1,38 @@
+package dk.atisa.hs07.model;
+
+import java.net.URL;
+
+import dk.atisa.hs07.Service;
+
+/**
+ * 
+ * Radiator service. A home may have a number of radiator services
+ * 
+ * @author Jesper Dalberg, jdalberg@daimi.au.dk
+ *
+ */
+public class ModelService extends Service {
+	public Object getController() {
+		return new Model();
+	}
+	
+	/**
+	 * Construct a radiator service and register the service with the gateway
+	 * as an observer
+	 * 
+	 * @param gatewayLocation
+	 * @param thisLocation
+	 * @throws Exception
+	 */
+	public ModelService(String thisLocation) throws Exception {
+		super(thisLocation);
+		URL url = new URL(thisLocation);
+		System.out.println("Started model service at " + url);
+	}
+
+	public static void main(String[] args) throws Exception {
+		URL baseUrl = new URL(args[0]);
+		String location = "http://" + baseUrl.getHost() + ":" + baseUrl.getPort() + "/";
+		new ModelService(location);
+	}
+}

### 2008-02-24 trunk/SAiP/Module1/H2/paper.tex
Active Objects figure added.

@@
 \begin{itemize}
-\item 2. and 7.
-\item 6. and 16.
-\item 10. and 17.
-\item 9. and 11. and 18.
-\item 1. and 20.
-\item 5. and 21.
+\item 2 and 7
+\item 6 and 16
+\item 10 and 17
+\item 9 and 11 and 18
+\item 1 and 20
+\item 5 and 21
 \end{itemize}
@@
 \hline
-Jesper&Developer&11&19&9&11&13&6/16\\
+Jesper&Developer&9/11/18&9/11/18&9/11/18&19&13&6/16\\
 Lars&User&1/20&4&6/16&15&4&5/21\\
-Anders&Producer&9&6/16&11&12&17&9/18\\
+Anders&Producer&9/11/18&6/16&5/21&5/21&12&10/17\\
 \hline
@@
 \begin{enumerate}
-\item 9/18 (6 votes)
-\item 11 (5 votes)
-\item 5/21 (4 votes)
+\item 9/11/18 (8 votes)
+\item 5/21 (6 votes)
 \item 6/16 (4 votes)
@@
 \item 2/7 (2 votes)
-\item 10/17 (2 votes)
+\item 10/17 (3 votes)
 \item 12 (2 votes)
@@
 A web interface must be provided to support the requirement that the user can change the temperature from a web page.
-In order to maintain the modifiability of the system, a Model-View-Controller pattern will be used to separate the web interface from the gateway.
+In order to maintain the modifiability of the system, a \emph{Model-View-Controller} pattern will be used to separate the web interface from the gateway.
 A web backend process will be added to limit the web access to the model thereby hiding irrelevant information.
-The model needs to maintain all information about the system that is relevant to the user.
+The model needs to maintain all information about the system that is relevant to the user - \emph{Maintain a model of the system}.
+
+\section{Active Objects}
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure="figures/Active object",scale=0.8 }}
+\caption{Active Objects in the redesigned HS07 system}
+\label{fig:activeobjects}
+\end{figure}
+
+\clearpage
 

### 2008-02-25 trunk/SAiP/Module1/H2/paper.tex
Added the figures.

@@
 
+\clearpage
+
+\section{Packages}
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure="figures/mv-model",scale=0.8 }}
+\caption{Module View of the web interface in the redesigned HS07 system}
+\label{fig:mv-model}
+\end{figure}
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure="figures/mv-factory",scale=0.8 }}
+\caption{Module View of the abstract factory in the redesigned HS07 system}
+\label{fig:mv-factory}
+\end{figure}
+
+\clearpage
+
 \section{Active Objects}
@@
 
+\section{Sequence Diagrams}
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure="figures/seq-stt",scale=0.8 }}
+\caption{Sequence Diagram for the web requests in the redesigned HS07 system}
+\label{fig:seq-stt}
+\end{figure}
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure="figures/seq-tfac",scale=0.8 }}
+\caption{Sequence Diagram for the Abstract Factory in the redesigned HS07 system}
+\label{fig:seq-tfac}
+\end{figure}
+
+\clearpage
+
+\section{Deployment}
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure="figures/deployment",scale=0.8 }}
+\caption{Deployment in the redesigned HS07 system}
+\label{fig:deployment}
+\end{figure}
+
+\clearpage
+
 % =====================================================================

### 2008-02-25 SAiP/Module1/H2/paper.tex
Comments for some of the figures.

@@
 \subsection{Testability}
+Tactics for testability include \emph{Test harness} which is essentially a needed setup for regression testing.
+The test harness for the HS07 system will consist of a series of JUnit test cases on a modular level.
+On a larger scale, the deployment on a PC, shown in /cite{SAiP1H1} is one big test scenario where sensors and actuators a stubbed using visual inspection for verification.
+This harness will continue to be the main test setup for introduction into the gateway/model parts of the system.
 \subsection{Refined Scenario \#3}
@@
 
+\clearpage
+
+\section{Packages}
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure="figures/mv-model",scale=0.8 }}
+\caption{Module View of the web interface in the redesigned HS07 system}
+\label{fig:mv-model}
+\end{figure}
+Figure~\ref{fig:mv-model} illustrates the new packages needed in HS07 to include a web interface for the user.
+The model package creates a service with witch to get and set the values  need by the user to make a decision an any new target temperature.
+Both the gateway and the webapp uses their invokers to contact the model service for parameter manipulation.
+
+The separation between the gateway and the model packages creates an opportunity for increase of security when that becomes a driving quality attribute, since the gateway information can be hidden within the system, letting the ModelService become the only exposed part.
+\begin{figure}[!htb]
+\centerline{\epsfig{figure="figures/mv-factory",scale=0.8 }}
+\caption{Module View of the abstract factory in the redesigned HS07 system}
+\label{fig:mv-factory}
+\end{figure}
+
+\clearpage
+
+\section{Active Objects}
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure="figures/Active object",scale=0.8 }}
+\caption{Active Objects in the redesigned HS07 system}
+\label{fig:activeobjects}
+\end{figure}
+Figure~\ref{fig:activeobjects} shows all the components of the HS07 system. The new parts are the ModelService and the HTTP Server, needed to create a suitable web interface for the user.
+\clearpage
+
+\section{Sequence Diagrams}
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure="figures/seq-stt",scale=0.8 }}
+\caption{Sequence Diagram for the web requests in the redesigned HS07 system}
+\label{fig:seq-stt}
+\end{figure}
+Figure~\ref{fig:seq-stt} shows the flow of contol when a web application uses the methods exposed throgh the model service.
+The amount of traffic on the model service will have no direct impact on the performance of the gateway service.
+
+The gateway uses the same http interface as the webapp to set the value of the current average room temperature, and to read the values of the target temperature requested by the user.
+
+If the ModelService is non responsive, it is important for the gateway service to continue to work, "guessing" the target temperature either by cache or by default values.
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure="figures/seq-tfac",scale=0.8 }}
+\caption{Sequence Diagram for the Abstract Factory in the redesigned HS07 system}
+\label{fig:seq-tfac}
+\end{figure}
+
+\clearpage
+
+\section{Deployment}
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure="figures/deployment",scale=0.8 }}
+\caption{Deployment in the redesigned HS07 system}
+\label{fig:deployment}
+\end{figure}
+
+\clearpage
+
 % =====================================================================

### 2008-02-25 trunk/SAiP/Module1/H2/paper.tex
Updated sectioning

@@
 \end{itemize}
-\subsection{The Vote}
 \subsection{First vote}
@@
 
-\subsection{Adding new sensor/actuator}
+\subsubsection{Adding new sensor/actuator}
 Localized to changes within the given sensor or actuator type. Update by modification, supported by the architecture now. Not in gateway, which is Update by addition in this case.
 
-\subsection{Adding different types}
+\subsubsection{Adding different types}
 Update by addition will need to be built into the architecture in the sensor and actuator subsystems to encompass new types apart from thermometers and radiators.
@@
 
-\subsection{Refined Scenario \#1}
-\subsection{Tactics}
-\subsection{Modifiablity}
+\subsection{Refined Scenario \#1 - Tactics}
+\subsubsection{Modifiablity}
 \begin{itemize}
@@
 \end{itemize}
-\subsection{Testability}
-\subsection{Refined Scenario \#3}
+\subsubsection{Testability}
+Tactics for testability include \emph{Test harness} which is essentially a needed setup for regression testing.
+The test harness for the HS07 system will consist of a series of JUnit test cases on a modular level.
+On a larger scale, the deployment on a PC, shown in /cite{SAiP1H1} is one big test scenario where sensors and actuators a stubbed using visual inspection for verification.
+This harness will continue to be the main test setup for introduction into the gateway/model parts of the system.
+\subsection{Refined Scenario \#3 - Tactics}
 Scenario \#3 requires the HS07 system to provide a web interface that can be used to change the temperature in the house.
 The response requirements sets some requirements to the availability of the http service.
-\subsection{Tactics}
-\subsection{Availability}
+\subsubsection{Availability}
 As the user may experience non-responsiveness 0,1\% of the time it is believed that this can be done with a single hardware unit (no HW-reduadance, e.g.
@@
 
-\subsection{Usability}
+\subsubsection{Usability}
 A web interface must be provided to support the requirement that the user can change the temperature from a web page.
@@
 \end{figure}
+Figure~\ref{fig:mv-model} illustrates the new packages needed in HS07 to include a web interface for the user.
+The model package creates a service with witch to get and set the values  need by the user to make a decision an any new target temperature.
+Both the gateway and the webapp uses their invokers to contact the model service for parameter manipulation.
 
+The separation between the gateway and the model packages creates an opportunity for increase of security when that becomes a driving quality attribute, since the gateway information can be hidden within the system, letting the ModelService become the only exposed part.
 \begin{figure}[!htb]
@@
 \end{figure}
-
+Figure~\ref{fig:activeobjects} shows all the components of the HS07 system. The new parts are the ModelService and the HTTP Server, needed to create a suitable web interface for the user.
 \clearpage
@@
 \end{figure}
+Figure~\ref{fig:seq-stt} shows the flow of contol when a web application uses the methods exposed throgh the model service.
+The amount of traffic on the model service will have no direct impact on the performance of the gateway service.
+
+The gateway uses the same http interface as the webapp to set the value of the current average room temperature, and to read the values of the target temperature requested by the user.
+
+If the ModelService is non responsive, it is important for the gateway service to continue to work, "guessing" the target temperature either by cache or by default values.
 

### 2008-02-25 trunk/SAiP/Module1/H2/paper.tex
The date was wrong
Sections reorganized

@@
 
-\date{2008-02-21}
+\date{2008-02-25}
 
@@
 
-% =====================================================================
-\section{Architectural Requirements}
-\label{sec:requirements}
-
-For our purposes there is one main use case for the HS07 system:
-\begin{quote}
-  \emph{Control Temperature}: The gateway collects measurements from
-  thermometers and reports this to radiators that then control the
-  temperature.
-\end{quote}
-
-The major driving quality attributes of the HS07 system
-are:
-
-\begin{itemize}
-\item \emph{Performance.} HS07 should be performant so that a large
-  number of thermometers and radiators may be part of the system.
-\item \emph{Modifiability.} It must be possible to modify HS07 to
-  include new types of sensors and actuators.
-\item \emph{Usability.} It must be easy to use the gateway user interface.
-  include new types of sensors and actuators.
-\item \emph{Availability.} The system must be responsive and available for 
-both users and hardware elements.
-\end{itemize}
+An architecture of the HS07 system was proposed in \cite{SAiP1H1}. In this
+report we will elaborate on the quality attributes of the system, define some
+quality attribute scenarios and propose a new architectural design to fulfill
+these scenarios.
 
 % =====================================================================
-\section{Quality Attributes}
-\subsection{Quality Attributes in the HS07 system}
+%\section{Architectural Requirements}
+%\label{sec:requirements}
+%
+%For our purposes there is one main use case for the HS07 system:
+%\begin{quote}
+%  \emph{Control Temperature}: The gateway collects measurements from
+%  thermometers and reports this to radiators that then control the
+%  temperature.
+%\end{quote}
+%
+%The major driving quality attributes of the HS07 system
+%are:
+%
+%\begin{itemize}
+%\item \emph{Performance.} HS07 should be performant so that a large
+%  number of thermometers and radiators may be part of the system.
+%\item \emph{Modifiability.} It must be possible to modify HS07 to
+%  include new types of sensors and actuators.
+%\item \emph{Usability.} It must be easy to use the gateway user interface.
+%  include new types of sensors and actuators.
+%\item \emph{Availability.} The system must be responsive and available for 
+%both users and hardware elements.
+%\end{itemize}
+
+% =====================================================================
+\section{Quality Attributes in the HS07 system}
 \subsection{Availability}
@@
 \subsection{Usability}
-The HS07 system must be used by "ordinary" users (e.g.
+The HS07 system must be used by "ordinary" users (i.e.
 non-technical), and a high degree of usability must therefore be inherent in the architecture to support rapid user interface modification as well as implementing functionality that may aid the user during use.
@@
 \subsection{Quality Attribute Selection}
+
+The major driving quality attributes of the HS07 system
+are:
+
+\begin{itemize}
+\item \emph{Performance.} HS07 should be performant so that a large
+  number of thermometers and radiators may be part of the system.
+\item \emph{Modifiability.} It must be possible to modify HS07 to
+  include new types of sensors and actuators.
+\item \emph{Usability.} It must be easy to use the gateway user interface
+  and include new types of sensors and actuators.
+\item \emph{Availability.} The system must be responsive and available for 
+both users and hardware elements.
+\end{itemize}
+
+\section{Quality Attribute Scenarios}
 The following chapters documents the selection of the Quality Attributes that we 
@@
 \end{itemize}
+
 \subsection{Brainstorming}
 In this part of the process we just attempt to think of as many possible scenarios for the HS07 system as we possibly can.
-
-\subsection{Scenarios}
 
@@
 interesting for us.
-\subsection{Stakeholders and interests}
+
+\subsubsection{Stakeholders and interests}
+
 \begin{itemize}
@@
 \end{itemize}
-\subsection{Identical scenarios}
+
+\subsubsection{Identical scenarios}
+
 Due to the distributed location of the participants in the brainstorming multiple "similar" scenarios was created.
@@
 \end{itemize}
-\subsection{First vote}
+
+\subsubsection{First vote}
+
 \begin{tabular}{|l|c|c|c|c|c|c|c|}
@@
 \end{tabular}
-\subsection{Second vote}
+
+\subsubsection{Second vote}
+
 \begin{tabular}{|l|c|c|c|c|c|c|c|}
@@
 \end{tabular}
-\subsection{Voting Results}
+
+\subsubsection{Voting Results}
+
 This is the result of the vote in a prioritized order.
@@
 \end{enumerate}
+
 \subsection{Refinement}

### 2008-02-25 trunk/SAiP/Module1/H2/paper.tex
Stavebugs + flere section-ændringer

@@
 parts of the system.
-\item After an average user has used the system for less than one hour e or she has mastered the basic functionality
+\item After an average user has used the system for less than one hour he or she has mastered the basic functionality.
 \item The user adds a new actuator and the system configures to use the new hardware automatically within 1 minute.
-\item The user logs on externally through the gateway and has access to the home control within 1 second of verifying 95\% of the time (5 seconds 4\% of the time, more than 5 seconds (unscheduled down-time) less than 1\%) is connected and has access
-\item The user attempts to turn radiator above normal safety-level from outside the house and receive an error back explaining the infraction.
+\item The user logs on externally through the gateway and has access to the home control within 5 second of verifying 98,9\% of the
+time (30 seconds 1\% of the time, more than 30 seconds (unscheduled down-time) less than 0,1\%) is connected and has access.
+\item The user attempts to turn radiator above normal safety-level from outside the house and receive an error explaining the infraction.
 The system stores a log of the episode and refuse to execute the command.
@@
 \item A technical installer installs a system at a customer in no more than 4 hours.
-\item Developer updates the code base to a new requirement in less than 30 man-hours
-\item Configuration manager installs an upgrade remotely at a customer in no more than 30min
+\item Developer updates the code base to a new requirement in less than 30 man-hours.
+\item Configuration manager installs an upgrade remotely at a customer in no more than 30 minutes.
 \item Tester/debugger extract debugging information directly from running system at customer.
@@
 \item The customer requests that the system runs on other hardware (e.g. less power consuming hardware).
-\item The customer requests that the system interfaces with other sensors/actuators?
+\item The customer requests that the system interfaces with other sensors/actuators.
 \item The developer modifies the sensor poll interval at build time.
@@
 
-\section{Packages}
+\subsection{Module View}
 
@@
 
-\section{Active Objects}
+\subsection{Component \& Connector View}
 
@@
 
-\section{Sequence Diagrams}
+% XXX \subsection{Sequence Diagrams}
+% Text til figur her
 
@@
 
-\section{Deployment}
+\subsection{Allocation View}
 

### 2008-02-25 trunk/SAiP/Module1/H2/paper.tex
Figure references added

@@
 \subsection{Selection}
+\label{sec:selection}
 Here we put the scenarios to the vote. This is done by emulation of a real-life situation where the actual stakeholder would
@@
 
+Figure \ref{scenario1}, \ref{scenario2} and \ref{scenario3} shows refined scenarios of the three
+quality attribute scenarios that were given the highest priorities in section \ref{sec:selection}.
+
 \begin{table}[!htp]
@@
 \caption{Refined Scenario \#1 - QA Scenario 9/11/18}
+\label{scenario1}
 \end{center}
@@
 \caption{Refined Scenario \#2 - QA Scenario 5/21}
+\label{scenario2}
 \end{center}
@@
 \caption{Refined Scenario \#3 - QA Scenario 6/16}
+\label{scenario3}
 \end{center}

### 2008-02-25 SAiP/Module1/H2/paper.tex
Figure -> Table

@@
 \end{itemize}
-\subsection{The Vote}
-\subsection{First vote}
+
+\subsubsection{First vote}
+
 \begin{tabular}{|l|c|c|c|c|c|c|c|}
@@
 \end{enumerate}
+
 \subsection{Refinement}
 The process of refining the chosen scenarios centers around specification of various elements of the scenario. Stimulus, Response, Response Measurement.
+
+Table \ref{scenario1}, \ref{scenario2} and \ref{scenario3} shows refined scenarios of the three
+quality attribute scenarios that were given the highest priorities in section \ref{sec:selection}.
 
@@
 \end{itemize}
-\subsection{Testability}
+\subsubsection{Testability}
 Tactics for testability include \emph{Test harness} which is essentially a needed setup for regression testing.
@@
 This harness will continue to be the main test setup for introduction into the gateway/model parts of the system.
-\subsection{Refined Scenario \#3}
+\subsection{Refined Scenario \#3 - Tactics}
 Scenario \#3 requires the HS07 system to provide a web interface that can be used to change the temperature in the house.
 The response requirements sets some requirements to the availability of the http service.
-\subsection{Tactics}
-\subsection{Availability}
+\subsubsection{Availability}
 As the user may experience non-responsiveness 0,1\% of the time it is believed that this can be done with a single hardware unit (no HW-reduadance, e.g.

### 2008-02-25 trunk/SAiP/Module1/H2/paper.tex
Adding comments to factory diagrams

--- 2008-02-25 trunk/SAiP/Module1/H2/paper.tex
@@
 \end{itemize}
+The actual brainstorming and selection (voting) was done in Wiki in order to allow for a decentralised brainstorming and vote. The setup may be seen at:
+
+http://code.google.com/p/pasoos2007hbc/wiki/SAiPM1H2.
+
+All information has been moved to the report so it is only the use of Wiki as a meeting-place that may be of interest.
 
@@
 
+Table \ref{scenario1}, \ref{scenario2} and \ref{scenario3} shows refined scenarios of the three
+quality attribute scenarios that were given the highest priorities in section \ref{sec:selection}.
+
 \begin{table}[!htp]
@@
 The separation between the gateway and the model packages creates an opportunity for increase of security when that becomes a driving quality attribute, since the gateway information can be hidden within the system, letting the ModelService become the only exposed part.
+\clearpage
+
 \begin{figure}[!htb]
@@
 \end{figure}
-
-\clearpage
+Figure~\ref{fig:mv-factory} shows the classes involved and their interdependencies for implementing the factory pattern for the sensor module.
+Please refer to Figure~\ref{fig:seq-tfac} for its runtime usage.
 
@@
 \end{figure}
-
+Figure~\ref{fig:seq-tfac} shows initialization sequence when using the SensorFactory illustrated in Figure~\ref{fig:mv-factory}.
+The sequence shows the use of the factory to create a decoupling between the actual Thermometer implementation and the ThermometerService.
 \clearpage

### 2008-02-25 trunk/SAiP/Module1/H2/paper.tex
Section 5.1 and 5.2 has been merged into 4.1 and 4.3 respectively. Needs more updating.

@@
 \subsection{Refinement}
+\label{sec:refinement}
+
 The process of refining the chosen scenarios centers around specification of various elements of the scenario. Stimulus, Response, Response Measurement.
@@
 \section{Evaluation}
-The compliance of the Architectural Description of HS07 \cite{SAiP1H1} for the selected refined scenarios
-is discussed in the following chapters.
+
+The following sections will elaborate on the refined scenarios described in section \ref{sec:refinement}.
+The compliance of the Architectural Description of HS07 \cite{SAiP1H1} for the selected refined
+scenarios will be described and tactics that can be applied to the architectural design in order to
+fulfill the quality attributes are identified.
+
 \subsection{Refined Scenario \#1}
@@
 
-\subsubsection{Adding new sensor/actuator}
-Localized to changes within the given sensor or actuator type. Update by modification, supported by the architecture now. Not in gateway, which is Update by addition in this case.
+\begin{description}
+\item[Adding new sensor/actuator:] Localized to changes within the given sensor or actuator type.
+Update by modification, supported by the architecture now. Not in gateway, which is Update by addition in this case. 
+\item[Adding different types:] Update by addition will need to be built into the architecture in the sensor
+and actuator subsystems to encompass new types apart from thermometers and radiators. Radiators and
+thermometers are hardcoded into the gateway, it will need to be more abstract.
+\end{description}
 
-\subsubsection{Adding different types}
-Update by addition will need to be built into the architecture in the sensor and actuator subsystems to encompass new types apart from thermometers and radiators.
-Radiators and thermometers are hardcoded into the gateway, it will need to be more abstract.
+We have chosen to focus in the integration of new types of thermometers and actuators and
+addition of different sensor and actuator types will not be described any further.
 
-We have chosen to focus in the integration of new types of thermometers and actuators.
-
-\subsection{Refined Scenario \#2}
-Because of the Client/Server architecture of the gateway, adding new actuators is currently possible without architectural changes.
-With the current design, only actuators of the radiator type can be added.
-
-\subsection{Refined Scenario \#3}
-In order to have the system encompass a user interface that allows a user to change the preferred temperature from a web page, the gateway subsystem needs to become the Model part of a MVC pattern.
-
-This functionality is not part of the current architecture, even though it supports commands on the gateway. At least one new command needs to be implemented on the gateway.
-
-% =====================================================================
-\section{Architectural Design}
-
-\subsection{Refined Scenario \#1 - Tactics}
 \subsubsection{Modifiablity}
@@
 This harness will continue to be the main test setup for introduction into the gateway/model parts of the system.
-\subsection{Refined Scenario \#3 - Tactics}
+
+\subsection{Refined Scenario \#2}
+Because of the Client/Server architecture of the gateway, adding new actuators is currently possible without architectural changes.
+With the current design, only actuators of the radiator type can be added.
+
+\subsection{Refined Scenario \#3}
+In order to have the system encompass a user interface that allows a user to change the preferred temperature from a web page, the gateway subsystem needs to become the Model part of a MVC pattern.
+
+This functionality is not part of the current architecture, even though it supports commands on the gateway. At least one new command needs to be implemented on the gateway.
+
 Scenario \#3 requires the HS07 system to provide a web interface that can be used to change the temperature in the house.
 The response requirements sets some requirements to the availability of the http service.
+
 \subsubsection{Availability}
@@
 
-\clearpage
+% =====================================================================
+\section{Architectural Design}
 

### 2008-02-25 SAiP/Module1/H2/paper.tex
Creating more text for refinement

@@
 \section{Evaluation}
-The compliance of the Architectural Description of HS07 \cite{SAiP1H1} for the selected refined scenarios
-is discussed in the following chapters.
+
+The following sections will elaborate on the refined scenarios described in section \ref{sec:refinement}.
+The compliance of the Architectural Description of HS07 \cite{SAiP1H1} for the selected refined
+scenarios will be described and tactics that can be applied to the architectural design in order to
+fulfill the quality attributes are identified.
+
 \subsection{Refined Scenario \#1}
-The new device type can either be a new thermometer/actuator or a completely new type of device, i.e. a door.
+The new device type can either be a new thermometer/actuator or a completely new type of device, e.g. a door.
 
-\subsubsection{Adding new sensor/actuator}
-Localized to changes within the given sensor or actuator type. Update by modification, supported by the architecture now. Not in gateway, which is Update by addition in this case.
+\begin{description}
+\item[Adding new sensor/actuator:] y
+\begin{itemize}
+\item Currently the Gateway <-> sensor/actuator architecture, being decoupled by Service, supports adding a new Sensor implementation as long as it meets the specifications.
+The sepcification lie only in semantics, and require that a method getTemperature exist for the sensors (and notify for actuators).
+There is no interface and therefore no syntax checking at compile time (only runtime).
+\item The present architecture of the sensor/actuator do not directly support replacing the implementation with a new as the Service new the Thermometer type directly (update by modification as opposed to update by addition).
+It also performs a call to a pure virtual method from a base-class constructor, which is so wrong it is unmentionable.
+\end{itemize}
+\item[Adding different types:] x
+\begin{itemize}
+\item A new sensor/actuator type must only implement the getController method in order to be elligable for use with the Service. The architecture holds.
+\item For the Gateway there is a hard-coded reference to the Thermometer and Ragiator, in the sense that it has registerThermometer methods and what is notified to the oberservers are the temperature.
+In a general architecture with generic sensors and actuators this hard-coding will greatly limit the possibility of adding new types, and certainly require modifying the existing code (with the required regression test to boot) in order to do so.
+\end{itemize}
+\end{description}
 
-\subsubsection{Adding different types}
-Update by addition will need to be built into the architecture in the sensor and actuator subsystems to encompass new types apart from thermometers and radiators.
-Radiators and thermometers are hardcoded into the gateway, it will need to be more abstract.
+We have chosen to focus in the integration of new types of thermometers and actuators and
+addition of different sensor and actuator types will not be described any further.
 
-We have chosen to focus in the integration of new types of thermometers and actuators.
-
-\subsection{Refined Scenario \#2}
-Because of the Client/Server architecture of the gateway, adding new actuators is currently possible without architectural changes.
-With the current design, only actuators of the radiator type can be added.
-
-\subsection{Refined Scenario \#3}
-In order to have the system encompass a user interface that allows a user to change the preferred temperature from a web page, the gateway subsystem needs to become the Model part of a MVC pattern.
-
-This functionality is not part of the current architecture, even though it supports commands on the gateway. At least one new command needs to be implemented on the gateway.
-
-% =====================================================================
-\section{Architectural Design}
-
-\subsection{Refined Scenario \#1 - Tactics}
 \subsubsection{Modifiablity}

### 2008-02-25 trunk/SAiP/Module1/H2/paper.tex
More "bla" added to section 5.

@@
 \subsection{Refined Scenario \#1}
-The new device type can either be a new thermometer/actuator or a completely new type of device, i.e. a door.
+The new device type can either be a new thermometer/actuator or a completely new type of device, e.g. a door.
 
 \begin{description}
-\item[Adding new sensor/actuator:] Localized to changes within the given sensor or actuator type.
-Update by modification, supported by the architecture now. Not in gateway, which is Update by addition in this case. 
-\item[Adding different types:] Update by addition will need to be built into the architecture in the sensor
-and actuator subsystems to encompass new types apart from thermometers and radiators. Radiators and
-thermometers are hardcoded into the gateway, it will need to be more abstract.
+\item[Adding new sensor/actuator:] \
+\begin{itemize}
+\item Currently the Gateway <-> sensor/actuator architecture, being decoupled by Service, supports adding a new Sensor implementation as long as it meets the specifications.
+The sepcification lie only in semantics, and require that a method getTemperature exist for the sensors (and notify for actuators).
+There is no interface and therefore no syntax checking at compile time (only runtime).
+\item The present architecture of the sensor/actuator do not directly support replacing the implementation with a new as the Service new the Thermometer type directly (update by modification as opposed to update by addition).
+It also performs a call to a pure virtual method from a base-class constructor, which is so wrong it is unmentionable.
+\end{itemize}
+\item[Adding different types:] \
+\begin{itemize}
+\item A new sensor/actuator type must only implement the getController method in order to be elligable for use with the Service. The architecture holds.
+\item For the Gateway there is a hard-coded reference to the Thermometer and Ragiator, in the sense that it has registerThermometer methods and what is notified to the oberservers are the temperature.
+In a general architecture with generic sensors and actuators this hard-coding will greatly limit the possibility of adding new types, and certainly require modifying the existing code (with the required regression test to boot) in order to do so.
+\end{itemize}
 \end{description}
@@
 \begin{figure}[!htb]
-\centerline{\epsfig{figure="figures/mv-model",scale=0.8 }}
+\centerline{\epsfig{figure="figures/mv-model",scale=0.7 }}
 \caption{Module View of the web interface in the redesigned HS07 system}
@@
 Figure~\ref{fig:mv-model} illustrates the new packages needed in HS07 to include a web interface for the user.
-The model package creates a service with witch to get and set the values  need by the user to make a decision an any new target temperature.
-Both the gateway and the webapp uses their invokers to contact the model service for parameter manipulation.
+A \emph{Model-View-Controller} pattern has been applied in order to seperate the model from the view and the controller.
+The model package creates a service with which to get and set the values needed by the user to make a decision an any new target temperature, that is, the Model \emph{maintains a model of the system}.
+Both the gateway and the WebApp uses their invokers to contact the model service for parameter manipulation.
 
 The separation between the gateway and the model packages creates an opportunity for increase of security when that becomes a driving quality attribute, since the gateway information can be hidden within the system, letting the ModelService become the only exposed part.
-\clearpage
+
+In order to archieve the availability requirements, the Apache web server can be used as HTTP server for the WepApp. This is not depicted in the figure.
 
@@
 \end{figure}
-Figure~\ref{fig:mv-factory} shows the classes involved and their interdependencies for implementing the factory pattern for the sensor module.
+Figure~\ref{fig:mv-factory} shows the classes involved and their interdependencies for implementing the \emph{factory} pattern for the sensor module.
 Please refer to Figure~\ref{fig:seq-tfac} for its runtime usage.
@@
 \begin{figure}[!htb]
-\centerline{\epsfig{figure="figures/Active object",scale=0.8 }}
+\centerline{\epsfig{figure="figures/Active object",scale=0.6 }}
 \caption{Active Objects in the redesigned HS07 system}
@@
 \end{figure}
-Figure~\ref{fig:activeobjects} shows all the components of the HS07 system. The new parts are the ModelService and the HTTP Server, needed to create a suitable web interface for the user.
+
+Figure~\ref{fig:activeobjects} shows all the active components of the HS07 system.
+The new parts are the ModelService and the HTTP Server, needed to create a suitable web interface for the user.
+As can be seen on the figure, the HTTP server only accesses the model and not any internal parts of the system.
+
 \clearpage
 
-% XXX \subsection{Sequence Diagrams}
-% Text til figur her
-
 \begin{figure}[!htb]
-\centerline{\epsfig{figure="figures/seq-stt",scale=0.8 }}
+\centerline{\epsfig{figure="figures/seq-stt",scale=0.6 }}
 \caption{Sequence Diagram for the web requests in the redesigned HS07 system}
@@
 \end{figure}
-Figure~\ref{fig:seq-stt} shows the flow of contol when a web application uses the methods exposed throgh the model service.
-The amount of traffic on the model service will have no direct impact on the performance of the gateway service.
+
+Figure~\ref{fig:seq-stt} shows the flow of contol when a web application uses the methods exposed through the model service.
+The amount of trafic on the model service will have no direct impact on the performance of the gateway service, and the response times on the web application will not be directly impacted when the gateway is polling the sensors.
 
@@
 
-If the ModelService is non responsive, it is important for the gateway service to continue to work, "guessing" the target temperature either by cache or by default values.
+If the ModelService is non-responsive, it is important for the gateway service to continue to work, "guessing" the target temperature either by cache or by default values.
 
@@
 The sequence shows the use of the factory to create a decoupling between the actual Thermometer implementation and the ThermometerService.
-\clearpage
 
@@
 
+Figure \ref{fig:deployment} shows how the HTTP Server and the ModelService can be deployed on the Home Gateway next to the GatewayService.
+As all tasks on the Home Gateway are communicating using HTTP, the HTTP Server or the ModelService can be deployed on different hardware if required.
+E.g.
+the HTTP Server can be moved to other hardware in order to minimize the impact on the more critical functions of the GatewayService when accessing the web interface.
+
 \begin{figure}[!htb]
-\centerline{\epsfig{figure="figures/deployment",scale=0.8 }}
+\centerline{\epsfig{figure="figures/deployment",scale=0.7 }}
 \caption{Deployment in the redesigned HS07 system}

### 2008-02-25 SAiP/Module1/H2/paper.tex
Fixing newline with \
Short comment about implementation added.

@@
 these scenarios.
+
+The proposed design has been implemented as part of exercise 2.
 
@@
 \begin{description}
-\item[Adding new sensor/actuator:] y
+\item[Adding new sensor/actuator:] \
 \begin{itemize}
@@
 \end{itemize}
-\item[Adding different types:] x
+\item[Adding different types:] \
 \begin{itemize}

### 2008-02-25 trunk/SAiP/Module1/H2/paper.tex
Arrows fixed

@@
 \begin{itemize}
-\item Currently the Gateway <-> sensor/actuator architecture, being decoupled by Service, supports adding a new Sensor implementation as long as it meets the specifications.
+\item Currently the Gateway $\leftrightarrow$ sensor/actuator architecture, being decoupled by Service, supports adding a new Sensor implementation as long as it meets the specifications.
 The sepcification lie only in semantics, and require that a method getTemperature exist for the sensors (and notify for actuators).

### 2008-02-28 trunk/SAiP/Module1/H3/doc/beerwebstore.tex
Report files added.

%%%%
%%%% Beer Web Store
%%%%
\section{Software Architecture in Practice Case: Beer Web Store}

%%%
% Architectural Description
%%%
\newpage
\subsection{Software Architecture -- Beer Web Store}
A new Beer Web Store for specialty beers has the following main use
case:
\begin{quote}
  The customer browses the beer catalogue for descriptions and
  reviews. The customer selects appropriate beers for purchase and
  these are added to his shopping basket. When the customer checks
  out, payment is validated and he is presented with a receipt
\end{quote}
As a first step, the software architect of this systems has decided to
use a layered architectural style (with clients, server, and a
database in separate layers). He has decided that the communication to
the database will be done via a Fa\c cade pattern. The system will be
built in Java.

Discuss:
\begin{itemize}
    \item Outline how you would approach the tasks of creating the
    architecture for the Beer Web Store. Consider, e.g., which steps
    would you takes and in which order?
    \item Relate the definition of 'software architecture' in [Bass et
    al., 2003] to the elements of your design as outlined above. Give
    concrete examples of elements, relations, and structures.
    \item Apply [Perry and Wolf, 1992]'s model of software architecture as
      \begin{quote}
	{\it Software Architecture = \{Elements, Form, Rationale\}}
      \end{quote}
      to the Beer Web System
    \item Reflect on what happens if the words ``software'' and
    ``computing'' are removed from the definition of 'software
    architecture' in [Bass et al., 2003]
    \item The architect decides to create a full architecture
    description before embarking on any implementation of the
    system. Discuss pros and cons of taking that approach
\end{itemize}

%%%
% Architectural Description
%%%
\newpage
\subsection{Architectural Description -- Beer Web Store}
The architect for a new Beer Web Store for specialty beers has drawn
the following figure of the architecture together with his customers.
\begin{figure}[h!]
  \centerline{\epsfig{figure=figures/beerstore,scale=0.30,clip=true}}
\end{figure}
The system is to support web users that may purchase beers using a web
browser. The system runs on top of a database (``DB''), contains
domain model (``Data''), an overview page (``Overview''), and a
``Basket'' subpart.

Discuss:
\begin{itemize}
    \item Which structures does a system such as the above exhibit?
    Give examples of elements and relations pertaining to each
    structure
    \item Which viewpoints are relevant when describing the example?
    Give examples of partials views for each viewpoint
    \item Argue for benefits and liabilities of describing software
    architecture via a box-and-line drawing such as the above
    \item Discuss what architectural description would be needed if
    the system was to be realized as a Service-Oriented Architecture
\end{itemize}

%%%
% Quality Attributes
%%%
\newpage
\subsection{Quality Attributes -- Beer Web Store}

Consider the following main use case for a Beer Web Store for specialty
beers:

\begin{quote}
  The customer browses the beer catalogue for descriptions and
  reviews. The customer selects appropriate beers for purchase and
  these are added to his shopping basket. When the customer checks
  out, payment is validated through the payment service
\end{quote}
The main architectural drivers are considered to be performance and
availability. Moreover, the software architect for the system has
created the following box-and-line drawing documenting his initial
ideas of the system:
\begin{figure}[h!]
  \centerline{\epsfig{figure=figures/beerstore,scale=0.30,clip=true}}
\end{figure}

Discuss:
\begin{itemize}
    \item Describe technique(s) for architectural requirements capture
    that are applicable to the above case
    \item Give feasible architectural requirements for availability
    and performance for the Beer Web Store using such techniques (at
    least one for each quality attribute)
    \item Argue how tactics and/or styles may be used to resolve the requirements
    \item Discuss where Service-Oriented Architecture can play a role
    with respect to the Beer Web Store? Consider the quality implications
    \item Reflect upon what the role of quality attributes in software
    architecture is
\end{itemize}

%%%
% Architectural Design
%%%
\newpage
\subsection{Architectural Design -- Beer Web Store}
Consider the following main use case for a Beer Web Store for specialty
beers:
\begin{quote}
  The customer browses the beer catalogue for descriptions and
  reviews. The customer selects appropriate beers for purchase and
  these are added to his shopping basket. When the customer checks
  out, payment is validated through the payment service
\end{quote}
The main architectural drivers are considered to be performance and
availability. The software architect and stakeholders of the system
have based on this outlined the following two central Quality
Attribute Scenarios:
\begin{itemize}
  \item A customer makes a payment to the system in normal mode. The
  payment is processed within 10 seconds
  \item The payment service fails during normal operation. The system
  detects this. The administrator is notified and the system continues
  in degraded mode until the payment service is made available
\end{itemize}

Discuss:
\begin{itemize}
\item Give examples of performance tactics respectively availability
tactics that may be applied to the scenario. What are the consequences
of applying these tactics?
\item The software architect decides to use a layered architectural
style for the system. Discuss quality implications of this choice
\item Outline major elements of a possible component and connector
structure of the Beer Web Store system
\end{itemize}

%%%
% Product Lines and Frameworks
%%%
\newpage
\subsection{Product Lines -- Beer Web Store}
Consider a Beer Web Store for buying specialty beers. The software
architect has designed a deployment structure of the system as shown
below:
\begin{figure}[h!]
  \centerline{\epsfig{figure=figures/bws-deployment,scale=0.80}}
  \label{fig:observation}
\end{figure}
The main architectural drivers for the system are performance and
modifiability and the following is the main use case for the system:
\begin{quote}
  The customer browses the beer catalogue for descriptions and
  reviews. The customer selects appropriate beers for purchase and
  these are added to his shopping basket. When the customer checks
  out, payment is validated through the payment service
\end{quote}

The company behind the Beer Web Store has decided to build a product
line based on the store so that, e.g., sodas or stamps may be sold in
a new shop or other delivery platforms such as mobile phones may be
used.

Discuss:
\begin{itemize}
\item How will the choice of a product line approach affect the
  architecture of the beer web store?
  \item Discuss the benefits and liabilities of using a product line
    approach in relation to the case?
  \item Relate Service-Oriented Architecture and Product Lines?
\end{itemize}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "document"
%%% End:

### 2008-02-28 trunk/SAiP/Module1/H3/doc/beerwebstore.tex
Each section moved into seperate .tex file.

@@
 \section{Software Architecture in Practice Case: Beer Web Store}
-
-%%%
-% Architectural Description
-%%%
-\newpage
-\subsection{Software Architecture -- Beer Web Store}
-A new Beer Web Store for specialty beers has the following main use
-case:
-\begin{quote}
-  The customer browses the beer catalogue for descriptions and
-  reviews. The customer selects appropriate beers for purchase and
-  these are added to his shopping basket. When the customer checks
-  out, payment is validated and he is presented with a receipt
-\end{quote}
-As a first step, the software architect of this systems has decided to
-use a layered architectural style (with clients, server, and a
-database in separate layers). He has decided that the communication to
-the database will be done via a Fa\c cade pattern. The system will be
-built in Java.
-
-Discuss:
-\begin{itemize}
-    \item Outline how you would approach the tasks of creating the
-    architecture for the Beer Web Store. Consider, e.g., which steps
-    would you takes and in which order?
-    \item Relate the definition of 'software architecture' in [Bass et
-    al., 2003] to the elements of your design as outlined above. Give
-    concrete examples of elements, relations, and structures.
-    \item Apply [Perry and Wolf, 1992]'s model of software architecture as
-      \begin{quote}
-	{\it Software Architecture = \{Elements, Form, Rationale\}}
-      \end{quote}
-      to the Beer Web System
-    \item Reflect on what happens if the words ``software'' and
-    ``computing'' are removed from the definition of 'software
-    architecture' in [Bass et al., 2003]
-    \item The architect decides to create a full architecture
-    description before embarking on any implementation of the
-    system. Discuss pros and cons of taking that approach
-\end{itemize}
-
-%%%
-% Architectural Description
-%%%
-\newpage
-\subsection{Architectural Description -- Beer Web Store}
-The architect for a new Beer Web Store for specialty beers has drawn
-the following figure of the architecture together with his customers.
-\begin{figure}[h!]
-  \centerline{\epsfig{figure=figures/beerstore,scale=0.30,clip=true}}
-\end{figure}
-The system is to support web users that may purchase beers using a web
-browser. The system runs on top of a database (``DB''), contains
-domain model (``Data''), an overview page (``Overview''), and a
-``Basket'' subpart.
-
-Discuss:
-\begin{itemize}
-    \item Which structures does a system such as the above exhibit?
-    Give examples of elements and relations pertaining to each
-    structure
-    \item Which viewpoints are relevant when describing the example?
-    Give examples of partials views for each viewpoint
-    \item Argue for benefits and liabilities of describing software
-    architecture via a box-and-line drawing such as the above
-    \item Discuss what architectural description would be needed if
-    the system was to be realized as a Service-Oriented Architecture
-\end{itemize}
-
-%%%
-% Quality Attributes
-%%%
-\newpage
-\subsection{Quality Attributes -- Beer Web Store}
-
-Consider the following main use case for a Beer Web Store for specialty
-beers:
-
-\begin{quote}
-  The customer browses the beer catalogue for descriptions and
-  reviews. The customer selects appropriate beers for purchase and
-  these are added to his shopping basket. When the customer checks
-  out, payment is validated through the payment service
-\end{quote}
-The main architectural drivers are considered to be performance and
-availability. Moreover, the software architect for the system has
-created the following box-and-line drawing documenting his initial
-ideas of the system:
-\begin{figure}[h!]
-  \centerline{\epsfig{figure=figures/beerstore,scale=0.30,clip=true}}
-\end{figure}
-
-Discuss:
-\begin{itemize}
-    \item Describe technique(s) for architectural requirements capture
-    that are applicable to the above case
-    \item Give feasible architectural requirements for availability
-    and performance for the Beer Web Store using such techniques (at
-    least one for each quality attribute)
-    \item Argue how tactics and/or styles may be used to resolve the requirements
-    \item Discuss where Service-Oriented Architecture can play a role
-    with respect to the Beer Web Store? Consider the quality implications
-    \item Reflect upon what the role of quality attributes in software
-    architecture is
-\end{itemize}
-
-%%%
-% Architectural Design
-%%%
-\newpage
-\subsection{Architectural Design -- Beer Web Store}
-Consider the following main use case for a Beer Web Store for specialty
-beers:
-\begin{quote}
-  The customer browses the beer catalogue for descriptions and
-  reviews. The customer selects appropriate beers for purchase and
-  these are added to his shopping basket. When the customer checks
-  out, payment is validated through the payment service
-\end{quote}
-The main architectural drivers are considered to be performance and
-availability. The software architect and stakeholders of the system
-have based on this outlined the following two central Quality
-Attribute Scenarios:
-\begin{itemize}
-  \item A customer makes a payment to the system in normal mode. The
-  payment is processed within 10 seconds
-  \item The payment service fails during normal operation. The system
-  detects this. The administrator is notified and the system continues
-  in degraded mode until the payment service is made available
-\end{itemize}
-
-Discuss:
-\begin{itemize}
-\item Give examples of performance tactics respectively availability
-tactics that may be applied to the scenario. What are the consequences
-of applying these tactics?
-\item The software architect decides to use a layered architectural
-style for the system. Discuss quality implications of this choice
-\item Outline major elements of a possible component and connector
-structure of the Beer Web Store system
-\end{itemize}
-
-%%%
-% Product Lines and Frameworks
-%%%
-\newpage
-\subsection{Product Lines -- Beer Web Store}
-Consider a Beer Web Store for buying specialty beers. The software
-architect has designed a deployment structure of the system as shown
-below:
-\begin{figure}[h!]
-  \centerline{\epsfig{figure=figures/bws-deployment,scale=0.80}}
-  \label{fig:observation}
-\end{figure}
-The main architectural drivers for the system are performance and
-modifiability and the following is the main use case for the system:
-\begin{quote}
-  The customer browses the beer catalogue for descriptions and
-  reviews. The customer selects appropriate beers for purchase and
-  these are added to his shopping basket. When the customer checks
-  out, payment is validated through the payment service
-\end{quote}
-
-The company behind the Beer Web Store has decided to build a product
-line based on the store so that, e.g., sodas or stamps may be sold in
-a new shop or other delivery platforms such as mobile phones may be
-used.
-
-Discuss:
-\begin{itemize}
-\item How will the choice of a product line approach affect the
-  architecture of the beer web store?
-  \item Discuss the benefits and liabilities of using a product line
-    approach in relation to the case?
-  \item Relate Service-Oriented Architecture and Product Lines?
-\end{itemize}
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "document"
-%%% End:

### 2008-02-28 trunk/SAiP/Module1/H3/doc/productlines.tex
Each section moved into seperate .tex file.

%%%
% Product Lines and Frameworks
%%%
\subsection{Product Lines -- Beer Web Store}
Consider a Beer Web Store for buying specialty beers. The software
architect has designed a deployment structure of the system as shown
below:
\begin{figure}[h!]
  \centerline{\epsfig{figure=figures/bws-deployment,scale=0.80}}
  \label{fig:observation}
\end{figure}
The main architectural drivers for the system are performance and
modifiability and the following is the main use case for the system:
\begin{quote}
  The customer browses the beer catalogue for descriptions and
  reviews. The customer selects appropriate beers for purchase and
  these are added to his shopping basket. When the customer checks
  out, payment is validated through the payment service
\end{quote}

The company behind the Beer Web Store has decided to build a product
line based on the store so that, e.g., sodas or stamps may be sold in
a new shop or other delivery platforms such as mobile phones may be
used.

Discuss:
\begin{itemize}
\item How will the choice of a product line approach affect the
  architecture of the beer web store?
  \item Discuss the benefits and liabilities of using a product line
    approach in relation to the case?
  \item Relate Service-Oriented Architecture and Product Lines?
\end{itemize}

### 2008-02-28 SAiP/Module1/H3/doc/architecturaldescription.tex
Each section moved into seperate .tex file.

%%%
% Architectural Description
%%%
\subsection{Architectural Description -- Beer Web Store}
The architect for a new Beer Web Store for specialty beers has drawn
the following figure of the architecture together with his customers.
\begin{figure}[h!]
  \centerline{\epsfig{figure=figures/beerstore,scale=0.30,clip=true}}
\end{figure}
The system is to support web users that may purchase beers using a web
browser. The system runs on top of a database (``DB''), contains
domain model (``Data''), an overview page (``Overview''), and a
``Basket'' subpart.

Discuss:
\begin{itemize}
    \item Which structures does a system such as the above exhibit?
    Give examples of elements and relations pertaining to each
    structure
    \item Which viewpoints are relevant when describing the example?
    Give examples of partials views for each viewpoint
    \item Argue for benefits and liabilities of describing software
    architecture via a box-and-line drawing such as the above
    \item Discuss what architectural description would be needed if
    the system was to be realized as a Service-Oriented Architecture
\end{itemize}

### 2008-02-28 SAiP/Module1/H3/doc/architecturaldesign.tex
Each section moved into seperate .tex file.

%%%
% Architectural Design
%%%
\subsection{Architectural Design -- Beer Web Store}
Consider the following main use case for a Beer Web Store for specialty
beers:
\begin{quote}
  The customer browses the beer catalogue for descriptions and
  reviews. The customer selects appropriate beers for purchase and
  these are added to his shopping basket. When the customer checks
  out, payment is validated through the payment service
\end{quote}
The main architectural drivers are considered to be performance and
availability. The software architect and stakeholders of the system
have based on this outlined the following two central Quality
Attribute Scenarios:
\begin{itemize}
  \item A customer makes a payment to the system in normal mode. The
  payment is processed within 10 seconds
  \item The payment service fails during normal operation. The system
  detects this. The administrator is notified and the system continues
  in degraded mode until the payment service is made available
\end{itemize}

Discuss:
\begin{itemize}
\item Give examples of performance tactics respectively availability
tactics that may be applied to the scenario. What are the consequences
of applying these tactics?
\item The software architect decides to use a layered architectural
style for the system. Discuss quality implications of this choice
\item Outline major elements of a possible component and connector
structure of the Beer Web Store system
\end{itemize}

### 2008-02-28 SAiP/Module1/H3/doc/qualityattributes.tex
Each section moved into seperate .tex file.

%%%
% Quality Attributes
%%%
\subsection{Quality Attributes -- Beer Web Store}

Consider the following main use case for a Beer Web Store for specialty
beers:

\begin{quote}
  The customer browses the beer catalogue for descriptions and
  reviews. The customer selects appropriate beers for purchase and
  these are added to his shopping basket. When the customer checks
  out, payment is validated through the payment service
\end{quote}
The main architectural drivers are considered to be performance and
availability. Moreover, the software architect for the system has
created the following box-and-line drawing documenting his initial
ideas of the system:
\begin{figure}[h!]
  \centerline{\epsfig{figure=figures/beerstore,scale=0.30,clip=true}}
\end{figure}

Discuss:
\begin{itemize}
    \item Describe technique(s) for architectural requirements capture
    that are applicable to the above case
    \item Give feasible architectural requirements for availability
    and performance for the Beer Web Store using such techniques (at
    least one for each quality attribute)
    \item Argue how tactics and/or styles may be used to resolve the requirements
    \item Discuss where Service-Oriented Architecture can play a role
    with respect to the Beer Web Store? Consider the quality implications
    \item Reflect upon what the role of quality attributes in software
    architecture is
\end{itemize}

### 2008-02-28 SAiP/Module1/H3/doc/softwarearchitecture.tex
Each section moved into seperate .tex file.

%%%
% Architectural Description
%%%
\subsection{Software Architecture -- Beer Web Store}
A new Beer Web Store for specialty beers has the following main use
case:
\begin{quote}
  The customer browses the beer catalogue for descriptions and
  reviews. The customer selects appropriate beers for purchase and
  these are added to his shopping basket. When the customer checks
  out, payment is validated and he is presented with a receipt
\end{quote}
As a first step, the software architect of this systems has decided to
use a layered architectural style (with clients, server, and a
database in separate layers). He has decided that the communication to
the database will be done via a Fa\c cade pattern. The system will be
built in Java.

Discuss:
\begin{itemize}
    \item Outline how you would approach the tasks of creating the
    architecture for the Beer Web Store. Consider, e.g., which steps
    would you takes and in which order?
    \item Relate the definition of 'software architecture' in [Bass et
    al., 2003] to the elements of your design as outlined above. Give
    concrete examples of elements, relations, and structures.
    \item Apply [Perry and Wolf, 1992]'s model of software architecture as
      \begin{quote}
	{\it Software Architecture = \{Elements, Form, Rationale\}}
      \end{quote}
      to the Beer Web System
    \item Reflect on what happens if the words ``software'' and
    ``computing'' are removed from the definition of 'software
    architecture' in [Bass et al., 2003]
    \item The architect decides to create a full architecture
    description before embarking on any implementation of the
    system. Discuss pros and cons of taking that approach
\end{itemize}

### 2008-03-01 SAiP/Module1/H3/doc/softwarearchitecture.tex
Updating with answer to first question

@@
 %%%
-\subsection{Software Architecture -- Beer Web Store}
-A new Beer Web Store for specialty beers has the following main use
-case:
+\section{Software Architecture -- Beer Web Store}
+
+%A new Beer Web Store for specialty beers has the following main use
+%case:
+%\begin{quote}
+%  The customer browses the beer catalogue for descriptions and
+%  reviews. The customer selects appropriate beers for purchase and
+%  these are added to his shopping basket. When the customer checks
+%  out, payment is validated and he is presented with a receipt
+%\end{quote}
+%As a first step, the software architect of this systems has decided to
+%use a layered architectural style (with clients, server, and a
+%database in separate layers). He has decided that the communication to
+%the database will be done via a Fa\c cade pattern. The system will be
+%built in Java.
+
+\subsection{Question 1}
 \begin{quote}
-  The customer browses the beer catalogue for descriptions and
-  reviews. The customer selects appropriate beers for purchase and
-  these are added to his shopping basket. When the customer checks
-  out, payment is validated and he is presented with a receipt
+\emph{Outline how you would approach the tasks of creating the
+architecture for the Beer Web Store. Consider, e.g., which steps
+would you takes and in which order?}
 \end{quote}
-As a first step, the software architect of this systems has decided to
-use a layered architectural style (with clients, server, and a
-database in separate layers). He has decided that the communication to
-the database will be done via a Fa\c cade pattern. The system will be
-built in Java.
 
-Discuss:
+Naturally there are many ways to create an architecture, yet as the focus is on the architecture then ADD (Architecture Driven Design) is an obvious choice.
+ADD is merely a method that employs many of the techniques related to Software Architecture in a predefined structure.
+
+\subsubsection{Choose the module to decompose}
+As this is the beginning of the architectual design then the module is the entire system. In later iterations the sub-components (the layers, the components in the layers, etc.) will be decomposed.
+
+\subsubsection{Refine the module according to the following steps}
+
+\begin{description}
+    \item[a. Choose the architectual drivers]
+  Here the quality attributes of the system would be defined by the stakeholders of the system and then a vote would be used to determine which of the quality attributes are architectual drivers.
+The quality attributes would be expressed as quality attribute scenarios.
+    \item[b. Choose an architectual pattern]
+  Based on the architectual drivers selected in 1, use the mapping in Bass et al.
+as well as other literature and experience to determine some tactics for achieving the architectual drivers for the system.
+Then select some architectual patterns to realize the tactics.
+    \item[c. Instantiate modules using multiple views]
+  Based on the tactics and patterns found above the architecture for the system may be expressed using UML.
+It is \emph{always} required to use multiple views in order to properly describe the architecture.
+A good approach is the "3 + 1" structure, which uses three viewpoints; Module viewpoint, Component \& Connector viewpoint and Allocation viewpoint.
+These three viewpoint is expressed in UML using Package diagrams, Object and sequence diagrams, and deployment diagrams respectively.
+These diagrams should in the first iteration only look at relations and interactions between the main components (the layers).
+    \item[d. Define interfaces]
+  Define the interfaces between the main components (the layers).
+    \item[e. Verify scenarios and use cases]
+  Make sure that the steps a - d covered all the parts of the scenarios. If not include the missing parts and updates the scenarios/architecture is inconsistencies are found.
+\end{description}
+\subsubsection{Repeat the steps above for every module that needs further decomposition}
+The first iteration only look at the overall architecture (the leation and communication between layers).
+The following iterations will decompose the individual layers and as needed the individual components in the layers.
+\subsection{Question 2}
+
+\begin{quote}
+\emph{Give concrete examples of what elements, relations, and structures as defined in
+\cite{bass2003sa} could be in relation to an architecture for the Beer Web Store.}
+\end{quote}
+
 \begin{itemize}
-    \item Outline how you would approach the tasks of creating the
-    architecture for the Beer Web Store. Consider, e.g., which steps
-    would you takes and in which order?
     \item Relate the definition of 'software architecture' in [Bass et

### 2008-03-01 SAiP/Module1/H3/doc/softwarearchitecture.tex
Adding fixme and question 2 begin

@@
 \subsection{Question 1}
-\begin{quote}
-\emph{Outline how you would approach the tasks of creating the
+\begin{question}
+Outline how you would approach the tasks of creating the
 architecture for the Beer Web Store. Consider, e.g., which steps
-would you takes and in which order?}
-\end{quote}
+would you takes and in which order?
+\end{question}
 
@@
 
+\begin{question}
+Give concrete examples of what elements, relations, and structures as defined in \cite{bass2003sa} could be in relation to an architecture for the Beer Web Store.
+\end{question}
+\subsubsection{Elements}
+Based on the present decomposition the elements that are know the layers, where each layer is an element as described below.
+\begin{itemize}
+    \item[Clients] The clients access the server inorder to perform some action
+    \item[Server] The server is accessed by the clients and when needed access the database via its Fa\c cade pattern.
+    \item[Database] The database exposes its interface to the server via a Fa\c cade pattern.
+\end{itemize}
+\fixme{Insert element (class) diagram for the above}
+\subsubsection{Relations}
+\subsubsection{Strustures}
+\subsection{Question 3}
+\begin{question}
+\emph{Apply \cite{perrywolf1992}'s model of software architecture as
       \begin{quote}
-\emph{Give concrete examples of what elements, relations, and structures as defined in
-\cite{bass2003sa} could be in relation to an architecture for the Beer Web Store.}
+	{\it Software Architecture = \{Elements, Form, Rationale\}}
       \end{quote}
+      to the Beer Web System}
+\end{question}
 
 \begin{itemize}
-    \item Relate the definition of 'software architecture' in [Bass et
-    al., 2003] to the elements of your design as outlined above. Give
-    concrete examples of elements, relations, and structures.
     \item Apply [Perry and Wolf, 1992]'s model of software architecture as

### 2008-03-01 trunk/SAiP/Module1/H3/doc/architecturaldesign.tex
Formatting of questions completed.

--- 2008-02-28 SAiP/Module1/H3/doc/architecturaldesign.tex
@@
 %%%
-\subsection{Architectural Design -- Beer Web Store}
+\section{Architectural Design -- Beer Web Store}
 Consider the following main use case for a Beer Web Store for specialty
@@
 
-Discuss:
-\begin{itemize}
-\item Give examples of performance tactics respectively availability
+\subsection{Architectural Design - Question 1}
+
+\begin{question}
+Give examples of performance tactics respectively availability
 tactics that may be applied to the scenario. What are the consequences
 of applying these tactics?
-\item The software architect decides to use a layered architectural
+\end{question}
+
+\fixme{Answer question 4.1}
+
+\subsection{Architectural Design - Question 2}
+
+\begin{question}
+The software architect decides to use a layered architectural
 style for the system. Discuss quality implications of this choice
-\item Outline major elements of a possible component and connector
+\end{question}
+
+\fixme{Answer question 4.2}
+
+\subsection{Architectural Design - Question 3}
+
+\begin{question}
+Outline major elements of a possible component and connector
 structure of the Beer Web Store system
-\end{itemize}
+\end{question}
+
+\fixme{Answer question 4.3}

### 2008-03-01 SAiP/Module1/H3/doc/architecturaldescription.tex
Formatting of questions completed.

@@
 %%%
-\subsection{Architectural Description -- Beer Web Store}
+\section{Architectural Description -- Beer Web Store}
+
 The architect for a new Beer Web Store for specialty beers has drawn
@@
 
-Discuss:
-\begin{itemize}
-    \item Which structures does a system such as the above exhibit?
+\subsection{Architectural Description - Question 1}
+
+\begin{question}
+Which structures does a system such as the above exhibit?
 Give examples of elements and relations pertaining to each
 structure
-    \item Which viewpoints are relevant when describing the example?
+\end{question}
+
+\fixme{Answer question 2.1}
+
+\subsection{Architectural Description - Question 2}
+
+\begin{question}
+Which viewpoints are relevant when describing the example?
 Give examples of partials views for each viewpoint
-    \item Argue for benefits and liabilities of describing software
+\end{question}
+
+\fixme{Answer question 2.2}
+
+\subsection{Architectural Description - Question 3}
+
+\begin{question}
+Argue for benefits and liabilities of describing software
 architecture via a box-and-line drawing such as the above
-    \item Discuss what architectural description would be needed if
+\end{question}
+
+\fixme{Answer question 2.3}
+
+\subsection{Architectural Description - Question 4}
+
+\begin{question}
+Discuss what architectural description would be needed if
 the system was to be realized as a Service-Oriented Architecture
-\end{itemize}
+\end{question}
+
+\fixme{Answer question 2.4}

### 2008-03-01 SAiP/Module1/H3/doc/productlines.tex
Formatting of questions completed.

--- 2008-02-28 trunk/SAiP/Module1/H3/doc/productlines.tex
@@
 %%%
-\subsection{Product Lines -- Beer Web Store}
+\section{Product Lines -- Beer Web Store}
 Consider a Beer Web Store for buying specialty beers. The software
@@
 
-Discuss:
-\begin{itemize}
-\item How will the choice of a product line approach affect the
+\subsection{Product Lines - Question 1}
+
+\begin{question}
+How will the choice of a product line approach affect the
 architecture of the beer web store?
-  \item Discuss the benefits and liabilities of using a product line
+\end{question}
+
+\fixme{Answer question 5.1}
+
+\subsection{Product Lines - Question 2}
+
+\begin{question}
+Discuss the benefits and liabilities of using a product line
 approach in relation to the case?
-  \item Relate Service-Oriented Architecture and Product Lines?
-\end{itemize}
+\end{question}
+
+\fixme{Answer question 5.2}
+
+\subsection{Product Lines - Question 3}
+
+\begin{question}
+Relate Service-Oriented Architecture and Product Lines?
+\end{question}
+
+\fixme{Answer question 5.3}

### 2008-03-01 SAiP/Module1/H3/doc/qualityattributes.tex
Formatting of questions completed.

@@
 %%%
-\subsection{Quality Attributes -- Beer Web Store}
+\section{Quality Attributes -- Beer Web Store}
 
@@
 
-Discuss:
-\begin{itemize}
-    \item Describe technique(s) for architectural requirements capture
+\subsection{Quality Attributes - Question 1}
+
+\begin{question}
+Describe technique(s) for architectural requirements capture
 that are applicable to the above case
-    \item Give feasible architectural requirements for availability
+\end{question}
+
+\fixme{Answer question 3.1}
+
+\subsection{Quality Attributes - Question 2}
+
+\begin{question}
+Give feasible architectural requirements for availability
 and performance for the Beer Web Store using such techniques (at
 least one for each quality attribute)
-    \item Argue how tactics and/or styles may be used to resolve the requirements
-    \item Discuss where Service-Oriented Architecture can play a role
+\end{question}
+
+\fixme{Answer question 2.2}
+
+\subsection{Quality Attributes - Question 3}
+
+\begin{question}
+Argue how tactics and/or styles may be used to resolve the requirements
+\end{question}
+
+\fixme{Answer question 3.3}
+
+\subsection{Quality Attributes - Question 4}
+
+\begin{question}
+Discuss where Service-Oriented Architecture can play a role
 with respect to the Beer Web Store? Consider the quality implications
-    \item Reflect upon what the role of quality attributes in software
+\end{question}
+
+\fixme{Answer question 3.4}
+
+\subsection{Quality Attributes - Question 5}
+
+\begin{question}
+Reflect upon what the role of quality attributes in software
 architecture is
-\end{itemize}
+\end{question}
+
+\fixme{Answer question 3.5}

### 2008-03-01 SAiP/Module1/H3/doc/softwarearchitecture.tex
Formatting of questions completed.

@@
 
-\subsection{Question 1}
+\subsection{Software Architecture - Question 1}
 \begin{question}
@@
 The following iterations will decompose the individual layers and as needed the individual components in the layers.
-\subsection{Question 2}
+
+\subsection{Software Architecture - Question 2}
 
@@
 \end{question}
+
 \subsubsection{Elements}
@@
 \fixme{Insert element (class) diagram for the above}
+
 \subsubsection{Relations}
+
+\fixme{Answer question 1.2}
+
 \subsubsection{Strustures}
-\subsection{Question 3}
+
+\fixme{Answer question 1.2}
+
+\subsection{Software Architecture - Question 3}
+
 \begin{question}
-\emph{Apply \cite{perrywolf1992}'s model of software architecture as
-      \begin{quote}
-	{\it Software Architecture = \{Elements, Form, Rationale\}}
-      \end{quote}
-      to the Beer Web System}
-\end{question}
-
-\begin{itemize}
-    \item Apply [Perry and Wolf, 1992]'s model of software architecture as
+Apply \cite{perrywolf1992}'s model of software architecture as
 \begin{quote}
@@
 to the Beer Web System
-    \item Reflect on what happens if the words ``software'' and
+\end{question}
+
+\fixme{Answer question 1.3}
+
+\subsection{Software Architecture - Question 4}
+
+\begin{question}
+Reflect on what happens if the words ``software'' and
 ``computing'' are removed from the definition of 'software
 architecture' in [Bass et al., 2003]
-    \item The architect decides to create a full architecture
+\end{question}
+
+\fixme{Answer question 1.4}
+
+\subsection{Software Architecture - Question 5}
+
+\begin{question}
+The architect decides to create a full architecture
 description before embarking on any implementation of the
 system. Discuss pros and cons of taking that approach
-\end{itemize}
+\end{question}
+
+\fixme{Answer question 1.5}

### 2008-03-01 SAiP/Module1/H3/doc/softwarearchitecture.tex
Introduction reinserted.

@@
 
-%A new Beer Web Store for specialty beers has the following main use
-%case:
-%\begin{quote}
-%  The customer browses the beer catalogue for descriptions and
-%  reviews. The customer selects appropriate beers for purchase and
-%  these are added to his shopping basket. When the customer checks
-%  out, payment is validated and he is presented with a receipt
-%\end{quote}
-%As a first step, the software architect of this systems has decided to
-%use a layered architectural style (with clients, server, and a
-%database in separate layers). He has decided that the communication to
-%the database will be done via a Fa\c cade pattern. The system will be
-%built in Java.
+A new Beer Web Store for specialty beers has the following main use
+case:
+\begin{quote}
+  The customer browses the beer catalogue for descriptions and
+  reviews. The customer selects appropriate beers for purchase and
+  these are added to his shopping basket. When the customer checks
+  out, payment is validated and he is presented with a receipt
+\end{quote}
+As a first step, the software architect of this systems has decided to
+use a layered architectural style (with clients, server, and a
+database in separate layers). He has decided that the communication to
+the database will be done via a Fa\c cade pattern. The system will be
+built in Java.
 

### 2008-03-04 SAiP/Module1/H3/doc/softwarearchitecture.tex
Updating with diagram

@@
 \begin{itemize}
-    \item[Clients] The clients access the server inorder to perform some action
+    \item[Clients] The clients access the server in order to perform some action.
     \item[Server] The server is accessed by the clients and when needed access the database via its Fa\c cade pattern.
@@
 \end{itemize}
-\fixme{Insert element (class) diagram for the above}
+The above is quickly sketched in the box-and-line drawing below, where the layering may also be seen.
+\begin{figure}[!htb]
+\centerline{\epsfig{figure=figures/module_view,scale=1.0 }}
+\caption{Module-View overview}
+\label{fig:module_view}
+\end{figure}
+\clearpage
 
 \subsubsection{Relations}
+The figure xx shows some of the relations, namely the statical, yet it is also important to know the dynamical behaviour.
 

### 2008-03-06 SAiP/Module1/H3/doc/softwarearchitecture.tex
Spelling error corrected.
Adding sequence diagram and comments about relation and structure

@@
 \subsubsection{Elements}
-Based on the present decomposition the elements that are know the layers, where each layer is an element as described below.
+Based on the present decomposition the elements that are know are the layers, where each layer is an element as described below.
 \begin{itemize}
@@
 \centerline{\epsfig{figure=figures/module_view,scale=1.0 }}
-\caption{Module-View overview}
+\caption{Element overview}
 \label{fig:module_view}
@@
 The figure xx shows some of the relations, namely the statical, yet it is also important to know the dynamical behaviour.
+These may be expressed by e.g.
+a sequence diagram, as shown in Figure yy.
+In these it may be seen that the Client and Server has a week semantic relation through the http protocol, and the Server and Database has a strong syntactic relation from a compiled compatibility.
 
-\fixme{Answer question 1.2}
+\begin{figure}[!htb]
+\centerline{\epsfig{figure=figures/module_view,scale=1.0 }}
+\caption{Relation overview}
+\label{fig:sequence}
+\end{figure}
 
-\subsubsection{Strustures}
+\subsubsection{Structures}
 
-\fixme{Answer question 1.2}
+There are many structures which may be used to model a software system.
+Common for all of them is that they present a certain view of the system.
+Very rarely is a single structure sufficient to sufficiently express the architecture of a system.
+For this reason many structures are combined, each structure being represented by one or more views, to create a sufficient representation of the software system.
+
+The most popular combination of structures are the Module, the Component-and-Connector and the Allocation.
+Each of these three structures focus on a specific aspect of the software architecture; the static elements and their relation, the elements dynamic relations and the elements deployment on hardware, respectively.
+
+In each structure there is a number of views.
+Which view, and indeed which structure, that apply is quite dependent on not only the system, but also the iteration of the architectual analysis.
+In the first iteration of the beer-store it would make sense to choose more overall views, e.g.
+the Module Layered or Module Decomposition, whereas e.g.
+the Module Class would be too fine-grained for the first iteration.
+For the C\&C it would make sense to start with the Client-Server, and finally for the Allocation the Deployment view may give a good indication of the physical location and requirements of the different elements.
 

### 2008-03-06 trunk/SAiP/Module1/H3/doc/architecturaldescription.tex
Question 1 completed.

--- 2008-03-01 SAiP/Module1/H3/doc/architecturaldescription.tex
@@
 
-\fixme{Answer question 2.1}
+The information depicted in the figure above shows several
+software structures of the Beer Web Store to some extent.
+
+\subsubsection{Uses}
+
+The \emph{uses structure} shows how elements are related with regard
+to usage. One perspective of the uses structure is that the
+elements in the figure are ''Overview创, ''Basket创 and ''Data创
+which are all related and the database which is only related
+to ''Data创. This shows that the overview page and the basket
+has been decoupled from the database which therefore can be
+replaced without affecting them.
+
+\subsubsection{Layered}
+
+Another perspective is that the users, the webserver and the
+database are elements where the relations show that the users
+use the webserver and the webserver uses the database. This
+also shows the that the architecture is a
+\emph{strictly layered structure} where the users cannot
+access the database directly.
+
+\subsubsection{Client-Server}
+
+As there are several users, the elements of the
+\emph{strictly layered structure} can also be seen as a
+\emph{client-server structure}, where the users are clients
+and the webserver is a server. The webserver can also be
+seen as a client to the database server. This would make it
+possible to archieve better performance if several webservers
+are used to make load balancing of the user requests.
+
+\subsubsection{Deployment}
+
+With regard to a \emph{deployment structure}, the figure shows
+that the overview page, the basket and the domain model (elements)
+are allocated to (relations) on the webserver and the database is
+allocated to its own server.
 

### 2008-03-06 SAiP/Module1/H3/doc/architecturaldescription.tex
Very preliminary start up of answering question 2.1
Answer to question 2 started. Figures still missing.

@@
 \subsection{Architectural Description - Question 1}
+\label{sec:arch_desc_q1}
 
@@
 
-\fixme{Answer question 2.1}
+The information depicted in the figure above shows several
+software structures of the Beer Web Store to some extent.
+
+\subsubsection{Uses}
+
+The \emph{uses structure} shows how elements are related with regard
+to usage. One perspective of the uses structure is that the
+elements in the figure are ``Overview'', ``Basket'' and ``Data''
+which are all related and the database which is only related
+to ``Data''. This shows that the overview page and the basket
+has been decoupled from the database which therefore can be
+replaced without affecting them.
+
+\subsubsection{Layered}
+
+Another perspective is that the users, the webserver and the
+database are elements where the relations show that the users
+use the webserver and the webserver uses the database. This
+also shows the that the architecture is a
+\emph{strictly layered structure} where the users cannot
+access the database directly.
+
+\subsubsection{Client-Server}
+
+As there are several users, the elements of the
+\emph{strictly layered structure} can also be seen as a
+\emph{client-server structure}, where the users are clients
+and the webserver is a server. The webserver can also be
+seen as a client to the database server. This would make it
+possible to archieve better performance if several webservers
+are used to make load balancing of the user requests.
+
+\subsubsection{Deployment}
+
+With regard to a \emph{deployment structure}, the figure shows
+that the overview page, the basket and the domain model (elements)
+are allocated to (relations) on the webserver and the database is
+allocated to its own server.
 
@@
 \end{question}
+
+Selecting the relevant views shall actually be performed on
+the basis of the architectural drivers which have not been
+determined for the Beer Web Store. There are however several
+models that provide a number of views that are often useful,
+e.g. the ``4+1'' model \cite{kruchten1995} and the ``3+1''
+model \cite{christensen2004archdesc} and many others.
+
+As described in \ref{sec:arch_desc_q1} the figure shows parts
+of a module viewpoint (uses, layered), component \& connector
+viewpoint (client-server) and an allocation viewpoint
+(deployment) which are the recommended viewpoints in
+\cite{christensen2004archdesc}.
+
+These viewpoints are all relevant to the Beer Web Store example.
+The Module view describes the internals of the webserver and how
+the functionality is decomposed into units. The Component \& connector
+viewpoint describes the runtime relations between the units and the
+Allocation viewpoint describes the environment of the running
+system.
+
+XXX: Udvælg views og lav figurer:
+
+- Layered model (refer to figure 1) and decomposition of the server?
+- Sequence diagram
+- Deployment view
 

### 2008-03-06 trunk/SAiP/Module1/H3/doc/softwarearchitecture.tex
Adding first part of Perry&Wolf

--- 2008-03-06 SAiP/Module1/H3/doc/softwarearchitecture.tex
@@
 \begin{figure}[!htb]
-\centerline{\epsfig{figure=figures/module_view,scale=1.0 }}
+\centerline{\epsfig{figure=figures/sequence,scale=1.0 }}
 \caption{Relation overview}
@@
 
+Perry and Wolf suggest seperating the model of the architecture into three components, as mentioned in the question.
+The components relate in the sense that the \emph{elements} has a particular \emph{form} and them and their relations are chosen based on a \emph{rational}.
+
+There are three class of architectual elements defined.
+\begin{itemize}
+    \item[Processing elements] Transforms data elements.
+    \item[Data elements] Information to be transformed or used.
+    \item[Connecting elements] The \emph{glue} that hold the architecture together.
+May at times be \emph{Data elements}, \emph{Connecting elements} or both.
+The connecting elements are the relations between the different elements, e.g.
+procedure calls or message parsing.
+\end{itemize}
+
+The architectual form describe the architectual properties and the relationships and their relative importance. This is important as some architectual properties
+may be vital to the design and others may be gold-plating.
+These properties are used to put constraints on the elements and also define the relative importanse of these constraints.
+The relationships are similar to the properties except they are used to put constraints and relative priority on the relations between elements.
+
+A software architecture is a selection and description of architectual choices.
+These choices are, however, difficult to convay without including the reason for the choices made, for the elemenents and form not chosen and for choices not made due to lack of information or because they were deemed irrelevant.
+
+When the above is applied to the information present for the Beer Web Store three processing elements, one data element and two connecting elements may be defined. These are as follow:
+
+\begin{itemize}
+    \item Processing elements.
+	 \begin{itemize}
+	    \item[Client]
+	    \item[Server]
+	    \item[DatabaseFacade]
+	 \end{itemize}
+    \item Data elements.
+	 \begin{itemize}
+	    \item[Request data]
+	 \end{itemize}
+    \item Connecting elements.
+	 \begin{itemize}
+	    \item[HTTP]
+	    \item[Database facade methods]
+	 \end{itemize}
+\end{itemize}
+
 \fixme{Answer question 1.3}

### 2008-03-12 SAiP/Module1/H3/doc/softwarearchitecture.tex
Fixing error in sequence diagram include
Finishing Perry&Wolf description

@@
 
-\fixme{Answer question 1.3}
+Perry and Wolf suggest seperating the model of the architecture into three components, as mentioned in the question.
+The components relate in the sense that the \emph{elements} has a particular \emph{form} and them and their relations are chosen based on a \emph{rational}.
+
+There are three class of architectual elements defined.
+\begin{itemize}
+    \item[Processing elements] Transforms data elements.
+    \item[Data elements] Information to be transformed or used.
+    \item[Connecting elements] The \emph{glue} that hold the architecture together.
+May at times be \emph{Data elements}, \emph{Connecting elements} or both.
+The connecting elements are the relations between the different elements, e.g.
+procedure calls or message parsing.
+\end{itemize}
+
+The architectual form describe the architectual properties and the relationships and their relative importance. This is important as some architectual properties
+may be vital to the design and others may be gold-plating.
+These properties are used to put constraints on the elements and also define the relative importanse of these constraints.
+The relationships are similar to the properties except they are used to put constraints and relative priority on the relations between elements.
+
+A software architecture is a selection and description of architectual choices.
+These choices are, however, difficult to convay without including the reason for the choices made, for the elemenents and form not chosen and for choices not made due to lack of information or because they were deemed irrelevant.
+
+When the above is applied to the information present for the Beer Web Store three processing elements, one data element and two connecting elements may be defined.
+
+\subsubsection{Elements in Beer Web Store}
+
+\begin{itemize}
+    \item Processing elements.
+	 \begin{itemize}
+	    \item[Client] Receive user entries and process them into http requests
+	    \item[Server] Receive http requests and transform them into high level database lookup
+	    \item[DatabaseFacade] Transform high level database lookup into proprietary database lookup
+	 \end{itemize}
+    \item Data elements.
+	 \begin{itemize}
+	    \item[Request data] The packets of data flowing from the client to the server and form the server to the database, e.g. Beer type, Credit card information, search refinement parameter, etc.
+	 \end{itemize}
+    \item Connecting elements.
+	 \begin{itemize}
+	    \item[HTTP] Protocol for connecting the server and client (loose coupling).
+	    \item[Database facade methods] Protocol for connecting the server to the database (medium-loose coupling).
+	 \end{itemize}
+\end{itemize}
+
+\subsubsection{Form in Beer Web Store}
+\begin{itemize}
+    \item[Properties] The Server and the database, hidden behind the Facade, is the backbone of the Beer Web Store.
+If either is unavailable then the other is irrelevant.
+Multiple instances of both Server and Database may be running simultaniously in order to supply hot-swapping.
+The importance of up-time (availability), security (payment), usability (transmission abilities), etc.
+must be defined here.
+    \item[Relationship] The Client may only see the server through its HTTP interface, and the Server may only access the database through the Database Facade.
+Other demands to the relationship (certain commands to the database may only be accessed when using https).
+\end{itemize}
+
+\subsubsection{Rational in Beer Web Store}
+
+Finally it is important to describe the decisions made when defining the form - why is availability important? Why is it OK to sacrifice availability for some (security) functionality.
 

### 2008-03-12 SAiP/Module1/H3/doc/softwarearchitecture.tex
Updating with contemplation for Bass definition

@@
 
-\fixme{Answer question 1.4}
+If the references to software and computing is removed the definition talks about elements, structures and properties, and their relations.
+This definition may be used about anything which may be broken down into generalised components.
+As an example could be used the construction of a rifle.
+Describing the elements that goes into making a rifle (hammer, barrol, ...), the structures that comes from putting them together, and the individual relationships that must be met (form fit, etc.).
+The elements may fit together in different ways and multiple structures may be required to complete the description.
+(the hammer up, the hammer down, etc.).
+The properties are in fact only the externally visible to the given element, as those are the only properties that are relevant to the architecture.
+These details include things like the strength of the metal, which may be different for the hammer and the pipe, the price range, produktion time, etc.
+while things like the cheak-pad may be optional.
+
+The description of an architecture, as defined by Bass et.
+al.
+is therefore very general and apart from the use of words like "software" and "computer", may apply to any architecture, be it a building, a produktion line for a rifle, or a embedded software system.
 

### 2008-03-13 trunk/SAiP/Module1/H3/doc/softwarearchitecture.tex
Adding question 5

@@
 
-When the above is applied to the information present for the Beer Web Store three processing elements, one data element and two connecting elements may be defined. These are as follow:
+When the above is applied to the information present for the Beer Web Store three processing elements, one data element and two connecting elements may be defined.
+
+\subsubsection{Elements in Beer Web Store}
 
@@
 	 \begin{itemize}
-	    \item[Client]
-	    \item[Server]
-	    \item[DatabaseFacade]
+	    \item[Client] Receive user entries and process them into http requests
+	    \item[Server] Receive http requests and transform them into high level database lookup
+	    \item[DatabaseFacade] Transform high level database lookup into proprietary database lookup
 	 \end{itemize}
@@
 	 \begin{itemize}
-	    \item[Request data]
+	    \item[Request data] The packets of data flowing from the client to the server and form the server to the database, e.g. Beer type, Credit card information, search refinement parameter, etc.
 	 \end{itemize}
@@
 	 \begin{itemize}
-	    \item[HTTP]
-	    \item[Database facade methods]
+	    \item[HTTP] Protocol for connecting the server and client (loose coupling).
+	    \item[Database facade methods] Protocol for connecting the server to the database (medium-loose coupling).
 	 \end{itemize}
@@
 
-\fixme{Answer question 1.3}
+\subsubsection{Form in Beer Web Store}
+\begin{itemize}
+    \item[Properties] The Server and the database, hidden behind the Facade, is the backbone of the Beer Web Store.
+If either is unavailable then the other is irrelevant.
+Multiple instances of both Server and Database may be running simultaniously in order to supply hot-swapping.
+The importance of up-time (availability), security (payment), usability (transmission abilities), etc.
+must be defined here.
+    \item[Relationship] The Client may only see the server through its HTTP interface, and the Server may only access the database through the Database Facade.
+Other demands to the relationship (certain commands to the database may only be accessed when using https).
+\end{itemize}
+
+\subsubsection{Rational in Beer Web Store}
+
+Finally it is important to describe the decisions made when defining the form - why is availability important? Why is it OK to sacrifice availability for some (security) functionality.
 
@@
 
-\fixme{Answer question 1.5}
+There are generally two approaches to creating an architecture, as defined y Bass et.
+al.
+and that is to define it in advance, and only define the overall structure and then let it grow from the following implementations.
+
+Creating an architecture only makes sense if it is believed that it may be used in other products, and the number of products in the pipeline and how well their commonalities may be predicted, helps to determine which approach to use.
+Naturally the situation is not black and white, as there are many possibilities between the fully evolutionary architecture to the fully initially defined, yet for the purpose of debating advantages and disadvantages the simple black and white is simply easier.
+
+Defining the architecture fully up front means a big investment in time and money in order to create the arcitecture and the first product based on it, so if it is the only product made, then it will go way over budget both in time and money.
+The idea is that creating the next product will be significantly cheaper, because a very large part of the products functionality is covered by the architecture and very little (if any) redesign has to be done.
+As more and more products are created based on the architecture their individual budget savings will help pay for the initial investment, eventually leading to a profit compared to building the products from scratch each time.
+
+Letting the architecture evolve means that only the basics are defined up front, only causing a small initial investment, before building the first product.
+When the second product is to be created more of the architecture is defined based on the parts of the first product, which may be used in the second.
+Naturally the second product is faster to build than the first, due to reuse, yet not as fast as the second product when the entire architectur was defined in advance.
+As more products come along the amount of reuse grow, and would eventually reach the same as the one where everything was defined in advance.
+
+If the architecture is used to define a Product Line it becomes even more pronounced, and the above two approaches are defined as Heavyweight and Lightweight, respectively.
+A statistical analysis of these two approaches, compared to simply building the products from scratch, can be shown in figure xx.
+
+\clearpage
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure=figures/product-line,scale=1.0 }}
+\caption{Heavy-weight vs. Light-weight}
+\label{fig:product-line}
+\end{figure}
+
+Deciding which approach to use, is more than just a matter of choosing between Heavyweight and Lightweight (or fully defined or evolutionary) based on the number of expected products, it is also a matter of how well the common architecture of the future products can be predicted, for if the architecture needs to be changed from product to product then all the advantage is lost.
+
+Based on this the Beer Web Store architect must have decided that the products in the pipe-line and the Beer Web Store share a common architecture and that the future products are so well know that the architecture can be fully specified in advance.

### 2008-03-15 SAiP/Module1/H3/doc/architecturaldescription.tex
Adding a little about discussion
Adding UML description

@@
 
-\fixme{Answer question 2.3}
+The benifits of a box-and-line drawing is that it can be written and "understood" by almost anyone, and in the initial phases of the architectual design they may be fine.
+They are fast to draw and can be shown to non-technical stakeholders.
+It is kind of like simple signlanguage - everyone understands nodding and shaking ones head or indicating a direction as long as it is for general information.
+
+Unfortunately the general information is where the advantages stop.
+For simple sign-language, it may quickly come to an argument if one wishes to know if the indication of "this way" means "a short way that way" or "a long way that way", and the same is true for box-and-line drawings.
+"How many clients does the system support?".
+"How strong is the binding indicated by the line - do all lines mean the same?" There is merely a line, so what does that line mean?
+
+Working as an developer/tester/engineer means that everything must be quantifiable in some way, and box-and-line drawings do not offer this unambigious quantifiability.
+For this a clearly defined modeling language, without ambiguaty as to the meaning of the model, is needed.
+An example of such a language is Unified Modelling Language(UML), which would allow the box-and-line drawing to be represented as one or more diagrams solving all the ambiguaty of the box-and-line drawing with only a single requirement; one must know the language, which is in deed one of the reasons while box-and-line drawings are still widely used - there is no single unambigious modelling language know by all stakeholders, and when that is the case it may be more confusing then benificial to spend time making sure the diagram is clear and unambigious, if that clarity is derived from symbolic notation that the readers (stakeholders) do not understand.
 

### 2008-03-15 SAiP/Module1/H3/doc/productlines.tex
Moving som description from Architectual description to here

@@
 
-\fixme{Answer question 5.2}
+Creating an architecture product line only makes sense if it is believed that it may be used in other products, and the number of products in the pipeline and how well their commonalities may be predicted, helps to determine which approach to use.
+Naturally the situation is not black and white, as there are many possibilities between the fully evolutionary architecture to the fully initially defined, yet for the purpose of debating advantages and disadvantages the simple black and white is simply easier.
+
+Defining the architecture fully up front means a big investment in time and money in order to create the arcitecture and the first product based on it, so if it is the only product made, then it will go way over budget both in time and money.
+The idea is that creating the next product will be significantly cheaper, because a very large part of the products functionality is covered by the architecture and very little (if any) redesign has to be done.
+As more and more products are created based on the architecture their individual budget savings will help pay for the initial investment, eventually leading to a profit compared to building the products from scratch each time.
+
+Letting the architecture evolve means that only the basics are defined up front, only causing a small initial investment, before building the first product.
+When the second product is to be created more of the architecture is defined based on the parts of the first product, which may be used in the second.
+Naturally the second product is faster to build than the first, due to reuse, yet not as fast as the second product when the entire architectur was defined in advance.
+As more products come along the amount of reuse grow, and would eventually reach the same as the one where everything was defined in advance.
+
+If the architecture is used to define a Product Line it becomes even more pronounced, and the above two approaches are defined as Heavyweight and Lightweight, respectively.
+A statistical analysis of these two approaches, compared to simply building the products from scratch, can be shown in figure xx.
+
+\clearpage
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure=figures/product-line,scale=1.0 }}
+\caption{Heavy-weight vs. Light-weight}
+\label{fig:product-line}
+\end{figure}
+
+Deciding which approach to use, is more than just a matter of choosing between Heavyweight and Lightweight (or fully defined or evolutionary) based on the number of expected products, it is also a matter of how well the common architecture of the future products can be predicted, for if the architecture needs to be changed from product to product then all the advantage is lost.
+
+Based on this the Beer Web Store architect must have decided that the products in the pipe-line and the Beer Web Store share a common architecture and that the future products are so well know that the architecture can be fully specified in advance.
 

### 2008-03-15 trunk/SAiP/Module1/H3/doc/softwarearchitecture.tex
Finishing section about fully specifying up front

@@
 
-There are generally two approaches to creating an architecture, as defined y Bass et.
+There are generally two approaches to creating an architecture, as defined by Bass et.
 al.
@@
 
-Creating an architecture only makes sense if it is believed that it may be used in other products, and the number of products in the pipeline and how well their commonalities may be predicted, helps to determine which approach to use.
-Naturally the situation is not black and white, as there are many possibilities between the fully evolutionary architecture to the fully initially defined, yet for the purpose of debating advantages and disadvantages the simple black and white is simply easier.
+There are also two reasons for creating an architecture; in order to create a product that exhibits some of the quality attributes defined in the architecture (modifiability, usability, ...) and in order to reuse the architecture (or part of the architecture) as a product line for creating other products.
+This latter reason is described in section yy.
 
-Defining the architecture fully up front means a big investment in time and money in order to create the arcitecture and the first product based on it, so if it is the only product made, then it will go way over budget both in time and money.
-The idea is that creating the next product will be significantly cheaper, because a very large part of the products functionality is covered by the architecture and very little (if any) redesign has to be done.
-As more and more products are created based on the architecture their individual budget savings will help pay for the initial investment, eventually leading to a profit compared to building the products from scratch each time.
+Creating an architecture up front allows for the definition of all properties of the architecture; the quality attributes and their individual priorities, the individual elements/components and how they interconnect both at runtime and statically, etc.
+This has the benifit that when the design and implementation phase is begun it may easily be broken down into parallel activities and integrated later, as everyone has a clear picture of what is important and what the interfaces are.
 
-Letting the architecture evolve means that only the basics are defined up front, only causing a small initial investment, before building the first product.
-When the second product is to be created more of the architecture is defined based on the parts of the first product, which may be used in the second.
-Naturally the second product is faster to build than the first, due to reuse, yet not as fast as the second product when the entire architectur was defined in advance.
-As more products come along the amount of reuse grow, and would eventually reach the same as the one where everything was defined in advance.
+Unfortunately defining the architecture fully up front only makes sense if there is enough information in advance to create the architecture form.
+Very often there are "dark horses", things that do not become apparent until the design and implementation phase.
+If it is not believed that sufficient information exist to fully define the architecture, then doing so may not only be a waste, it may lead to the very dangerous loose-loose principle, where e.g.
+an overhead is introduced to service modifiability, only to be hacked in order to service a performance issue that was not apparent at the architectual design pahse.
+This means that part of the overhead is preserved (in memory-footprint if nowhere else), without preserving any of the modifiability advantages.
 
-If the architecture is used to define a Product Line it becomes even more pronounced, and the above two approaches are defined as Heavyweight and Lightweight, respectively.
-A statistical analysis of these two approaches, compared to simply building the products from scratch, can be shown in figure xx.
+Another requirement to defining the architecture fully up front has to do with the organization.
+It may be good and dandy to create a fully defined architecture with focus on parallel implementation, if there is only one in-house developer to implement the product.
+Based on this the Beer Web Store architect must have decided that there are sufficient information to fully define the architecture and also that there is a reason (parallel implementation, out-sourceing of the design/implementation, etc.).
 
-\clearpage
-
-\begin{figure}[!htb]
-\centerline{\epsfig{figure=figures/product-line,scale=1.0 }}
-\caption{Heavy-weight vs. Light-weight}
-\label{fig:product-line}
-\end{figure}
-
-Deciding which approach to use, is more than just a matter of choosing between Heavyweight and Lightweight (or fully defined or evolutionary) based on the number of expected products, it is also a matter of how well the common architecture of the future products can be predicted, for if the architecture needs to be changed from product to product then all the advantage is lost.
-
-Based on this the Beer Web Store architect must have decided that the products in the pipe-line and the Beer Web Store share a common architecture and that the future products are so well know that the architecture can be fully specified in advance.
+Asuming that the architect in question has followed the course Software Architecture in Practice, he has obviously looked at the architectual properties, the stake holders, the organization and the requirement and deemed that there is sufficient grounds to fully specify the architecture in advance, with the risks and possible gains that it has.

### 2008-03-15 SAiP/Module1/H3/doc/architecturaldescription.tex
Adding beginning of SOA

@@
 
-\fixme{Answer question 2.2}
+\fixme{Finish question 2.2}
 
@@
 
-\fixme{Answer question 2.4}
+In a sense there are already aspects of SOA in the architecture, as the validation of the credit card information is most likely performed using a service on the internet - not that web services and SOA is equivalent, but they do complement each other nicely.
+
+As SOA consists of a collection of (preferrably stateless) well-defined services what needs to be desctibed is, which existing (external) services are needed and which new services has to be created.
+This may be done using the MacKenzie Topic maps, which describe the static dependencies between the services.
+
+The static description is naturally not sufficient, and two other views may be presented to show the Dynamic interrelations; Visibility and Interaction.
+
+\fixme{Finish question 2.4}

### 2008-03-16 SAiP/Module1/H3/doc/qualityattributes.tex
Adding more text about architectural properties

@@
 
-\fixme{Answer question 3.1}
+Defining an architectures requirements is not like defining the functional requirements; actually many software professionals believe that they are indeed orthogonal.
+This is supported by the fact that a set of functional requirements may be implemented by a number of different architectures.
+
+Finding out the requirements for the architecture is therefore more a matter of finding out what is important, more than it is to define what the product must do.
+The people who are able to define what is important to the architecture is naturally the stakeholders that the architecture affects, i.e.
+the owner, the user, the developer, the special interest groups, etc.
+Getting all of these peoples, often conflicting, requirements assembled and, even more importantly, prioritised, can be a difficult task.
+One way of doing so is to define scenarios, like the one mentioned above, and then having the stakeholders vote at the importance of the scenario.
+
+Naturally in order for the scenarios to be able to be used to clarify the architecture it must be possible quantify them, and place them in different categories.
+This requires that the above scenario(s) are extended with e.g.
+response time, how many times the user should be able to perform an operation before experiencing a fault (down-town).
+
+Once the scenarios are clarified and voted on, then the top priority scenarios may be formalized and it may be determined which quality attribute(s) they involve.
+For examples sake lets asume that the above scenario was selected as top priority.
+
+\fixme{Insert table with quality attribute breakdown}
+
+As it may be seen, the scenario(s) focus primarily on \emph{Availability} (how rare should it be that a user experience non-resonsiveness), \emph{Usability} (how easy should it be for the user to cancel a transaction halfway through and how easy should it be to pick up a previous transaction), \emph{Performance} (how big a delay can the user expect when performing an operation like a lookup in the database), \emph{Security} (how to avoid someone using stolen credit cards and also prevent a "hacker" from stealing the credit card information either durig the transaction or from the Beer Web Store's database.
+
+The scenario does not include all aspects of the architecture, though.
+Another scenario may deal with how long it may take to alter a product built on the architecture (\emph{Modifiability}) - this may be particularly interesting to the owner.
+Another quality that has not been dealt with in the mentioned scenario(s), is how easy it is to check a product developed based on the architecture for faults (\emph{Testability}), which is of great interest to the owner and the developers.
+
+Naturally there are many more scenario required in order to properly cover all aspects of the quality properties.
 

### 2008-03-16 trunk/SAiP/Module1/H3/doc/qualityattributes.tex
Update with more comments about scenario

--- 2008-03-16 SAiP/Module1/H3/doc/qualityattributes.tex
@@
 
+The scenarios may be further clarified in order to include the aspects of availability and performance, as shown below.
+
+\begin{quote}
+The customer browses the beer catalogue for descriptions and reviews and receives the result within 5 seconds 99\% of the time, 30 seconds 0,9\% of the time and more than 30 seconds 0,1\% of the time.
+
+The customer selects appropriate beers for purchase and these are added to his shopping basket. When the customer checks out, payment is validated through the payment service within 30 seconds.
+\end{quote}
+
 \fixme{Answer question 2.2}

### 2008-03-17 trunk/SAiP/Module1/H3/doc/architecturaldescription.tex
Figures and text added to the architectural descriptions.

@@
 to usage. One perspective of the uses structure is that the
-elements in the figure are ''Overview创, ''Basket创 and ''Data创
+elements in the figure are ``Overview'', ``Basket'' and ``Data''
 which are all related and the database which is only related
-to ''Data创. This shows that the overview page and the basket
+to ``Data''. This shows that the overview page and the basket
 has been decoupled from the database which therefore can be
@@
 
-\fixme{Answer question 2.2}
+Selecting the relevant views shall actually be performed on
+the basis of the architectural drivers which have not been
+determined for the Beer Web Store. There are however several
+models that provide a number of views that are often useful,
+e.g. the ``4+1'' model \cite{kruchten1995} and the ``3+1''
+model \cite{christensen2004archdesc} and many others.
+
+As described in section \ref{sec:arch_desc_q1} the figure shows
+parts of a module viewpoint (uses, layered), component \&
+connector viewpoint (client-server) and an allocation viewpoint
+(deployment) which are the recommended viewpoints in
+\cite{christensen2004archdesc}.
+
+These viewpoints are all relevant to the Beer Web Store example.
+The Module viewpoint describes the internals of the webserver and how
+the functionality is decomposed into units. The Component \& connector
+viewpoint describes the runtime relations between the units and the
+Allocation viewpoint describes the environment of the running
+system.
+
+In the following sections we will give examples on views for each
+of the three viewpoints.
+
+\subsubsection{Module Viewpoint}
+
+Figure \ref{fig:ad_package} shows a package diagram of the Beer Web
+Store and a decomposition of the web package. The Web package
+contains the elements used for the web interface, 
+the Overview page, the basket and the domain model. The domain model
+depends on a Database package that contains the database facade and the
+database facade uses some database package library, e.g. MySQL.
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure=figures/ad_package,scale=0.8 }}
+\caption{Package diagram for the Beer Web Store}
+\label{fig:ad_package}
+\end{figure}
+
+\subsubsection{Component \& Connector Viewpoint}
+
+There are only two active components in the Beer Web Store, a web server,
+e.g. Apache, and a database server. An active objects diagram will
+not be included here.
+
+Figure \ref{fig:ad_sequence} shows a sequence diagram where a user searches
+for a an item ``tuborg'' and adds one of the search results to his basket.
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure=figures/ad_sequence,scale=0.6 }}
+\caption{Sequence diagram showing a search and adding of an item}
+\label{fig:ad_sequence}
+\end{figure}
+
+The domain model accesses the database through the database facade
+when needed. This can be expressed in another sequence diagram.
+Sequence diagrams showing the checkout including payment should
+also be made.
+
+\subsubsection{Allocation Viewpoint}
+
+Figure \ref{fig:ad_allocation} shows a deployment view of the Beer
+Web Store where the web server and the database is located on two
+different servers. Clients can connect to the Beer Web Server
+using a http web browser. The web server and database server
+are connected through a local area network.
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure=figures/ad_allocation,scale=0.6 }}
+\caption{Deployment view of the Beer Web Store}
+\label{fig:ad_allocation}
+\end{figure}
+
+\fixme{Skal vi tilføje applikationen til Web serveren? Den indeholder
+ikke nogen aktive komponenter.}
 

### 2008-03-17 SAiP/Module1/H3/doc/softwarearchitecture.tex
Grammar fix

@@
 A good approach is the "3 + 1" structure, which uses three viewpoints; Module viewpoint, Component \& Connector viewpoint and Allocation viewpoint.
-These three viewpoint is expressed in UML using Package diagrams, Object and sequence diagrams, and deployment diagrams respectively.
+These three viewpoints are expressed in UML using Package diagrams, Object and sequence diagrams, and deployment diagrams respectively.
 These diagrams should in the first iteration only look at relations and interactions between the main components (the layers).
@@
 
-\fixme{Answer question 1.5}
+There are generally two approaches to creating an architecture, as defined by Bass et.
+al.
+and that is to define it in advance, and only define the overall structure and then let it grow from the following implementations.
+
+There are also two reasons for creating an architecture; in order to create a product that exhibits some of the quality attributes defined in the architecture (modifiability, usability, ...) and in order to reuse the architecture (or part of the architecture) as a product line for creating other products.
+This latter reason is described in section yy.
+
+Creating an architecture up front allows for the definition of all properties of the architecture; the quality attributes and their individual priorities, the individual elements/components and how they interconnect both at runtime and statically, etc.
+This has the benifit that when the design and implementation phase is begun it may easily be broken down into parallel activities and integrated later, as everyone has a clear picture of what is important and what the interfaces are.
+
+Unfortunately defining the architecture fully up front only makes sense if there is enough information in advance to create the architecture form.
+Very often there are "dark horses", things that do not become apparent until the design and implementation phase.
+If it is not believed that sufficient information exist to fully define the architecture, then doing so may not only be a waste, it may lead to the very dangerous loose-loose principle, where e.g.
+an overhead is introduced to service modifiability, only to be hacked in order to service a performance issue that was not apparent at the architectual design pahse.
+This means that part of the overhead is preserved (in memory-footprint if nowhere else), without preserving any of the modifiability advantages.
+
+Another requirement to defining the architecture fully up front has to do with the organization.
+It may be good and dandy to create a fully defined architecture with focus on parallel implementation, if there is only one in-house developer to implement the product.
+Based on this the Beer Web Store architect must have decided that there are sufficient information to fully define the architecture and also that there is a reason (parallel implementation, out-sourceing of the design/implementation, etc.).
+
+Asuming that the architect in question has followed the course Software Architecture in Practice, he has obviously looked at the architectual properties, the stake holders, the organization and the requirement and deemed that there is sufficient grounds to fully specify the architecture in advance, with the risks and possible gains that it has.

### 2008-03-17 SAiP/Module1/H3/doc/qualityattributes.tex
Updating with Availability tactics

@@
 
-As it may be seen, the scenario(s) focus primarily on \emph{Availability} (how rare should it be that a user experience non-resonsiveness), \emph{Usability} (how easy should it be for the user to cancel a transaction halfway through and how easy should it be to pick up a previous transaction), \emph{Performance} (how big a delay can the user expect when performing an operation like a lookup in the database), \emph{Security} (how to avoid someone using stolen credit cards and also prevent a "hacker" from stealing the credit card information either durig the transaction or from the Beer Web Store's database.
+As it may be seen, the scenario(s) focus primarily on \emph{Availability} (how rare should it be that a user experience non-resonsiveness and how many simultanious users can the system support), \emph{Usability} (how easy should it be for the user to cancel a transaction halfway through and how easy should it be to pick up a previous transaction), \emph{Performance} (how big a delay can the user expect when performing an operation like a lookup in the database), \emph{Security} (how to avoid someone using stolen credit cards and also prevent a "hacker" from stealing the credit card information either durig the transaction or from the Beer Web Store's database.
 
@@
 \end{question}
+
+As we have only defined the Availability and Performance requirements, only these will be dealt with here, yet naturally requirements concerning the other quality attributes must also be handled.
+
+\subsubsection{Availability}
+Availability has two main concerns; finding out that a resource is unavailable and if it is, then restart or switch to another resource.
+
+Finding out if a resource is down may be done using one of the fault detection tactice; ping/echo, heartbeat or exception.
+Due to the loose coupling between the client (browser) and the server, the exception solution is not benificial.
+Since it may be assumed that the supervising process (remotely deployed) (the process monitoring if the server is down) is stable, then a heartbeat is sufficient to ensure availability, meaning that the web-server will issue a heart-beat every e.g.
+1 second to a maintance process on the same or another machine, and if one (or perhaps more) heartbeats is missed then the maintanance service can act.
+
+The proper response to a faulty server may be many things, as defined by the tactics group \emph{fault recovery}.
+As the requirement for availability stipulated a three stage fault recovery, the server or servers (as it may be required to have multiple serves to do load-balancing, depending on how many users must be supported) can service the 5 second requirement, and tehn if an error occur (as detected by the heartbeat) then restart or swap to another server/database in 30 seconds.
+As 30 seconds is insufficient to bott a webserver and a database, the Spare tactic would be insufficient.
+At the same tiem having Active redundancy would be over-kill, as a this kind of immidiat switch-over is not required.
+For these reasons Passive Redundancy is sufficient.
+
+As it is an internet system, it will also be sufficient to simply restart the session, and no \emph{Shadow}, \emph{State resynchronization} or \emph{Rollback} is needed in the main web server.
+The user may experience beeing thrown back to the main page, yet this is deemed acceptable, under availability and performance, yet it may not be in Usability.
+During a payment transaction, however, it must be ensured that the purchase is either completed or not-completed - halfway completed is not acceptable.
+for this it will be benificial to use the \emph{Rollback} tactic on the \emph{Basket} component.
+
+Finally a web-server may be deployed multiple times on teh same machine, allowing for the recovery or prevention of the lesser faults, by simply re-spawning the process that has become non-responsive, utilizing the \emph{Process Monitor} tactic.
+This is e.g.
+done by the Apache-server.
+
+As it is not only the web server which may become non-responsive, it is also the database, it is also required to monitor the database.
+However, as a non-responsive data-base may be discovered by a non-responsive server, it is possible to use the ping-echo tactic with the database, and only perform the ping-echo if the server becomes non-responsive, to find out how much must be restarted (only the server or both the server and the database).
+Depending on the binding between teh database and the server, it may also be possibel to use the exceptions tactic between teh server and the database, so the server itself can restart the database, if an exception occur in a call to the database.
+
+\subsubsection{Performance}
 

### 2008-03-17 trunk/SAiP/Module1/H3/doc/qualityattributes.tex
Adding performance quality attribute

@@
 
+As we have only defined the Availability and Performance requirements, only these will be dealt with here, yet naturally requirements concerning the other quality attributes must also be handled.
+
+\subsubsection{Availability}
+Availability has two main concerns; finding out that a resource is unavailable and if it is, then restart or switch to another resource.
+
+Finding out if a resource is down may be done using one of the fault detection tactice; ping/echo, heartbeat or exception.
+Due to the loose coupling between the client (browser) and the server, the exception solution is not benificial.
+Since it may be assumed that the supervising process (remotely deployed) (the process monitoring if the server is down) is stable, then a heartbeat is sufficient to ensure availability, meaning that the web-server will issue a heart-beat every e.g.
+1 second to a maintance process on the same or another machine, and if one (or perhaps more) heartbeats is missed then the maintanance service can act.
+
+The proper response to a faulty server may be many things, as defined by the tactics group \emph{fault recovery}.
+As the requirement for availability stipulated a three stage fault recovery, the server or servers (as it may be required to have multiple serves to do load-balancing, depending on how many users must be supported) can service the 5 second requirement, and tehn if an error occur (as detected by the heartbeat) then restart or swap to another server/database in 30 seconds.
+As 30 seconds is insufficient to bott a webserver and a database, the Spare tactic would be insufficient.
+At the same tiem having Active redundancy would be over-kill, as a this kind of immidiat switch-over is not required.
+For these reasons Passive Redundancy is sufficient.
+
+As it is an internet system, it will also be sufficient to simply restart the session, and no \emph{Shadow}, \emph{State resynchronization} or \emph{Rollback} is needed in the main web server.
+The user may experience beeing thrown back to the main page, yet this is deemed acceptable, under availability and performance, yet it may not be in Usability.
+During a payment transaction, however, it must be ensured that the purchase is either completed or not-completed - halfway completed is not acceptable.
+for this it will be benificial to use the \emph{Rollback} tactic on the \emph{Basket} component.
+
+Finally a web-server may be deployed multiple times on teh same machine, allowing for the recovery or prevention of the lesser faults, by simply re-spawning the process that has become non-responsive, utilizing the \emph{Process Monitor} tactic.
+This is e.g.
+done by the Apache-server.
+
+As it is not only the web server which may become non-responsive, it is also the database, it is also required to monitor the database.
+However, as a non-responsive data-base may be discovered by a non-responsive server, it is possible to use the ping-echo tactic with the database, and only perform the ping-echo if the server becomes non-responsive, to find out how much must be restarted (only the server or both the server and the database).
+Depending on the binding between teh database and the server, it may also be possibel to use the exceptions tactic between teh server and the database, so the server itself can restart the database, if an exception occur in a call to the database.
+
+\subsubsection{Performance}
+
+As we have all experience the irritation of a slow web-site, and as the competition on the internet is growing, it is important to achieve propper performance.
+
+Due to the lowing price of hardware and the increasing price of developement time, it is often cheapest and certainly simper to use the tactic \emph{Increase available resourcez}; introduce load balancing with multiple parallel servers thereby increase performance during heavy-load situations, ensure that the serves have sufficient network bandwidth, RAM and CPU strength to handle many simultanious database look-ups - here the tactic of \emph{Maintain multiple copies of data or computations} may be used to speed up database-look-up by caching the most popular searches, or cach the entire result in preperation for a search refinement.
+If the database is remotely located the caching of results becomes even more important.
+
+There are many other tactics which may be employed, yet given the nature to the application, it is not believed that the biggest resource consumers are under developement control, for building the web-server from scratch to support better concurrency or using non-standard performance tweaking algorithms for the database-lookup, will most likely become inconcivably expensive compare to simply increasing the computational resources.
+
 \fixme{Answer question 3.3}

### 2008-03-17 trunk/SAiP/Module1/H3/doc/qualityattributes.tex
Adding some text about SOA

@@
 
-\fixme{Answer question 3.3}
-
 \subsection{Quality Attributes - Question 4}
@@
 
-\fixme{Answer question 3.4}
+As SOA is all about standardised interfaces supplying a functionality which may be used by many different products, the areas where SOA makes sens is naturally between teh components.
+
+Instead of creating a web-server which access a database through a specific facade, one may argue whether the desired functionality may not be generalised.
+"Look-up in database".
+This may easily be hidden behind a standardised service, e.g.
+a web-service, meaning that anyone on the WWW who supplies a batabase service may be used to host the beer database, and if one becomes dissatisfied with the database provider, it is transparent to simply switch to another.
+
+Taking this one step further, perhaps the database service may be extended with a data-base type, meaning that the Beer Web Store's domain model is exposed in a service, which would allow for not only displaying ones own brands of beer, but through a discovery service, also beer-brands by other brewers on the net who implement this service.
+This will also point the way for a more standardised lookk-and-feel when shopping for beer.
+
+Naturally just because a web-service (or other service) is used it does not mean that it has to be public, jet it does have some advabtages.
+
+Finally the WS-Security web-service may be used to handle payment, and also possibly, if the database or Beer Web Store Domain model is exposed through a web-service, it may also be benificial that not everyone may have the ability to update the domain model or database, so here the WS-Security service may also come in handy.
 

### 2008-03-17 trunk/SAiP/Module1/H3/doc/qualityattributes.tex
Adding section about SOA penalties

@@
 
+Naturally this generality do not necesarrliy come free.
+A look-up in a database behind a service, especially a web-service, is far slower than a direct look-up in a local database, thereby influencing Performance poorly.
+On the other hand the use of web-services may increase Availability, as the functionality is deployed on multiple servers, thereby redusing the chance of the system becomming fully non-responsive.
+Naturally the critical path must be evaluated, as if a link on the critical path chain become non-responsive then the entire system becomes non-responsive.
+
 \subsection{Quality Attributes - Question 5}

### 2008-03-17 trunk/SAiP/Module1/H3/doc/architecturaldescription.tex
Architectural description updated.

@@
 
-\fixme{Skal vi tilføje applikationen til Web serveren? Den indeholder
-ikke nogen aktive komponenter.}
+The application code (Beer Web) and the database facade are
+both deployed on the web server.
 
@@
 They are fast to draw and can be shown to non-technical stakeholders.
-It is kind of like simple signlanguage - everyone understands nodding and shaking ones head or indicating a direction as long as it is for general information.
+It is kind of like simple sign-language - everyone understands nodding and shaking ones head or indicating a direction as long as it is for general information.
 
@@
 Working as an developer/tester/engineer means that everything must be quantifiable in some way, and box-and-line drawings do not offer this unambigious quantifiability.
-For this a clearly defined modeling language, without ambiguaty as to the meaning of the model, is needed.
-An example of such a language is Unified Modelling Language(UML), which would allow the box-and-line drawing to be represented as one or more diagrams solving all the ambiguaty of the box-and-line drawing with only a single requirement; one must know the language, which is in deed one of the reasons while box-and-line drawings are still widely used - there is no single unambigious modelling language know by all stakeholders, and when that is the case it may be more confusing then benificial to spend time making sure the diagram is clear and unambigious, if that clarity is derived from symbolic notation that the readers (stakeholders) do not understand.
+For this a clearly defined modeling language, without ambiguity as to the meaning of the model, is needed.
+An example of such a language is Unified Modelling Language(UML), which would allow the box-and-line drawing to be represented as one or more diagrams solving all the ambiguaty of the box-and-line drawing with only a single requirement; one must know the language, which is indeed one of the reasons while box-and-line drawings are still widely used - there is no single unambigious modelling language known by all stakeholders, and when that is the case it may be more confusing then benificial to spend time making sure the diagram is clear and unambigious, if that clarity is derived from symbolic notation that the readers (stakeholders) do not understand.
 
@@
 
-As SOA consists of a collection of (preferrably stateless) well-defined services what needs to be desctibed is, which existing (external) services are needed and which new services has to be created.
+As SOA consists of a collection of (preferrably stateless) well-defined services what needs to be described is, which existing (external) services are needed and which new services has to be created.
 This may be done using the MacKenzie Topic maps, which describe the static dependencies between the services.
 
-The static description is naturally not sufficient, and two other views may be presented to show the Dynamic interrelations; Visibility and Interaction.
-
-\fixme{Finish question 2.4}
+The static description is naturally not sufficient, and two other views may be presented to show the Dynamic interrelations for the services that are provided; Visibility and Interaction.
+The visibility view describes the awareness, willingness and reachability of a service and the interaction view describes how the service can be used (behaviour model) and the structure and semantics of the information that is provided (information model).

### 2008-03-17 SAiP/Module1/H3/doc/qualityattributes.tex
Finishing Quality attributes

@@
 
-\fixme{Answer question 3.3}
+As we have all experience the irritation of a slow web-site, and as the competition on the internet is growing, it is important to achieve propper performance.
+
+Due to the lowing price of hardware and the increasing price of developement time, it is often cheapest and certainly simper to use the tactic \emph{Increase available resourcez}; introduce load balancing with multiple parallel servers thereby increase performance during heavy-load situations, ensure that the serves have sufficient network bandwidth, RAM and CPU strength to handle many simultanious database look-ups - here the tactic of \emph{Maintain multiple copies of data or computations} may be used to speed up database-look-up by caching the most popular searches, or cach the entire result in preperation for a search refinement.
+If the database is remotely located the caching of results becomes even more important.
+
+There are many other tactics which may be employed, yet given the nature to the application, it is not believed that the biggest resource consumers are under developement control, for building the web-server from scratch to support better concurrency or using non-standard performance tweaking algorithms for the database-lookup, will most likely become inconcivably expensive compare to simply increasing the computational resources.
 
@@
 
-\fixme{Answer question 3.4}
+As SOA is all about standardised interfaces supplying a functionality which may be used by many different products, the areas where SOA makes sens is naturally between teh components.
+
+Instead of creating a web-server which access a database through a specific facade, one may argue whether the desired functionality may not be generalised.
+"Look-up in database".
+This may easily be hidden behind a standardised service, e.g.
+a web-service, meaning that anyone on the WWW who supplies a batabase service may be used to host the beer database, and if one becomes dissatisfied with the database provider, it is transparent to simply switch to another.
+
+Taking this one step further, perhaps the database service may be extended with a data-base type, meaning that the Beer Web Store's domain model is exposed in a service, which would allow for not only displaying ones own brands of beer, but through a discovery service, also beer-brands by other brewers on the net who implement this service.
+This will also point the way for a more standardised lookk-and-feel when shopping for beer.
+
+Naturally just because a web-service (or other service) is used it does not mean that it has to be public, jet it does have some advabtages.
+
+Finally the WS-Security web-service may be used to handle payment, and also possibly, if the database or Beer Web Store Domain model is exposed through a web-service, it may also be benificial that not everyone may have the ability to update the domain model or database, so here the WS-Security service may also come in handy.
+
+Naturally this generality do not necesarrliy come free.
+A look-up in a database behind a service, especially a web-service, is far slower than a direct look-up in a local database, thereby influencing Performance poorly.
+On the other hand the use of web-services may increase Availability, as the functionality is deployed on multiple servers, thereby redusing the chance of the system becomming fully non-responsive.
+Naturally the critical path must be evaluated, as if a link on the critical path chain become non-responsive then the entire system becomes non-responsive.
 
@@
 
-\fixme{Answer question 3.5}
+The quality attributes are vital to the software architecture, as they not only define the requirements to the software architecture (along with the scenarios), but also describe their relative importance.
+The quality attributes therefore allow the architecture the be evaluated in terms of how well it supports or enables a system's desired external (e.g.
+availability and performance) and internal behaviour (e.g.
+testability and modifiability).
+\cite{h_sarjoughian2002}

### 2008-03-17 SAiP/Module1/H3/doc/architecturaldesign.tex
Adding further tactic discussion

@@
 
-Discuss:
-\begin{itemize}
-\item Give examples of performance tactics respectively availability
+\subsection{Architectural Design - Question 1}
+
+\begin{question}
+Give examples of performance tactics respectively availability
 tactics that may be applied to the scenario. What are the consequences
 of applying these tactics?
-\item The software architect decides to use a layered architectural
+\end{question}
+
+This topic has already been breached in Section \ref{sec:qa_q3}, so this section may be seen as a continuation of that.
+
+Asuming that the 10 seconds is an increase in the requirements to the performance, and that the payment service is one that we can actually control (if it is a web-service supplied by a third party (e.g.
+PBS) then we may have no influence on its performance).
+
+If this is the case then the performance tactics which may be employed here could be \emph{introduce concurrency}, as some of the validations may be done in parallel, e.g.
+validate the credit card, prepare shipping information, access logistics system.
+Naturally this will require a roll-back system if it turns out that the payment did not go through.
+This approach is acceptable as the performance requirement is only valid in the instance where the payment is valid - how long it takes if the user enteres invalid information is less important.
+
+As there is no clarification of the response measure, it it could be asumed that the requirement must be met 100\% of the time, yet naturally this is not possible.
+It may however be requried that multiple instance of the payment service or validating server are running in order to servcie multiple purchases simultanious, as required by the \emph{Availability}.
+
+If the payment is considered of higher priority (\emph{Scheduling Policy} tactic), so that other functionality (e.g.
+browsing) may be handled slower while the payment is performed, then handling the scheduling of these tasks may also help to meet the 10 second requirement.
+
+Naturally implementing these tactics do not come for free; increasing hardware performance and bandwidth cost money, controlling the scheduling adds complexity and thereby reduce readability and modifiability.
+
+The second scenario focuses mainly on Availability, and actually goes against some of the previos scenarios by stating that the payment service is allowed to be "down" for a longer period of time, pending an administrator.
+This approach has one big benefit; it is simple.
+No hot- or cold-swapping of payment services, and the tactic \emph{Spare} may be employed.
+
+If the failure is part of an attack, it is easier to prevent by simply degrading the system and waiting for an administrator (may increase \emph{Security}).
+
+The second scenario does include requirements to fault detection, yet these are already described in Section \ref{sec:qa_q3}, and the price of fault detection is also increased cost of hardware and readability/modifiability.
+
+\subsection{Architectural Design - Question 2}
+
+\begin{question}
+The software architect decides to use a layered architectural
 style for the system. Discuss quality implications of this choice
-\item Outline major elements of a possible component and connector
+\end{question}
+
+\fixme{Answer question 4.2}
+
+\subsection{Architectural Design - Question 3}
+
+\begin{question}
+Outline major elements of a possible component and connector
 structure of the Beer Web Store system
-\end{itemize}
+\end{question}
+
+\fixme{Answer question 4.3}

### 2008-03-17 trunk/SAiP/Module1/H3/doc/architecturaldesign.tex
Adding text about Layering

@@
 
-\fixme{Answer question 4.1}
+This topic has already been breached in Section \ref{sec:qa_q3}, so this section may be seen as a continuation of that.
+
+Asuming that the 10 seconds is an increase in the requirements to the performance, and that the payment service is one that we can actually control (if it is a web-service supplied by a third party (e.g.
+PBS) then we may have no influence on its performance).
+
+If this is the case then the performance tactics which may be employed here could be \emph{introduce concurrency}, as some of the validations may be done in parallel, e.g.
+validate the credit card, prepare shipping information, access logistics system.
+Naturally this will require a roll-back system if it turns out that the payment did not go through.
+This approach is acceptable as the performance requirement is only valid in the instance where the payment is valid - how long it takes if the user enteres invalid information is less important.
+
+As there is no clarification of the response measure, it it could be asumed that the requirement must be met 100\% of the time, yet naturally this is not possible.
+It may however be requried that multiple instance of the payment service or validating server are running in order to servcie multiple purchases simultanious, as required by the \emph{Availability}.
+
+If the payment is considered of higher priority (\emph{Scheduling Policy} tactic), so that other functionality (e.g.
+browsing) may be handled slower while the payment is performed, then handling the scheduling of these tasks may also help to meet the 10 second requirement.
+
+Naturally implementing these tactics do not come for free; increasing hardware performance and bandwidth cost money, controlling the scheduling adds complexity and thereby reduce readability and modifiability.
+
+The second scenario focuses mainly on Availability, and actually goes against some of the previos scenarios by stating that the payment service is allowed to be "down" for a longer period of time, pending an administrator.
+This approach has one big benefit; it is simple.
+No hot- or cold-swapping of payment services, and the tactic \emph{Spare} may be employed.
+
+If the failure is part of an attack, it is easier to prevent by simply degrading the system and waiting for an administrator (may increase \emph{Security}).
+
+The second scenario does include requirements to fault detection, yet these are already described in Section \ref{sec:qa_q3}, and the price of fault detection is also increased cost of hardware and readability/modifiability.
 
@@
 
-\fixme{Answer question 4.2}
+Layerd styles come in two variants; the strictly layerd and the layered.
+The strictly layered means that the components in a layer may only access otyher components in the same layer or the layer directly below it.
+The Layered may access components in the same layer or any layer beneath it.
+
+When a system is Strictly Layered it gives rise to high \emph{Portability} (a for of \emph{Modifiability}), as it is possible to abstract all dependencies of a given layer by simply abstracting the layer directly beneath it, and if anything in a layer changes then only the layer directly above it must be updated.
+
+This also enables simpler \emph{Testability}, as the introduction of test doubles (fake objects, stubs, spys and/or mocks) must only be performed on the in a single layer; the one directly below the one under test.
+
+It may however be difficult to adher to a Strictly Layered approach, as some functionality may be required on all levels, e.g.
+the Java Collections, and very often vertical "layers" are introduced to support this all-layer access, which reduces \emph{Modifiability} if the change is in the veritcal "layer".
 

### 2008-03-17 SAiP/Module1/H3/doc/qualityattributes.tex
Adding lable
Quality attribute scenario refinement tables added.

@@
 
-The scenarios may be further clarified in order to include the aspects of availability and performance, as shown below.
+The scenarios may be further clarified in order to include the aspects of availability and performance, as shown in tables \ref{tab:availability} and \ref{tab:performance} below:
 
-\begin{quote}
-The customer browses the beer catalogue for descriptions and reviews and receives the result within 5 seconds 99\% of the time, 30 seconds 0,9\% of the time and more than 30 seconds 0,1\% of the time.
+\begin{table}[H]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The customer browses the beer catalogue for descriptions and reviews and receives a timely response \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Availability, (Performance) \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The user \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Browsing the beer catalogue \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The system \\
+  \cline{2-3}
+  & {\bfseries Environment:} & During normal operation \\
+  \cline{2-3}
+  & {\bfseries Response:} & Web page presented to user \\
+  \cline{2-3}
+  & {\bfseries Response Measure:} & The result is presented within 5 seconds 99\% of the time, 30 seconds 0,9\% of the time and more than 30 seconds 0,1\% of the time \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} & The response measure is also dependent on the speed of the users internet connection and computer hardware. \\
+  \hline
+\end{tabular}
+\caption{Availability scenario for the Beer Web Store}
+\label{tab:availability}
+\end{center}
+\end{table}
 
-The customer selects appropriate beers for purchase and these are added to his shopping basket. When the customer checks out, payment is validated through the payment service within 30 seconds.
-\end{quote}
+\begin{table}[H]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & When the customer checks out, payment is validated through the payment service within 30 seconds. \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Performance, (Availability) \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The user \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & The customer checks out \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The system \\
+  \cline{2-3}
+  & {\bfseries Environment:} & During normal operation \\
+  \cline{2-3}
+  & {\bfseries Response:} & Result of payment validation if presented to the user \\
+  \cline{2-3}
+  & {\bfseries Response Measure:} & Payment is validated through the payment service within 30 seconds \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} & The response measure is also dependent on the performance of the payment service. \\
+  \hline
+\end{tabular}
+\caption{Performance scenario for the Beer Web Store}
+\label{tab:performance}
+\end{center}
+\end{table}
 
-\fixme{Answer question 2.2}
+\fixme{Answer question 3.2}
 
 \subsection{Quality Attributes - Question 3}
+\label{sec:qa_q3}
 

### 2008-03-17 SAiP/Module1/H3/doc/architecturaldesign.tex
Adding text to C&C - diagrams missing

@@
 
-\fixme{Answer question 4.3}
+There are four major structures in Component-and-Connector:
+
+\begin{itemize}
+  \item Process or communicating processes.
+	\begin{itemize}
+	  \item This structure shows the execution path through a system, and is used to give the engineer an understand about the dynamic behaviour of the system.
+These structures may help ingeneers increase Performance and Availability.
+	\end{itemize}
+  \item Concurrency.
+	\begin{itemize}
+	  \item This structure is used to see the parallelism in the system, or part of the system, and when a system, or part of a system, is build around this structure it promotes performance. 
+	\end{itemize}
+  \item Shared data or Repository
+	\begin{itemize}
+	  \item This is a very good structure to use to illustarte how the data is processed and consumed to/from the database (Database Facade).
+When a system, or part of a system, is build around this structure it promotes performance and data entegrity.
+	\end{itemize}
+  \item Client-server
+	\begin{itemize}
+  		\item This is a very good structure to see the protocols used between the clients and the server, and the messages they communicate.
+When a system, or part of a system, is build around this structure it promotes seperation of concern (i.e.
+Modifiability) and allows for Availability tactics like Load balancing.
+	\end{itemize}
+\end{itemize}
+
+\fixme{Insert diagrams for C&C}

### 2008-03-17 SAiP/Module1/H3/doc/architecturaldesign.tex
Adding last part about C&C structures

@@
 	  \item This structure shows the execution path through a system, and is used to give the engineer an understand about the dynamic behaviour of the system.
-These structures may help ingeneers increase Performance and Availability.
+These structures may help engineers increase Performance and Availability.
 	\end{itemize}
@@
 	\begin{itemize}
-	  \item This is a very good structure to use to illustarte how the data is processed and consumed to/from the database (Database Facade).
+	  \item This is a very good structure to use to illustarte how the data is processed and consumed to/from the database (Server to/from Database Facade).
 When a system, or part of a system, is build around this structure it promotes performance and data entegrity.
@@
 	\begin{itemize}
-  		\item This is a very good structure to see the protocols used between the clients and the server, and the messages they communicate.
+  		\item This is a very good structure to see the protocols used between the Clients and the Server, and the messages they communicate.
 When a system, or part of a system, is build around this structure it promotes seperation of concern (i.e.
@@
 
-\fixme{Insert diagrams for C&C}
+In the Beer Web Store has two clear C\&C structures, the Shared data structure between the Server and the Database Facade and the Client-server between the Client and Server (see Figure \ref{fig:module_view}).
+
+The Process or Communicating processes is a more fine-grained structure, and an example of this may be seen in Figure \ref{fig:ad_sequence}.
+
+The Concurrency structure is not used in the overall elements, yet may be employed in a given element to achieve more performance.
+
+\fixme{Possible extention - insert diagrams}

### 2008-03-17 trunk/SAiP/Module1/H3/doc/productlines.tex
Adding Figure reference

@@
 
-Discuss:
-\begin{itemize}
-\item How will the choice of a product line approach affect the
+\subsection{Product Lines - Question 1}
+
+\begin{question}
+How will the choice of a product line approach affect the
 architecture of the beer web store?
-  \item Discuss the benefits and liabilities of using a product line
+\end{question}
+
+\fixme{Answer question 5.1}
+
+\subsection{Product Lines - Question 2}
+
+\begin{question}
+Discuss the benefits and liabilities of using a product line
 approach in relation to the case?
-  \item Relate Service-Oriented Architecture and Product Lines?
-\end{itemize}
+\end{question}
+
+Creating an architecture product line only makes sense if it is believed that it may be used in other products, and the number of products in the pipeline and how well their commonalities may be predicted, helps to determine which approach to use.
+Naturally the situation is not black and white, as there are many possibilities between the fully evolutionary architecture to the fully initially defined, yet for the purpose of debating advantages and disadvantages the simple black and white is simply easier.
+
+Defining the architecture fully up front means a big investment in time and money in order to create the arcitecture and the first product based on it, so if it is the only product made, then it will go way over budget both in time and money.
+The idea is that creating the next product will be significantly cheaper, because a very large part of the products functionality is covered by the architecture and very little (if any) redesign has to be done.
+As more and more products are created based on the architecture their individual budget savings will help pay for the initial investment, eventually leading to a profit compared to building the products from scratch each time.
+
+Letting the architecture evolve means that only the basics are defined up front, only causing a small initial investment, before building the first product.
+When the second product is to be created more of the architecture is defined based on the parts of the first product, which may be used in the second.
+Naturally the second product is faster to build than the first, due to reuse, yet not as fast as the second product when the entire architectur was defined in advance.
+As more products come along the amount of reuse grow, and would eventually reach the same as the one where everything was defined in advance.
+
+If the architecture is used to define a Product Line it becomes even more pronounced, and the above two approaches are defined as Heavyweight and Lightweight, respectively.
+A statistical analysis of these two approaches, compared to simply building the products from scratch, can be shown in Figure \ref{fig:product-line}.
+
+\clearpage
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure=figures/product-line,scale=1.0 }}
+\caption{Heavy-weight vs. Light-weight}
+\label{fig:product-line}
+\end{figure}
+
+Deciding which approach to use, is more than just a matter of choosing between Heavyweight and Lightweight (or fully defined or evolutionary) based on the number of expected products, it is also a matter of how well the common architecture of the future products can be predicted, for if the architecture needs to be changed from product to product then all the advantage is lost.
+
+Based on this the Beer Web Store architect must have decided that the products in the pipe-line and the Beer Web Store share a common architecture and that the future products are so well know that the architecture can be fully specified in advance.
+
+\subsection{Product Lines - Question 3}
+
+\begin{question}
+Relate Service-Oriented Architecture and Product Lines?
+\end{question}
+
+\fixme{Answer question 5.3}

### 2008-03-17 SAiP/Module1/H3/doc/qualityattributes.tex
Finishing question 2 (at least sort of :-( )

@@
 
-Defining an architectures requirements is not like defining the functional requirements; actually many software professionals believe that they are indeed orthogonal.
+Defining requirements for an architecture is not like defining the functional requirements; actually many software professionals believe that they are indeed orthogonal.
 This is supported by the fact that a set of functional requirements may be implemented by a number of different architectures.
@@
 
-Naturally in order for the scenarios to be able to be used to clarify the architecture it must be possible quantify them, and place them in different categories.
+Naturally in order for the scenarios to be able to be used to clarify the architecture it must be possible to quantify them, and place them in different categories.
 This requires that the above scenario(s) are extended with e.g.
-response time, how many times the user should be able to perform an operation before experiencing a fault (down-town).
+response time, how many times the user should be able to perform an operation before experiencing a fault (down-time).
 
@@
 
-\fixme{Insert table with quality attribute breakdown}
-
-As it may be seen, the scenario(s) focus primarily on \emph{Availability} (how rare should it be that a user experience non-resonsiveness and how many simultanious users can the system support), \emph{Usability} (how easy should it be for the user to cancel a transaction halfway through and how easy should it be to pick up a previous transaction), \emph{Performance} (how big a delay can the user expect when performing an operation like a lookup in the database), \emph{Security} (how to avoid someone using stolen credit cards and also prevent a "hacker" from stealing the credit card information either durig the transaction or from the Beer Web Store's database.
+As it may be seen, the scenario(s) focus primarily on \emph{Availability} (how rare should it be that a user experience non-resonsiveness and how many simultanious users can the system support), \emph{Usability} (how easy should it be for the user to cancel a transaction halfway through and how easy should it be to pick up a previous transaction), \emph{Performance} (how big a delay can the user expect when performing an operation like a lookup in the database), \emph{Security} (how to avoid someone using stolen credit cards and also prevent a "hacker" from stealing the credit card information either during the transaction or from the Beer Web Store's database.
 
@@
 
-The scenarios may be further clarified in order to include the aspects of availability and performance, as shown in tables \ref{tab:availability} and \ref{tab:performance} below:
+The following scenarios (table \ref{tab:availability} and \ref{tab:performance}) is a possible outcome of prioritization and clarification of quality attribute scenarios as described above:
 
@@
 
-\fixme{Answer question 3.2}
+In a real life system there will of course be many more scenarios to describe the architectural requirements.
 

### 2008-03-17 SAiP/Module1/H3/doc/architecturaldescription.tex
Adding label

--- 2008-03-06 SAiP/Module1/H3/doc/architecturaldescription.tex
@@
 
-As described in \ref{sec:arch_desc_q1} the figure shows parts
-of a module viewpoint (uses, layered), component \& connector
-viewpoint (client-server) and an allocation viewpoint
+As described in section \ref{sec:arch_desc_q1} the figure shows
+parts of a module viewpoint (uses, layered), component \&
+connector viewpoint (client-server) and an allocation viewpoint
 (deployment) which are the recommended viewpoints in
@@
 These viewpoints are all relevant to the Beer Web Store example.
-The Module view describes the internals of the webserver and how
+The Module viewpoint describes the internals of the webserver and how
 the functionality is decomposed into units. The Component \& connector
@@
 
-XXX: Udvælg views og lav figurer:
+In the following sections we will give examples on views for each
+of the three viewpoints.
 
-- Layered model (refer to figure 1) and decomposition of the server?
-- Sequence diagram
-- Deployment view
+\subsubsection{Module Viewpoint}
 
-\fixme{Answer question 2.2}
+Figure \ref{fig:ad_package} shows a package diagram of the Beer Web
+Store and a decomposition of the web package. The Web package
+contains the elements used for the web interface, 
+the Overview page, the basket and the domain model. The domain model
+depends on a Database package that contains the database facade and the
+database facade uses some database package library, e.g. MySQL.
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure=figures/ad_package,scale=0.8 }}
+\caption{Package diagram for the Beer Web Store}
+\label{fig:ad_package}
+\end{figure}
+
+\subsubsection{Component \& Connector Viewpoint}
+
+There are only two active components in the Beer Web Store, a web server,
+e.g. Apache, and a database server. An active objects diagram will
+not be included here.
+
+Figure \ref{fig:ad_sequence} shows a sequence diagram where a user searches
+for a an item ``tuborg'' and adds one of the search results to his basket.
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure=figures/ad_sequence,scale=0.6 }}
+\caption{Sequence diagram showing a search and adding of an item}
+\label{fig:ad_sequence}
+\end{figure}
+
+The domain model accesses the database through the database facade
+when needed. This can be expressed in another sequence diagram.
+Sequence diagrams showing the checkout including payment should
+also be made.
+
+\subsubsection{Allocation Viewpoint}
+
+Figure \ref{fig:ad_allocation} shows a deployment view of the Beer
+Web Store where the web server and the database is located on two
+different servers. Clients can connect to the Beer Web Server
+using a http web browser. The web server and database server
+are connected through a local area network.
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure=figures/ad_allocation,scale=0.6 }}
+\caption{Deployment view of the Beer Web Store}
+\label{fig:ad_allocation}
+\end{figure}
+
+The application code (Beer Web) and the database facade are
+both deployed on the web server.
 
@@
 
-\fixme{Answer question 2.3}
+The benifits of a box-and-line drawing is that it can be written and "understood" by almost anyone, and in the initial phases of the architectual design they may be fine.
+They are fast to draw and can be shown to non-technical stakeholders.
+It is kind of like simple sign-language - everyone understands nodding and shaking ones head or indicating a direction as long as it is for general information.
+
+Unfortunately the general information is where the advantages stop.
+For simple sign-language, it may quickly come to an argument if one wishes to know if the indication of "this way" means "a short way that way" or "a long way that way", and the same is true for box-and-line drawings.
+"How many clients does the system support?".
+"How strong is the binding indicated by the line - do all lines mean the same?" There is merely a line, so what does that line mean?
+
+Working as an developer/tester/engineer means that everything must be quantifiable in some way, and box-and-line drawings do not offer this unambigious quantifiability.
+For this a clearly defined modeling language, without ambiguity as to the meaning of the model, is needed.
+An example of such a language is Unified Modelling Language(UML), which would allow the box-and-line drawing to be represented as one or more diagrams solving all the ambiguaty of the box-and-line drawing with only a single requirement; one must know the language, which is indeed one of the reasons while box-and-line drawings are still widely used - there is no single unambigious modelling language known by all stakeholders, and when that is the case it may be more confusing then benificial to spend time making sure the diagram is clear and unambigious, if that clarity is derived from symbolic notation that the readers (stakeholders) do not understand.
 
 \subsection{Architectural Description - Question 4}
-
+\label{sec:ad_q4}
 \begin{question}
@@
 
-\fixme{Answer question 2.4}
+In a sense there are already aspects of SOA in the architecture, as the validation of the credit card information is most likely performed using a service on the internet - not that web services and SOA is equivalent, but they do complement each other nicely.
+
+As SOA consists of a collection of (preferrably stateless) well-defined services what needs to be described is, which existing (external) services are needed and which new services has to be created.
+This may be done using the MacKenzie Topic maps, which describe the static dependencies between the services.
+
+The static description is naturally not sufficient, and two other views may be presented to show the Dynamic interrelations for the services that are provided; Visibility and Interaction.
+The visibility view describes the awareness, willingness and reachability of a service and the interaction view describes how the service can be used (behaviour model) and the structure and semantics of the information that is provided (information model).

### 2008-03-17 trunk/SAiP/Module1/H3/doc/productlines.tex
Adding SOA vs. Product-line

@@
 
-Defining the architecture fully up front means a big investment in time and money in order to create the arcitecture and the first product based on it, so if it is the only product made, then it will go way over budget both in time and money.
-The idea is that creating the next product will be significantly cheaper, because a very large part of the products functionality is covered by the architecture and very little (if any) redesign has to be done.
-As more and more products are created based on the architecture their individual budget savings will help pay for the initial investment, eventually leading to a profit compared to building the products from scratch each time.
+Defining the product line architecture fully up front means a big investment in time and money in order to create the arcitecture and the first product based on it, so if it is the only product made, then it will go way over budget both in time and money.
+The idea is that creating the next product will be significantly cheaper, because a very large part of the products functionality is covered by the product line and very little (if any) redesign has to be done.
+As more and more products are created based on the product line their individual budget savings will help pay for the initial investment, eventually leading to a profit compared to building the products from scratch each time.
 
-Letting the architecture evolve means that only the basics are defined up front, only causing a small initial investment, before building the first product.
-When the second product is to be created more of the architecture is defined based on the parts of the first product, which may be used in the second.
-Naturally the second product is faster to build than the first, due to reuse, yet not as fast as the second product when the entire architectur was defined in advance.
+Letting the product line evolve means that only the basics are defined up front, only causing a small initial investment, before building the first product.
+When the second product is to be created more of the product line is defined based on the parts of the first product, which may be used in the second.
+Naturally the second product is faster to build than the first, due to reuse, yet not as fast as the second product when the entire product line was defined in advance.
 As more products come along the amount of reuse grow, and would eventually reach the same as the one where everything was defined in advance.
 
-If the architecture is used to define a Product Line it becomes even more pronounced, and the above two approaches are defined as Heavyweight and Lightweight, respectively.
+The above two approaches are defined as Heavyweight and Lightweight, respectively.
 A statistical analysis of these two approaches, compared to simply building the products from scratch, can be shown in Figure \ref{fig:product-line}.
@@
 
-\fixme{Answer question 5.3}
+As mentioned in Section \ref{sec:qa_q4} part of SOA is about generalization, just like product lines are.
+Taking the example given above, having the architecture support mobile phones (i.e.
+WAB) means that the server must be able to supply a scaled down result of a search as well as the fully blown.
+The server must thereby handle different types of clients.
+
+The requirement that other product's such as other stamps or soda-pop should be suppoprted, means that the server's domain model and database must support general elements as opposed to just beer elements.
+
+This generalization fits well with SOA, as it is not that clever to make a "Beer purchasing Service", but far better to create a "Purchasing service", which operate on elements.
+This will allow any company, who has products they wish to sell, to use this service.
+This is the same generalization requirement that apply to the Product line architecture, where the architecture must handle the purchase of an arbitrary product, be it stamps, soda-pop or beer.
+It is possible that the architecture requires that it is a quantifiable product, yet this is not an unlikely requirement for a service either.
+
+It may now seem like SOA is the answer to creatign a product line, yet this is not the case.
+It is very possible to create very specific services, like the "Beer look-up service", and this may be just fine for a Beer Web Store based on SOA, yet this is not a product line.
+The same architectual consideration must be done when creating the elemenets of a product line and define their interrelations, must also be done when defining the services and their interrelations.
+The only "structure" that SOA imposes by definition is the loose coupling between the user and the service.

### 2008-03-17 trunk/SAiP/Module1/H3/doc/softwarearchitecture.tex
Fixing spelling error and Figure references

@@
 
-A new Beer Web Store for specialty beers has the following main use
+A new Beer Web Store for speciality beers has the following main use
 case:
@@
 \subsubsection{Choose the module to decompose}
-As this is the beginning of the architectual design then the module is the entire system. In later iterations the sub-components (the layers, the components in the layers, etc.) will be decomposed.
+As this is the beginning of the architectural design then the module is the entire system. In later iterations the sub-components (the layers, the components in the layers, etc.) will be decomposed.
 
@@
 \begin{description}
-    \item[a. Choose the architectual drivers]
-  Here the quality attributes of the system would be defined by the stakeholders of the system and then a vote would be used to determine which of the quality attributes are architectual drivers.
+    \item[a. Choose the architectural drivers]
+  Here the quality attributes of the system would be defined by the stakeholders of the system and then a vote would be used to determine which of the quality attributes are architectural drivers.
 The quality attributes would be expressed as quality attribute scenarios.
-    \item[b. Choose an architectual pattern]
-  Based on the architectual drivers selected in 1, use the mapping in Bass et al.
-as well as other literature and experience to determine some tactics for achieving the architectual drivers for the system.
-Then select some architectual patterns to realize the tactics.
+    \item[b. Choose an architectural pattern]
+  Based on the architectural drivers selected in 1, use the mapping in Bass et al.
+as well as other literature and experience to determine some tactics for achieving the architectural drivers for the system.
+Then select some architectural patterns to realize the tactics.
     \item[c. Instantiate modules using multiple views]
@@
 \subsubsection{Repeat the steps above for every module that needs further decomposition}
-The first iteration only look at the overall architecture (the leation and communication between layers).
+The first iteration only look at the overall architecture (the location and communication between layers).
 The following iterations will decompose the individual layers and as needed the individual components in the layers.
@@
 \subsubsection{Relations}
-The figure xx shows some of the relations, namely the statical, yet it is also important to know the dynamical behaviour.
+The figure \ref{fig:module_view} shows some of the relations, namely the statical, yet it is also important to know the dynamical behaviour.
 These may be expressed by e.g.
-a sequence diagram, as shown in Figure yy.
+a sequence diagram, as shown in Figure \ref{fig:sequence}.
 In these it may be seen that the Client and Server has a week semantic relation through the http protocol, and the Server and Database has a strong syntactic relation from a compiled compatibility.
@@
 
-Perry and Wolf suggest seperating the model of the architecture into three components, as mentioned in the question.
+Perry and Wolf suggest separating the model of the architecture into three components, as mentioned in the question.
 The components relate in the sense that the \emph{elements} has a particular \emph{form} and them and their relations are chosen based on a \emph{rational}.
 
-There are three class of architectual elements defined.
+There are three class of architectural elements defined.
 \begin{itemize}
@@
 
-The architectual form describe the architectual properties and the relationships and their relative importance. This is important as some architectual properties
+The architectural form describe the architectural properties and the relationships and their relative importance. This is important as some architectural properties
 may be vital to the design and others may be gold-plating.
-These properties are used to put constraints on the elements and also define the relative importanse of these constraints.
+These properties are used to put constraints on the elements and also define the relative importance of these constraints.
 The relationships are similar to the properties except they are used to put constraints and relative priority on the relations between elements.
 
-A software architecture is a selection and description of architectual choices.
-These choices are, however, difficult to convay without including the reason for the choices made, for the elemenents and form not chosen and for choices not made due to lack of information or because they were deemed irrelevant.
+A software architecture is a selection and description of architectural choices.
+These choices are, however, difficult to convey without including the reason for the choices made, for the elements and form not chosen and for choices not made due to lack of information or because they were deemed irrelevant.
 
@@
 	    \item[HTTP] Protocol for connecting the server and client (loose coupling).
-	    \item[Database facade methods] Protocol for connecting the server to the database (medium-loose coupling).
+	    \item[Database fa\c ade methods] Protocol for connecting the server to the database (medium-loose coupling).
 	 \end{itemize}
@@
 \begin{itemize}
-    \item[Properties] The Server and the database, hidden behind the Facade, is the backbone of the Beer Web Store.
+    \item[Properties] The Server and the database, hidden behind the Fa\c ade, is the backbone of the Beer Web Store.
 If either is unavailable then the other is irrelevant.
-Multiple instances of both Server and Database may be running simultaniously in order to supply hot-swapping.
+Multiple instances of both Server and Database may be running simultaneously in order to supply hot-swapping.
 The importance of up-time (availability), security (payment), usability (transmission abilities), etc.
 must be defined here.
-    \item[Relationship] The Client may only see the server through its HTTP interface, and the Server may only access the database through the Database Facade.
-Other demands to the relationship (certain commands to the database may only be accessed when using https).
+    \item[Relationship] The Client may only see the server through its
+ HTTP interface, and the Server may only access the database through the Database Fa\c ade. Other demands to the relationship (certain commands to the database may only be accessed when using https).
 \end{itemize}
@@
 The properties are in fact only the externally visible to the given element, as those are the only properties that are relevant to the architecture.
-These details include things like the strength of the metal, which may be different for the hammer and the pipe, the price range, produktion time, etc.
-while things like the cheak-pad may be optional.
+These details include things like the strength of the metal, which may be different for the hammer and the pipe, the price range, production time, etc.
+while things like the cheek-pad may be optional.
 
@@
 al.
-is therefore very general and apart from the use of words like "software" and "computer", may apply to any architecture, be it a building, a produktion line for a rifle, or a embedded software system.
+is therefore very general and apart from the use of words like "software" and "computer", may apply to any architecture, be it a building, a production line for a rifle, or a embedded software system.
 
@@
 There are also two reasons for creating an architecture; in order to create a product that exhibits some of the quality attributes defined in the architecture (modifiability, usability, ...) and in order to reuse the architecture (or part of the architecture) as a product line for creating other products.
-This latter reason is described in section yy.
+This latter reason is described in section \ref{sec:pl_q2}.
 
 Creating an architecture up front allows for the definition of all properties of the architecture; the quality attributes and their individual priorities, the individual elements/components and how they interconnect both at runtime and statically, etc.
-This has the benifit that when the design and implementation phase is begun it may easily be broken down into parallel activities and integrated later, as everyone has a clear picture of what is important and what the interfaces are.
+This has the benefit that when the design and implementation phase is begun it may easily be broken down into parallel activities and integrated later, as everyone has a clear picture of what is important and what the interfaces are.
 
@@
 If it is not believed that sufficient information exist to fully define the architecture, then doing so may not only be a waste, it may lead to the very dangerous loose-loose principle, where e.g.
-an overhead is introduced to service modifiability, only to be hacked in order to service a performance issue that was not apparent at the architectual design pahse.
+an overhead is introduced to service modifiability, only to be hacked in order to service a performance issue that was not apparent at the architectural design phase.
 This means that part of the overhead is preserved (in memory-footprint if nowhere else), without preserving any of the modifiability advantages.
@@
 It may be good and dandy to create a fully defined architecture with focus on parallel implementation, if there is only one in-house developer to implement the product.
-Based on this the Beer Web Store architect must have decided that there are sufficient information to fully define the architecture and also that there is a reason (parallel implementation, out-sourceing of the design/implementation, etc.).
+Based on this the Beer Web Store architect must have decided that there are sufficient information to fully define the architecture and also that there is a reason (parallel implementation, out-sourcing of the design/implementation, etc.).
 
-Asuming that the architect in question has followed the course Software Architecture in Practice, he has obviously looked at the architectual properties, the stake holders, the organization and the requirement and deemed that there is sufficient grounds to fully specify the architecture in advance, with the risks and possible gains that it has.
+Assuming that the architect in question has followed the course Software Architecture in Practice, he has obviously looked at the architectural properties, the stake holders, the organization and the requirement and deemed that there is sufficient grounds to fully specify the architecture in advance, with the risks and possible gains that it has.

### 2008-03-17 SAiP/Module1/H3/doc/architecturaldescription.tex
Adding spell-check

@@
 
-The architect for a new Beer Web Store for specialty beers has drawn
+The architect for a new Beer Web Store for speciality beers has drawn
 the following figure of the architecture together with his customers.
@@
 
-Another perspective is that the users, the webserver and the
+Another perspective is that the users, the web server and the
 database are elements where the relations show that the users
-use the webserver and the webserver uses the database. This
+use the web server and the web server uses the database. This
 also shows the that the architecture is a
@@
 \emph{client-server structure}, where the users are clients
-and the webserver is a server. The webserver can also be
+and the web server is a server. The web server can also be
 seen as a client to the database server. This would make it
-possible to archieve better performance if several webservers
+possible to achieve better performance if several web servers
 are used to make load balancing of the user requests.
@@
 that the overview page, the basket and the domain model (elements)
-are allocated to (relations) on the webserver and the database is
+are allocated to (relations) on the web server and the database is
 allocated to its own server.
@@
 These viewpoints are all relevant to the Beer Web Store example.
-The Module viewpoint describes the internals of the webserver and how
+The Module viewpoint describes the internals of the web server and how
 the functionality is decomposed into units. The Component \& connector
@@
 the Overview page, the basket and the domain model. The domain model
-depends on a Database package that contains the database facade and the
-database facade uses some database package library, e.g. MySQL.
+depends on a Database package that contains the database fa\c ade and the
+database fa\c ade uses some database package library, e.g. MySQL.
 
@@
 
-The domain model accesses the database through the database facade
+The domain model accesses the database through the database fa\c ade
 when needed. This can be expressed in another sequence diagram.
@@
 
-The application code (Beer Web) and the database facade are
+The application code (Beer Web) and the database fa\c ade are
 both deployed on the web server.
@@
 
-The benifits of a box-and-line drawing is that it can be written and "understood" by almost anyone, and in the initial phases of the architectual design they may be fine.
+The benefits of a box-and-line drawing is that it can be written and "understood" by almost anyone, and in the initial phases of the architectural design they may be fine.
 They are fast to draw and can be shown to non-technical stakeholders.
@@
 
-Working as an developer/tester/engineer means that everything must be quantifiable in some way, and box-and-line drawings do not offer this unambigious quantifiability.
-For this a clearly defined modeling language, without ambiguity as to the meaning of the model, is needed.
-An example of such a language is Unified Modelling Language(UML), which would allow the box-and-line drawing to be represented as one or more diagrams solving all the ambiguaty of the box-and-line drawing with only a single requirement; one must know the language, which is indeed one of the reasons while box-and-line drawings are still widely used - there is no single unambigious modelling language known by all stakeholders, and when that is the case it may be more confusing then benificial to spend time making sure the diagram is clear and unambigious, if that clarity is derived from symbolic notation that the readers (stakeholders) do not understand.
+Working as an developer/tester/engineer means that everything must be quantifiable in some way, and box-and-line drawings do not offer this unambiguous quantifiable.
+For this a clearly defined modelling language, without ambiguity as to the meaning of the model, is needed.
+An example of such a language is Unified Modelling Language(UML), which would allow the box-and-line drawing to be represented as one or more diagrams solving all the ambiguity of the box-and-line drawing with only a single requirement; one must know the language, which is indeed one of the reasons while box-and-line drawings are still widely used - there is no single unambiguous modelling language known by all stakeholders, and when that is the case it may be more confusing then beneficial to spend time making sure the diagram is clear and unambiguous, if that clarity is derived from symbolic notation that the readers (stakeholders) do not understand.
 
@@
 
-As SOA consists of a collection of (preferrably stateless) well-defined services what needs to be described is, which existing (external) services are needed and which new services has to be created.
+As SOA consists of a collection of (preferably stateless) well-defined services what needs to be described is, which existing (external) services are needed and which new services has to be created.
 This may be done using the MacKenzie Topic maps, which describe the static dependencies between the services.

### 2008-03-17 trunk/SAiP/Module1/H3/doc/qualityattributes.tex
Updating with spell-check

--- 2008-03-17 trunk/SAiP/Module1/H3/doc/qualityattributes.tex
@@
 
-Consider the following main use case for a Beer Web Store for specialty
+Consider the following main use case for a Beer Web Store for speciality
 beers:
@@
 
-Naturally in order for the scenarios to be able to be used to clarify the architecture it must be possible quantify them, and place them in different categories.
+Naturally in order for the scenarios to be able to be used to clarify the architecture it must be possible to quantify them, and place them in different categories.
 This requires that the above scenario(s) are extended with e.g.
-response time, how many times the user should be able to perform an operation before experiencing a fault (down-town).
+response time, how many times the user should be able to perform an operation before experiencing a fault (down-time).
 
 Once the scenarios are clarified and voted on, then the top priority scenarios may be formalized and it may be determined which quality attribute(s) they involve.
-For examples sake lets asume that the above scenario was selected as top priority.
+For examples sake lets assume that the above scenario was selected as top priority.
 
-\fixme{Insert table with quality attribute breakdown}
-
-As it may be seen, the scenario(s) focus primarily on \emph{Availability} (how rare should it be that a user experience non-resonsiveness and how many simultanious users can the system support), \emph{Usability} (how easy should it be for the user to cancel a transaction halfway through and how easy should it be to pick up a previous transaction), \emph{Performance} (how big a delay can the user expect when performing an operation like a lookup in the database), \emph{Security} (how to avoid someone using stolen credit cards and also prevent a "hacker" from stealing the credit card information either durig the transaction or from the Beer Web Store's database.
+As it may be seen, the scenario(s) focus primarily on \emph{Availability} (how rare should it be that a user experience non-responsiveness and how many simultaneous users can the system support), \emph{Usability} (how easy should it be for the user to cancel a transaction halfway through and how easy should it be to pick up a previous transaction), \emph{Performance} (how big a delay can the user expect when performing an operation like a lookup in the database), \emph{Security} (how to avoid someone using stolen credit cards and also prevent a "hacker" from stealing the credit card information either during the transaction or from the Beer Web Store's database.
 
@@
 
-The scenarios may be further clarified in order to include the aspects of availability and performance, as shown below.
+The following scenarios (table \ref{tab:availability} and \ref{tab:performance}) is a possible outcome of prioritization and clarification of quality attribute scenarios as described above:
 
-\begin{quote}
-The customer browses the beer catalogue for descriptions and reviews and receives the result within 5 seconds 99\% of the time, 30 seconds 0,9\% of the time and more than 30 seconds 0,1\% of the time.
+\begin{table}[H]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & The customer browses the beer catalogue for descriptions and reviews and receives a timely response \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Availability, (Performance) \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The user \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & Browsing the beer catalogue \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The system \\
+  \cline{2-3}
+  & {\bfseries Environment:} & During normal operation \\
+  \cline{2-3}
+  & {\bfseries Response:} & Web page presented to user \\
+  \cline{2-3}
+  & {\bfseries Response Measure:} & The result is presented within 5 seconds 99\% of the time, 30 seconds 0,9\% of the time and more than 30 seconds 0,1\% of the time \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} & The response measure is also dependent on the speed of the users internet connection and computer hardware. \\
+  \hline
+\end{tabular}
+\caption{Availability scenario for the Beer Web Store}
+\label{tab:availability}
+\end{center}
+\end{table}
 
-The customer selects appropriate beers for purchase and these are added to his shopping basket. When the customer checks out, payment is validated through the payment service within 30 seconds.
-\end{quote}
+\begin{table}[H]
+\begin{center}
+\begin{tabular}{|p{0.3cm}|p{2.5cm}|p{8cm}|}
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Scenario(s):} & When the customer checks out, payment is validated through the payment service within 30 seconds. \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Relevant Quality Attributes:} & Performance, (Availability) \\
+  \hline
+  \multirow{6}{*}{\begin{sideways}{\bfseries Scenario Parts}\end{sideways}}
+  & {\bfseries Source:} & The user \\
+  \cline{2-3}
+  & {\bfseries Stimulus:} & The customer checks out \\
+  \cline{2-3}
+  & {\bfseries Artifact} & The system \\
+  \cline{2-3}
+  & {\bfseries Environment:} & During normal operation \\
+  \cline{2-3}
+  & {\bfseries Response:} & Result of payment validation if presented to the user \\
+  \cline{2-3}
+  & {\bfseries Response Measure:} & Payment is validated through the payment service within 30 seconds \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Questions:} & \\
+  \hline
+  \multicolumn{2}{|p{3cm}|}{\bfseries Issues:} & The response measure is also dependent on the performance of the payment service. \\
+  \hline
+\end{tabular}
+\caption{Performance scenario for the Beer Web Store}
+\label{tab:performance}
+\end{center}
+\end{table}
 
-\fixme{Answer question 2.2}
+In a real life system there will of course be many more scenarios to describe the architectural requirements.
 
 \subsection{Quality Attributes - Question 3}
+\label{sec:qa_q3}
 
@@
 
-Finding out if a resource is down may be done using one of the fault detection tactice; ping/echo, heartbeat or exception.
-Due to the loose coupling between the client (browser) and the server, the exception solution is not benificial.
+Finding out if a resource is down may be done using one of the fault detection tactics; ping/echo, heartbeat or exception.
+Due to the loose coupling between the client (browser) and the server, the exception solution is not beneficial.
 Since it may be assumed that the supervising process (remotely deployed) (the process monitoring if the server is down) is stable, then a heartbeat is sufficient to ensure availability, meaning that the web-server will issue a heart-beat every e.g.
-1 second to a maintance process on the same or another machine, and if one (or perhaps more) heartbeats is missed then the maintanance service can act.
+1 second to a maintenance process on the same or another machine, and if one (or perhaps more) heartbeats is missed then the maintenance service can act.
 
 The proper response to a faulty server may be many things, as defined by the tactics group \emph{fault recovery}.
-As the requirement for availability stipulated a three stage fault recovery, the server or servers (as it may be required to have multiple serves to do load-balancing, depending on how many users must be supported) can service the 5 second requirement, and tehn if an error occur (as detected by the heartbeat) then restart or swap to another server/database in 30 seconds.
-As 30 seconds is insufficient to bott a webserver and a database, the Spare tactic would be insufficient.
-At the same tiem having Active redundancy would be over-kill, as a this kind of immidiat switch-over is not required.
+As the requirement for availability stipulated a three stage fault recovery, the server or servers (as it may be required to have multiple serves to do load-balancing, depending on how many users must be supported) can service the 5 second requirement, and then if an error occur (as detected by the heartbeat) then restart or swap to another server/database in 30 seconds.
+As 30 seconds is insufficient to boot a web server and a database, the Spare tactic would be insufficient.
+At the same time having Active redundancy would be over-kill, as a this kind of immediate switch-over is not required.
 For these reasons Passive Redundancy is sufficient.
@@
 As it is an internet system, it will also be sufficient to simply restart the session, and no \emph{Shadow}, \emph{State resynchronization} or \emph{Rollback} is needed in the main web server.
-The user may experience beeing thrown back to the main page, yet this is deemed acceptable, under availability and performance, yet it may not be in Usability.
+The user may experience being thrown back to the main page, yet this is deemed acceptable, under availability and performance, yet it may not be in Usability.
 During a payment transaction, however, it must be ensured that the purchase is either completed or not-completed - halfway completed is not acceptable.
-for this it will be benificial to use the \emph{Rollback} tactic on the \emph{Basket} component.
+for this it will be beneficial to use the \emph{Rollback} tactic on the \emph{Basket} component.
 
@@
 As it is not only the web server which may become non-responsive, it is also the database, it is also required to monitor the database.
-However, as a non-responsive data-base may be discovered by a non-responsive server, it is possible to use the ping-echo tactic with the database, and only perform the ping-echo if the server becomes non-responsive, to find out how much must be restarted (only the server or both the server and the database).
-Depending on the binding between teh database and the server, it may also be possibel to use the exceptions tactic between teh server and the database, so the server itself can restart the database, if an exception occur in a call to the database.
+However, as a non-responsive data-base may be discovered by a non-responsive server, it is possible to use the ping-echo
+ tactic with the database, and only perform the ping-echo if the server becomes non-responsive, to find out how much must be restarted (only the server or both the server and the database). 
+Depending on the binding between the database and the server, it may also be possible to use the exceptions tactic between the server and the database, so the server itself can restart the database, if an exception occur in a call to the database.
 
@@
 
-As we have all experience the irritation of a slow web-site, and as the competition on the internet is growing, it is important to achieve propper performance.
+As we have all experience the irritation of a slow web-site, and as the competition on the internet is growing, it is important to achieve popper performance.
 
-Due to the lowing price of hardware and the increasing price of developement time, it is often cheapest and certainly simper to use the tactic \emph{Increase available resourcez}; introduce load balancing with multiple parallel servers thereby increase performance during heavy-load situations, ensure that the serves have sufficient network bandwidth, RAM and CPU strength to handle many simultanious database look-ups - here the tactic of \emph{Maintain multiple copies of data or computations} may be used to speed up database-look-up by caching the most popular searches, or cach the entire result in preperation for a search refinement.
+Due to the lowing price of hardware and the increasing price of development time, it is often cheapest and certainly simper to use the tactic \emph{Increase available resource}; introduce load balancing with multiple parallel servers thereby increase performance during heavy-load situations, ensure that the serves have sufficient network bandwidth, RAM and CPU strength to handle many simultanious database look-ups - here the tactic of \emph{Maintain multiple copies of data or computations} may be used to speed up database-look-up by caching the most popular searches, or cache the entire result in preparation for a search refinement.
 If the database is remotely located the caching of results becomes even more important.
 
-There are many other tactics which may be employed, yet given the nature to the application, it is not believed that the biggest resource consumers are under developement control, for building the web-server from scratch to support better concurrency or using non-standard performance tweaking algorithms for the database-lookup, will most likely become inconcivably expensive compare to simply increasing the computational resources.
-
-\fixme{Answer question 3.3}
+There are many other tactics which may be employed, yet given the nature to the application, it is not believed that the biggest resource consumers are under development control, for building the web-server from scratch to support better concurrency or using non-standard performance tweaking algorithms for the database-lookup, will most likely become inconceivably expensive compare to simply increasing the computational resources.
 
 \subsection{Quality Attributes - Question 4}
-
+\label{sec:qa_q4}
 \begin{question}
@@
 
-\fixme{Answer question 3.4}
+As SOA is all about standardised interfaces supplying a functionality which may be used by many different products, the areas where SOA makes sense is naturally between the components.
+
+Instead of creating a web-server which access a database through a specific fa\c ade, one may argue whether the desired functionality may not be generalised.
+"Look-up in database".
+This may easily be hidden behind a standardised service, e.g.
+a web-service, meaning that anyone on the WWW who supplies a database service may be used to host the beer database, and if one becomes dissatisfied with the database provider, it is transparent to simply switch to another.
+
+Taking this one step further, perhaps the database service may be extended with a data-base type, meaning that the Beer Web Store's domain model is exposed in a service, which would allow for not only displaying ones own brands of beer, but through a discovery service, also beer-brands by other brewers on the net who implement this service.
+This will also point the way for a more standardised look-and-feel when shopping for beer.
+
+Naturally just because a web-service (or other service) is used it does not mean that it has to be public, jet it does have some advantages.
+
+Finally the WS-Security web-service may be used to handle payment, and also possibly, if the database or Beer Web Store Domain model is exposed through a web-service, it may also be beneficial that not everyone may have the ability to update the domain model or database, so here the WS-Security service may also come in handy.
+
+Naturally this generality do not necessarily come free.
+A look-up in a database behind a service, especially a web-service, is far slower than a direct look-up in a local database, thereby influencing Performance poorly.
+On the other hand the use of web-services may increase Availability, as the functionality is deployed on multiple servers, thereby reducing the chance of the system becoming fully non-responsive.
+Naturally the critical path must be evaluated, as if a link on the critical path chain become non-responsive then the entire system becomes non-responsive.
 

### 2008-03-17 SAiP/Module1/H3/doc/architecturaldesign.tex
Adding spell-check

@@
 \section{Architectural Design -- Beer Web Store}
-Consider the following main use case for a Beer Web Store for specialty
+Consider the following main use case for a Beer Web Store for speciality
 beers:
@@
 
-Asuming that the 10 seconds is an increase in the requirements to the performance, and that the payment service is one that we can actually control (if it is a web-service supplied by a third party (e.g.
+Assuming that the 10 seconds is an increase in the requirements to the performance, and that the payment service is one that we can actually control (if it is a web-service supplied by a third party (e.g.
 PBS) then we may have no influence on its performance).
@@
 Naturally this will require a roll-back system if it turns out that the payment did not go through.
-This approach is acceptable as the performance requirement is only valid in the instance where the payment is valid - how long it takes if the user enteres invalid information is less important.
+This approach is acceptable as the performance requirement is only valid in the instance where the payment is valid - how long it takes if the user enters invalid information is less important.
 
-As there is no clarification of the response measure, it it could be asumed that the requirement must be met 100\% of the time, yet naturally this is not possible.
-It may however be requried that multiple instance of the payment service or validating server are running in order to servcie multiple purchases simultanious, as required by the \emph{Availability}.
+As there is no clarification of the response measure, it it could be assumed that the requirement must be met 100\% of the time, yet naturally this is not possible.
+It may however be required that multiple instance of the payment service or validating server are running in order to service multiple purchases simultaneous, as required by the \emph{Availability}.
 
@@
 
-Layerd styles come in two variants; the strictly layerd and the layered.
-The strictly layered means that the components in a layer may only access otyher components in the same layer or the layer directly below it.
+Layered styles come in two variants; the strictly layered and the layered.
+The strictly layered means that the components in a layer may only access other components in the same layer or the layer directly below it.
 The Layered may access components in the same layer or any layer beneath it.
@@
 
-It may however be difficult to adher to a Strictly Layered approach, as some functionality may be required on all levels, e.g.
-the Java Collections, and very often vertical "layers" are introduced to support this all-layer access, which reduces \emph{Modifiability} if the change is in the veritcal "layer".
+It may however be difficult to adhere to a Strictly Layered approach, as some functionality may be required on all levels, e.g.
+the Java Collections, and very often vertical "layers" are introduced to support this all-layer access, which reduces \emph{Modifiability} if the change is in the vertical "layer".
 
@@
 	\begin{itemize}
-	  \item This is a very good structure to use to illustarte how the data is processed and consumed to/from the database (Server to/from Database Facade).
-When a system, or part of a system, is build around this structure it promotes performance and data entegrity.
+	  \item This is a very good structure to use to illustrate how the data is processed and consumed to/from the database (Server to/from Database Fa\c ade).
+When a system, or part of a system, is build around this structure it promotes performance and data integrity.
 	\end{itemize}
@@
   		\item This is a very good structure to see the protocols used between the Clients and the Server, and the messages they communicate.
-When a system, or part of a system, is build around this structure it promotes seperation of concern (i.e.
+When a system, or part of a system, is build around this structure it promotes separation of concern (i.e.
 Modifiability) and allows for Availability tactics like Load balancing.
@@
 
-In the Beer Web Store has two clear C\&C structures, the Shared data structure between the Server and the Database Facade and the Client-server between the Client and Server (see Figure \ref{fig:module_view}).
+In the Beer Web Store has two clear C\&C structures, the Shared data structure between the Server and the Database Fa\c ade and the Client-server between the Client and Server (see Figure \ref{fig:module_view}).
 
@@
 
-\fixme{Possible extention - insert diagrams}
+\fixme{Possible extension - insert diagrams}

### 2008-03-17 trunk/SAiP/Module1/H3/doc/productlines.tex
Adding label
Adding spell-check

@@
 \section{Product Lines -- Beer Web Store}
-Consider a Beer Web Store for buying specialty beers. The software
+Consider a Beer Web Store for buying speciality beers. The software
 architect has designed a deployment structure of the system as shown
@@
 \subsection{Product Lines - Question 2}
-
+\label{sec:pl_q2}
 \begin{question}
@@
 
-Defining the product line architecture fully up front means a big investment in time and money in order to create the arcitecture and the first product based on it, so if it is the only product made, then it will go way over budget both in time and money.
+Defining the product line architecture fully up front means a big investment in time and money in order to create the architecture and the first product based on it, so if it is the only product made, then it will go way over budget both in time and money.
 The idea is that creating the next product will be significantly cheaper, because a very large part of the products functionality is covered by the product line and very little (if any) redesign has to be done.
@@
 
-The requirement that other product's such as other stamps or soda-pop should be suppoprted, means that the server's domain model and database must support general elements as opposed to just beer elements.
+The requirement that other product's such as other stamps or soda-pop should be supported, means that the server's domain model and database must support general elements as opposed to just beer elements.
 
@@
 
-It may now seem like SOA is the answer to creatign a product line, yet this is not the case.
+It may now seem like SOA is the answer to creating a product line, yet this is not the case.
 It is very possible to create very specific services, like the "Beer look-up service", and this may be just fine for a Beer Web Store based on SOA, yet this is not a product line.
-The same architectual consideration must be done when creating the elemenets of a product line and define their interrelations, must also be done when defining the services and their interrelations.
+The same architectural consideration must be done when creating the elements of a product line and define their interrelations, must also be done when defining the services and their interrelations.
 The only "structure" that SOA imposes by definition is the loose coupling between the user and the service.

### 2008-03-17 trunk/SAiP/Module1/H3/doc/architecturaldescription.tex
Should -> Must

@@
 
-The domain model accesses the database through the database facade
+The domain model accesses the database through the database fa\c ade
 when needed. This can be expressed in another sequence diagram.
-Sequence diagrams showing the checkout including payment should
+Sequence diagrams showing the checkout including payment must
 also be made.
@@
 
-Working as an developer/tester/engineer means that everything must be quantifiable in some way, and box-and-line drawings do not offer this unambigious quantifiability.
-For this a clearly defined modeling language, without ambiguity as to the meaning of the model, is needed.
-An example of such a language is Unified Modelling Language(UML), which would allow the box-and-line drawing to be represented as one or more diagrams solving all the ambiguaty of the box-and-line drawing with only a single requirement; one must know the language, which is indeed one of the reasons while box-and-line drawings are still widely used - there is no single unambigious modelling language known by all stakeholders, and when that is the case it may be more confusing then benificial to spend time making sure the diagram is clear and unambigious, if that clarity is derived from symbolic notation that the readers (stakeholders) do not understand.
+Working as an developer/tester/engineer means that everything must be quantifiable in some way, and box-and-line drawings do not offer this unambiguous quantifiable.
+For this a clearly defined modelling language, without ambiguity as to the meaning of the model, is needed.
+An example of such a language is Unified Modelling Language(UML), which would allow the box-and-line drawing to be represented as one or more diagrams solving all the ambiguity of the box-and-line drawing with only a single requirement; one must know the language, which is indeed one of the reasons while box-and-line drawings are still widely used - there is no single unambiguous modelling language known by all stakeholders, and when that is the case it may be more confusing then beneficial to spend time making sure the diagram is clear and unambiguous, if that clarity is derived from symbolic notation that the readers (stakeholders) do not understand.
 
 \subsection{Architectural Description - Question 4}
-
+\label{sec:ad_q4}
 \begin{question}

### 2008-03-17 trunk/SAiP/Module1/H3/doc/productlines.tex
Initial answer to question 5.1

@@
 \end{question}
+
+When considering a product line approach for our Beer Web Store, we first need to identify the scope of our product line.
+In this case the product line is defined to have a narrow scope.
+The product line supports a web enabled store that sells quantifiable products (i.e.
+not services).
+The products may however have different details associated, e.g.
+beers have a alcohol volume, stamps do not.
+The web stores must support different user interfaces (but all web based) and online payment methods.
+
+All these variations and commonalities needs to be identified.
+The variations needs to be seperated so that they either can be created specifically for the actual web store, or so that a common component can be configured to handle the specific product.
+
+The following variation points can be identified from the description above:
+
+\begin{description}
+	\item[Products:] The products of the web stores are the main variation point when creating this product line.
+Some product details will be relevant to all stores, e.g.
+a product price, and others will be specific to the type of product sold, e.g.
+alcohol volume as stated above.
+	\item[User interface:] It must be possible to vary the user interface layout with regard to colors and structure of the web page in order to sell the specific product the best possible way.
+Futhermore the user interface must be decoupled from the handling of products in order to support several types of web interfaces such as regular browsers and mobile devices.
+	\item[Payment form:] Online payment methods must be supported, but all stores may not want to support all types of credit cards.
+The supported payment forms must therefore be a variation point in the product line.
+\end{description}
 

### 2008-03-17 SAiP/Module1/H3/doc/softwarearchitecture.tex
Adding corrections

@@
 \begin{description}
-    \item[a. Choose the architectual drivers]
-  Here the quality attributes of the system would be defined by the stakeholders of the system and then a vote would be used to determine which of the quality attributes are architectual drivers.
+    \item[a. Choose the architectural drivers]
+  Here the quality attributes of the system would be defined by the stakeholders of the system and then a vote would be used to determine which of the quality attributes are architectural drivers.
 The quality attributes would be expressed as quality attribute scenarios.
-    \item[b. Choose an architectual pattern]
-  Based on the architectual drivers selected in 1, use the mapping in Bass et al.
-as well as other literature and experience to determine some tactics for achieving the architectual drivers for the system.
-Then select some architectual patterns to realize the tactics.
+    \item[b. Choose an architectural pattern]
+  Based on the architectural drivers selected in a, use the mapping in Bass et al.
+as well as other literature and experience to determine some tactics for achieving the architectural drivers for the system.
+Then select some architectural patterns to realize the tactics.
     \item[c. Instantiate modules using multiple views]
@@
     \item[e. Verify scenarios and use cases]
-  Make sure that the steps a - d covered all the parts of the scenarios. If not include the missing parts and updates the scenarios/architecture is inconsistencies are found.
+  Make sure that the steps a - d covered all the parts of the scenarios. If not, include the missing parts and updates the scenarios/architecture if inconsistencies are found.
 \end{description}
 \subsubsection{Repeat the steps above for every module that needs further decomposition}
-The first iteration only look at the overall architecture (the leation and communication between layers).
+The first iteration only look at the overall architecture (the location and communication between layers).
 The following iterations will decompose the individual layers and as needed the individual components in the layers.
@@
 \subsubsection{Relations}
-The figure xx shows some of the relations, namely the statical, yet it is also important to know the dynamical behaviour.
+The figure \ref{fig:module_view} shows some of the relations, namely the statical, yet it is also important to know the dynamical behaviour.
 These may be expressed by e.g.
-a sequence diagram, as shown in Figure yy.
-In these it may be seen that the Client and Server has a week semantic relation through the http protocol, and the Server and Database has a strong syntactic relation from a compiled compatibility.
+a sequence diagram, as shown in Figure \ref{fig:sequence}.
+In this it may be seen that the Client and Server has a week semantic relation through the http protocol, and the Server and Database has a strong syntactic relation from a compiled compatibility.
 
@@
 Common for all of them is that they present a certain view of the system.
-Very rarely is a single structure sufficient to sufficiently express the architecture of a system.
+Very rarely is a single structure sufficient to express the architecture of a system.
 For this reason many structures are combined, each structure being represented by one or more views, to create a sufficient representation of the software system.
@@
 
-Perry and Wolf suggest seperating the model of the architecture into three components, as mentioned in the question.
-The components relate in the sense that the \emph{elements} has a particular \emph{form} and them and their relations are chosen based on a \emph{rational}.
+Perry and Wolf suggest separating the model of the architecture into three components, as mentioned in the question.
+The components relate in the sense that the \emph{elements} has a particular \emph{form} and the elements and their relations are chosen based on a \emph{rational}.
 
-There are three class of architectual elements defined.
+There are three class of architectural elements defined.
 \begin{itemize}
@@
 
-The architectual form describe the architectual properties and the relationships and their relative importance. This is important as some architectual properties
-may be vital to the design and others may be gold-plating.
-These properties are used to put constraints on the elements and also define the relative importanse of these constraints.
+The architectural form describe the architectural properties and the relationships and their relative importance.
+This is important as some architectural properties may be vital to the design and others may be gold-plating.
+These properties are used to put constraints on the elements and also define the relative importance of these constraints.
 The relationships are similar to the properties except they are used to put constraints and relative priority on the relations between elements.
 
-A software architecture is a selection and description of architectual choices.
-These choices are, however, difficult to convay without including the reason for the choices made, for the elemenents and form not chosen and for choices not made due to lack of information or because they were deemed irrelevant.
+A software architecture is a selection and description of architectural choices.
+These choices are, however, difficult to convey without including the reason for the choices made, for the elements and form not chosen and for choices not made due to lack of information or because they were deemed irrelevant.
+This information is kept in the Rational.
 
@@
 	    \item[HTTP] Protocol for connecting the server and client (loose coupling).
-	    \item[Database facade methods] Protocol for connecting the server to the database (medium-loose coupling).
+	    \item[Database fa\c cade methods] Protocol for connecting the server to the database (medium-loose coupling).
 	 \end{itemize}
@@
 \begin{itemize}
-    \item[Properties] The Server and the database, hidden behind the Facade, is the backbone of the Beer Web Store.
+    \item[Properties] The Server and the database, hidden behind the Fa\c cade, is the backbone of the Beer Web Store.
 If either is unavailable then the other is irrelevant.
-Multiple instances of both Server and Database may be running simultaniously in order to supply hot-swapping.
+Multiple instances of both Server and Database may be running simultaneously in order to supply hot-swapping.
 The importance of up-time (availability), security (payment), usability (transmission abilities), etc.
 must be defined here.
-    \item[Relationship] The Client may only see the server through its HTTP interface, and the Server may only access the database through the Database Facade.
-Other demands to the relationship (certain commands to the database may only be accessed when using https).
+    \item[Relationship] The Client may only see the server through its
+ HTTP interface, and the Server may only access the database through the Database Fa\c cade. Other demands to the relationship (certain commands to the database may only be accessed when using https).
 \end{itemize}

### 2008-03-17 trunk/SAiP/Module1/H3/doc/architecturaldescription.tex
Fixing c in facade

@@
 the Overview page, the basket and the domain model. The domain model
-depends on a Database package that contains the database fa\c ade and the
-database fa\c ade uses some database package library, e.g. MySQL.
+depends on a Database package that contains the database fa\c cade and the
+database fa\c cade uses some database package library, e.g. MySQL.
 
@@
 
-The domain model accesses the database through the database fa\c ade
+The domain model accesses the database through the database fa\c cade
 when needed. This can be expressed in another sequence diagram.
@@
 
-The application code (Beer Web) and the database fa\c ade are
+The application code (Beer Web) and the database fa\c cade are
 both deployed on the web server.

### 2008-03-17 trunk/SAiP/Module1/H3/doc/architecturaldesign.tex
Fixing \c in Facade

@@
 
-\fixme{Answer question 4.3}
+There are four major structures in Component-and-Connector:
+
+\begin{itemize}
+  \item Process or communicating processes.
+	\begin{itemize}
+	  \item This structure shows the execution path through a system, and is used to give the engineer an understand about the dynamic behaviour of the system.
+These structures may help engineers increase Performance and Availability.
+	\end{itemize}
+  \item Concurrency.
+	\begin{itemize}
+	  \item This structure is used to see the parallelism in the system, or part of the system, and when a system, or part of a system, is build around this structure it promotes performance. 
+	\end{itemize}
+  \item Shared data or Repository
+	\begin{itemize}
+	  \item This is a very good structure to use to illustrate how the data is processed and consumed to/from the database (Server to/from Database Fa\c cade).
+When a system, or part of a system, is build around this structure it promotes performance and data integrity.
+	\end{itemize}
+  \item Client-server
+	\begin{itemize}
+  		\item This is a very good structure to see the protocols used between the Clients and the Server, and the messages they communicate.
+When a system, or part of a system, is build around this structure it promotes separation of concern (i.e.
+Modifiability) and allows for Availability tactics like Load balancing.
+	\end{itemize}
+\end{itemize}
+
+In the Beer Web Store has two clear C\&C structures, the Shared data structure between the Server and the Database Fa\c cade and the Client-server between the Client and Server (see Figure \ref{fig:module_view}).
+
+The Process or Communicating processes is a more fine-grained structure, and an example of this may be seen in Figure \ref{fig:ad_sequence}.
+
+The Concurrency structure is not used in the overall elements, yet may be employed in a given element to achieve more performance.
+
+\fixme{Possible extension - insert diagrams}

### 2008-03-17 trunk/SAiP/Module1/H3/doc/qualityattributes.tex
Fixing \c in Facade

@@
 
-Instead of creating a web-server which access a database through a specific fa\c ade, one may argue whether the desired functionality may not be generalised.
+Instead of creating a web-server which access a database through a specific fa\c cade, one may argue whether the desired functionality may not be generalised.
 "Look-up in database".

### 2008-03-17 SAiP/Module1/H3/doc/softwarearchitecture.tex
Correktur reading

@@
 
-Finally it is important to describe the decisions made when defining the form - why is availability important? Why is it OK to sacrifice availability for some (security) functionality.
+Finally it is important to describe the decisions made when defining the form - why is availability important? Why is it OK to sacrifice availability for some (security) functionality?
 
@@
 
-Unfortunately defining the architecture fully up front only makes sense if there is enough information in advance to create the architecture form.
-Very often there are "dark horses", things that do not become apparent until the design and implementation phase.
+Unfortunately defining the architecture fully up front only makes sense if there is enough information in advance to create the architecture from.
+Very often there are "dark horses"; things that do not become apparent until the design and implementation phase.
 If it is not believed that sufficient information exist to fully define the architecture, then doing so may not only be a waste, it may lead to the very dangerous loose-loose principle, where e.g.
@@
 Another requirement to defining the architecture fully up front has to do with the organization.
-It may be good and dandy to create a fully defined architecture with focus on parallel implementation, if there is only one in-house developer to implement the product.
-Based on this the Beer Web Store architect must have decided that there are sufficient information to fully define the architecture and also that there is a reason (parallel implementation, out-sourcing of the design/implementation, etc.).
+It is a waste to create a fully defined architecture with focus on parallel implementation, if there is only one in-house developer to implement the product.
 
-Assuming that the architect in question has followed the course Software Architecture in Practice, he has obviously looked at the architectural properties, the stake holders, the organization and the requirement and deemed that there is sufficient grounds to fully specify the architecture in advance, with the risks and possible gains that it has.
+Assuming that the architect in question has followed the course Software Architecture in Practice, he has obviously looked at the architectural properties, the stake holders, the organization and the requirement and deemed that there is sufficient grounds to fully specify the architecture in advance, with the risks and possible gains that it has (parallel implementation,   out-sourcing of the design/implementation, etc.).

### 2008-03-17 trunk/SAiP/Module1/H3/doc/architecturaldescription.tex
correction reading

@@
 to usage. One perspective of the uses structure is that the
-elements in the figure are ``Overview'', ``Basket'' and ``Data''
-which are all related and the database which is only related
+elements in the figure are ``Overview'', ``Basket'' and ``Data'',
+which are all related and the database, which is only related
 to ``Data''. This shows that the overview page and the basket
-has been decoupled from the database which therefore can be
+has been decoupled from the database, which therefore can be
 replaced without affecting them.
@@
 Another perspective is that the users, the web server and the
-database are elements where the relations show that the users
+database, are elements where the relations show that the users
 use the web server and the web server uses the database. This
@@
 The Module viewpoint describes the internals of the web server and how
-the functionality is decomposed into units. The Component \& connector
+the functionality is decomposed into units. The Component \& Connector
 viewpoint describes the runtime relations between the units and the
@@
 
-In the following sections we will give examples on views for each
+In the following sections we will give examples of views for each
 of the three viewpoints.
@@
 
-Working as an developer/tester/engineer means that everything must be quantifiable in some way, and box-and-line drawings do not offer this unambiguous quantifiable.
+Working as a developer/tester/engineer means that everything must be quantifiable in some way, and box-and-line drawings do not offer this unambiguous quantifiability.
 For this a clearly defined modelling language, without ambiguity as to the meaning of the model, is needed.
-An example of such a language is Unified Modelling Language(UML), which would allow the box-and-line drawing to be represented as one or more diagrams solving all the ambiguity of the box-and-line drawing with only a single requirement; one must know the language, which is indeed one of the reasons while box-and-line drawings are still widely used - there is no single unambiguous modelling language known by all stakeholders, and when that is the case it may be more confusing then beneficial to spend time making sure the diagram is clear and unambiguous, if that clarity is derived from symbolic notation that the readers (stakeholders) do not understand.
+An example of such a language is Unified Modelling Language (UML), which would allow the box-and-line drawing to be represented as one or more diagrams solving all the ambiguity of the box-and-line drawing with only a single requirement; one must know the language, which is indeed one of the reasons while box-and-line drawings are still widely used - there is no single unambiguous modelling language known by all stakeholders, and when that is the case it may be more confusing then beneficial to spend time making sure the diagram is clear and unambiguous, if that clarity is derived from symbolic notation that the readers (stakeholders) do not understand.
 

### 2008-03-17 trunk/SAiP/Module1/H3/doc/qualityattributes.tex
Adding correctional reading

@@
 Getting all of these peoples, often conflicting, requirements assembled and, even more importantly, prioritised, can be a difficult task.
-One way of doing so is to define scenarios, like the one mentioned above, and then having the stakeholders vote at the importance of the scenario.
+One way of doing so is to define scenarios, like the one mentioned above, and then having the stakeholders vote at the importance of the scenarios.
 
@@
 Once the scenarios are clarified and voted on, then the top priority scenarios may be formalized and it may be determined which quality attribute(s) they involve.
-For examples sake lets assume that the above scenario was selected as top priority.
+For examples lets assume that the above scenario was selected as top priority.
 
@@
 
-Naturally there are many more scenario required in order to properly cover all aspects of the quality properties.
+Naturally there are many more scenario required in order to properly cover all aspects of the quality attributes.
 
@@
 Since it may be assumed that the supervising process (remotely deployed) (the process monitoring if the server is down) is stable, then a heartbeat is sufficient to ensure availability, meaning that the web-server will issue a heart-beat every e.g.
-1 second to a maintenance process on the same or another machine, and if one (or perhaps more) heartbeats is missed then the maintenance service can act.
+1 second to a maintenance process, and if one (or perhaps more) heartbeats is missed then the maintenance service can act accordingly.
 
 The proper response to a faulty server may be many things, as defined by the tactics group \emph{fault recovery}.
-As the requirement for availability stipulated a three stage fault recovery, the server or servers (as it may be required to have multiple serves to do load-balancing, depending on how many users must be supported) can service the 5 second requirement, and then if an error occur (as detected by the heartbeat) then restart or swap to another server/database in 30 seconds.
-As 30 seconds is insufficient to boot a web server and a database, the Spare tactic would be insufficient.
-At the same time having Active redundancy would be over-kill, as a this kind of immediate switch-over is not required.
-For these reasons Passive Redundancy is sufficient.
+As the requirement for availability stipulated a three stage fault recovery, the server or servers (as it may be required to have multiple serves to do load-balancing, depending on how many users must be supported) can service the 5 second requirement, and then if an error occur (as detected by the heartbeat), a restart or swap to another server/database can be performed in 30 seconds.
+As 30 seconds is insufficient to boot a web server and a database, the \emph{Spare} tactic would be insufficient.
+At the same time having \emph{Active redundancy} would be over-kill, as this kind of immediate switch-over is not required.
+For these reasons \emph{Passive Redundancy} is sufficient.
 
-As it is an internet system, it will also be sufficient to simply restart the session, and no \emph{Shadow}, \emph{State resynchronization} or \emph{Rollback} is needed in the main web server.
+As it is an Internet system, it will also be sufficient to simply restart the session, and no \emph{Shadow}, \emph{State resynchronization} or \emph{Rollback} is needed in the main web server.
 The user may experience being thrown back to the main page, yet this is deemed acceptable, under availability and performance, yet it may not be in Usability.
 During a payment transaction, however, it must be ensured that the purchase is either completed or not-completed - halfway completed is not acceptable.
-for this it will be beneficial to use the \emph{Rollback} tactic on the \emph{Basket} component.
+For this it will be beneficial to use the \emph{Rollback} tactic on the \emph{Basket} component.
 
-Finally a web-server may be deployed multiple times on teh same machine, allowing for the recovery or prevention of the lesser faults, by simply re-spawning the process that has become non-responsive, utilizing the \emph{Process Monitor} tactic.
+Finally a web-server may be deployed multiple times on the same machine, allowing for the recovery or prevention of the lesser faults, by simply re-spawning the process that has become non-responsive, utilizing the \emph{Process Monitor} tactic.
 This is e.g.
@@
 
-As it is not only the web server which may become non-responsive, it is also the database, it is also required to monitor the database.
-However, as a non-responsive data-base may be discovered by a non-responsive server, it is possible to use the ping-echo
- tactic with the database, and only perform the ping-echo if the server becomes non-responsive, to find out how much must be restarted (only the server or both the server and the database). 
+As it is not only the web server which may become non-responsive, it is also required to monitor the database.
+However, as a non-responsive data-base may be discovered by a non-responsive server, it is possible to use the \emph{ping-echo} tactic with the database, and only perform the ping-echo if the server becomes non-responsive, to find out how much must be restarted (only the server or both the server and the database).
 Depending on the binding between the database and the server, it may also be possible to use the exceptions tactic between the server and the database, so the server itself can restart the database, if an exception occur in a call to the database.
@@
 
-As we have all experience the irritation of a slow web-site, and as the competition on the internet is growing, it is important to achieve popper performance.
+As we have all experience the irritation of a slow web-site, and as the competition on the internet is growing, it is important to achieve propper performance.
 
-Due to the lowing price of hardware and the increasing price of development time, it is often cheapest and certainly simper to use the tactic \emph{Increase available resource}; introduce load balancing with multiple parallel servers thereby increase performance during heavy-load situations, ensure that the serves have sufficient network bandwidth, RAM and CPU strength to handle many simultanious database look-ups - here the tactic of \emph{Maintain multiple copies of data or computations} may be used to speed up database-look-up by caching the most popular searches, or cache the entire result in preparation for a search refinement.
+Due to the decreasing price of hardware and the increasing price of development time, it is often cheapest and certainly simpler, to use the tactic \emph{Increase available resource}; introduce load balancing with multiple parallel servers thereby increase performance during heavy-load situations, ensure that the serves have sufficient network bandwidth, RAM and CPU strength to handle many simultanious database look-ups - here the tactic of \emph{Maintain multiple copies of data or computations} may be used to speed up database-look-up by caching the most popular searches, or cache the entire result in preparation for a search refinement.
 If the database is remotely located the caching of results becomes even more important.

### 2008-03-17 trunk/SAiP/Module1/H3/doc/softwarearchitecture.tex
Figure made smaller and clearpage removed.

@@
     \item[b. Choose an architectural pattern]
-  Based on the architectural drivers selected in 1, use the mapping in Bass et al.
+  Based on the architectural drivers selected in a, use the mapping in Bass et al.
 as well as other literature and experience to determine some tactics for achieving the architectural drivers for the system.
@@
     \item[e. Verify scenarios and use cases]
-  Make sure that the steps a - d covered all the parts of the scenarios. If not include the missing parts and updates the scenarios/architecture is inconsistencies are found.
+  Make sure that the steps a - d covered all the parts of the scenarios. If not, include the missing parts and updates the scenarios/architecture if inconsistencies are found.
 \end{description}
@@
 \end{itemize}
-The above is quickly sketched in the box-and-line drawing below, where the layering may also be seen. 
+The above is quickly sketched in the box-and-line drawing below in figure \ref{fig:module_view}, where the layering may also be seen. 
 \begin{figure}[!htb]
-\centerline{\epsfig{figure=figures/module_view,scale=1.0 }}
+\centerline{\epsfig{figure=figures/module_view,scale=0.9 }}
 \caption{Element overview}
@@
 \end{figure}
-\clearpage
 
 \subsubsection{Relations}
-The figure \ref{fig:module_view} shows some of the relations, namely the statical, yet it is also important to know the dynamical behaviour.
+Figure \ref{fig:module_view} shows some of the relations, namely the statical, yet it is also important to know the dynamical behaviour.
 These may be expressed by e.g.
 a sequence diagram, as shown in Figure \ref{fig:sequence}.
-In these it may be seen that the Client and Server has a week semantic relation through the http protocol, and the Server and Database has a strong syntactic relation from a compiled compatibility.
+In this it may be seen that the Client and Server has a week semantic relation through the http protocol, and the Server and Database has a strong syntactic relation from a compiled compatibility.
 
@@
 Perry and Wolf suggest separating the model of the architecture into three components, as mentioned in the question.
-The components relate in the sense that the \emph{elements} has a particular \emph{form} and them and their relations are chosen based on a \emph{rational}.
+The components relate in the sense that the \emph{elements} has a particular \emph{form} and the elements and their relations are chosen based on a \emph{rational}.
 
@@
 
-The architectural form describe the architectural properties and the relationships and their relative importance. This is important as some architectural properties
-may be vital to the design and others may be gold-plating.
+The architectural form describe the architectural properties and the relationships and their relative importance.
+This is important as some architectural properties may be vital to the design and others may be gold-plating.
 These properties are used to put constraints on the elements and also define the relative importance of these constraints.
@@
 These choices are, however, difficult to convey without including the reason for the choices made, for the elements and form not chosen and for choices not made due to lack of information or because they were deemed irrelevant.
+This information is kept in the Rational.
 
@@
 	    \item[HTTP] Protocol for connecting the server and client (loose coupling).
-	    \item[Database fa\c ade methods] Protocol for connecting the server to the database (medium-loose coupling).
+	    \item[Database fa\c cade methods] Protocol for connecting the server to the database (medium-loose coupling).
 	 \end{itemize}
@@
 \begin{itemize}
-    \item[Properties] The Server and the database, hidden behind the Fa\c ade, is the backbone of the Beer Web Store.
+    \item[Properties] The Server and the database, hidden behind the Fa\c cade, is the backbone of the Beer Web Store.
 If either is unavailable then the other is irrelevant.
@@
     \item[Relationship] The Client may only see the server through its
- HTTP interface, and the Server may only access the database through the Database Fa\c ade. Other demands to the relationship (certain commands to the database may only be accessed when using https).
+ HTTP interface, and the Server may only access the database through the Database Fa\c cade. Other demands to the relationship (certain commands to the database may only be accessed when using https).
 \end{itemize}

### 2008-03-17 trunk/SAiP/Module1/H3/doc/qualityattributes.tex
Table texts removed to preserve space.

@@
 
-The following scenarios (table \ref{tab:availability} and \ref{tab:performance}) is a possible outcome of prioritization and clarification of quality attribute scenarios as described above:
+The following scenarios are a possible outcome of prioritization and clarification of quality attribute scenarios as described above:
 
@@
 \end{tabular}
-\caption{Availability scenario for the Beer Web Store}
-\label{tab:availability}
 \end{center}
@@
 \end{tabular}
-\caption{Performance scenario for the Beer Web Store}
-\label{tab:performance}
 \end{center}

### 2008-03-17 SAiP/Module1/H3/doc/productlines.tex
Question 5.1 finished.

--- 2008-03-01 SAiP/Module1/H3/doc/productlines.tex
@@
 
-\fixme{Answer question 5.1}
+When considering a product line approach for our Beer Web Store, we first need to identify the scope of our product line.
+In this case the product line is defined to have a narrow scope.
+The product line supports a web enabled store that sells quantifiable products (i.e.
+not services).
+The products may however have different details associated, e.g.
+beers have a alcohol volume, stamps do not.
+The web stores must support different user interfaces (but all web based) and on line payment methods.
+
+All these variations and commonalities needs to be identified.
+The variations needs to be separated so that they either can be created specifically for the actual web store, or so that a common component can be configured to handle the specific product.
+
+The following variation points can be identified from the description above:
+
+\begin{description}
+	\item[Products:] The products of the web stores are the main variation point when creating this product line.
+Some product details will be relevant to all stores, e.g.
+a product price, and others will be specific to the type of product sold, e.g.
+alcohol volume as stated above.
+	\item[User interface:] It must be possible to vary the user interface layout with regard to colors and structure of the web page in order to sell the specific product the best possible way.
+Furthermore the user interface must be decoupled from the handling of products in order to support several types of web interfaces such as regular browsers and mobile devices.
+	\item[Payment form:] On line payment methods must be supported, but all stores may not want to support all types of credit cards.
+The supported payment forms must therefore be a variation point in the product line.
+\end{description}
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure=figures/pl_package_overview,scale=0.6 }}
+\caption{Package diagram for the web store product line}
+\label{fig:pl_package_overview}
+\end{figure}
+
+Figure \ref{fig:pl_package_overview} shows an example of a package diagram for a web
+store product line. It shows that the user interface and the product information has
+been separated from the product handling and put in separate packages. The user
+interface may be entirely specific to each web store or the package may contain
+predefined user interfaces that can be configured e.g. using style sheets. A reactive
+approach can be used to develop more web interfaces when the need for more features
+arises. The user interface depends on the product handling package that can handle
+all kinds of products. The products package can provide an interface that the user
+interfaces and product handling module can depend on. There need to be some way to
+define product specific information that can be retrieved from and searched for in
+the database and presented on the user interface. This concept is not defined any
+further in this report.
+
+The dependencies to from the user interface and product handling to the products package can e.g. be solved at compile time.
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure=figures/pl_package_ph,scale=0.6 }}
+\caption{Decomposition of the Product Handling Package for the web store product line}
+\label{fig:pl_package_ph}
+\end{figure}
+
+In figure \ref{fig:pl_package_ph} the Product Handling package has been decomposed
+into smaller components. The ``Basket'', ``Overview'' and ``Domain Model'' from the
+original Beer Web Store has been generalized so that they do not contain any product
+specific handling. A ``Payment'' package has been created that shall support all
+needed payment methods. When need for a new payment method arises, this must be added
+to the package. The supported payment methods for a specific web store must be
+configured, e.g. at compile time.
 
 \subsection{Product Lines - Question 2}
-
+\label{sec:pl_q2}
 \begin{question}
@@
 
-\fixme{Answer question 5.2}
+Creating an architecture product line only makes sense if it is believed that it may be used in other products, and the number of products in the pipeline and how well their commonalities may be predicted, helps to determine which approach to use.
+Naturally the situation is not black and white, as there are many possibilities between the fully evolutionary architecture to the fully initially defined, yet for the purpose of debating advantages and disadvantages the simple black and white is simply easier.
+
+Defining the product line architecture fully up front means a big investment in time and money in order to create the architecture and the first product based on it, so if it is the only product made, then it will go way over budget both in time and money.
+The idea is that creating the next product will be significantly cheaper, because a very large part of the products functionality is covered by the product line and very little (if any) redesign has to be done.
+As more and more products are created based on the product line their individual budget savings will help pay for the initial investment, eventually leading to a profit compared to building the products from scratch each time.
+
+Letting the product line evolve means that only the basics are defined up front, only causing a small initial investment, before building the first product.
+When the second product is to be created more of the product line is defined based on the parts of the first product, which may be used in the second.
+Naturally the second product is faster to build than the first, due to reuse, yet not as fast as the second product when the entire product line was defined in advance.
+As more products come along the amount of reuse grow, and would eventually reach the same as the one where everything was defined in advance.
+
+The above two approaches are defined as Heavyweight and Lightweight, respectively.
+A statistical analysis of these two approaches, compared to simply building the products from scratch, can be shown in Figure \ref{fig:product-line}.
+
+\begin{figure}[!htb]
+\centerline{\epsfig{figure=figures/product-line,scale=0.9 }}
+\caption{Heavy-weight vs. Light-weight}
+\label{fig:product-line}
+\end{figure}
+
+Deciding which approach to use, is more than just a matter of choosing between Heavyweight and Lightweight (or fully defined or evolutionary) based on the number of expected products, it is also a matter of how well the common architecture of the future products can be predicted, for if the architecture needs to be changed from product to product then all the advantage is lost.
+
+Based on this the Beer Web Store architect must have decided that the products in the pipe-line and the Beer Web Store share a common architecture and that the future products are so well know that the architecture can be fully specified in advance.
 
@@
 
-\fixme{Answer question 5.3}
+As mentioned in Section \ref{sec:qa_q4} part of SOA is about generalization, just like product lines are.
+Taking the example given above, having the architecture support mobile phones (i.e.
+WAB) means that the server must be able to supply a scaled down result of a search as well as the fully blown.
+The server must thereby handle different types of clients.
+
+The requirement that other product's such as other stamps or soda-pop should be supported, means that the server's domain model and database must support general elements as opposed to just beer elements.
+
+This generalization fits well with SOA, as it is not that clever to make a ``Beer purchasing Service'', but far better to create a ``Purchasing service'', which operate on elements.
+This will allow any company, who has products they wish to sell, to use this service.
+This is the same generalization requirement that apply to the Product line architecture, where the architecture must handle the purchase of an arbitrary product, be it stamps, soda-pop or beer.
+It is possible that the architecture requires that it is a quantifiable product, yet this is not an unlikely requirement for a service either.
+
+It may now seem like SOA is the answer to creating a product line, yet this is not the case.
+It is very possible to create very specific services, like the ``Beer look-up service'', and this may be just fine for a Beer Web Store based on SOA, yet this is not a product line.
+The same architectural consideration must be done when creating the elements of a product line and define their interrelations, must also be done when defining the services and their interrelations.
+The only ``structure'' that SOA imposes by definition is the loose coupling between the user and the service.

### 2008-03-17 trunk/SAiP/Module1/H3/doc/softwarearchitecture.tex
Last corrections.

@@
     \item[b. Choose an architectural pattern]
-  Based on the architectural drivers selected in a, use the mapping in Bass et al.
-as well as other literature and experience to determine some tactics for achieving the architectural drivers for the system.
+  Based on the architectural drivers selected in \emph{a}, use the mapping in \cite{bass2003sa} as well as other literature and experience to determine some tactics for achieving the architectural drivers for the system.
 Then select some architectural patterns to realize the tactics.
@@
 \subsubsection{Elements}
-Based on the present decomposition the elements that are know are the layers, where each layer is an element as described below.
+Based on the present decomposition the elements that are known are the layers, where each layer is an element as described below.
 \begin{itemize}
@@
 \subsubsection{Relations}
-Figure \ref{fig:module_view} shows some of the relations, namely the statical, yet it is also important to know the dynamical behaviour.
+Figure \ref{fig:module_view} shows some of the relations, namely the statical, yet it is also important to know the dynamical behavior.
 These may be expressed by e.g.
 a sequence diagram, as shown in Figure \ref{fig:sequence}.
-In this it may be seen that the Client and Server has a week semantic relation through the http protocol, and the Server and Database has a strong syntactic relation from a compiled compatibility.
+In this it may be seen that the Client and Server has a weak semantic relation through the http protocol, and the Server and Database has a strong syntactic relation from a compiled compatibility.
 
@@
 
-There are three class of architectural elements defined.
+There are three classes of architectural elements defined.
 \begin{itemize}
@@
 ``computing'' are removed from the definition of 'software
-architecture' in [Bass et al., 2003]
+architecture' in \cite{bass2003sa}
 \end{question}
@@
 As an example could be used the construction of a rifle.
-Describing the elements that goes into making a rifle (hammer, barrol, ...), the structures that comes from putting them together, and the individual relationships that must be met (form fit, etc.).
+Describing the elements that goes into making a rifle (hammer, barrel, ...), the structures that comes from putting them together, and the individual relationships that must be met (form fit, etc.).
 The elements may fit together in different ways and multiple structures may be required to complete the description.
@@
 
-The description of an architecture, as defined by Bass et.
-al.
-is therefore very general and apart from the use of words like "software" and "computer", may apply to any architecture, be it a building, a production line for a rifle, or a embedded software system.
+The description of an architecture, as defined by \cite{bass2003sa} is therefore very general and apart from the use of words like "software" and "computer", may apply to any architecture, be it a building, a production line for a rifle, or a embedded software system.
 

### 2008-03-17 SAiP/Module1/H3/doc/architecturaldesign.tex
Last corrections.

@@
 \section{Architectural Design -- Beer Web Store}
-Consider the following main use case for a Beer Web Store for speciality
+Consider the following main use case for a Beer Web Store for specialty
 beers:
@@
 
-When a system is Strictly Layered it gives rise to high \emph{Portability} (a for of \emph{Modifiability}), as it is possible to abstract all dependencies of a given layer by simply abstracting the layer directly beneath it, and if anything in a layer changes then only the layer directly above it must be updated.
+When a system is Strictly Layered it gives rise to high \emph{Portability} (a form of \emph{Modifiability}), as it is possible to abstract all dependencies of a given layer by simply abstracting the layer directly beneath it, and if anything in a layer changes then only the layer directly above it must be updated.
 
-This also enables simpler \emph{Testability}, as the introduction of test doubles (fake objects, stubs, spys and/or mocks) must only be performed on the in a single layer; the one directly below the one under test.
+This also enables simpler \emph{Testability}, as the introduction of test doubles (fake objects, stubs, spys and/or mocks) must only be performed in a single layer; the one directly below the layer under test.
 
@@
 	\begin{itemize}
-	  \item This structure shows the execution path through a system, and is used to give the engineer an understand about the dynamic behaviour of the system.
+	  \item This structure shows the execution path through a system, and is used to give the engineer an understanding about the dynamic behavior of the system.
 These structures may help engineers increase Performance and Availability.
@@
 	\begin{itemize}
-	  \item This is a very good structure to use to illustrate how the data is processed and consumed to/from the database (Server to/from Database Fa\c ade).
+	  \item This is a very good structure to use to illustrate how the data is processed and consumed to/from the database (Server to/from Database Fa\c cade).
 When a system, or part of a system, is build around this structure it promotes performance and data integrity.
@@
 
-In the Beer Web Store has two clear C\&C structures, the Shared data structure between the Server and the Database Fa\c ade and the Client-server between the Client and Server (see Figure \ref{fig:module_view}).
+The Beer Web Store has two clear C\&C structures, the Shared data structure between the Server and the Database Fa\c cade and the Client-server between the Client and Server (see Figure \ref{fig:module_view}).
 
@@
 The Concurrency structure is not used in the overall elements, yet may be employed in a given element to achieve more performance.
-
-\fixme{Possible extension - insert diagrams}

### 2008-03-17 SAiP/Module1/H3/doc/productlines.tex
Last corrections.

@@
 The above two approaches are defined as Heavyweight and Lightweight, respectively.
-A statistical analysis of these two approaches, compared to simply building the products from scratch, can be shown in Figure \ref{fig:product-line}.
+A statistical analysis of these two approaches, compared to simply building the products from scratch, is shown in Figure \ref{fig:product-line}.
 
@@
 \centerline{\epsfig{figure=figures/product-line,scale=0.9 }}
-\caption{Heavy-weight vs. Light-weight}
+\caption{Heavy-weight vs. Light-weight \cite{mcgregor2002}}
 \label{fig:product-line}
@@
 
-Based on this the Beer Web Store architect must have decided that the products in the pipe-line and the Beer Web Store share a common architecture and that the future products are so well know that the architecture can be fully specified in advance.
+Based on this the Beer Web Store architect must have decided that the products in the pipe-line and the Beer Web Store share a common architecture and that the future products are so well known that the architecture can be fully specified in advance.
 
@@
 It is very possible to create very specific services, like the ``Beer look-up service'', and this may be just fine for a Beer Web Store based on SOA, yet this is not a product line.
-The same architectural consideration must be done when creating the elements of a product line and define their interrelations, must also be done when defining the services and their interrelations.
+The same architectural consideration must be done when creating the elements of a product line and define their interrelations, as when defining the services and their interrelations.
 The only ``structure'' that SOA imposes by definition is the loose coupling between the user and the service.

### 2008-03-17 SAiP/Module1/H3/doc/qualityattributes.tex
Adding label
Last corrections.

@@
 Once the scenarios are clarified and voted on, then the top priority scenarios may be formalized and it may be determined which quality attribute(s) they involve.
-For examples sake lets asume that the above scenario was selected as top priority.
+For examples lets assume that the above scenario was selected as top priority.
 
-As it may be seen, the scenario(s) focus primarily on \emph{Availability} (how rare should it be that a user experience non-resonsiveness and how many simultanious users can the system support), \emph{Usability} (how easy should it be for the user to cancel a transaction halfway through and how easy should it be to pick up a previous transaction), \emph{Performance} (how big a delay can the user expect when performing an operation like a lookup in the database), \emph{Security} (how to avoid someone using stolen credit cards and also prevent a "hacker" from stealing the credit card information either during the transaction or from the Beer Web Store's database.
+As it may be seen, the scenario(s) focus primarily on \emph{Availability} (how rare should it be that a user experience non-responsiveness and how many simultaneous users can the system support), \emph{Usability} (how easy should it be for the user to cancel a transaction halfway through and how easy should it be to pick up a previous transaction), \emph{Performance} (how big a delay can the user expect when performing an operation like a lookup in the database), \emph{Security} (how to avoid someone using stolen credit cards and also prevent a "hacker" from stealing the credit card information either during the transaction or from the Beer Web Store's database.
 
@@
 
-Finding out if a resource is down may be done using one of the fault detection tactice; ping/echo, heartbeat or exception.
-Due to the loose coupling between the client (browser) and the server, the exception solution is not benificial.
+Finding out if a resource is down may be done using one of the fault detection tactics; ping/echo, heartbeat or exception.
+Due to the loose coupling between the client (browser) and the server, the exception solution is not beneficial.
 Since it may be assumed that the supervising process (remotely deployed) (the process monitoring if the server is down) is stable, then a heartbeat is sufficient to ensure availability, meaning that the web-server will issue a heart-beat every e.g.
-1 second to a maintance process on the same or another machine, and if one (or perhaps more) heartbeats is missed then the maintanance service can act.
+1 second to a maintenance process, and if one (or perhaps more) heartbeats is missed then the maintenance service can act accordingly.
 
 The proper response to a faulty server may be many things, as defined by the tactics group \emph{fault recovery}.
-As the requirement for availability stipulated a three stage fault recovery, the server or servers (as it may be required to have multiple serves to do load-balancing, depending on how many users must be supported) can service the 5 second requirement, and tehn if an error occur (as detected by the heartbeat) then restart or swap to another server/database in 30 seconds.
-As 30 seconds is insufficient to bott a webserver and a database, the Spare tactic would be insufficient.
-At the same tiem having Active redundancy would be over-kill, as a this kind of immidiat switch-over is not required.
-For these reasons Passive Redundancy is sufficient.
+As the requirement for availability stipulated a three stage fault recovery, the server or servers (as it may be required to have multiple serves to do load-balancing, depending on how many users must be supported) can service the 5 second requirement, and then if an error occur (as detected by the heartbeat), a restart or swap to another server/database can be performed in 30 seconds.
+As 30 seconds is insufficient to boot a web server and a database, the \emph{Spare} tactic would be insufficient.
+At the same time having \emph{Active redundancy} would be over-kill, as this kind of immediate switch-over is not required.
+For these reasons \emph{Passive Redundancy} is sufficient.
 
-As it is an internet system, it will also be sufficient to simply restart the session, and no \emph{Shadow}, \emph{State resynchronization} or \emph{Rollback} is needed in the main web server.
-The user may experience beeing thrown back to the main page, yet this is deemed acceptable, under availability and performance, yet it may not be in Usability.
+As it is an Internet system, it will also be sufficient to simply restart the session, and no \emph{Shadow}, \emph{State resynchronization} or \emph{Rollback} is needed in the main web server.
+The user may experience being thrown back to the main page, yet this is deemed acceptable, under availability and performance, yet it may not be in Usability.
 During a payment transaction, however, it must be ensured that the purchase is either completed or not-completed - halfway completed is not acceptable.
-for this it will be benificial to use the \emph{Rollback} tactic on the \emph{Basket} component.
+For this it will be beneficial to use the \emph{Rollback} tactic on the \emph{Basket} component.
 
-Finally a web-server may be deployed multiple times on teh same machine, allowing for the recovery or prevention of the lesser faults, by simply re-spawning the process that has become non-responsive, utilizing the \emph{Process Monitor} tactic.
+Finally a web-server may be deployed multiple times on the same machine, allowing for the recovery or prevention of the lesser faults, by simply re-spawning the process that has become non-responsive, utilizing the \emph{Process Monitor} tactic.
 This is e.g.
@@
 
-As it is not only the web server which may become non-responsive, it is also the database, it is also required to monitor the database.
-However, as a non-responsive data-base may be discovered by a non-responsive server, it is possible to use the ping-echo tactic with the database, and only perform the ping-echo if the server becomes non-responsive, to find out how much must be restarted (only the server or both the server and the database).
-Depending on the binding between teh database and the server, it may also be possibel to use the exceptions tactic between teh server and the database, so the server itself can restart the database, if an exception occur in a call to the database.
+As it is not only the web server which may become non-responsive, it is also required to monitor the database.
+However, as a non-responsive data-base may be discovered by a non-responsive server, it is possible to use the \emph{ping-echo} tactic with the database, and only perform the ping-echo if the server becomes non-responsive, to find out how much must be restarted (only the server or both the server and the database).
+Depending on the binding between the database and the server, it may also be possible to use the exceptions tactic between the server and the database, so the server itself can restart the database, if an exception occur in a call to the database.
 
@@
 
-Due to the lowing price of hardware and the increasing price of developement time, it is often cheapest and certainly simper to use the tactic \emph{Increase available resourcez}; introduce load balancing with multiple parallel servers thereby increase performance during heavy-load situations, ensure that the serves have sufficient network bandwidth, RAM and CPU strength to handle many simultanious database look-ups - here the tactic of \emph{Maintain multiple copies of data or computations} may be used to speed up database-look-up by caching the most popular searches, or cach the entire result in preperation for a search refinement.
+Due to the decreasing price of hardware and the increasing price of development time, it is often cheapest and certainly simpler, to use the tactic \emph{Increase available resource}; introduce load balancing with multiple parallel servers thereby increase performance during heavy-load situations, ensure that the serves have sufficient network bandwidth, RAM and CPU strength to handle many simultanious database look-ups - here the tactic of \emph{Maintain multiple copies of data or computations} may be used to speed up database-look-up by caching the most popular searches, or cache the entire result in preparation for a search refinement.
 If the database is remotely located the caching of results becomes even more important.
 
-There are many other tactics which may be employed, yet given the nature to the application, it is not believed that the biggest resource consumers are under developement control, for building the web-server from scratch to support better concurrency or using non-standard performance tweaking algorithms for the database-lookup, will most likely become inconcivably expensive compare to simply increasing the computational resources.
+There are many other tactics which may be employed, yet given the nature to the application, it is not believed that the biggest resource consumers are under development control, for building the web-server from scratch to support better concurrency or using non-standard performance tweaking algorithms for the database-lookup, will most likely become inconceivably expensive compare to simply increasing the computational resources.
 
 \subsection{Quality Attributes - Question 4}
-
+\label{sec:qa_q4}
 \begin{question}
@@
 
-As SOA is all about standardised interfaces supplying a functionality which may be used by many different products, the areas where SOA makes sens is naturally between teh components.
+As SOA is all about standardised interfaces supplying a functionality which may be used by many different products, the areas where SOA makes sense is naturally between the components.
 
-Instead of creating a web-server which access a database through a specific facade, one may argue whether the desired functionality may not be generalised.
+Instead of creating a web-server which access a database through a specific fa\c cade, one may argue whether the desired functionality may not be generalised.
 "Look-up in database".
 This may easily be hidden behind a standardised service, e.g.
-a web-service, meaning that anyone on the WWW who supplies a batabase service may be used to host the beer database, and if one becomes dissatisfied with the database provider, it is transparent to simply switch to another.
+a web-service, meaning that anyone on the WWW who supplies a database service may be used to host the beer database, and if one becomes dissatisfied with the database provider, it is transparent to simply switch to another.
 
 Taking this one step further, perhaps the database service may be extended with a data-base type, meaning that the Beer Web Store's domain model is exposed in a service, which would allow for not only displaying ones own brands of beer, but through a discovery service, also beer-brands by other brewers on the net who implement this service.
-This will also point the way for a more standardised lookk-and-feel when shopping for beer.
+This will also point the way for a more standardised look-and-feel when shopping for beer.
 
-Naturally just because a web-service (or other service) is used it does not mean that it has to be public, jet it does have some advabtages.
+Naturally just because a web-service (or other service) is used it does not mean that it has to be public, jet it does have some advantages.
 
-Finally the WS-Security web-service may be used to handle payment, and also possibly, if the database or Beer Web Store Domain model is exposed through a web-service, it may also be benificial that not everyone may have the ability to update the domain model or database, so here the WS-Security service may also come in handy.
+Finally the WS-Security web-service may be used to handle payment, and also possibly, if the database or Beer Web Store Domain model is exposed through a web-service, it may also be beneficial that not everyone may have the ability to update the domain model or database, so here the WS-Security service may also come in handy.
 
-Naturally this generality do not necesarrliy come free.
+Naturally this generality do not necessarily come free.
 A look-up in a database behind a service, especially a web-service, is far slower than a direct look-up in a local database, thereby influencing Performance poorly.
-On the other hand the use of web-services may increase Availability, as the functionality is deployed on multiple servers, thereby redusing the chance of the system becomming fully non-responsive.
+On the other hand the use of web-services may increase Availability, as the functionality is deployed on multiple servers, thereby reducing the chance of the system becoming fully non-responsive.
 Naturally the critical path must be evaluated, as if a link on the critical path chain become non-responsive then the entire system becomes non-responsive.
@@
 The quality attributes are vital to the software architecture, as they not only define the requirements to the software architecture (along with the scenarios), but also describe their relative importance.
-The quality attributes therefore allow the architecture the be evaluated in terms of how well it supports or enables a system's desired external (e.g.
+The quality attributes therefore allow the architecture to be evaluated in terms of how well it supports or enables a system's desired external (e.g.
 availability and performance) and internal behaviour (e.g.
